### 2019-06-06 manuscript/ideas.tex
Inital commit

Typically, microbial community structure-function research questions are framed
as Does microbial composition matter to ecosystem function? The answer to this
question is obviously "yes" given that microorganisms mediate many of the
processes of interest. The question then becomes does microbial community
variation within a functional group or overall matter to ecosystem function? 
Generally it is assumed that microbial composition is functionally irrelevant, 
though studies of environmental treatments, common gardens, and reciprocal
transplants have shown that different community compositions are correlated with
different functions. The challenge with these studies is separating the close
connection between environmental variation, community composition, and ecosystem
function. Microorganisms cannot be useful to ecosystem modeling efforts if there
is no identifiable variation in composition that consistently alters function
across space, time, and environmental gradients. The question "do we need to
include microorganisms in ecosystem models?" depends on identifying variation in
the community associated with function independent of the environment. If that's
true, then the next step in understanding or modeling these processes is 
identify what components of the community alter ecosystem function.

It has been demonstrated that both geographic and ecological distance can lead
to population differentiation. For geography, local genetic
drift and reduced gene flow due to geographic separation can lead to population
differentation (Wright 1943). In addition, ecological distance can lead to both selection
against migrants (Hendry 2004) and local selection on endemic populations to
generate variation. While
these processes have primarily been described for macrobial populations, the
same principles apply to microorganisms (Martiny et al. ).

Genome-wide association studies are used to identify the genetic variation
underlying phenotypic traits in populations of organisms. A classic example is
identifying genomic regions associated with genetic diseases in human
populations. Generally there is a correspondence between disease states, shared
ancestry, and geographic proximity of populations. If the goal were only to
build a model that would
predict the distribution of disease states, then controlling for this covariance
structure would be unnecessary. However, investigators are specifically
interested in identify loci to use as disease indicators or to identify
mechanisms for the genetic basis of disease. Therefore, they use statistical
techniques to control for population stratification.

Outline

1. We don't know whether microbes are important for function
2. BEF and CAFE research has focused on terrestrial plant communities where
richness and composition can alter ecosystem function
3. Microbial communities are studied in different ways and so these approaches
don't apply. Instead, we try to correlate functional gene (lots of assumptions)
or Prokaryote diversity (few assumptions) with function, neither of which work
very well.
4. A more analogous problem is GWAS
5. GWAS studies demonstrate population strucutre can produce false positives
6. Things exhibit isolation by distance (Sewall Wright) including microbes
(Martiny), which produce structure
7. Therefore, we should consider population structure when 
8. Reframing the question

Things I still want to talk about:
- 
- GWAS studies have demonstrated population structure due to historical effects
or selection can produce false positives 
- A more analogous approach

### 2019-06-06 manuscript/main.tex
Inital commit

\documentclass{article} \usepackage[utf8]{inputenc}
%\usepackage[utf8]{inputenc}
%\usepackage{amsmath}
%\usepackage{amsfonts}
%\usepackage{amssymb}
%\usepackage{textalpha}
\usepackage{natbib}

\bibpunct{(}{)}{,}{a}{}{;}

\title{Linking microbial communities to ecosystem functions: what we can learn
from genotype-phenotype mapping in organisms} \author{Andrew H. Morris, Brendan
J. M. Bohannan}

\begin{document}

\maketitle

\section*{Tasks}

Deadline: August 1st

Theme: Conceptual Challenges in Microbial Community Ecology

Journal: Philosophical Transactions B

Max length: 8-9 pages

Data must be submitted as supplementary information in a repository with
accession number, link, or DOI

Supplement: They have encouraged us to include additional information including
data, tables, figures, text, images, videos, or recordings.

What is novel about this? The first anyone has treated an ecosystem like an
organism. The first time anyone has done this correlative approach well
by considering environmental and spatial structure in the data.

\section*{Outline} 
\begin{enumerate} 
  \item Problem statement: Microorganisms putatively mediate biogeochemical
    cycles and yet we have been unable to demonstrate consistent BEF or CAFE
  relationships.   
\item  Opening: Biodiversity ecosystem function research has demonstrated the
  importance of protecting biodiversity to maintain ecosystem function.
  However, more recently the importance of community assembly processes such as
  dispersal have shown that regional diversity is potentially as important as
  local processes. Therefore, metacommunity dynamics should play a role in local
  ecosystem functioning.  
  \item Challenge: Applying CAFE to microbes would be
  convenient, however microbes pose several unique challenges. 1. Microbial
  communities are orders of magnitude more complex than macrobial communities.
  2. We don’t know who performs each function of interest. 3. We cannot
  attribute a portion of the rate of an ecosystem function to a particular
  individual or species like we can in plant communities.  
  \item Solution: The
  microbial community assembly - ecosystem function relationship is more
  analogous to the problem of mapping complex phenotypic traits to genetic
  variation in organisms. In this case we have thousands of genetic loci with
  unknown functions that may contribute positively or negatively (or not at all)
  to a particular trait.  
  \item Studies linking genotype to phenotype have to
  consider similar processes as metacommunities within the
  metapopulation such as confounding geography with genotypic similarity.
  Applying this to microbial metacommunities, associations between microbes and
  function could be due to a number of factors other than the “genetic” effect
  of microbial community structure on ecosystem function. By accounting for the
  covariance structure (space, environment, community interactions) we can
  identify true structure-function connections.  
  \item Approach: Demonstrate the
  previous approaches to microbial structure-function (e.g. abundance-function,
  diversity-function, PCA or indicator species analysis) and demonstrate our new
  approach, which accounts for metacommunity dynamics in local ecosystem
  function. Step 1. Variance partitioning between community, environment, and
  space. Step 2. GWAS-esque association mapping to identify microbes.  
  \item Take home: The way we study the functioning of microbial communities is
  inherently limiting in our ability to apply BEF or CAFE approaches. Instead,
  we need to learn from GWAS to identify robust structure-function
  relationships. What we learn from this is that 1. Contrary to past results,
  microbes do explain a substantial portion of the variation in ecosystem
  function when we account for other sources of variation. 2. Stochastic
  processes can alter the relationship between environment/community and
  function 3.  we cannot assume that we know what part of the community controls
  ecosystem function (e.g. functional genes) i.e. non-intuitive features can
  explain a lot of residual variance.  
  \item Conclusion: This work presents a
  framework for testing the relationship between microbial community structure
  and ecosystem function and demonstrates the importance of understanding
  microbial genetic variation in predicting ecosystem function.  
  \end{enumerate}

\section{Introduction}

Microorganisms mediate many biogeochemical processes, but despite 
much effort to link microbial communities
to these functions, the mechanism by which they regulate these
processes remains unknown. Ecologists have investigated the effect of changing
biodiversity on ecosystem function for decades, but many of these approaches
are not applicable in microbial ecology due to the nature of microbial data. To
address this challenge, we propose integrating current theory on community
assembly and the functioning of ecosystems with lessons from genotype-phenotype
mapping in quantitative
genetics. This will address the
question of how does variation in microbial community structure contribute to
variation in the rate of ecosystem function?

\subsection{Biodiversity-ecosystem function in macrobial communities}

The history of biodiversity-ecosystem function (BEF) research has demonstrated
that local species richness is correlated with ecosystem function
\citep{hooper2005}. Most studies
have tested this idea in terrestrial plant communities where plant productivity
exhibits a positive, asymptotic relationship with species richness
\citep{naeem2003}. There are
two major explanations for this relationship. The first is the sampling effect
where a more diverse community is more likely to have one particularly
productive species \citep{cardinale2006}. The second is niche complementarity where more species are
more likely to occupy separate niches within the same ecosystem resulting in
more efficient resource use and greater function.

More recently, regional species diversity has been shown to be important for
local ecosystem functioning where metacommunity dynamics can play a
role in local BEF relationships \citep{leibold2017}. Typical BEF research focuses on local selection
as a driver of ecosystem function. However, if regional
diversity is important, then dispersal processes could alter local species diversity and composition resulting in
a change in ecosystem function. This model is referred to as Community Assembly
and the Functioning of Ecosystems (CAFE). Community assembly can alter ecosystem
function if limiting dispersal reduces the likelihood of a highly productive
species from dispersing to a local site or if high dispersal of an unproductive
species overwhelms local selection.

The CAFE approach has been demonstrated using the Price equation where a change in ecosystem function between two, matched
observations can be attributed to species additions, species losses, and
changes in abundance of shared species \citep{fox2006,
bannar-martin2018}. To
calculate these vectors, function must be attributed at the species level (e.g.
biomass measurements of individual plant species). In microbial ecology, we
could look at species turnover or changes in relative abundance between matched
sites in space, time, or across treatments. However, we cannot apply the Price
equation, because it is not typically possible to attribute a portion of ecosystem
function to individual microbial taxa. We can only measure microbial ecosystem
functions at the whole-ecosystem scale.

\subsection{How are microbial communities different?}

The trajectory of microbial BEF research has differed from that of macrobial BEF
for three reasons. First, microbial communities are
immensely complex with orders of magnitude more taxa than macrobial communities
and great intra-taxon variation for certain traits. Second, the way we measure
microbes using marker gene sequences does not allow us to attribute portions of
a function of interest to individual taxonomic groups. We can only
measure function in aggregate across an entire community. Finally, we cannot be
sure that the taxonomic groups that putatively perform a function are the ones
contributing in this moment. Therefore, microbial ecologists have needed
alternative approaches to assessing microbial BEF relationships.

As a result of these limitations, microbial ecologists have taken a different
approach to assessing BEF relationships for microbial communities than others
have for macrobial communities. This is generally a correlative approach where
certain community attributes that can be derived from a community matrix are
tested in terms of their relationship with function. These approaches make
different assumptions about how microbial communities sohuld map onto function.
On one end of the spectrum, it is hypothesized that functional gene abundance - genes that
code for an enzyme that performs one step of a process - should be correlated
with the rate of the process associated with that gene. In this case, qPCR
is used to quantify the abundance of a functional gene to test whether it is
correlated with the rate of a process. This approach has had varied success and even the abundance of a
functional gene that putatively performs a process is rarely correlated with
that process \citep{rocca2015}. On the other end of the
spectrum, investigators make few assumptions about which members of the community should
alter function and so compute
taxonomic richness or other diversity metrics from the entire Prokaryotic
community inferred from 16S rRNA gene sequences. Taxonomic or functional
diversity of the Prokaryote community generally adds little variance explained
to models of ecosystem function \citep{graham2016}.

The correlative approach does not take into account many of the fundamental
theories in ecology. For example, the history of biodiversity research has shown
that diversity effects are often hidden variables such as treatment effects like
nitrogen additions or differences in community composition. The recent CAFE
literature, and more broadly the theory of metacommunity ecology, shows the
importance of regional biodiversity in local dynamics in terms of immigration
and emigration. Biogeography in general has demonstrated spatial patterns such
as distance-decay and species-area relationships that could produce false
positives when attempting to correlate structure with function. In the way that
CAFE integrates diversity and composition of communities in addition to
dispersal effects, microbial community structure-ecosystem function research
should incorporate the underlying drivers of community assembly such as
selection and dispersal.  One limitation of the CAFE framework is that it
explicitly omits spatial processes. However, this could be incorporate by
including geographic distance as a covariate.

\subsection{Insights from genotype-phenotype mapping in population genetics}

In addition to metacommunity theory, microbial ecosystem function research could
gain from considerations in organismal biology addressing the genetic variation
underlying complex traits. These studies use data with a similar nature to
microbial data: a matrix of genotype levels and a vector of values of a
phenotype where the link between genotype and phenotype is unknown. When mapping
genotype to phenotype, population geneticists incorporate our understanding of
genetic architecture and population stratification into genotype-phenotype
correlations. For example, human genetic variation is structured such that
populations closer together will tend to be more similar than populations
farther apart \citep{novembre2008}. In case-control studies of genetic diseases, one population might
be more likely to have the disease and also share certain genotypes even if
those two facts are unrelated. Therefore, spatial autocorrelation must be
considered in models of complex traits. Generally, these studies do not take
into account environmental or spatial variability directly, but correct for
population stratification by controlling for shared ancestry.

There are a number of ways that environmental or spatial covariates could alter
or obscure connections between microbial community structure and ecoystem
function. For example, microbial communities exhibit strong biogeographic
patterns \citep{martiny2006}. The historical effects of geographic proximity,
such as dispersal limitation, could lead to spurious associations between taxa
and function if taxa are unevenly distributed across the functional landscape.
Environmental selection can also shape microbial community structure, so
selection on the microbial community that is independent of the
process could also lead to erroneous conclusions about the
underlying drivers of ecosystem function \citep{thompson2017}. For
example, pH is known to regulate both microbial biodiversity patterns
\citep{fierer2006} and methane cycling dynamics \citep{ye2012} which could lead
to microbial taxa correlated with methane cycling, but mechanistically
unrelated. Environmental factors that directly influence function, such as
moisture, could alter function with or without affecting the community. 
Finally, community stratification where taxa co-occur in certain ecosystems due to
selection or historical effects for reasons unrelated to the ecosystem function
could lead to spurious associations
(e.g. \cite{novembre2008} in humans).

\subsection{A new framework for structure-function mapping in microbial
communities}

Microbial ecosystem function could be assessed using statistical techniques
borrowed from quantitative genetics along with the conceptual framework for
community assembly processes provided by metacommunity theory and CAFE. This
would be a two-step process to first identify the portion of variation in
ecosystem function attributable to different aspects of an ecosystem: the
community, the underlying environmental variation, and to spatial processes.

\begin{equation}
V_F=V_E+V_S+V_C+\epsilon
\end{equation}

Ecosystem function is represented by $V_F$, $V_E$ is variation in the
environment and covariation between the community and the environment, $V_C$ is
variation in spatial distance between samples and covariation between the
community and space, and $V_C$ is variation in the community independent of
local selection and historical effects. The parameter $V_C$ also incorporates
unmeasured environmental and spatial effects.

Each of these variance components also correspond to alternative hypotheses for
what drives ecosystem functioning. If most of the variation in $V_F$ is
attributable to $V_E$, then the environment drives function either directly
(abiotic processes) or indirectly by selecting for a particular microbial
community that maximizes function. In this model, the community is simply a 
conduit through which the environment acts on function. Alternatively, if $V_S$
explains most of the variation in $V_F$ then this suggests that function is
primarily a result of historical contingency and probabilistic dispersal.
Finally, $V_C$ could explain most of the variation in $V_F$. This would be
consistent with reciprocal transplant and common garden experiments showing that
community composition drivers ecosystem function independent of the environment.

This framework allows us to address the following questions: 1. Does variation
in microbial community structure explain a measurable portion of the variance in
ecosystem function or can all of that variation be attributed to the underlying
effect of the environmental and spatial processes? If so, 2. Which members of
the community are significantly correlated with function after controlling for
the covariances of space and environment? Finally, 3. Do these taxa match our hypotheses about
what controls ecosystem function (e.g. functional group abundance) or are there
alternative mechanisms that need to be explored?

\section{Methods}

Sample 16S rRNA gene sequences, environmental covariates, geographic distance,
ecosystem function. Use the DADA2 pipeline to assign ASVs. 
phylogenetic and taxonomic assignment Null
Models \citep{chase2011, stegen2015}

Would it be valuable to use a null model (as in Chase or Stegen) where we
generate random communities based on the distribution across samples and
abundance within samples of each taxon. Then try to correlate those abundances
with our observed rates of ecosystem function to see what distribution of
structure-function correlations to expect given our community data. These sort
of null community generating techniques can omit certain community assembly
processes (like homogenizing selection or stochastic processes)

An alternative testing approach would be to use Hannah’s permutation test idea
where we randomly permute the process rates ~1000 times and rebuild the model
each time to see what the distribution of significant taxa is.

VC - Correlations between taxa and process determined by accounting for spatial
structure, environmental covariates, and community structure. Represents effect
of taxonomy, unmeasured environmental variables VE - Account for community
structure and correlate environmental variables cov(C, E) - The part of the venn
diagram that disappears when controlling for environment and space VC:VE - Test
the correlations between community and process rate at a variety of
environmental conditions.  epsilon/error term - Effects of natural variation,
unmeasured variables

Another modeling options is HMSC from \cite{ovaskainen2017}

\section{Results}

For high-affinity methane oxidation in Gabon: We experimentally controlled some
environmental variables and our measured environmental variables are not
significantly correlated with process rate (C, N, WFPS) Space is correlated with
function

Be careful with attributing a certain \% of the variance to an individual taxa
when performing partial regression on the residuals. If a taxa explains 60\% of
the variance in the residuals after controlling for environment and space that
doesn’t mean it explains 60\% of the overall variance.

\section{Discussion}
Problems that need to be addressed: We shouldn’t assume that functional groups
that putatively perform a process are the rate-limiting step or the strongest
correlate.  There may be unknown members of the community that perform or limit
a process The final step of the process may not be the rate-limiting step: it
may be substrate production or nutrient acquisition or something else (e.g.
Qusheng Jin’s work on modeling methanogenesis) We need to consider the
underlying biogeographic patterns shaping the distribution of microbial
communities (à la GWAS).

\section{Conclusion}

\section*{Possible things to include:}
Why do we think that communities will be important for understanding ecosystem
function?

Talk to Andy, CJ, and Jeff about recent work directly

\begin{equation}
V_F=V_E+V_C+V_EV_C+cov(C,E)+\epsilon
\end{equation}
\begin{equation}
F_{ij}=C_i+E_j+CE_{ij}+\epsilon_{ijk}
\end{equation}

In this equation, VF represents variation in the rate of an ecosystem function.
VE is variation in the environment that independently contributes to the rate of
the ecosystem function. VC is variation in microbial community structure (i.e.
composition, abundance, or diversity) that independently contributes to
ecosystem function. The interaction term VE:VC represents situations in which
the difference in function between two communities varies by environment. For
example, community A could outperform community B in environment 1, but not in
environment 2. Alternatively, community A could outperform B in environment 1
and environment 2, but the magnitude of the difference varies by environment.
The term covC,E captures covariance between community and environment where a
change in the environment produces an associated change in the community.
Finally,  is an error term for natural variation in these relationships.

In the context of deterministic versus stochastic/neutral community assembly,
each portion of the variance represents different assembly processes. The cov(C,
E) represents selection whereas VC represents neutral processes of drift,
homogenizing dispersal, or dispersal-limitation. The interaction term VC:VE
represents differences in the rate constants/Q10s for different communities or
neutral processes of community assembly under a particular environmental
condition. VE is independent of the community and so doesn’t represent a
community assembly process, but is simply a physicochemical driver of VF such as
gas diffusion through water, tortuosity of soil pores, redox conditions, … I’m
not sure this paragraph is correct

VE = Direct physicochemical effects on process rates including abiotic
reactions, diffusivity, temperature limitation/Q10 (that doesn’t affect taxon
abundance or activity but does affect e.g. reaction rates), soil stabilization
of chemicals (sorption), tortuosity of soil pores, wind speed (affecting
differential pressure between soil air and atmosphere), concentration gradient
between soil and atmosphere.

cov(C, E) = Environmental selection on soil microorganisms that directly
influence function. Environmental variables that directly affect function and
also select for a microbe unrelated to function. An environmental variable that
both directly affects function and indirectly affects function via a microbe.

VC:VE = Context-dependent organismal traits including different temperature
optima, substrate affinities, enzyme specificity, redox status/oxygen tolerance,
pH optima.

Epsilon = Stochastic processes. Taxa occurrences or
environmental variables completely unrelated to function.

I can imagine several mechanisms by which a significant correlation between a
taxon and a process rate might be detected.  An environmental variable directly
affects function and selects for a microbe unrelated to function (false
positive). (VE) This would be removed by correcting for environment, but would
remain if it was due to an unmeasured environmental variable (and so would
appear in VC).  An environmental variable selects for a microbe that directly
affects function (cov(C, E)). This would be removed by correcting for the right
environmental variable, but would be revealed by controlling for everything
except the environment.  A taxon directly affects function (true positive)  and
co-occurs with another microbe unrelated to function (false positive) (This
could be VC or cov(C,E)). False positive would be removed by correcting for
community similarity. True positive would be lost if perfect covariance between
function, true positive, and false positive.  Function exhibits a spatial signal
(due to underlying spatial VE or VC) and a taxon exhibits a similar spatial
signal, but for different reasons (selection on different VE or VC or historical
effects (e.g. dispersal limitation)). Would be removed by correcting for space,
but would not be removed if there was a perfect covariance.

Why is it important to understand VC?  At this point in the development of
microbial ecology, we know that both selection and dispersal limitation shape
microbial community assembly. At the extremes, a single bacterial phylotype
cannot instantly be present in all environments. Likewise, oxygen will kill
obligate anaerobes. It also appears, to the best of our knowledge, that
microbial communities of different origins exhibit different functional rates
when introduced to identical environmental conditions. Therefore, environmental
selection on microbial communities cannot fully explain variation in ecosystem
functions if they are mediated by microorganisms. Put another way, even if we
could perfectly measure VE, we cannot fully understand VF.

The question that follows from this is: Does VC matter on the scales relevant to
humans? To answer that question, we have to understand the magnitude of

Null models of community assembly

VC or VE:VC could also represent unmeasured environmental variables.

\bibliographystyle{ecology} 
\bibliography{library}

\end{document}

### 2019-06-06 presentation/presentation.tex
Inital commit

\documentclass{beamer}
\usepackage[version=3]{mhchem}
\newcommand\blfootnote[1]{%
  \begingroup
  \renewcommand\thefootnote{}\footnote{#1}%
  \addtocounter{footnote}{-1}%
  \endgroup
}
\graphicspath {{../Figures/}}
\title{Linking microbial communities to ecosystem functions: }
\subtitle{What can we learn
from genotype-phenotype mapping in organisms?}
\author{Andrew Morris and Brendan Bohannan}
\date{May 9, 2019}
\institute{Philosophical Transactions B: Conceptual Challenges in Microbial Community Ecology}
\begin{document}

\frame{
\titlepage
}

\frame{
  \frametitle{My goals for today}
  \begin{enumerate}
    \item Present my latest ideas for this conceptual framework
    \item Get feedback on the logic/conceptual ideas. 
      \begin{itemize}
        \item cool things
        \item plot holes
      \end{itemize}
    \item Get your ideas on what analyses to include
  \end{enumerate}
}

\frame{
  \frametitle{Problem Statement}
  \centering
  \begin{itemize}
    \item Microorganisms putatively mediate biogeochemical cycles
    \item However, we have not yet demonstrated microbial biodiversity-ecosystem
      function relationships\footnotemark
    \item This limits our ability to understand and model these processes
  \end{itemize}
\includegraphics[width=0.8\textwidth]{cycle_crobes.pdf}
\footnotetext{Rocca et al. 2015, Graham et al. 2016}
}
\frame{
  \frametitle{Questions}
  \begin{enumerate}
    \item Do microbial communities explain a unique portion of the variance in
      ecosystem function (independent of environmental and spatial
      factors)?
    \item Which taxa are significantly correlated with function after controlling
      for space and environment?
    \item Do the identified taxa fit into expected functional groups?
  \end{enumerate}
}
\note{
Question 1 is if we control for the underlying envrionmental variables and for
the spatial proximity of sites, does the structure of microbial communities
alter ecosystem function? The reason this is important is if there is some
environmental variable that can reasonably predict function, it's a whole lot
easier to measure pH than to sequence microbial genes. So microbes will really
only be important for ecosystem models if there's something about microbial
physiology that is important for function independent of selection.

Assuming there is some variance there...2.

Question 3 is kind of a sub-question to 2. So, do methanogens explain most of
the variation in methane production or is it some other group e.g. substrate
producers?
}

\frame{
  \frametitle{Biodiversity-Ecosystem Function (BEF)}
\centering
Traditional BEF research has demonstrated for macrobial communities that greater
local richness can lead to greater overall ecosystem function. 
\includegraphics[width=0.8\textwidth]{hooper_biodiversity.png}\blfootnote{Hooper
et al. 2005}
}

\frame{
  \frametitle{Two explanations for BEF}
  \begin{itemize}
    \item Sampling/Selection Effect
    \item Niche Complementarity
  \end{itemize}
  \blfootnote{Hooper et al. 2005}
}

\frame{
  \frametitle{Community Assembly and the Functioning of Ecosystem (CAFE)}
  \centering
  More recently, the importance of regional biodiversity has been demonstrated
  in biodiversity experiments. In this case, when selection for a highly
  productive species is strong, there is a negative local richness-function
  relationship. \\~\\

  \includegraphics[width=0.6\textwidth]{cafe}\blfootnote{Leibold et al. 2017}
}
\note{
  Because the competitive dominant outcompetes other members of the community
}

\frame{
  \frametitle{Implication: Regional biodiversity can influence local ecosystem
    function via metacommunity dynamics. 
}
\pause
  \begin{itemize}
    \item e.g. At intermediate dispersal, when local selection is maximized, maximum ecosystem function is
      associated with a decrease in diversity (\textbf{species-sorting}). 
      \pause
    \item At lower dispersal rates, dispersal limitation weakens the selection
      effect and decreases
  function while increasing complementarity (\textbf{patch-dynamics}). 
  \pause
\item At higher dispersal rates, source-sink dynamics overwhelm local communities with less productive
  species resulting in an increase in diversity and a decrease in function
  (\textbf{mass-effects}).
\end{itemize}
\pause
\textbf{Therefore, spatial processes (metacommunity dynamics) could play an
  important role in local ecosystem function.}\blfootnote{Leibold et al. 2004,
Leibold et al. 2016}
}

\frame{
  \frametitle{Implementing CAFE: The Price Equation}
  \begin{itemize}
    \item The ecological version of the Price equation is used within the CAFE framework to partition ecosystem
  function into the effect of species gains, species losses, and changes in
  abundance (composition effects). 
\item This depends on assigning a portion of ecosystem function to individual species
(e.g. biomass of a single grass plant from a quadrat)
\end{itemize}
\textbf{CAFE demonstrates that not only is local richness important for
  functioning, but composition also plays a role in the functioning of
ecosystems}
  \blfootnote{Fox et al. 2006, Bannar-Martin et
  al. 2018}

}
\frame{
  \frametitle{How we measure microbes is different from macrobial BEF and CAFE}
  \pause
  \begin{itemize}
    \item We do not know who is performing the function at this time (e.g. dormancy,
    incomplete knowledge of functional groups)
    \pause
  \item We cannot directly attribute function to individual organisms or taxa with our
    current sampling methods (metabarcoding, metagenomics)
    \pause
\end{itemize}
\textbf{Therefore, we cannot (yet) use the Price equation for microbes}
}
\note{
The positive BEF relationships from the literature are looking at plant species
richness as it relates to ecosystem function NOT overall diversity. To put it in
microbial terms: They're correlating the diversity of photoautotrophs with the
rate of of photoautotrophy. We don't
know all of the taxa that necessarily contribute to function (or limit it
upstream).

A plant ecologist can pick up a single grass plant and say "This plant fixed thi
smuch carbon", but we cannot pick up a single methanotroph and say "This microbe
oxidized this many molecules of methane".
}

\frame{
  \frametitle{Microbial BEF so far}
  \framesubtitle{What \textit{have} we done?}
  \pause
  \begin{itemize}
    \item Assume functional group or functional gene abundance is correlated
      with the process performed by that group. \textbf{For most functions, that
      doesn't work}\footnotemark \\~\\
  \footnotetext{Rocca et al. 2015}
  \pause
    \item Explain variation in function with \textbf{functional diversity or
taxonomic diversity} \textit{of the entire prokaryotic community} inferred 
from 16S rRNA genes, which generally \textbf{achieves only a small amount of
explanatory power}\footnotemark.
\end{itemize}
  \footnotetext{Graham et al. 2016}
}

\frame{
  \frametitle{Microbial BEF so far, cont'd.}
  Generally ignore:
  \begin{itemize}
    \item Environmental covariates\footnote{except in Graham et al. 2016, but then microbes
      weren't very explanatory}
    \item Spatial proximity
    \item Community structure/co-occurrence patterns\footnote{One host-microbiome
      diabetes study does consider this (Qin et al. 2012)}
  \end{itemize}

}
\note{
  These studies generally do simple correlations between a single taxon or gene
  or diversity metric and function. 
}
\frame{
  \frametitle{Goal: Develop a framework that achieves the following}
  \begin{itemize}
    \item Use the CAFE framework to consider both effects of diversity and
      composition
    \item Incorporate spatial dynamics \textit{\'a la} Metacommunity Theory
    \item Do not assume anything about who is perfoming the function
\end{itemize}
}
\note{
  To achieve these goals, I actually think the BEF literature is not the right
  place to look
}

\frame{
  \frametitle{Biology through Analogy: microbial ecosystem function is less like
    BEF and more like genotype-phenotype mapping}

\begin{table}[ht]
\centering  
\begin{tabular}{rllllll}
  \hline
  Function & & Taxon & Taxon& Taxon& Taxon& Taxon \\ 
  \hline
  6.4 & & 0 & 10 & 5 & 3 & 0  \\
  12.1 & & 2 & 0 & 15 & 2 & 0  \\
  8.3 & & 3 & 0 & 17 & 1 & 0  \\
  0.4 & & 0 & 3 & 7 & 0 & 3  \\
  \\
  \hline
  Phenotype & & SNP & SNP & SNP & SNP & SNP \\ 
  \hline
  6.4 & & 0 & 1 & 1 & 1 & 0  \\
  12.1 & & 1 & 0 & 1 & 1 & 0  \\
  8.3 & & 1 & 0 & 1 & 1 & 0  \\
  0.4 & & 0 & 1 & 1 & 0 & 1  \\
   \hline       
\end{tabular} 
\end{table}

}
\note{

  The nature of microbial community data (community matrix of sequence variants
for thousands of taxa) and microbial function data (single measurement for an
entire ecosystem) makes this problem analogous to identifying the genetic
variation that underlies complex traits in organisms (i.e. a single phenotype
value per individual with thousands of SNPs).

In each of these cases the data are quite similar.
}
\frame{
  \frametitle{Genome-wide association studies}
  \begin{itemize}
    \item Robust correlations (adjusting for multiple comparisons)
    \item Correct for population structure
  \end{itemize}
  \centerline{
  \includegraphics[width=0.7\textwidth]{europe}
  \blfootnote{Novembre et al. 2008}
}
}
\frame{
  \frametitle{Underlying covariances could obscure structure-function
  relationships}

	\begin{columns}
		\begin{column}{0.3\textwidth}
			\centering
			\pause
			\includegraphics[width=\textwidth]{distance_decay.pdf}
		\end{column}
		\begin{column}{0.3\textwidth} \centering
			\pause
			\includegraphics[width=\textwidth]{community_environment.pdf}
		\end{column}
		\begin{column}{0.3\textwidth} \centering
			\pause
			\includegraphics[width=\textwidth]{similarity_of_function.pdf}
		\end{column}
	\end{columns}
}
\frame{
  \frametitle{Ways in which we could get false positives}
  \begin{enumerate}
    \item An environmental variable selects for a crobe that is correlated with
      function
    \item Sites closer together are more likely to have similar function and
      similar communities due to historical effects
    \item Co-occurrences among taxa due to species interactions. i.e. if taxon A
      is correlated with function and correlated with taxon B, then taxon B
      would be incorrectly correlated with function.

  \end{enumerate}
}

\frame{

	\frametitle{How does variation in microbial community structure contribute to
	variation in ecosystem function?}
	\centering
	\large
\pause	
  $V_F = V_E + V_C +$ 
  $V_EV_C + Cov(E, C) + \epsilon$ \\~\\

  \pause
	$V_F = V_E + V_C + V_S + \epsilon$
}
\note{
  I'm thinking about simplifying that equation by combining some of the terms
  and talking to Mathew Leibold helped influence me in this regard. I liked how
  he laid out his variance partitioning framework.  }

\note{
Okay, imagine a venn diagram with three circles. The overlap between Environment
and Space and covariance of environment and space. As per Liebold's suggestion,
incorporate 50\% of that overlap into each of E and S. The overlap between C-E
and C-S is environmental selection and spatial autocorrelation. Incorporates
those into E and S, respectively. The remainder of the third circle is variance
uniquely associated with variation in the community. The reponse variable is a
vector of ecosystem function (or a matrix if you consider multiple functions at
one)
}
\note{
  Variance in ecosystem function is a complex set of interactions between local
  environmental conditions, local community structure, spatial processes like
  dispersal, and stochastic processes. For example, the environment has direct
  effects on funtion via abiotic processes, but also influences function by
  selecting on the community. Spatial processes influence function through
  patterns like distance-decay where closer ecosystems tend to be more similar
  and organisms are more likely to disperse close-by. Stochastic processes like
  community drift and diversification influence the community. I'm choosing to
  simplify that variance partitioning problem by focusing on the three main
  sources of variation: effects of the environment (directly or indirectly via
  selection), effects of spatial proximity, and effects of co-occurrence
  patterns among taxa unrelated to function (community structure).
}

\frame{
  \begin{itemize}
    \item $V_F$ = Variation in ecosystem function
    \item $V_E$ = Direct effect of environment on function (abiotic) plus
      indirect effect via selection on the community
    \item $V_S$ = Effect of spatial proximity e.g. distance-decay
    \item $V_C$ = Community effect independent of the environment and spatial
      proximity, unmeasured environmental variables
    \item $\epsilon$ = error term, stochastic processes 
  \end{itemize}
}
\frame{
  \frametitle{How to do this?}
  \begin{itemize}
    \item varComp?
    \item HMSC
    \item ...?
  \end{itemize}
}

\frame{
  \frametitle{Which taxa are significantly correlated with function after controlling
  for space and environment?}
  My current approach:
  \begin{itemize}
    \item Partial regression on function and abundance to control for community
      similarity, environmental similarity, and spatial proximity
    \item Perform individual correlations between each taxon and function
    \item Correct for multiple comparisons using Bonferroni (alpha/n)
  \end{itemize}
}
\frame{
  \frametitle{Things I don't like about my current approach}
  \begin{itemize}
    \item Have to choose an arbitrary number of PC axes for partial regression
    \item Cannot estimate the variance components for Ve, Vc, and Vs
    \item People say it's out-of-date and it would be worth using a more modern
      mixed modeling approach
  \end{itemize}
  What should I do instead?
  \begin{itemize}
    \item Mixed modeling
    \item Matrix regression
  \end{itemize}
}
\frame{
  \frametitle{Possible Research Plan}
  \begin{enumerate}
    \item Variance partitioning between community, environment,
      space, and error.
    \item Identify significant taxa using statistical techniques from GWAS
    \item Use one or two datasets: Gabon and Brazil; One or more functions:
      High-affinity methane oxidation, low-affinity, methane production
    \item Compare results to other analyses that people typically do: indicator
      species analysis, PC regression...
  \end{enumerate}
}
\frame{
  \frametitle{Results}
  \begin{enumerate}
        \item \textasciitilde80 taxa correlated with function, decreases to \textasciitilde50 if I
          control for environmental/spatial covariates, decreases to \textasciitilde3 if I
          control for community structure
        \item Number of taxa varies dramatically depending on the number of PC
          axes I include
    \item This identified three taxa, but we don't know anything about their
      metabolism:
      \begin{itemize}
        \item Group 2 Acidobacterium
        \item Conexibacter (Actinobacteria)
        \item Armatimonas (Group 1 Armatimonadetes)
      \end{itemize}
    \item No methanotrophs!
  \end{enumerate}
}
\note{
  Possible explanations: Methane oxidizers are selected for under the
  environmental conditions and so act as conduits for the environment.
  Methanotrophs perfectly covary with some environmental variable or other
  community member or space. 
}
\end{document}

### 2019-06-06 PC_correction.R

@@ -1,0 +1,196 @@
+
+#### Calculate community and environment similarity scores using PCA
+#### with the vegan::rda function. Then adjust the community matrix
+#### and ecosystem function rates by correcting for the regression
+#### coefficients calculated from genotype ~ community similarity,
+#### phenotype ~ community similarity, genotype ~ environmental
+#### similarity, and phenotype ~ environmental similarity
+#### See: Price et al. 2006 in Nature Genetics
+
+# Whether to correct by each factor
+com_bool <- T
+env_bool <- T
+spa_bool <- T
+# Set the number of axes to correct by
+n_com_axes <- 6 
+n_env_axes <- 3 
+n_spa_axes <- 1
+# Whether to center/scale the data
+scale_com <- T
+scale_env <- T
+scale_spa <- T
+# File names
+input_tsv <- "output/cleaned_data.tsv"
+output_csv <- "output/adj_data.csv"
+com_output_csv <- 'output/com_adj_data.csv'
+env_output_csv <- 'output/env_adj_data.csv'
+spa_output_csv <- 'output/spa_adj_data.csv'
+# Function column names
+fun_cols <- c('CH4')
+# Environment column names
+env_cols <- c('pH', 'Organic_Matter', 'P', 'S', 'Ca', 'Mg', 'Al', 'H_plus_Al_SMP', 
+              'Sum_of_bases', 'CEC', 'V', 'Al_saturation', 'B', 'Cu', 'Fe',
+              'Mn', 'Zn', 'N_total')
+# Spatial column names
+spa_cols <- c('Lat', 'Long')
+# For community data, this assumes the ASV columns all start with 'asv'
+com_cols <- 'asv'
+
+library(tidyverse)
+library(vegan)
+
+all_data <- read_tsv(input_tsv)
+
+all_data <- all_data[complete.cases(all_data$Lat), ]
+all_data <- all_data[complete.cases(all_data$Long), ]
+
+asv_mat <- as.matrix(select(all_data, starts_with('asv')))
+
+asv_mat[asv_mat > 0] <- 1
+dim(asv_mat)
+asv_mat <- asv_mat[, colSums(asv_mat) > 1]
+colSums(asv_mat)
+n_samples <- nrow(all_data)
+
+# Extract ASV table
+community <- select(all_data, starts_with(com_cols))
[…]

### 2019-06-06 brazil_cleaning.R

@@ -1,0 +1,34 @@
+library(tidyverse)
+
+## Cleaning up metadata
+metadata <- read_tsv('data/Dimensions_meta_data.txt')
+metadata <- as.data.frame(metadata)
+## Converting below detection limit to the detection limt
+## Controversial decision - consider other options (zeros?)
+metadata <- sapply(metadata,function(x) 
+               gsub("<","",x) )
+metadata <- as.data.frame(metadata)
+metadata <- sapply(metadata,function(x) gsub(",",".",x))
+metadata <- as_tibble(metadata)
+metadata <- mutate_at(metadata, vars(pH:Error_N2O), as.numeric)
+as.data.frame(metadata)
+
+## ASVs
+asvs <- read_tsv('data/16S_dada2_table_RDP_tax.txt')
+colnames(asvs)
+taxa_abund <- 
+  asvs %>% 
+  mutate(OTU_ID = paste0("asv_", OTU_ID)) %>% 
+  select(-ConsensusLineage) %>% 
+  gather(Sample_ID, Abundance, 2:75) %>% 
+  spread(OTU_ID, Abundance)
+gps <- read.table('data/Dimensions_GPS_dec_deg.txt', header = FALSE)
+colnames(gps) <- c('Sample_ID', 'Lat', 'Long')
+
+all_data <-
+  gps %>% 
+  right_join(metadata, by = 'Sample_ID') %>% 
+  right_join(taxa_abund, by = 'Sample_ID')
+
+all_data$library_size <- rowSums(taxa_abund[, -1])
+write_tsv(all_data, 'output/cleaned_data.tsv')

### 2019-06-06 combine_data.R

@@ -1,0 +1,35 @@
+#### Input: troph_attr_table and rare_asv_table.csv
+#### Output: troph_total.csv
+
+library(tidyverse)
+
+source('R/misc_functions.R')
+
+#gen_attr_table <- read_csv('output/gen_attr_table.csv')
+
+troph_attr_table <- read_csv('output/troph_attr_table.csv')
+troph_attr_table$Site <- paste0(troph_attr_table$Site, 
+				reverse_substr(troph_attr_table$Sample, 3, 3))
+## Rarefied ASV Table
+troph_asv_table <- read_csv('output/rare_asv_table.csv')
+
+## DESeq2 variance stabilized asv table
+# troph_asv_table <- read_csv('output/vst_asv_table.csv')
+
+## Raw sequence counts
+#troph_asv_table <- read_csv('output/asv_table.csv')
+#asvs <- troph_asv_table[, -1] 
+### Presence Absence
+#asvs[asvs>0] <- 1
+#troph_asv_table[, -1] <- asvs
+
+#left_join(troph_attr_table, troph_asv_table, by = "Sample") %>%
+#  write_csv('output/troph_total.csv')
+geodist <- read.table('output/geodist.tsv')
+troph_attr_table <- left_join(troph_attr_table, geodist, by = "Site")
+troph_attr_table <- left_join(troph_attr_table, troph_asv_table, by = "Sample")
+  write_csv(troph_attr_table, 'output/troph_total.csv')
+
+# gen_attr_table %>% 
+# left_join(asv_table, by = "Sample") %>%
+# write_csv('output/gen_total.csv')

### 2019-06-06 id_taxa.R

@@ -1,0 +1,139 @@
+library(tidyverse)
+library(xtable)
+# import model output
+by_asv_all <- readRDS('output/model_fits.rds')
+by_asv_raw <- readRDS('output/model_fits_raw.rds')
+by_asv_com <- readRDS('output/model_fits_com.rds')
+by_asv_env <- readRDS('output/model_fits_env.rds')
+by_asv_spa <- readRDS('output/model_fits_spa.rds')
+
+new_data <- read_csv('output/adj_data.csv')
+taxon_table <- read_csv('output/taxon_table.csv')
+
+alpha <- 0.05 / nrow(by_asv_all)
+# Identify ASVs with -log(p value) > 
+id_asvs <- function(by_asv, alpha, taxon_table, new_data, file_name) {
+lowk_asvs <- by_asv %>% 
+  mutate(tidy = map(model, broom::tidy)) %>% 
+  unnest(tidy) %>% 
+  filter(term == 'abund') %>% 
+  filter(-log10(p.value) > -log10(alpha)) %>% 
+  pull(asv)
+
+# Pull out those ASVs
+important_asvs <- new_data %>% 
+  select(Low_final_k, one_of(lowk_asvs))%>% 
+  gather(asv, abund, starts_with('asv'))
+taxon_table %>% 
+  filter(asv %in% lowk_asvs) %>% 
+  select(-asv) %>%
+  xtable() %>% 
+  print.xtable(file = file_name)
+}
+
+id_asvs(by_asv_raw, alpha, taxon_table, new_data, file_name = "tables/table_rare_raw.tex")
+id_asvs(by_asv_env, alpha, taxon_table, new_data, file_name = "tables/table_rare_env.tex")
+id_asvs(by_asv_spa, alpha, taxon_table, new_data, file_name = "tables/table_rare_spa.tex")
+id_asvs(by_asv_com, alpha, taxon_table, new_data, file_name = "tables/table_rare_com.tex")
+id_asvs(by_asv_all, alpha, taxon_table, new_data, file_name = "tables/table_rare_all.tex")
+
+
+## Individual correlation
+imp_tax <- taxon_table %>% 
+  filter(asv %in% important_asvs$asv)
+imp_tax$taxon <- apply(imp_tax, 1, function(r) { r[ 
+                      if( any(is.na(r))){ 
+                          (which(is.na(r))[1]-1) 
[…]

### 2019-06-06 lm_model.R

@@ -1,0 +1,45 @@
+#### Input: adj_data.csv
+#### Output: model_fits.rds
+
+library(tidyverse)
+library(purrr)
+library(broom)
+library(modelr)
+
+adj_data <- read_csv('output/adj_data.csv')
+com_adj_data <- read_csv('output/com_adj_data.csv')
+env_adj_data <- read_csv('output/env_adj_data.csv')
+spa_adj_data <- read_csv('output/spa_adj_data.csv')
+troph_total <- read_csv('output/troph_total.csv')
+
+
+
+##### Functions
+
+lowk_model <- function(df) {
+  lm(Low_final_k ~ abund, data = df)
+}
+vmax_model <- function(df) {
+  lm(Vmax ~ abund, data = df)
+}
+# Create nested data set by asv
+  # Calculate linear model for each asv between abundance and function
+fit_models <- function(data, model) {
+  data %>% 
+    gather(asv, abund, starts_with('asv_')) %>% 
+    group_by(asv) %>% 
+    nest() %>% 
+    mutate(model = map(data, model)) %>% 
+    return()
+}
+
+saveRDS(fit_models(adj_data, lowk_model), file = "output/lowk_fits.rds")
+saveRDS(fit_models(com_adj_data, lowk_model), file = "output/lowk_fits_com.rds")
+saveRDS(fit_models(env_adj_data, lowk_model), file = "output/lowk_fits_env.rds")
+saveRDS(fit_models(spa_adj_data, lowk_model), file = "output/lowk_fits_spa.rds")
+saveRDS(fit_models(troph_total, lowk_model), file = "output/lowk_fits_raw.rds")
+saveRDS(fit_models(adj_data, vmax_model), file = "output/vmax_fits.rds")
+saveRDS(fit_models(com_adj_data, vmax_model), file = "output/vmax_fits_com.rds")
+saveRDS(fit_models(env_adj_data, vmax_model), file = "output/vmax_fits_env.rds")
+saveRDS(fit_models(spa_adj_data, vmax_model), file = "output/vmax_fits_spa.rds")
+saveRDS(fit_models(troph_total, vmax_model), file = "output/vmax_fits_raw.rds")

### 2019-06-06 man_plot.R

@@ -1,0 +1,79 @@
+library(tidyverse)
+taxon_table <- read_csv('output/taxon_table.csv')
+lowk_by_asv <- readRDS('output/lowk_fits.rds')
+lowk_by_asv_com <- readRDS('output/lowk_fits_com.rds')
+lowk_by_asv_env <- readRDS('output/lowk_fits_env.rds')
+lowk_by_asv_raw <- readRDS('output/lowk_fits_raw.rds')
+lowk_by_asv_spa <- readRDS('output/lowk_fits_spa.rds')
+
+vmax_by_asv <- readRDS('output/vmax_fits.rds')
+vmax_by_asv_com <- readRDS('output/vmax_fits_com.rds')
+vmax_by_asv_env <- readRDS('output/vmax_fits_env.rds')
+vmax_by_asv_raw <- readRDS('output/vmax_fits_raw.rds')
+vmax_by_asv_spa <- readRDS('output/vmax_fits_spa.rds')
+# Create plotting function
+
+man_plot <- function(by_asv, taxon_table) {
+
+by_asv %>% 
+  mutate(tidy = map(model, broom::tidy)) %>% 
+  unnest(tidy) %>% 
+  filter(term == 'abund') %>%
+  left_join(taxon_table, by = 'asv') %>% 
+  ggplot(aes(x = reorder(asv, Phylum, median), y = -log10(p.value))) +
+  geom_point(aes(color=Phylum), size = 1) +
+  geom_hline(yintercept = -log10(0.05/nrow(by_asv))) +
+  scale_color_discrete(guide = F) +
+  labs(x = 'Amplicon Sequence Variant (ASV)', 
+       y = expression(Significance ~ (-log[10](p.value)))) +
+    #  geom_text_repel(aes(label=ifelse(-log10(p.value)>-log10(alpha),as.character(genus),'')), size = 1) +
+    # geom_text(aes(label=ifelse(-log10(p.value)>-log10(alpha),as.character(genus),'')), position = position_dodge(width = 10)) +
+  theme_classic() +
+  theme(axis.text.x = element_blank(),
+        axis.ticks.x = element_blank()) +
+  ylim(0, 10)  %>% 
+  return()
+}
+
+# Create man plots for adjusted and un-adjusted abundances
+ggsave('figures/man_rare_all.pdf', plot = man_plot(by_asv, taxon_table), 
+       width = 4*1.2, height = 3*1.2)
[…]

### 2019-06-06 rarefy_otu_table.R

@@ -1,0 +1,44 @@
+library(tidyverse)
+source('R/Rarefy_mean.R')
+source('R/misc_functions.R')
+
+# Import ASV table
+asv_table <- read_tsv('data/16S_dada2_table_RDP_tax_edit.txt', n_max = 100000)
+
+# Pull out 
+taxon_table <- 
+  asv_table %>% 
+  select(OTU_ID, ConsensusLineage) %>% 
+  mutate(OTU_ID = paste('asv', OTU_ID, sep = "_")) %>% 
+  select(asv = OTU_ID, ConsensusLineage) %>% 
+  separate(ConsensusLineage, c('Domain', 'Phylum', 'Class', 'Order', 'Family',
+                               'Genus', 'Species'), ";") %>% 
+  select(-Species)
+write_csv(taxon_table, 'output/taxon_table.csv')
+
+# Rename ASVs to remove backticks in column names, add Sample ID column
+# Convert all abundances to doubles
+wide_asv <- asv_table %>% 
+  mutate(asv = paste('asv', OTU_ID, sep = "_")) %>% 
+  select(asv, starts_with('dna')) %>% 
+  select(asv, ends_with('MO')) %>% 
+  gather(Sample, abund, 2:length(.)) %>% 
+  spread(asv, abund) %>% 
+  mutate(Sample = substring(Sample, 5))
+wide_asv <- wide_asv[, colSums(wide_asv != 0, na.rm = T) > 0]
+
+comm_mat <- as.matrix(wide_asv[, -1])
+
+samples <- tibble(Sample = wide_asv$Sample)
+
+rare_mat <- rarefy.mean(comm_mat, n = 1000)
+
+rare_asv <- cbind(samples, as_tibble(rare_mat))
+
+wide_asv <- cbind(samples, as_tibble(comm_mat))
+
+
+write_csv(wide_asv, 'output/asv_table.csv')
+
+write_csv(rare_asv, 'output/rare_asv_table.csv')
+

### 2019-06-06 var_comp.R

@@ -1,0 +1,121 @@
+library(tidyverse)
+library(varComp)
+library(vegan)
+library(broom)
+source('R/misc_functions.R')
+
+
+asv_table <- read_csv('output/asv_table.csv')
+
+i = 2
+j = ncol(asv_table)
+  asv_table[, c(1, i:j)] %>% 
+  gather(ASV, ABU, 2:ncol(asv_table[, c(1, i:j)])) %>% 
+ggplot(aes(y = ABU, x = ASV)) +
+  geom_jitter()  
+
+
+asvs <- asv_table[, -1] 
+## Presence Absence
+asvs[asvs>0] <- 1
+asv_table[, -1] <- asvs
+asv_mat <- as.matrix(asv_table[, 2:ncol(asv_table)])
+ColVar <- function(x, ...) {
+  colSums((x - colMeans(x, ...))^2, ...)/(dim(x)[2] - 1)
+}
+
+asv_mat_var <- ColVar(asv_mat)
+asv_mat_high_var<-asv_mat[,log10(asv_mat_var) > 0]
+ncol(asv_mat_high_var)
+plot(density(log10(asv_mat_var)))
+rowSums(asv_mat)
+colSums(asv_mat)
+asv_mat_ns <- asv_mat[, colSums(asv_mat) > 1]
+
+colSums(asv_mat_ns)
+taxon <- asv_mat_high_var[, 3]
+com_dissim <- vegdist(asv_mat_high_var[, -3])
+com_dissim <- as.matrix(com_dissim)
+com_sim <- 1-com_dissim
+
+
+troph_attr_table <- read_csv('output/troph_attr_table.csv')
+troph_attr_table$Site <- paste0(troph_attr_table$Site, 
+				reverse_substr(troph_attr_table$Sample, 3, 3))
+lowk <- troph_attr_table$Low_final_k
+
+
+geodist <- read.table('output/geodist.tsv')
+troph_attr_table <- left_join(troph_attr_table, geodist, by = "Site")
+xy <- troph_attr_table[, c('X', 'Y')]
+xy <- scale(xy)
+geo_dist <- dist(as.matrix(xy))
+geo_dist <- as.matrix(geo_dist)
+geo_sim <-max(geo_dist) - geo_dist
+
+env <- select(troph_attr_table, WFPS, N_percent, C_percent)
[…]

### 2019-06-06 var_part_brazil.R

@@ -1,0 +1,24 @@
+library(tidyverse)
+library(vegan)
+
+all_data <- read_tsv('output/cleaned_data.tsv')
+dim(all_data)
+all_data <- all_data[complete.cases(all_data$Lat), ]
+dim(all_data)
+all_data <- all_data[complete.cases(all_data$Long), ]
+dim(all_data)
+colnames(all_data[, 1:40])
+
+spa <- select(all_data, Lat, Long)
+geo_dist <- as.matrix(dist(scale(as.matrix(spa))))
+
+env <- select(all_data, pH:Total_moisture)
+env_dissim <- as.matrix(dist(scale(as.matrix(env))))
+
+com <- select(all_data, starts_with('asv'))
+com_dissim <- as.matrix(vegdist(as.matrix(com)))
+
+fun <- select(all_data, CH4)
+fit <- varpart(fun$CH4, ~ geo_dist, ~ env_dissim)
+fit
+plot(fit)

### 2019-06-06 var_stab_transform.R

@@ -1,0 +1,33 @@
+
+library(DESeq2)
+library(tidyverse)
+asv_data <- read.delim('data/16S_dada2_table_RDP_tax_edit.txt', header = TRUE)
+
+
+asv_mat <- as.matrix(asv_data[,-c(1, ncol(asv_data))])
+rownames(asv_mat) <- asv_data$OTU_ID
+# colnames(asv_mat) <- colnames(asv_data[, -1])
+col <- data.frame(site = colnames(asv_data[, -c(1, ncol(asv_data))]))
+
+dds <- DESeqDataSetFromMatrix(countData = asv_mat, design = ~ 1, colData = col)
+
+dds
+dds <- estimateSizeFactors(dds, type = 'poscounts')
+dds <- estimateDispersions(dds)
+vsd <- varianceStabilizingTransformation(dds)
+
+varstab_commat <- assay(vsd)
+vsd_data <- 
+	cbind(tibble(asv = rownames(varstab_commat)), as_tibble(varstab_commat)) %>%
+  mutate(asv = paste('asv', asv, sep = "_")) 
+
+
+asv_table <- vsd_data %>%
+  select(asv, starts_with('dna')) %>% 
+  select(asv, ends_with('MO')) %>% 
+  gather(Sample, abund, 2:length(.)) %>% 
+  spread(asv, abund) %>% 
+  mutate(Sample = substring(Sample, 5))
+
+
+write_csv(asv_table, 'output/vst_asv_table.csv')

### 2019-06-06 geodist_processing.R
Add spatial data processing

@@ -1,0 +1,20 @@
+library(sp)
+library(rgdal)
+
+geodist <- read.table('data/Geodist_gabon_methane_samples_lat_long.txt')
+
+
+colnames(geodist) <- c('Site', 'X', 'Y')
+head(geodist)
+tail(geodist)
+coordinates(geodist) <- c('X', 'Y')
+proj4string(geodist) <- CRS("+proj=longlat +datum=WGS84")  ## for example
+res <- spTransform(geodist, CRS("+proj=utm +zone=32M ellps=WGS84"))
+geodist <- as.data.frame(res)
+
+ plot(geodist$X, geodist$Y)
+(geodist$X <- geodist$X - min(geodist$X))
+(geodist$Y <- geodist$Y - min(geodist$Y))
+ plot(geodist$X, geodist$Y)
+geodist$Site
+write.table(geodist, 'output/geodist.tsv')

### 2019-06-07 PC_correction.R
Add universal pc correction function

@@ -1,4 +1,3 @@
-
 #### Calculate community and environment similarity scores using PCA
 #### with the vegan::rda function. Then adjust the community matrix
 #### and ecosystem function rates by correcting for the regression
@@ -78,6 +77,21 @@ pc_adjust_mat <- function(raw_mat, aj, n_axes) {
   }
   return(raw_mat)
 }
+pc_adjust_mat2 <- function(raw_mat, aj, n_axes) {
+  # com_mat is the community matrix
+  # aj is the matrix of site scores from PCA
+  # n_axes is the number of PC axes to correct by
+  gamma <- vector(length = nrow(raw_mat))
+	for (p in 1:n_axes) {
+    a <- aj[, p] # select axis n for correction
+    for (i in 1:ncol(raw_mat)) { # adjust each taxon abundance
+      tax <- raw_mat[, i] # select taxon abundances for j samples
+      gamma[i] <- sum(a * tax)/sum(a^2) # calculate gamma (regression coef)
+      raw_mat[, i] <- tax - gamma[i] * a 
+    }
+  }
+  return(raw_mat)
+}
 pc_adjust_vec <- function(raw_vec, aj, n_axes) {
   for (p in 1:n_axes) {
     a <- aj[, p]
@@ -97,22 +111,7 @@ com_pca <- rda(community, scale = scale_com) # Calculate covariance matrix
 com_aj <- summary(com_pca)$sites # pull out site scores, which represent
 # the community similarity matrix (shared ancestry in GWAS language)
 # for community similarity predicting genotype
-com_adj <- pc_adjust_mat(com_mat, com_aj, n_com_axes) 
-
-# Same correction as the com matrix now for vector of low final k
-for (i in 1:ncol(func)) {
-  func_adj[, i] <- pc_adjust_vec(func[, i], com_aj, n_com_axes) 
 }
-
-# Recombine data
-func_adj_a <- as.data.frame(func_adj)
-com_adj_a <- as.data.frame(t(com_adj))
-new_data <- cbind(func_adj_a, com_adj_a)
-
-
-write.table(new_data, com_output_csv, sep = ",")
-}
-
 #### Correct for environmental similarity
 if (env_bool) {
 	
[…]

### 2019-06-07 PC_correction.R
Fix pc correction to work with one matrix

@@ -53,13 +53,15 @@ all_data <- cbind(all_data[, 1:non_asv_cols],
 n_samples <- nrow(all_data)
 
 # Extract ASV table
-community <- all_data[, grepl( "asv" , names(all_data))]
+community <- as.matrix(all_data[, grepl( "asv" , names(all_data))])
 com_mat <- t(community)
 
 # Initialize function vector
-func <- as.matrix(all_data[, fun_cols])
+fun_mat <- as.matrix(all_data[, fun_cols])
 
-func_adj <- func
+fun_com <- cbind(fun_mat, community)
+class(fun_com)
+
 #### Functions
 
 pc_adjust_mat <- function(raw_mat, aj, n_axes) {
@@ -139,6 +141,8 @@ com_adj <- pc_adjust_mat(com_mat, com_aj, n_com_axes)
 # Function correction
 func_adj <- apply(func, 2, pc_adjust_vec, com_aj, n_com_axes) 
 func_adj2 <- pc_adjust_mat2(func, com_aj, n_com_axes) 
+com_adj2 <- pc_adjust_mat2(community, com_aj, n_com_axes) 
+all_adj <- pc_adjust_mat2(fun_com, com_aj, n_com_axes) 
 cbind(func_adj, func_adj2)
 sum(func_adj != func_adj2)
 # Recombine data

### 2019-06-07 PC_correction.R
Update PC correction for all combinations

@@ -11,8 +11,8 @@ com_bool <- T
 env_bool <- T
 spa_bool <- T
 # Set the number of axes to correct by
-n_com_axes <- 6 
-n_env_axes <- 3 
+n_com_axes <- 5 
+n_env_axes <- 2 
 n_spa_axes <- 1
 # Whether to center/scale the data
 scale_com <- T
@@ -21,10 +21,7 @@ scale_spa <- T
 # File names
 input_sep <- ","
 input_file <- "output/troph_total.csv"
-output_csv <- "output/adj_data.csv"
-com_output_csv <- 'output/com_adj_data.csv'
-env_output_csv <- 'output/env_adj_data.csv'
-spa_output_csv <- 'output/spa_adj_data.csv'
+output_folder <- "output/"
 # Function column names
 fun_cols <- c('Low_final_k', 'Vmax')
 # Environment column names
@@ -36,16 +33,16 @@ com_cols <- 'asv'
 # Extra grouping variables to preserve
 group_cols <- c('Sample', 'Wetland', 'Site')
 
-#library(tidyverse)
 library(vegan)
 
 all_data <- read.table(input_file, header = TRUE, sep = input_sep)
 
+# Remove observations with missing location information
 all_data <- all_data[complete.cases(all_data[, spa_cols[1]]), ]
 all_data <- all_data[complete.cases(all_data[, spa_cols[2]]), ]
 
-non_asv_cols <- ncol(all_data[, !grepl( "asv" , names(all_data))])
-asv_mat <- as.matrix(all_data[, grepl( "asv" , names(all_data))])
+non_asv_cols <- ncol(all_data[, !grepl(com_cols, names(all_data))])
+asv_mat <- as.matrix(all_data[, grepl(com_cols, names(all_data))])
 asv_mat[asv_mat > 0] <- 1
 asv_mat <- asv_mat[, colSums(asv_mat) > 1]
 all_data <- cbind(all_data[, 1:non_asv_cols], 
@@ -53,34 +50,14 @@ all_data <- cbind(all_data[, 1:non_asv_cols],
 n_samples <- nrow(all_data)
 
 # Extract ASV table
-community <- as.matrix(all_data[, grepl( "asv" , names(all_data))])
-com_mat <- t(community)
+com_mat <- as.matrix(all_data[, grepl(com_cols, names(all_data))])
 
-# Initialize function vector
[…]

### 2019-06-07 PC_correction.R
Fix output file names

@@ -21,7 +21,7 @@ scale_spa <- T
 # File names
 input_sep <- ","
 input_file <- "output/troph_total.csv"
-output_folder <- "output/"
+output_file <- "output/gab_adj"
 # Function column names
 fun_cols <- c('Low_final_k', 'Vmax')
 # Environment column names
@@ -103,44 +103,45 @@ spa_pca <- rda(spa, scale = scale_spa)
 spa_aj <- summary(env_pca)$sites # pull out site scores, which represent
 ## the community similarity matrix (shared ancestry in GWAS language)
 }
+
 if (com_bool) {
 # Community correction
 all_adj <- pc_adjust_mat(fun_com, com_aj, n_com_axes) 
-write.table(all_adj, paste0(output_folder, 'com_adj_data.tsv'))
+write.table(all_adj, paste0(output_file, '_com.tsv'))
 }
 
 if (env_bool) {
 all_adj <- pc_adjust_mat(fun_com, env_aj, n_env_axes) 
-write.table(all_adj, paste0(output_folder, 'env_adj_data.tsv'))
+write.table(all_adj, paste0(output_file, '_env.tsv'))
 }
 
 if (spa_bool) {
 all_adj <- pc_adjust_mat(fun_com, spa_aj, n_spa_axes) 
-write.table(all_adj, paste0(output_folder, 'spa_adj_data.tsv'))
+write.table(all_adj, paste0(output_file, '_spa.tsv'))
 }
 
 if (com_bool & env_bool) {
 all_adj <- pc_adjust_mat(fun_com, com_aj, n_com_axes) 
 all_adj <- pc_adjust_mat(all_adj, env_aj, n_env_axes) 
 all_adj <- cbind(all_data[, group_cols], all_adj)
-write.table(all_adj, paste0(output_folder, 'com_env_adj_data.tsv'))
+write.table(all_adj, paste0(output_file, '_com_env.tsv'))
 }
 if (com_bool & spa_bool) {
 all_adj <- pc_adjust_mat(fun_com, com_aj, n_com_axes) 
 all_adj <- pc_adjust_mat(all_adj, spa_aj, n_spa_axes) 
 all_adj <- cbind(all_data[, group_cols], all_adj)
-write.table(all_adj, paste0(output_folder, 'com_spa_adj_data.tsv'))
+write.table(all_adj, paste0(output_file, '_com_spa.tsv'))
 }
[…]

### 2019-06-07 brazil_cleaning.R
Update readme

@@ -1,7 +1,7 @@
 library(tidyverse)
 
 ## Cleaning up metadata
-metadata <- read_tsv('data/Dimensions_meta_data.txt')
+metadata <- read_tsv('data/brazil/Dimensions_meta_data.txt')
 metadata <- as.data.frame(metadata)
 ## Converting below detection limit to the detection limt
 ## Controversial decision - consider other options (zeros?)
@@ -14,7 +14,7 @@ metadata <- mutate_at(metadata, vars(pH:Error_N2O), as.numeric)
 as.data.frame(metadata)
 
 ## ASVs
-asvs <- read_tsv('data/16S_dada2_table_RDP_tax.txt')
+asvs <- read_tsv('data/brazil/16S_dada2_table_RDP_tax.txt')
 colnames(asvs)
 taxa_abund <- 
   asvs %>% 
@@ -22,7 +22,7 @@ taxa_abund <-
   select(-ConsensusLineage) %>% 
   gather(Sample_ID, Abundance, 2:75) %>% 
   spread(OTU_ID, Abundance)
-gps <- read.table('data/Dimensions_GPS_dec_deg.txt', header = FALSE)
+gps <- read.table('data/brazil/Dimensions_GPS_dec_deg.txt', header = FALSE)
 colnames(gps) <- c('Sample_ID', 'Lat', 'Long')
 
 all_data <-
@@ -31,4 +31,4 @@ all_data <-
   right_join(taxa_abund, by = 'Sample_ID')
 
 all_data$library_size <- rowSums(taxa_abund[, -1])
-write_tsv(all_data, 'output/cleaned_data.tsv')
+write_tsv(all_data, 'output/dim_cleaned_data.tsv')

### 2019-06-10 var_comp_brazil.R
Add Brazil Var Comp script

@@ -1,0 +1,137 @@
+library(tidyverse)
+library(varComp)
+library(vegan)
+#library(broom)
+#library(data.table)
+library(amorris)
+
+
+## Import and organize data
+
+all_data <- read.table('output/dim_cleaned_data.tsv', header = TRUE)
+class(all_data)
+#all_data <- fread('output/dim_cleaned_data.tsv')
+all_data[1:nrow(all_data), 1:40]
+colnames(all_data[, 1:40])
+sum(is.na(all_data))
+all_data <- all_data[complete.cases(all_data$Lat), ]
+all_data <- all_data[complete.cases(all_data$Long), ]
+asv_table <- select(all_data, starts_with("asv"))
+asv_table[asv_table>0] <- 1 # Presence/Absence
+asv_mat <- as.matrix(asv_table)
+all_data <- all_data[rowSums(asv_mat) > 1, ]
+all_data <- cbind(all_data[, 1:39], all_data[, colnames(asv_mat[, colSums(asv_mat) > 1])])
+all_data[1:nrow(all_data), 1:ncol(all_data)]
+all_data
+
+### Plot ASV abundances
+#
+#i = 2
+#j = ncol(asv_table)
+#  asv_table[, c(1, i:j)] %>% 
+#  gather(ASV, ABU, 2:ncol(asv_table[, c(1, i:j)])) %>% 
+#ggplot(aes(y = ABU, x = ASV)) +
+#  geom_jitter()  
+
+
+### Estimate and plot variation in ASV abundance
+#
+#ColVar <- function(x, ...) {
+#  colSums((x - colMeans(x, ...))^2, ...)/(dim(x)[2] - 1)
+#}
+#asv_mat_var <- ColVar(asv_mat)
+#plot(density(log10(asv_mat_var)))
+#asv_mat_high_var<-asv_mat[,log10(asv_mat_var) > 0]
+#ncol(asv_mat_high_var)
+#rowSums(asv_mat)
+#colSums(asv_mat)
+#asv_mat_ns <- asv_mat[rowSums(asv_mat) > 1, ]
+#asv_mat_ns <- asv_mat[, colSums(asv_mat) > 1]
+
+
+
+# ASV matrix
+taxa <- select(select(all_data, starts_with('asv')), -1)
+# Pull out a single predictor taxon
+taxon <- taxa[, 5437]
[…]

### 2019-06-10 var_comp_brazil.R
Clean up var comp for braz and gab

@@ -22,7 +22,6 @@ asv_mat <- as.matrix(asv_table)
 all_data <- all_data[rowSums(asv_mat) > 1, ]
 all_data <- cbind(all_data[, 1:39], all_data[, colnames(asv_mat[, colSums(asv_mat) > 1])])
 all_data[1:nrow(all_data), 1:ncol(all_data)]
-all_data
 
 ### Plot ASV abundances
 #

### 2019-06-10 var_comp_gab.R

@@ -2,65 +2,67 @@ library(tidyverse)
 library(varComp)
 library(vegan)
 library(broom)
-source('R/misc_functions.R')
+library(amorris)
 
+asv_table <- read.csv('output/asv_table.csv')
 
-asv_table <- read_csv('output/asv_table.csv')
+### Plot ASV abundances
+#i = 2
+#j = ncol(asv_table)
+#  asv_table[, c(1, i:j)] %>% 
+#  gather(ASV, ABU, 2:ncol(asv_table[, c(1, i:j)])) %>% 
+#ggplot(aes(y = ABU, x = ASV)) +
+#  geom_jitter()  
 
-i = 2
-j = ncol(asv_table)
-  asv_table[, c(1, i:j)] %>% 
-  gather(ASV, ABU, 2:ncol(asv_table[, c(1, i:j)])) %>% 
-ggplot(aes(y = ABU, x = ASV)) +
-  geom_jitter()  
-
-
+## Remove ASVs only present in 1 sample
 asvs <- asv_table[, -1] 
-## Presence Absence
 asvs[asvs>0] <- 1
-asv_table[, -1] <- asvs
+asvs <- as.matrix(asvs)
+asv_table <- asv_table[rowSums(asvs) > 1, ]
 asv_mat <- as.matrix(asv_table[, 2:ncol(asv_table)])
-ColVar <- function(x, ...) {
-  colSums((x - colMeans(x, ...))^2, ...)/(dim(x)[2] - 1)
-}
 
-asv_mat_var <- ColVar(asv_mat)
-asv_mat_high_var<-asv_mat[,log10(asv_mat_var) > 0]
-ncol(asv_mat_high_var)
-plot(density(log10(asv_mat_var)))
-rowSums(asv_mat)
-colSums(asv_mat)
-asv_mat_ns <- asv_mat[, colSums(asv_mat) > 1]
+### High variance ASVs
+#ColVar <- function(x, ...) {
+#  colSums((x - colMeans(x, ...))^2, ...)/(dim(x)[2] - 1)
+#}
+#asv_mat_var <- ColVar(asv_mat)
+#asv_mat_high_var<-asv_mat[,log10(asv_mat_var) > 0]
+#ncol(asv_mat_high_var)
+#plot(density(log10(asv_mat_var)))
+#rowSums(asv_mat)
+#colSums(asv_mat)
[…]

### 2019-06-10 var_comp_brazil.R
Update var comp scripts

@@ -9,11 +9,10 @@ library(amorris)
 ## Import and organize data
 
 all_data <- read.table('output/dim_cleaned_data.tsv', header = TRUE)
-class(all_data)
 #all_data <- fread('output/dim_cleaned_data.tsv')
-all_data[1:nrow(all_data), 1:40]
-colnames(all_data[, 1:40])
-sum(is.na(all_data))
+#all_data[1:nrow(all_data), 1:40]
+#colnames(all_data[, 1:40])
+#sum(is.na(all_data))
 all_data <- all_data[complete.cases(all_data$Lat), ]
 all_data <- all_data[complete.cases(all_data$Long), ]
 asv_table <- select(all_data, starts_with("asv"))
@@ -21,7 +20,7 @@ asv_table[asv_table>0] <- 1 # Presence/Absence
 asv_mat <- as.matrix(asv_table)
 all_data <- all_data[rowSums(asv_mat) > 1, ]
 all_data <- cbind(all_data[, 1:39], all_data[, colnames(asv_mat[, colSums(asv_mat) > 1])])
-all_data[1:nrow(all_data), 1:ncol(all_data)]
+#all_data[1:nrow(all_data), 1:ncol(all_data)]
 
 ### Plot ASV abundances
 #
@@ -60,7 +59,7 @@ com_sim <- 1-com_dissim
 
 # Methane flux response variable
 CH4 <- all_data$CH4
-
+hist(CH4)
 # Calculate geographic similarity from lat/long
 xy <- all_data[, c('Lat', 'Long')]
 geo_dis <- as.matrix(dist(as.matrix(scale(xy))))

### 2019-06-10 var_comp_gab.R

@@ -60,8 +60,25 @@ env <- select(troph_attr_table, WFPS, N_percent, C_percent)
 env_dis <- as.matrix(dist(as.matrix(scale(env))))
 env_sim <- 1/(1+env_dis)
 
+
 model <- varComp(lowk ~ taxon, varcov = list(com = com_sim, geo = geo_sim, env = env_sim) )
 summary(model)
+# Zero residual variance association with community similarity
+
+mantel(com_sim, geo_sim)
+qplot(x = xy$X, y = xy$Y)
+# Community and geography are correlated and sites cluster around three locations
+troph_attr_table[troph_attr_table$Y > 30000, 'geocode'] <- 'A'
+troph_attr_table[troph_attr_table$Y < 30000 & troph_attr_table$Y > 10000, 'geocode'] <- 'B'
+troph_attr_table[troph_attr_table$Y < 10000, 'geocode'] <- 'C'
+troph_attr_table$geocode
+qplot(x = xy$X, y = xy$Y) + geom_point(aes(color = troph_attr_table$geocode))
+sample <- troph_attr_table$Sample
+model <- varComp(Low_final_k ~ taxon, 
+                 data = troph_attr_table, 
+                 varcov = list(com = com_sim, env = env_sim) )
+summary(model)
+
 
 for (i in 11:20) {
   taxon <- asv_mat_ns[, i]

### 2019-06-10 main.tex
add make

--- 2019-06-06 manuscript/main.tex
@@
-\documentclass{article} \usepackage[utf8]{inputenc}
+\documentclass{article}
 %\usepackage[utf8]{inputenc}
@@
 
-\section*{Tasks}
-
 Deadline: August 1st
@@
 
-What is novel about this? The first anyone has treated an ecosystem like an
+\section*{Tasks} 
+\begin{enumerate}
+\item What is novel about this? The first anyone has treated an ecosystem like an
 organism. The first time anyone has done this correlative approach well
 by considering environmental and spatial structure in the data. 
+\item Figure out variance partitioning (with Hannah)
+\item Maybe variance partitioning plus correlations (varPart)
+\item Do PC correction in all combinations
+\item Identify significant taxa and build a venn diagram figure (or
+something)
+\item 
+\end{enumerate}
+\section{Introduction}
+
+Microorganisms mediate many biogeochemical processes, but despite 
+much effort to link microbial communities
+to these functions the mechanism by which they regulate these
+processes remains unknown. Ecologists have investigated the effects of changing
+biodiversity on ecosystem function for decades, but many of their approaches
+are not applicable in microbial ecology due to the nature of microbial data. 
+This limits our ability to predict and manage globally relevant biogeochemical cycles. To
+address this challenge, we propose integrating current theory on community
+assembly and the functioning of ecosystems with lessons from genotype-phenotype
+mapping in quantitative genetics. This will address the
+question of how does variation in microbial community structure contribute to
+variation in the rate of ecosystem function?
+
+\subsection{Biodiversity increases ecosystem function in macrobial communities}
+
+Local species richness is correlated with ecosystem function
+\citep{hooper2005}. Most BEF research has focused on terrestrial plant 
+communities where species losses lead to decreases in community abundance or biomass
+\citep{naeem2003}. There are
+two major explanations for this phenomenon. The first is the sampling effect
+where a more diverse community is more likely to have one particularly
+productive species \citep{cardinale2006}. The second is niche complementarity where more species are
+more likely to occupy separate niches within the same ecosystem resulting in
+more efficient resource use and greater biomass.
+
+Regional species diversity may also be important for
+local ecosystem functioning when metacommunity dynamics regulate the
+diversity and composition of communities \citep{leibold2017}. 
+Typical BEF research focuses on local selection
+as a driver of ecosystem function. However, if regional
+diversity is important, then dispersal processes could alter local species diversity and composition resulting in
+a change in ecosystem function. This model is referred to as Community Assembly
+and the Functioning of Ecosystems (CAFE). Community assembly can alter ecosystem
+function if limiting dispersal reduces the likelihood of a highly productive
+species from dispersing to a local site or if high dispersal of an unproductive
+species overwhelms local selection.
+
+The CAFE approach has been demonstrated using the Price equation where a change in ecosystem function between two, matched
+observations can be attributed to species additions, species losses, and
+changes in abundance of shared species \citep{fox2006,
+bannar-martin2018}. To
+calculate these vectors, function must be attributed at the species level (e.g.
+biomass measurements of individual plant species). The CAFE framework demonstrates
+that not only local richness, 
+but also the composition of communities is important to the functioning of ecosystems.
+
+\subsection{Current approaches to microbial BEF reveal little about 
+structure-function mapping}
+
+The trajectory of microbial BEF research has differed from that of macrobial BEF
+for three reasons. First, microbial communities are
+immensely complex with orders of magnitude more taxa than macrobial communities
+and great intra-taxon variation for certain traits. Second, the way we measure
+microbes using marker gene sequences does not allow us to attribute portions of
+a function of interest to individual taxonomic groups . We can only
+measure function in aggregate across an entire community. Finally, we cannot be
+sure that the taxonomic groups that putatively perform a function are the ones
+contributing in this moment.
+
+As a result of these limitations, microbial ecologists have taken a different
+approach to assessing BEF relationships for microbial communities than others
+have for macrobial communities. This is generally a correlative approach where
+certain community attributes that can be derived from a community matrix are
+tested in terms of their relationship with function. These approaches make
+different assumptions about how microbial communities should map onto function.
+On one end of the spectrum, it is hypothesized that functional gene abundance - genes that
+code for an enzyme that performs one step of a process - should be correlated
+with the rate of the process associated with that gene. In this case, qPCR
+is used to quantify the abundance of a functional gene to test whether it is
+correlated with the rate of a process. This approach has had varied success and even the abundance of a
+functional gene that putatively performs a process is rarely correlated with
+that process \citep{rocca2015}. On the other end of the
+spectrum, investigators make few assumptions about which members of the community should
+alter function and so compute
+taxonomic richness or other diversity metrics from the entire Prokaryotic
+community inferred from 16S rRNA gene sequences. Taxonomic or functional
+diversity of the Prokaryote community generally adds little variance explained
+to models of ecosystem function \citep{graham2016}.
+
+\subsection{Biology through analogy: structure-function mapping is similar
+to genotype-phenotype mapping}
+
+The challenges described above for mapping microbial community structure to
+ecosystem function are analogous to the problem of identifying the
+genetic basis of complex traits in organismal populations. In that case,
+each individual organism has a level for a phenotype (e.g. presence
+or absence of a disease state) as well as a vector of potentially
+thousands of genetic loci with different alleles or single nucleotide
+polymorphisms (SNPs). To identify the genetic basis for this trait, we sample
+a population of organisms and try to correlate a matrix of SNPs with a
+vector of phenotypes. In most cases, there are many more loci than
+individuals, we cannot directly attribute a phenotypic state to a particular
+SNP, and we do not know the function of genes associated with these
+SNPs. This mirrors the problem of microbial structure-function mapping
+and presents a potential set of analyses that could be applied in
+microbial ecology.
+
+In addition to the nature of the data, population geneticists have other
+considerations when trying to make robust associations between phenotype
+and SNPs. For example, loci within the genome that are close together tend
+not to assort independently due to linkage disequilibrium. In addition,
+organisms exhibit population stratification due to a variety of historical
+and selective effects. For example, shared ancestry among European
+populations of humans results in an association between genotype and
+geography across Europe \citep{novembre2008}. This population structure
+can lead to false positives when trying to correlate human genotypes
+with genetic diseases or other phenotypes such as height (citation).
+
+In general, organisms exhibit population differentiation due to spatial
+and environmental patterns \citep{wright1943}. For geography, local genetic
+drift and reduced gene flow due to geographic separation can lead to population
+differentiation \citep{wright1943}. In addition, environmental distance can lead to both selection
+against migrants \citep{hendry2004} and local selection on endemic populations to
+generate variation. While
+these processes have primarily been described for macrobial populations, the
+same principles apply to microorganisms and should be considered
+in microbial structure-function mapping \citep{martiny2006}.
+
+\subsection{Reframing the question: Is variation in microbial community
+structure associated with differences in ecosystem function across space,
+time, and environments?}
+
+Typical microbial ecosystem function research focuses on whether microbial
+community composition is important at the ecosystem level
+\citep{schimel1998, cavigelli2000, balser2005, graham2016}. Given that
+microbial functional groups typically mediate biogeochemical processes
+and so the loss of a functional group will certainly alter function,
+the answer is "yes". However, this question does not address whether
+microbial community variation is actually useful for predicting or 
+managing ecosystem functions. To address that problem, we need to know
+whether variation within a functional group matters to variation in
+ecosystem function. If local selection always optimizes the available
+biodiversity to maximize process rates given the underlying environmental
+variation, then composition should not matter, because we should be able
+to predict ecosystem function from abiotic factors alone. Environmental
+treatments, common gardens, and reciprocal transplants have shown
+different functional potential for different community compositions.
+In addition, spatial patterns for microbial communities suggest that
+there should be observable diversification due to environmental and
+spatial processes \citep{wright1943, martiny2006}. Therefore, variation
+in microbial community composition seems to result in variation in 
+ecosystem function. However, in order for microbial community data to be
+useful for ecosystem modeling, the challenge is to identify specifically what variation in community composition
+is associated with function across space, time, and environmental gradients.
+To accomplish that, we need to separate the close association between environmental variation, community
+composition, and ecosystem function to identify the aspects of microbial
+biodiversity that are important for ecosystem function.
+
+\subsection*{Extra text for 1.3/1.4}
+
+\begin{enumerate}
+    \item Things exhibit population differentiation due to spatial and environmental 
+    patterns \citep{wright1943}
+    \item Microbes exhibit similar patterns \citep{martiny2006}
+    \item Pop gen tries to identify regions of the genome correlated with
+    traits
+    \item Considering ecosystems as individuals within a population presents
+    an analogous problem of determining community markers of ecosystem traits (functions)
+    \item Thus the goal of microbial BEF is to identify generalizable patterns
+    of genetic variation in microbes to feed into hypothesis testing about
+    potential mechanisms or to incorporate into ecosystem models
+\end{enumerate}
+
+Typically, microbial community structure-function research questions are framed
+as does microbial composition matter to ecosystem function? The answer to this
+question is obviously "yes" given that microorganisms mediate many of the
+processes of interest. The question then becomes does microbial community
+variation within a functional group or overall matter to ecosystem function? 
+Generally it is assumed that microbial composition is functionally irrelevant, 
+though studies of environmental treatments, common gardens, and reciprocal
+transplants have shown that different community compositions are correlated with
+different functions. The challenge with these studies is separating the close
+connection between environmental variation, community composition, and ecosystem
+function. Microorganisms cannot be useful to ecosystem modeling if there
+is no identifiable variation in composition that consistently alters function
+across space, time, and environmental gradients. The question "do we need to
+include microorganisms in ecosystem models?" depends on identifying variation in
+the community associated with function independent of the environment. If that's
+true, then the next step in understanding or modeling these processes is 
+identify what components of the community alter ecosystem function.
+
+There are a number of ways that environmental or spatial covariates could alter
+or obscure connections between microbial community structure and ecoystem
+function. For example, microbial communities exhibit strong biogeographic
+patterns \citep{martiny2006}. The historical effects of geographic proximity,
+such as dispersal limitation, could lead to spurious associations between taxa
+and function if taxa are unevenly distributed across the functional landscape.
+Environmental selection can also shape microbial community structure, so
+selection on the microbial community that is independent of the
+process could lead to erroneous conclusions about the
+underlying drivers of ecosystem function \citep{thompson2017}. For
+example, pH is known to regulate both microbial biodiversity patterns
+\citep{fierer2006} and methane cycling dynamics \citep{ye2012}, which could lead
+to microbial taxa correlated with methane cycling, but mechanistically
+unrelated. Environmental factors that directly influence function, such as
+moisture, could alter function with or without affecting the community. 
+Finally, community stratification where taxa co-occur in certain ecosystems due to
+selection or historical effects for reasons unrelated to the ecosystem function
+could lead to spurious associations.
+
+\subsection{A new framework for structure-function mapping in microbial
+communities}
+
+Microbial ecosystem function could be assessed using statistical techniques
+borrowed from quantitative genetics along with the conceptual framework for
+community assembly processes provided by metacommunity theory and CAFE. This
+would be a two-step process to first identify the portion of variation in
+ecosystem function attributable to different aspects of an ecosystem: the
+community, the underlying environmental variation, and to spatial processes.
+
+\begin{equation}
+V_F=V_E+V_S+V_C+\epsilon
+\end{equation}
+
+Ecosystem function is represented by $V_F$, $V_E$ is variation in the
+environment and covariation between the community and the environment, $V_S$ is
+variation in spatial distance between samples and covariation between the
+community and space, and $V_C$ is variation in the community independent of
+local selection and historical effects. The parameter $V_C$ also incorporates
+unmeasured environmental and spatial effects.
+
+Each of these variance components also correspond to alternative hypotheses for
+what drives ecosystem function. If most of the variation in $V_F$ is
+attributable to $V_E$, then the environment drives function either directly
+(abiotic processes) or indirectly by selecting for a particular microbial
+community that maximizes function. In this model, the community is simply a 
+conduit through which the environment acts on function. Alternatively, if $V_S$
+explains most of the variation in $V_F$ then this suggests that function is
+primarily a result of historical contingency and differences in dispersal traits.
+Finally, $V_C$ could explain most of the variation in $V_F$. This would be
+consistent with reciprocal transplant and common garden experiments showing that
+community composition drivers ecosystem function independent of the environment.
+
+This framework allows us to address the following questions: 1. Does variation
+in microbial community structure explain a measurable portion of the variation in
+ecosystem function or can all of that variation be attributed to the underlying
+effect of the environmental and spatial processes? If so, 2. Which members of
+the community are significantly correlated with function after controlling for
+the covariances of space and environment? Finally, 3. Do these taxa match our hypotheses about
+what controls ecosystem function (e.g. functional group abundance) or are there
+alternative mechanisms that need to be explored?
+
+\section{Methods}
+
+Sample 16S rRNA gene sequences, environmental covariates, geographic distance,
+ecosystem function. Use the DADA2 pipeline to assign ASVs. 
+phylogenetic and taxonomic assignment Null
+Models \citep{chase2011, stegen2015}
+
+Would it be valuable to use a null model (as in Chase or Stegen) where we
+generate random communities based on the distribution across samples and
+abundance within samples of each taxon. Then try to correlate those abundances
+with our observed rates of ecosystem function to see what distribution of
+structure-function correlations to expect given our community data. These sort
+of null community generating techniques can omit certain community assembly
+processes (such as homogenizing selection or stochastic processes)
+
+An alternative testing approach would be to use Hannah’s permutation test idea
+where we randomly permute the process rates ~1000 times and rebuild the model
+each time to see what the distribution of significant taxa is.
+
+VC - Correlations between taxa and process determined by accounting for spatial
+structure, environmental covariates, and community structure. Represents effect
+of taxonomy, unmeasured environmental variables VE - Account for community
+structure and correlate environmental variables cov(C, E) - The part of the venn
+diagram that disappears when controlling for environment and space VC:VE - Test
+the correlations between community and process rate at a variety of
+environmental conditions.  epsilon/error term - Effects of natural variation,
+unmeasured variables
+
+Another modeling options is HMSC from \cite{ovaskainen2017}
+
+\section{Results}
+
+For high-affinity methane oxidation in Gabon: We experimentally controlled some
+environmental variables and our measured environmental variables are not
+significantly correlated with process rate (C, N, WFPS) Space is correlated with
+function
+
+Be careful with attributing a certain \% of the variance to an individual taxa
+when performing partial regression on the residuals. If a taxa explains 60\% of
+the variance in the residuals after controlling for environment and space that
+doesn’t mean it explains 60\% of the overall variance.
+
+\section{Discussion}
+Combination of qPCR and pathway partinioing, qSIP \citep{hungate2015},
+and NanoSIMS \citep{mayali2012} may allow us to assign portions of
+function to particular taxa or functional groups.
+Problems that need to be addressed: We shouldn’t assume that functional groups
+that putatively perform a process are the rate-limiting step or the strongest
+correlate.  There may be unknown members of the community that perform or limit
+a process The final step of the process may not be the rate-limiting step: it
+may be substrate production or nutrient acquisition or something else (e.g.
+Qusheng Jin’s work on modeling methanogenesis) We need to consider the
+underlying biogeographic patterns shaping the distribution of microbial
+communities (à la GWAS).
+
+\section{Conclusion}
 
@@
 
-\section{Introduction}
-
-Microorganisms mediate many biogeochemical processes, but despite 
-much effort to link microbial communities
-to these functions, the mechanism by which they regulate these
-processes remains unknown. Ecologists have investigated the effect of changing
-biodiversity on ecosystem function for decades, but many of these approaches
-are not applicable in microbial ecology due to the nature of microbial data. To
-address this challenge, we propose integrating current theory on community
-assembly and the functioning of ecosystems with lessons from genotype-phenotype
-mapping in quantitative
-genetics. This will address the
-question of how does variation in microbial community structure contribute to
-variation in the rate of ecosystem function?
-
-\subsection{Biodiversity-ecosystem function in macrobial communities}
-
-The history of biodiversity-ecosystem function (BEF) research has demonstrated
-that local species richness is correlated with ecosystem function
-\citep{hooper2005}. Most studies
-have tested this idea in terrestrial plant communities where plant productivity
-exhibits a positive, asymptotic relationship with species richness
-\citep{naeem2003}. There are
-two major explanations for this relationship. The first is the sampling effect
-where a more diverse community is more likely to have one particularly
-productive species \citep{cardinale2006}. The second is niche complementarity where more species are
-more likely to occupy separate niches within the same ecosystem resulting in
-more efficient resource use and greater function.
-
-More recently, regional species diversity has been shown to be important for
-local ecosystem functioning where metacommunity dynamics can play a
-role in local BEF relationships \citep{leibold2017}. Typical BEF research focuses on local selection
-as a driver of ecosystem function. However, if regional
-diversity is important, then dispersal processes could alter local species diversity and composition resulting in
-a change in ecosystem function. This model is referred to as Community Assembly
-and the Functioning of Ecosystems (CAFE). Community assembly can alter ecosystem
-function if limiting dispersal reduces the likelihood of a highly productive
-species from dispersing to a local site or if high dispersal of an unproductive
-species overwhelms local selection.
-
-The CAFE approach has been demonstrated using the Price equation where a change in ecosystem function between two, matched
-observations can be attributed to species additions, species losses, and
-changes in abundance of shared species \citep{fox2006,
-bannar-martin2018}. To
-calculate these vectors, function must be attributed at the species level (e.g.
-biomass measurements of individual plant species). In microbial ecology, we
-could look at species turnover or changes in relative abundance between matched
-sites in space, time, or across treatments. However, we cannot apply the Price
-equation, because it is not typically possible to attribute a portion of ecosystem
-function to individual microbial taxa. We can only measure microbial ecosystem
-functions at the whole-ecosystem scale.
-
-\subsection{How are microbial communities different?}
-
-The trajectory of microbial BEF research has differed from that of macrobial BEF
-for three reasons. First, microbial communities are
-immensely complex with orders of magnitude more taxa than macrobial communities
-and great intra-taxon variation for certain traits. Second, the way we measure
-microbes using marker gene sequences does not allow us to attribute portions of
-a function of interest to individual taxonomic groups. We can only
-measure function in aggregate across an entire community. Finally, we cannot be
-sure that the taxonomic groups that putatively perform a function are the ones
-contributing in this moment. Therefore, microbial ecologists have needed
-alternative approaches to assessing microbial BEF relationships.
-
-As a result of these limitations, microbial ecologists have taken a different
-approach to assessing BEF relationships for microbial communities than others
-have for macrobial communities. This is generally a correlative approach where
-certain community attributes that can be derived from a community matrix are
-tested in terms of their relationship with function. These approaches make
-different assumptions about how microbial communities sohuld map onto function.
-On one end of the spectrum, it is hypothesized that functional gene abundance - genes that
-code for an enzyme that performs one step of a process - should be correlated
-with the rate of the process associated with that gene. In this case, qPCR
-is used to quantify the abundance of a functional gene to test whether it is
-correlated with the rate of a process. This approach has had varied success and even the abundance of a
-functional gene that putatively performs a process is rarely correlated with
-that process \citep{rocca2015}. On the other end of the
-spectrum, investigators make few assumptions about which members of the community should
-alter function and so compute
-taxonomic richness or other diversity metrics from the entire Prokaryotic
-community inferred from 16S rRNA gene sequences. Taxonomic or functional
-diversity of the Prokaryote community generally adds little variance explained
-to models of ecosystem function \citep{graham2016}.
-
-The correlative approach does not take into account many of the fundamental
-theories in ecology. For example, the history of biodiversity research has shown
-that diversity effects are often hidden variables such as treatment effects like
-nitrogen additions or differences in community composition. The recent CAFE
-literature, and more broadly the theory of metacommunity ecology, shows the
-importance of regional biodiversity in local dynamics in terms of immigration
-and emigration. Biogeography in general has demonstrated spatial patterns such
-as distance-decay and species-area relationships that could produce false
-positives when attempting to correlate structure with function. In the way that
-CAFE integrates diversity and composition of communities in addition to
-dispersal effects, microbial community structure-ecosystem function research
-should incorporate the underlying drivers of community assembly such as
-selection and dispersal.  One limitation of the CAFE framework is that it
-explicitly omits spatial processes. However, this could be incorporate by
-including geographic distance as a covariate.
-
-\subsection{Insights from genotype-phenotype mapping in population genetics}
-
-In addition to metacommunity theory, microbial ecosystem function research could
-gain from considerations in organismal biology addressing the genetic variation
-underlying complex traits. These studies use data with a similar nature to
-microbial data: a matrix of genotype levels and a vector of values of a
-phenotype where the link between genotype and phenotype is unknown. When mapping
-genotype to phenotype, population geneticists incorporate our understanding of
-genetic architecture and population stratification into genotype-phenotype
-correlations. For example, human genetic variation is structured such that
-populations closer together will tend to be more similar than populations
-farther apart \citep{novembre2008}. In case-control studies of genetic diseases, one population might
-be more likely to have the disease and also share certain genotypes even if
-those two facts are unrelated. Therefore, spatial autocorrelation must be
-considered in models of complex traits. Generally, these studies do not take
-into account environmental or spatial variability directly, but correct for
-population stratification by controlling for shared ancestry.
-
-There are a number of ways that environmental or spatial covariates could alter
-or obscure connections between microbial community structure and ecoystem
-function. For example, microbial communities exhibit strong biogeographic
-patterns \citep{martiny2006}. The historical effects of geographic proximity,
-such as dispersal limitation, could lead to spurious associations between taxa
-and function if taxa are unevenly distributed across the functional landscape.
-Environmental selection can also shape microbial community structure, so
-selection on the microbial community that is independent of the
-process could also lead to erroneous conclusions about the
-underlying drivers of ecosystem function \citep{thompson2017}. For
-example, pH is known to regulate both microbial biodiversity patterns
-\citep{fierer2006} and methane cycling dynamics \citep{ye2012} which could lead
-to microbial taxa correlated with methane cycling, but mechanistically
-unrelated. Environmental factors that directly influence function, such as
-moisture, could alter function with or without affecting the community. 
-Finally, community stratification where taxa co-occur in certain ecosystems due to
-selection or historical effects for reasons unrelated to the ecosystem function
-could lead to spurious associations
-(e.g. \cite{novembre2008} in humans).
-
-\subsection{A new framework for structure-function mapping in microbial
-communities}
-
-Microbial ecosystem function could be assessed using statistical techniques
-borrowed from quantitative genetics along with the conceptual framework for
-community assembly processes provided by metacommunity theory and CAFE. This
-would be a two-step process to first identify the portion of variation in
-ecosystem function attributable to different aspects of an ecosystem: the
-community, the underlying environmental variation, and to spatial processes.
-
-\begin{equation}
-V_F=V_E+V_S+V_C+\epsilon
-\end{equation}
-
-Ecosystem function is represented by $V_F$, $V_E$ is variation in the
-environment and covariation between the community and the environment, $V_C$ is
-variation in spatial distance between samples and covariation between the
-community and space, and $V_C$ is variation in the community independent of
-local selection and historical effects. The parameter $V_C$ also incorporates
-unmeasured environmental and spatial effects.
-
-Each of these variance components also correspond to alternative hypotheses for
-what drives ecosystem functioning. If most of the variation in $V_F$ is
-attributable to $V_E$, then the environment drives function either directly
-(abiotic processes) or indirectly by selecting for a particular microbial
-community that maximizes function. In this model, the community is simply a 
-conduit through which the environment acts on function. Alternatively, if $V_S$
-explains most of the variation in $V_F$ then this suggests that function is
-primarily a result of historical contingency and probabilistic dispersal.
-Finally, $V_C$ could explain most of the variation in $V_F$. This would be
-consistent with reciprocal transplant and common garden experiments showing that
-community composition drivers ecosystem function independent of the environment.
-
-This framework allows us to address the following questions: 1. Does variation
-in microbial community structure explain a measurable portion of the variance in
-ecosystem function or can all of that variation be attributed to the underlying
-effect of the environmental and spatial processes? If so, 2. Which members of
-the community are significantly correlated with function after controlling for
-the covariances of space and environment? Finally, 3. Do these taxa match our hypotheses about
-what controls ecosystem function (e.g. functional group abundance) or are there
-alternative mechanisms that need to be explored?
-
-\section{Methods}
-
-Sample 16S rRNA gene sequences, environmental covariates, geographic distance,
-ecosystem function. Use the DADA2 pipeline to assign ASVs. 
-phylogenetic and taxonomic assignment Null
-Models \citep{chase2011, stegen2015}
-
-Would it be valuable to use a null model (as in Chase or Stegen) where we
-generate random communities based on the distribution across samples and
-abundance within samples of each taxon. Then try to correlate those abundances
-with our observed rates of ecosystem function to see what distribution of
-structure-function correlations to expect given our community data. These sort
-of null community generating techniques can omit certain community assembly
-processes (like homogenizing selection or stochastic processes)
-
-An alternative testing approach would be to use Hannah’s permutation test idea
-where we randomly permute the process rates ~1000 times and rebuild the model
-each time to see what the distribution of significant taxa is.
-
-VC - Correlations between taxa and process determined by accounting for spatial
-structure, environmental covariates, and community structure. Represents effect
-of taxonomy, unmeasured environmental variables VE - Account for community
-structure and correlate environmental variables cov(C, E) - The part of the venn
-diagram that disappears when controlling for environment and space VC:VE - Test
-the correlations between community and process rate at a variety of
-environmental conditions.  epsilon/error term - Effects of natural variation,
-unmeasured variables
-
-Another modeling options is HMSC from \cite{ovaskainen2017}
-
-\section{Results}
-
-For high-affinity methane oxidation in Gabon: We experimentally controlled some
-environmental variables and our measured environmental variables are not
-significantly correlated with process rate (C, N, WFPS) Space is correlated with
-function
-
-Be careful with attributing a certain \% of the variance to an individual taxa
-when performing partial regression on the residuals. If a taxa explains 60\% of
-the variance in the residuals after controlling for environment and space that
-doesn’t mean it explains 60\% of the overall variance.
-
-\section{Discussion}
-Problems that need to be addressed: We shouldn’t assume that functional groups
-that putatively perform a process are the rate-limiting step or the strongest
-correlate.  There may be unknown members of the community that perform or limit
-a process The final step of the process may not be the rate-limiting step: it
-may be substrate production or nutrient acquisition or something else (e.g.
-Qusheng Jin’s work on modeling methanogenesis) We need to consider the
-underlying biogeographic patterns shaping the distribution of microbial
-communities (à la GWAS).
-
-\section{Conclusion}
-
 \section*{Possible things to include:}
@@
 function?
-
-Talk to Andy, CJ, and Jeff about recent work directly
 

### 2019-06-10 combine_data.R
housekeeping

@@ -1,35 +1,0 @@
-#### Input: troph_attr_table and rare_asv_table.csv
-#### Output: troph_total.csv
-
-library(tidyverse)
-
-source('R/misc_functions.R')
-
-#gen_attr_table <- read_csv('output/gen_attr_table.csv')
-
-troph_attr_table <- read_csv('output/troph_attr_table.csv')
-troph_attr_table$Site <- paste0(troph_attr_table$Site, 
-				reverse_substr(troph_attr_table$Sample, 3, 3))
-## Rarefied ASV Table
-troph_asv_table <- read_csv('output/rare_asv_table.csv')
-
-## DESeq2 variance stabilized asv table
-# troph_asv_table <- read_csv('output/vst_asv_table.csv')
-
-## Raw sequence counts
-#troph_asv_table <- read_csv('output/asv_table.csv')
-#asvs <- troph_asv_table[, -1] 
-### Presence Absence
-#asvs[asvs>0] <- 1
-#troph_asv_table[, -1] <- asvs
-
-#left_join(troph_attr_table, troph_asv_table, by = "Sample") %>%
-#  write_csv('output/troph_total.csv')
-geodist <- read.table('output/geodist.tsv')
-troph_attr_table <- left_join(troph_attr_table, geodist, by = "Site")
-troph_attr_table <- left_join(troph_attr_table, troph_asv_table, by = "Sample")
-  write_csv(troph_attr_table, 'output/troph_total.csv')
-
-# gen_attr_table %>% 
-# left_join(asv_table, by = "Sample") %>%
-# write_csv('output/gen_total.csv')

### 2019-06-10 rarefy_otu_table.R

@@ -1,44 +1,0 @@
-library(tidyverse)
-source('R/Rarefy_mean.R')
-source('R/misc_functions.R')
-
-# Import ASV table
-asv_table <- read_tsv('data/16S_dada2_table_RDP_tax_edit.txt', n_max = 100000)
-
-# Pull out 
-taxon_table <- 
-  asv_table %>% 
-  select(OTU_ID, ConsensusLineage) %>% 
-  mutate(OTU_ID = paste('asv', OTU_ID, sep = "_")) %>% 
-  select(asv = OTU_ID, ConsensusLineage) %>% 
-  separate(ConsensusLineage, c('Domain', 'Phylum', 'Class', 'Order', 'Family',
-                               'Genus', 'Species'), ";") %>% 
-  select(-Species)
-write_csv(taxon_table, 'output/taxon_table.csv')
-
-# Rename ASVs to remove backticks in column names, add Sample ID column
-# Convert all abundances to doubles
-wide_asv <- asv_table %>% 
-  mutate(asv = paste('asv', OTU_ID, sep = "_")) %>% 
-  select(asv, starts_with('dna')) %>% 
-  select(asv, ends_with('MO')) %>% 
-  gather(Sample, abund, 2:length(.)) %>% 
-  spread(asv, abund) %>% 
-  mutate(Sample = substring(Sample, 5))
-wide_asv <- wide_asv[, colSums(wide_asv != 0, na.rm = T) > 0]
-
-comm_mat <- as.matrix(wide_asv[, -1])
-
-samples <- tibble(Sample = wide_asv$Sample)
-
-rare_mat <- rarefy.mean(comm_mat, n = 1000)
-
-rare_asv <- cbind(samples, as_tibble(rare_mat))
-
-wide_asv <- cbind(samples, as_tibble(comm_mat))
-
-
-write_csv(wide_asv, 'output/asv_table.csv')
-
-write_csv(rare_asv, 'output/rare_asv_table.csv')
-

### 2019-06-11 R/gab_cleaning.R
Add Gabon raw metadata and processing scripts

@@ -1,0 +1,128 @@
+
+library(tidyverse)
+
+source('R/misc_functions.R')
+
+# Import ------------------------------------------
+raw_troph_data <- read_tsv('data/Gabon_methanotrophy_table_scaled.txt')
+raw_gen_data <- read_tsv('data/Gabon_methanogenesis_table_scaled.txt')
+
+# Organize Methanotroph ------------------------------------------
+troph_data <- raw_troph_data %>% 
+  dplyr::select(-Low_r2, -High_r2, -High_final_k) %>% 
+  dplyr::select(Vmax, Low_final_k, Bulk_dens:PC3_16S_rna) %>% 
+  mutate(Rep = reverse_substr(Sample_names, 3, 3)) %>% 
+  mutate_at(vars(Site, Description, Rep, Land_type, Sample_names, Experiment), funs(factor)) %>% 
+  select(Sample = Sample_names, Site, Rep, Experiment, Description, Land_type, 
+         Vmax:C_percent, pmoa_copy_num:PC3_16S_rna)
+
+# Add Wetland variable
+troph_data %<>% left_join(data_frame(Land_type = unique(troph_data$Land_type), 
+           Wetland = factor(c('Wetland', 'Upland', 'Upland', 'Wetland', 'Wetland',
+                              'Wetland', 'Upland', 'Upland', 'Upland'))), by = 'Land_type') %>% 
+  select(Sample:Experiment, Wetland, Description:PC3_16S_rna)
+
+# Add WFPS variable
+Site_Pd <- data_frame(
+  Site = factor(unique(troph_data$Site)),
+  Pd = c(2.65, 2.65, 2.65, 2.65, 1.55, 2.65, 1.55, 2.65, 2.65, 2.65, 2.65, 2.65, 2.65, 2.65, 2.65)
+)
+
+troph_data %<>% left_join(Site_Pd, by = 'Site') %>% 
+  mutate(WFPS = calc_wfps(Bulk_dens, Mois_cont, Pd)) %>% 
+  select(Sample:Low_final_k, WFPS, Bulk_dens:PC3_16S_rna, -Pd)
+
+# Separate numeric data (remove factors)
+troph_num_data <- troph_data %>% 
+  select_if(function(col) !is.factor(col))
+
+# Remove response variables
+num_predictors <- troph_num_data %>% dplyr::select(-Vmax, -Low_final_k)
+
+# Attribute table
[…]

### 2019-06-11 R/gab_combine_data.R

@@ -2,8 +2,7 @@
 #### Output: troph_total.csv
 
 library(tidyverse)
-
-source('R/misc_functions.R')
+library(amorris)
 
 #gen_attr_table <- read_csv('output/gen_attr_table.csv')
 
@@ -28,7 +27,7 @@ troph_asv_table <- read_csv('output/rare_asv_table.csv')
 geodist <- read.table('output/geodist.tsv')
 troph_attr_table <- left_join(troph_attr_table, geodist, by = "Site")
 troph_attr_table <- left_join(troph_attr_table, troph_asv_table, by = "Sample")
-  write_csv(troph_attr_table, 'output/troph_total.csv')
+  write_csv(troph_attr_table, '../output/gab_troph_total.csv')
 
 # gen_attr_table %>% 
 # left_join(asv_table, by = "Sample") %>%

### 2019-06-11 R/gab_clean_asvs.R
Add and revise data cleaning scripts

@@ -1,0 +1,31 @@
+library(tidyverse)
+library(amorris)
+
+# Import ASV table
+asv_table <- read_tsv('../data/gabon/16S_dada2_table_RDP_tax_edit.txt', n_max = 100000)
+
+# Pull out taxonomy
+taxon_table <- 
+  asv_table %>% 
+  select(OTU_ID, ConsensusLineage) %>% 
+  mutate(OTU_ID = paste('asv', OTU_ID, sep = "_")) %>% 
+  select(asv = OTU_ID, ConsensusLineage) %>% 
+  separate(ConsensusLineage, c('Domain', 'Phylum', 'Class', 'Order', 'Family',
+                               'Genus', 'Species'), ";") %>% 
+  select(-Species)
+write_csv(taxon_table, '../output/gab_taxon_table.csv')
+
+# Rename ASVs to remove backticks in column names, add Sample ID column
+# Convert all abundances to doubles
+wide_asv <- asv_table %>% 
+  mutate(asv = paste('asv', OTU_ID, sep = "_")) %>% 
+  select(asv, starts_with('dna')) %>% 
+  select(asv, ends_with('MO')) %>% 
+  gather(Sample, abund, 2:length(.)) %>% 
+  spread(asv, abund) %>% 
+  mutate(Sample = substring(Sample, 5))
+
+# Remove zero abundance taxa and NA taxa
+wide_asv <- wide_asv[, colSums(wide_asv != 0, na.rm = T) > 0]
+
+write_csv(wide_asv, '../output/gab_asv_table.csv')

### 2019-06-11 R/gab_cleaning.R

@@ -1,11 +1,10 @@
 
 library(tidyverse)
-
-source('R/misc_functions.R')
+library(amorris)
 
 # Import ------------------------------------------
-raw_troph_data <- read_tsv('data/Gabon_methanotrophy_table_scaled.txt')
-raw_gen_data <- read_tsv('data/Gabon_methanogenesis_table_scaled.txt')
+raw_troph_data <- read_tsv('../data/gabon/Gabon_methanotrophy_table_scaled.txt')
+raw_gen_data <- read_tsv('../data/gabon/Gabon_methanogenesis_table_scaled.txt')
 
 # Organize Methanotroph ------------------------------------------
 troph_data <- raw_troph_data %>% 
@@ -123,6 +122,6 @@ num_predictors <- gen_num_data %>% dplyr::select(-CH4, -H_CH4_percent)
 gen_attr_table <- gen_data %>% 
   select(-Rep, -Experiment, -Description, -Land_type, -Bulk_dens, -Mois_cont)
 
-write_csv(gen_attr_table, 'output/gen_attr_table.csv')
+write_csv(gen_attr_table, '../output/gab_gen_attr_table.csv')
 
-write_csv(troph_attr_table, 'output/troph_attr_table.csv')
+write_csv(troph_attr_table, '../output/gab_troph_attr_table.csv')

### 2019-06-11 R/gab_combine_data.R

@@ -1,22 +1,19 @@
-#### Input: troph_attr_table and rare_asv_table.csv
-#### Output: troph_total.csv
-
 library(tidyverse)
 library(amorris)
 
-#gen_attr_table <- read_csv('output/gen_attr_table.csv')
+#gen_attr_table <- read_csv('../output/gab_gen_attr_table.csv')
 
-troph_attr_table <- read_csv('output/troph_attr_table.csv')
+troph_attr_table <- read_csv('../output/gab_troph_attr_table.csv')
 troph_attr_table$Site <- paste0(troph_attr_table$Site, 
 				reverse_substr(troph_attr_table$Sample, 3, 3))
 ## Rarefied ASV Table
-troph_asv_table <- read_csv('output/rare_asv_table.csv')
+troph_asv_table <- read_csv('../output/rare_asv_table.csv')
 
 ## DESeq2 variance stabilized asv table
-# troph_asv_table <- read_csv('output/vst_asv_table.csv')
+# troph_asv_table <- read_csv('../output/vst_asv_table.csv')
 
 ## Raw sequence counts
-#troph_asv_table <- read_csv('output/asv_table.csv')
+#troph_asv_table <- read_csv('output/../asv_table.csv')
 #asvs <- troph_asv_table[, -1] 
 ### Presence Absence
 #asvs[asvs>0] <- 1
@@ -24,7 +21,7 @@ troph_asv_table <- read_csv('output/rare_asv_table.csv')
 
 #left_join(troph_attr_table, troph_asv_table, by = "Sample") %>%
 #  write_csv('output/troph_total.csv')
-geodist <- read.table('output/geodist.tsv')
+geodist <- read.table('../output/geodist.tsv')
 troph_attr_table <- left_join(troph_attr_table, geodist, by = "Site")
 troph_attr_table <- left_join(troph_attr_table, troph_asv_table, by = "Sample")
   write_csv(troph_attr_table, '../output/gab_troph_total.csv')

### 2019-06-11 R/rarefy_asv_table.R

@@ -1,44 +1,18 @@
 library(tidyverse)
-source('R/Rarefy_mean.R')
-source('R/misc_functions.R')
+library(amorris)
 
 # Import ASV table
-asv_table <- read_tsv('data/16S_dada2_table_RDP_tax_edit.txt', n_max = 100000)
+asv_table <- read.csv('../output/gab_asv_table.csv')
 
-# Pull out 
-taxon_table <- 
-  asv_table %>% 
-  select(OTU_ID, ConsensusLineage) %>% 
-  mutate(OTU_ID = paste('asv', OTU_ID, sep = "_")) %>% 
-  select(asv = OTU_ID, ConsensusLineage) %>% 
-  separate(ConsensusLineage, c('Domain', 'Phylum', 'Class', 'Order', 'Family',
-                               'Genus', 'Species'), ";") %>% 
-  select(-Species)
-write_csv(taxon_table, 'output/taxon_table.csv')
+comm_mat <- as.matrix(asv_table[, -1])
 
-# Rename ASVs to remove backticks in column names, add Sample ID column
-# Convert all abundances to doubles
-wide_asv <- asv_table %>% 
-  mutate(asv = paste('asv', OTU_ID, sep = "_")) %>% 
-  select(asv, starts_with('dna')) %>% 
-  select(asv, ends_with('MO')) %>% 
-  gather(Sample, abund, 2:length(.)) %>% 
-  spread(asv, abund) %>% 
-  mutate(Sample = substring(Sample, 5))
-wide_asv <- wide_asv[, colSums(wide_asv != 0, na.rm = T) > 0]
+hist(rowSums(comm_mat))
 
-comm_mat <- as.matrix(wide_asv[, -1])
+samples <- tibble(Sample = asv_table$Sample)
 
-samples <- tibble(Sample = wide_asv$Sample)
-
-rare_mat <- rarefy.mean(comm_mat, n = 1000)
+rare_mat <- rarefy_mean(comm_mat, n = 1000)
 
 rare_asv <- cbind(samples, as_tibble(rare_mat))
 
-wide_asv <- cbind(samples, as_tibble(comm_mat))
-
-
-write_csv(wide_asv, 'output/asv_table.csv')
-
-write_csv(rare_asv, 'output/rare_asv_table.csv')
+write_csv(rare_asv, '../output/gab_rare_asv_table.csv')
 

### 2019-06-11 R/gab_rarefy_asv_table.R
Drop 2 samples from rarefy

@@ -4,9 +4,13 @@ library(amorris)
 # Import ASV table
 asv_table <- read.csv('../output/gab_asv_table.csv')
 
-comm_mat <- as.matrix(asv_table[, -1])
+# Check library size
+rowSums(as.matrix(asv_table[, -1]))
 
-hist(rowSums(comm_mat))
+# Remove two samples with shallow sequencing depth
+asv_table <- asv_table[rowSums(as.matrix(asv_table[, -1])) > 10000, ]
+
+comm_mat <- as.matrix(asv_table[, -1])
 
 samples <- tibble(Sample = asv_table$Sample)
 

### 2019-06-12 R/PC_correction.R
Move functions to one script

@@ -6,6 +6,8 @@
 #### similarity, and phenotype ~ environmental similarity
 #### See: Price et al. 2006 in Nature Genetics
 
+library(vegan)
+source('functions.R')
 # Whether to correct by each factor
 com_bool <- T
 env_bool <- T
@@ -20,8 +22,8 @@ scale_env <- T
 scale_spa <- T
 # File names
 input_sep <- ","
-input_file <- "output/troph_total.csv"
-output_file <- "output/gab_adj"
+input_file <- "../output/gab_troph_total.csv"
+output_file <- "../output/gab_adj"
 # Function column names
 fun_cols <- c('Low_final_k', 'Vmax')
 # Environment column names
@@ -33,7 +35,6 @@ com_cols <- 'asv'
 # Extra grouping variables to preserve
 group_cols <- c('Sample', 'Wetland', 'Site')
 
-library(vegan)
 
 all_data <- read.table(input_file, header = TRUE, sep = input_sep)
 
@@ -55,22 +56,6 @@ com_mat <- as.matrix(all_data[, grepl(com_cols, names(all_data))])
 fun_mat <- as.matrix(all_data[, fun_cols])
 fun_com <- cbind(fun_mat, com_mat)
 
-#### Functions
-pc_adjust_mat <- function(raw_mat, aj, n_axes) {
-  # raw_mat is the matrix of function and community
-  # aj is the matrix of site scores from PCA
-  # n_axes is the number of PC axes to correct by
-  gamma <- vector(length = nrow(raw_mat))
-	for (p in 1:n_axes) {
-    a <- aj[, p] # select axis n for correction
-    for (i in 1:ncol(raw_mat)) { # adjust each taxon abundance
-      tax <- raw_mat[, i] # select taxon abundances for j samples
-      gamma[i] <- sum(a * tax)/sum(a^2) # calculate gamma (regression coef)
-      raw_mat[, i] <- tax - gamma[i] * a 
-    }
-  }
-  return(raw_mat)
-}
 
 ### Community similarity correction
 if (com_bool) {

### 2019-06-12 R/functions.R

@@ -1,0 +1,57 @@
+
+##### Functions
+
+lowk_model <- function(df) {
+  lm(Low_final_k ~ abund, data = df)
+}
+vmax_model <- function(df) {
+  lm(Vmax ~ abund, data = df)
+}
+# Create nested data set by asv
+  # Calculate linear model for each asv between abundance and function
+fit_models <- function(data, model) {
+  data %>% 
+    gather(asv, abund, starts_with('asv_')) %>% 
+    group_by(asv) %>% 
+    nest() %>% 
+    mutate(model = map(data, model)) %>% 
+    return()
+}
+#### Functions
+pc_adjust_mat <- function(raw_mat, aj, n_axes) {
+  # raw_mat is the matrix of function and community
+  # aj is the matrix of site scores from PCA
+  # n_axes is the number of PC axes to correct by
+  gamma <- vector(length = nrow(raw_mat))
+	for (p in 1:n_axes) {
+    a <- aj[, p] # select axis n for correction
+    for (i in 1:ncol(raw_mat)) { # adjust each taxon abundance
+      tax <- raw_mat[, i] # select taxon abundances for j samples
+      gamma[i] <- sum(a * tax)/sum(a^2) # calculate gamma (regression coef)
+      raw_mat[, i] <- tax - gamma[i] * a 
+    }
+  }
+  return(raw_mat)
+}
+
+man_plot <- function(by_asv, taxon_table) {
+
+by_asv %>% 
+  mutate(tidy = map(model, broom::tidy)) %>% 
+  unnest(tidy) %>% 
+  filter(term == 'abund') %>%
+  left_join(taxon_table, by = 'asv') %>% 
+  ggplot(aes(x = reorder(asv, Phylum, median), y = -log10(p.value))) +
+  geom_point(aes(color=Phylum), size = 1) +
+  geom_hline(yintercept = -log10(0.05/nrow(by_asv))) +
+  scale_color_discrete(guide = F) +
+  labs(x = 'Amplicon Sequence Variant (ASV)', 
+       y = expression(Significance ~ (-log[10](p.value)))) +
+    #  geom_text_repel(aes(label=ifelse(-log10(p.value)>-log10(alpha),as.character(genus),'')), size = 1) +
+    # geom_text(aes(label=ifelse(-log10(p.value)>-log10(alpha),as.character(genus),'')), position = position_dodge(width = 10)) +
+  theme_classic() +
[…]

### 2019-06-12 R/gab_lm_model.R

@@ -1,0 +1,32 @@
+library(tidyverse)
+library(purrr)
+library(broom)
+library(modelr)
+library(amorris)
+source('functions.R')
+
+
+adj_data_c <- read.table('../output/gab_adj_com.tsv')
+adj_data_e <- read.table('../output/gab_adj_env.tsv')
+adj_data_s <- read.table('../output/gab_adj_spa.tsv')
+adj_data_ce <- read.table('../output/gab_adj_com_env.tsv')
+adj_data_cs <- read.table('../output/gab_adj_com_spa.tsv')
+adj_data_es <- read.table('../output/gab_adj_env_spa.tsv')
+adj_data_ces <- read.table('../output/gab_adj_com_env_spa.tsv')
+
+adj_data <- list(c = adj_data_c, e = adj_data_e, s = adj_data_s, 
+                 ce = adj_data_ce, cs = adj_data_cs, es = adj_data_es,
+                 ces = adj_data_ces)
+lapply(names(adj_data), function(x) saveRDS(fit_models(adj_data[[x]], lowk_model), 
+                                     file = paste0("../output/lowk_fits_", x, 
+                                                   ".rds")))
+saveRDS(fit_models(adj_data_c, lowk_model), file = "../output/lowk_fits_c.rds")
+saveRDS(fit_models(com_adj_data, lowk_model), file = "../output/lowk_fits_com.rds")
+saveRDS(fit_models(env_adj_data, lowk_model), file = "../output/lowk_fits_env.rds")
+saveRDS(fit_models(spa_adj_data, lowk_model), file = "../output/lowk_fits_spa.rds")
+saveRDS(fit_models(troph_total, lowk_model), file = "../output/lowk_fits_raw.rds")
+saveRDS(fit_models(adj_data, vmax_model), file = "../output/vmax_fits.rds")
+saveRDS(fit_models(com_adj_data, vmax_model), file = "output/vmax_fits_com.rds")
+saveRDS(fit_models(env_adj_data, vmax_model), file = "output/vmax_fits_env.rds")
+saveRDS(fit_models(spa_adj_data, vmax_model), file = "output/vmax_fits_spa.rds")
+saveRDS(fit_models(troph_total, vmax_model), file = "output/vmax_fits_raw.rds")

### 2019-06-12 R/lm_model.R

@@ -1,45 +1,0 @@
-#### Input: adj_data.csv
-#### Output: model_fits.rds
-
-library(tidyverse)
-library(purrr)
-library(broom)
-library(modelr)
-
-adj_data <- read_csv('output/adj_data.csv')
-com_adj_data <- read_csv('output/com_adj_data.csv')
-env_adj_data <- read_csv('output/env_adj_data.csv')
-spa_adj_data <- read_csv('output/spa_adj_data.csv')
-troph_total <- read_csv('output/troph_total.csv')
-
-
-
-##### Functions
-
-lowk_model <- function(df) {
-  lm(Low_final_k ~ abund, data = df)
-}
-vmax_model <- function(df) {
-  lm(Vmax ~ abund, data = df)
-}
-# Create nested data set by asv
-  # Calculate linear model for each asv between abundance and function
-fit_models <- function(data, model) {
-  data %>% 
-    gather(asv, abund, starts_with('asv_')) %>% 
-    group_by(asv) %>% 
-    nest() %>% 
-    mutate(model = map(data, model)) %>% 
-    return()
-}
-
-saveRDS(fit_models(adj_data, lowk_model), file = "output/lowk_fits.rds")
-saveRDS(fit_models(com_adj_data, lowk_model), file = "output/lowk_fits_com.rds")
-saveRDS(fit_models(env_adj_data, lowk_model), file = "output/lowk_fits_env.rds")
-saveRDS(fit_models(spa_adj_data, lowk_model), file = "output/lowk_fits_spa.rds")
-saveRDS(fit_models(troph_total, lowk_model), file = "output/lowk_fits_raw.rds")
-saveRDS(fit_models(adj_data, vmax_model), file = "output/vmax_fits.rds")
-saveRDS(fit_models(com_adj_data, vmax_model), file = "output/vmax_fits_com.rds")
-saveRDS(fit_models(env_adj_data, vmax_model), file = "output/vmax_fits_env.rds")
-saveRDS(fit_models(spa_adj_data, vmax_model), file = "output/vmax_fits_spa.rds")
-saveRDS(fit_models(troph_total, vmax_model), file = "output/vmax_fits_raw.rds")

### 2019-06-12 R/gab_combine_data.R
Clean up and update makefile

@@ -3,11 +3,11 @@ library(amorris)
 
 #gen_attr_table <- read_csv('../output/gab_gen_attr_table.csv')
 
-troph_attr_table <- read_csv('../output/gab_troph_attr_table.csv')
+troph_attr_table <- read.csv('../output/gab_troph_attr_table.csv', stringsAsFactors=F)
 troph_attr_table$Site <- paste0(troph_attr_table$Site, 
 				reverse_substr(troph_attr_table$Sample, 3, 3))
 ## Rarefied ASV Table
-troph_asv_table <- read_csv('../output/rare_asv_table.csv')
+troph_asv_table <- read.csv('../output/gab_rare_asv_table.csv')
 
 ## DESeq2 variance stabilized asv table
 # troph_asv_table <- read_csv('../output/vst_asv_table.csv')
@@ -23,8 +23,8 @@ troph_asv_table <- read_csv('../output/rare_asv_table.csv')
 #  write_csv('output/troph_total.csv')
 geodist <- read.table('../output/geodist.tsv')
 troph_attr_table <- left_join(troph_attr_table, geodist, by = "Site")
-troph_attr_table <- left_join(troph_attr_table, troph_asv_table, by = "Sample")
-  write_csv(troph_attr_table, '../output/gab_troph_total.csv')
+troph_attr_table <- right_join(troph_attr_table, troph_asv_table, by = "Sample")
+write_csv(troph_attr_table, '../output/gab_troph_total.csv')
 
 # gen_attr_table %>% 
 # left_join(asv_table, by = "Sample") %>%

### 2019-06-12 R/gab_var_comp.R

@@ -4,7 +4,7 @@ library(vegan)
 library(broom)
 library(amorris)
 
-asv_table <- read.csv('output/asv_table.csv')
+asv_table <- read.csv('../output/gab_asv_table.csv')
 
 ### Plot ASV abundances
 #i = 2
@@ -41,15 +41,14 @@ com_dissim <- as.matrix(com_dissim)
 com_sim <- 1-com_dissim
 
 
-troph_attr_table <- read.csv('output/troph_attr_table.csv', stringsAsFactors = F)
-str(troph_attr_table)
+troph_attr_table <- read.csv('../output/troph_attr_table.csv', stringsAsFactors = F)
 troph_attr_table$Site <- paste0(troph_attr_table$Site, 
 				reverse_substr(troph_attr_table$Sample, 3, 3))
 # Function
 lowk <- troph_attr_table$Low_final_k
 
 # Geographic similarity
-geodist <- read.table('output/geodist.tsv')
+geodist <- read.table('../output/geodist.tsv')
 troph_attr_table <- left_join(troph_attr_table, geodist, by = "Site")
 xy <- troph_attr_table[, c('X', 'Y')]
 geo_dis <- as.matrix(dist(as.matrix(scale(xy))))

### 2019-06-13 R/plot_sig_taxa.R
Separate plotting script from id taxa script

@@ -1,0 +1,124 @@
+
+new_data <- read_csv('output/adj_data.csv')
+
+alpha <- 0.05 / nrow(by_asv_all)
+
+subset_sig_data <- function(
+, taxon_table, new_data, file_name
+# Pull out those ASVs
+important_asvs <- new_data %>% 
+  select(Low_final_k, one_of(lowk_asvs))%>% 
+  gather(asv, abund, starts_with('asv'))
+taxon_table %>% 
+  filter(asv %in% lowk_asvs) %>% 
+  select(-asv) %>%
+  xtable() %>% 
+  print.xtable(file = file_name)
+
+
+id_asvs(by_asv_raw, alpha, taxon_table, new_data, file_name = "tables/table_rare_raw.tex")
+id_asvs(by_asv_env, alpha, taxon_table, new_data, file_name = "tables/table_rare_env.tex")
+id_asvs(by_asv_spa, alpha, taxon_table, new_data, file_name = "tables/table_rare_spa.tex")
+id_asvs(by_asv_com, alpha, taxon_table, new_data, file_name = "tables/table_rare_com.tex")
+id_asvs(by_asv_all, alpha, taxon_table, new_data, file_name = "tables/table_rare_all.tex")
+
+
+## Individual correlation
+imp_tax <- taxon_table %>% 
+  filter(asv %in% important_asvs$asv)
+imp_tax$taxon <- apply(imp_tax, 1, function(r) { r[ 
+                      if( any(is.na(r))){ 
+                          (which(is.na(r))[1]-1) 
+                       }else{
+                          (length(r)-0)}
+                                      ] }
+       )
+imp_tax <- imp_tax %>% select(asv, taxon)
+imp_tax$index <- seq(from=1, to=nrow(imp_tax))
+#imp_tax$level <- as.character(1:5)
+imp_data <-
+	new_data %>% 
+	select(Low_final_k, one_of(imp_asv_list$asv))
+p <- imp_data %>% 
+	gather(asv, abund, 2:ncol(imp_data)) %>% 
+	left_join(imp_tax) %>% 
+	ggplot(aes(x = abund, y = Low_final_k)) + 
+	       geom_point() +
+	       facet_wrap(. ~ index)+
+	       theme_bw() +
+	       labs(x = 'ASV Abundance', y = 'Methane Oxidation')
+p
[…]

### 2019-06-13 R/PC_correction.R
Make the modeling pipeline more awesome

@@ -92,41 +92,43 @@ spa_aj <- summary(env_pca)$sites # pull out site scores, which represent
 if (com_bool) {
 # Community correction
 all_adj <- pc_adjust_mat(fun_com, com_aj, n_com_axes) 
-write.table(all_adj, paste0(output_file, '_com.tsv'))
+write.table(all_adj, paste0(output_file, '_c.tsv'))
 }
 
 if (env_bool) {
 all_adj <- pc_adjust_mat(fun_com, env_aj, n_env_axes) 
-write.table(all_adj, paste0(output_file, '_env.tsv'))
+write.table(all_adj, paste0(output_file, '_e.tsv'))
 }
 
 if (spa_bool) {
 all_adj <- pc_adjust_mat(fun_com, spa_aj, n_spa_axes) 
-write.table(all_adj, paste0(output_file, '_spa.tsv'))
+write.table(all_adj, paste0(output_file, '_s.tsv'))
 }
 
 if (com_bool & env_bool) {
 all_adj <- pc_adjust_mat(fun_com, com_aj, n_com_axes) 
 all_adj <- pc_adjust_mat(all_adj, env_aj, n_env_axes) 
 all_adj <- cbind(all_data[, group_cols], all_adj)
-write.table(all_adj, paste0(output_file, '_com_env.tsv'))
+write.table(all_adj, paste0(output_file, '_ce.tsv'))
 }
 if (com_bool & spa_bool) {
 all_adj <- pc_adjust_mat(fun_com, com_aj, n_com_axes) 
 all_adj <- pc_adjust_mat(all_adj, spa_aj, n_spa_axes) 
 all_adj <- cbind(all_data[, group_cols], all_adj)
-write.table(all_adj, paste0(output_file, '_com_spa.tsv'))
+write.table(all_adj, paste0(output_file, '_cs.tsv'))
 }
 if (env_bool & spa_bool) {
 all_adj <- pc_adjust_mat(fun_com, env_aj, n_env_axes) 
 all_adj <- pc_adjust_mat(all_adj, spa_aj, n_spa_axes) 
 all_adj <- cbind(all_data[, group_cols], all_adj)
-write.table(all_adj, paste0(output_file, '_env_spa.tsv'))
+write.table(all_adj, paste0(output_file, '_es.tsv'))
 }
 if (com_bool & env_bool & spa_bool) {
 all_adj <- pc_adjust_mat(fun_com, com_aj, n_com_axes) 
[…]

### 2019-06-13 R/functions.R

@@ -55,3 +55,18 @@ by_asv %>%
   ylim(0, 10)  %>% 
   return()
 }
+# Identify ASVs with -log(p value) > 
+id_asvs <- function(model_output, alpha = 0.05/nrow(model_output)) {
+lowk_asvs <- model_output %>% 
+  mutate(tidy = map(model, broom::tidy)) %>% 
+  unnest(tidy) %>% 
+  filter(term == 'abund') %>% 
+  filter(-log10(p.value) > -log10(alpha)) %>% 
+  pull(asv)
+}
+
+# Pull out taxonomy for significant taxa
+id_taxonomy <- function(sig_asvs, taxon_table) {
+  taxon_table %>% 
+    filter(asv %in% sig_asvs) 
+}

### 2019-06-13 R/gab_lm_model.R

@@ -5,28 +5,12 @@ library(modelr)
 library(amorris)
 source('functions.R')
 
-
-adj_data_c <- read.table('../output/gab_adj_com.tsv')
-adj_data_e <- read.table('../output/gab_adj_env.tsv')
-adj_data_s <- read.table('../output/gab_adj_spa.tsv')
-adj_data_ce <- read.table('../output/gab_adj_com_env.tsv')
-adj_data_cs <- read.table('../output/gab_adj_com_spa.tsv')
-adj_data_es <- read.table('../output/gab_adj_env_spa.tsv')
-adj_data_ces <- read.table('../output/gab_adj_com_env_spa.tsv')
-
-adj_data <- list(c = adj_data_c, e = adj_data_e, s = adj_data_s, 
-                 ce = adj_data_ce, cs = adj_data_cs, es = adj_data_es,
-                 ces = adj_data_ces)
+# Import all model output
+temp <- list.files(path="../output", pattern="gab_adj*", full.names = TRUE)
+adj_data <- sapply(temp, read.table, simplify = FALSE, USE.NAMES = TRUE)
+# Rename data subsets
+names(adj_data) <- reverse_substr(substr(names(adj_data), 19, 25), 5, 7)
+# Fit models and save model outputs
 lapply(names(adj_data), function(x) saveRDS(fit_models(adj_data[[x]], lowk_model), 
                                      file = paste0("../output/lowk_fits_", x, 
                                                    ".rds")))
-saveRDS(fit_models(adj_data_c, lowk_model), file = "../output/lowk_fits_c.rds")
-saveRDS(fit_models(com_adj_data, lowk_model), file = "../output/lowk_fits_com.rds")
-saveRDS(fit_models(env_adj_data, lowk_model), file = "../output/lowk_fits_env.rds")
-saveRDS(fit_models(spa_adj_data, lowk_model), file = "../output/lowk_fits_spa.rds")
-saveRDS(fit_models(troph_total, lowk_model), file = "../output/lowk_fits_raw.rds")
-saveRDS(fit_models(adj_data, vmax_model), file = "../output/vmax_fits.rds")
-saveRDS(fit_models(com_adj_data, vmax_model), file = "output/vmax_fits_com.rds")
-saveRDS(fit_models(env_adj_data, vmax_model), file = "output/vmax_fits_env.rds")
[…]

### 2019-06-13 R/id_taxa.R

@@ -1,139 +1,24 @@
 library(tidyverse)
 library(xtable)
+library(amorris)
+source('functions.R')
+
 # import model output
-by_asv_all <- readRDS('output/model_fits.rds')
-by_asv_raw <- readRDS('output/model_fits_raw.rds')
-by_asv_com <- readRDS('output/model_fits_com.rds')
-by_asv_env <- readRDS('output/model_fits_env.rds')
-by_asv_spa <- readRDS('output/model_fits_spa.rds')
+temp <- list.files(path="../output", pattern="lowk_fits*", full.names = TRUE)
+all_models <- sapply(temp, readRDS, simplify = FALSE, USE.NAMES = TRUE)
+names(all_models) <- reverse_substr(substr(names(all_models), 21, 27), 5, 8)
 
-new_data <- read_csv('output/adj_data.csv')
-taxon_table <- read_csv('output/taxon_table.csv')
+taxon_table <- read_csv('../output/taxon_table.csv')
 
-alpha <- 0.05 / nrow(by_asv_all)
-# Identify ASVs with -log(p value) > 
-id_asvs <- function(by_asv, alpha, taxon_table, new_data, file_name) {
-lowk_asvs <- by_asv %>% 
-  mutate(tidy = map(model, broom::tidy)) %>% 
-  unnest(tidy) %>% 
-  filter(term == 'abund') %>% 
-  filter(-log10(p.value) > -log10(alpha)) %>% 
-  pull(asv)
+# Pull out sig ASVs
+asvs <- lapply(all_models, id_asvs)
 
-# Pull out those ASVs
-important_asvs <- new_data %>% 
-  select(Low_final_k, one_of(lowk_asvs))%>% 
-  gather(asv, abund, starts_with('asv'))
-taxon_table %>% 
-  filter(asv %in% lowk_asvs) %>% 
-  select(-asv) %>%
-  xtable() %>% 
-  print.xtable(file = file_name)
-}
+# Number of significant ASVs
+lapply(asvs, length)
 
-id_asvs(by_asv_raw, alpha, taxon_table, new_data, file_name = "tables/table_rare_raw.tex")
-id_asvs(by_asv_env, alpha, taxon_table, new_data, file_name = "tables/table_rare_env.tex")
-id_asvs(by_asv_spa, alpha, taxon_table, new_data, file_name = "tables/table_rare_spa.tex")
[…]

### 2019-06-13 R/gab_clean_asvs.R
Add gabon methanogen data cleaning

@@ -20,12 +20,12 @@ write_csv(taxon_table, '../output/gab_taxon_table.csv')
 wide_asv <- asv_table %>% 
   mutate(asv = paste('asv', OTU_ID, sep = "_")) %>% 
   select(asv, starts_with('dna')) %>% 
-  select(asv, ends_with('MO')) %>% 
   gather(Sample, abund, 2:length(.)) %>% 
   spread(asv, abund) %>% 
   mutate(Sample = substring(Sample, 5))
 
 # Remove zero abundance taxa and NA taxa
 wide_asv <- wide_asv[, colSums(wide_asv != 0, na.rm = T) > 0]
+dim(wide_asv)
 
 write_csv(wide_asv, '../output/gab_asv_table.csv')

### 2019-06-13 R/gab_combine_data.R

@@ -1,13 +1,15 @@
 library(tidyverse)
 library(amorris)
 
-#gen_attr_table <- read_csv('../output/gab_gen_attr_table.csv')
+gen_attr_table <- read_csv('../output/gab_gen_attr_table.csv')
 
+gen_attr_table$Site <- paste0(gen_attr_table$Site, 
+				reverse_substr(gen_attr_table$Sample, 3, 3))
 troph_attr_table <- read.csv('../output/gab_troph_attr_table.csv', stringsAsFactors=F)
 troph_attr_table$Site <- paste0(troph_attr_table$Site, 
 				reverse_substr(troph_attr_table$Sample, 3, 3))
 ## Rarefied ASV Table
-troph_asv_table <- read.csv('../output/gab_rare_asv_table.csv')
+asv_table <- read.csv('../output/gab_rare_asv_table.csv')
 
 ## DESeq2 variance stabilized asv table
 # troph_asv_table <- read_csv('../output/vst_asv_table.csv')
@@ -23,9 +25,9 @@ troph_asv_table <- read.csv('../output/gab_rare_asv_table.csv')
 #  write_csv('output/troph_total.csv')
 geodist <- read.table('../output/geodist.tsv')
 troph_attr_table <- left_join(troph_attr_table, geodist, by = "Site")
-troph_attr_table <- right_join(troph_attr_table, troph_asv_table, by = "Sample")
+troph_attr_table <- inner_join(troph_attr_table, asv_table, by = "Sample")
 write_csv(troph_attr_table, '../output/gab_troph_total.csv')
 
-# gen_attr_table %>% 
-# left_join(asv_table, by = "Sample") %>%
-# write_csv('output/gen_total.csv')
+gen_attr_table <- left_join(gen_attr_table, geodist, by = "Site")
+gen_attr_table <- inner_join(gen_attr_table, asv_table, by = "Sample")
+write_csv(gen_attr_table, '../output/gab_gen_total.csv')

### 2019-06-13 R/gab_rarefy_asv_table.R

@@ -5,10 +5,10 @@ library(amorris)
 asv_table <- read.csv('../output/gab_asv_table.csv')
 
 # Check library size
-rowSums(as.matrix(asv_table[, -1]))
+sort(rowSums(as.matrix(asv_table[, -1])))
 
 # Remove two samples with shallow sequencing depth
-asv_table <- asv_table[rowSums(as.matrix(asv_table[, -1])) > 10000, ]
+asv_table <- asv_table[rowSums(as.matrix(asv_table[, -1])) > 12000, ]
 
 comm_mat <- as.matrix(asv_table[, -1])
 

### 2019-06-13 R/gab_lm_model.R
Clean up modeling

@@ -12,5 +12,8 @@ adj_data <- sapply(temp, read.table, simplify = FALSE, USE.NAMES = TRUE)
 names(adj_data) <- reverse_substr(substr(names(adj_data), 19, 25), 5, 7)
 # Fit models and save model outputs
 lapply(names(adj_data), function(x) saveRDS(fit_models(adj_data[[x]], lowk_model), 
-                                     file = paste0("../output/lowk_fits_", x, 
+                                     file = paste0("../output/fits_lowk_", x, 
+                                                   ".rds")))
+lapply(names(adj_data), function(x) saveRDS(fit_models(adj_data[[x]], vmax_model), 
+                                     file = paste0("../output/fits_vmax_", x, 
                                                    ".rds")))

### 2019-06-13 R/id_taxa.R

@@ -4,9 +4,9 @@ library(amorris)
 source('functions.R')
 
 # import model output
-temp <- list.files(path="../output", pattern="lowk_fits*", full.names = TRUE)
+temp <- list.files(path="../output", pattern="fits*", full.names = TRUE)
 all_models <- sapply(temp, readRDS, simplify = FALSE, USE.NAMES = TRUE)
-names(all_models) <- reverse_substr(substr(names(all_models), 21, 27), 5, 8)
+names(all_models) <- reverse_substr(substr(names(all_models), 16, 27), 5, 12)
 
 taxon_table <- read_csv('../output/taxon_table.csv')
 
@@ -20,5 +20,4 @@ lapply(asvs, length)
 sig_taxa <- lapply(asvs, id_taxonomy, taxon_table) 
 
 # Save taxa
-sapply(names(sig_taxa), 
- function (x) write.table(sig_taxa[[x]], file = paste0('../tables/sig_taxa_', x, ".txt")))
+sapply(names(sig_taxa), function (x) write.table(sig_taxa[[x]], file = paste0('../tables/sig_taxa_', x, ".txt")))

### 2019-06-14 R/gab_combine_data.R
Clean up combine data script

@@ -9,25 +9,36 @@ troph_attr_table <- read.csv('../output/gab_troph_attr_table.csv', stringsAsFact
 troph_attr_table$Site <- paste0(troph_attr_table$Site, 
 				reverse_substr(troph_attr_table$Sample, 3, 3))
 ## Rarefied ASV Table
-asv_table <- read.csv('../output/gab_rare_asv_table.csv')
+rare_asv_table <- read.csv('../output/gab_rare_asv_table.csv')
 
 ## DESeq2 variance stabilized asv table
 # troph_asv_table <- read_csv('../output/vst_asv_table.csv')
 
 ## Raw sequence counts
-#troph_asv_table <- read_csv('output/../asv_table.csv')
-#asvs <- troph_asv_table[, -1] 
-### Presence Absence
-#asvs[asvs>0] <- 1
-#troph_asv_table[, -1] <- asvs
+pa_asv_table <- read_csv('../output/gab_asv_table.csv')
+asvs <- pa_asv_table[, -1] 
+## Presence Absence
+asvs[asvs>0] <- 1
+pa_asv_table[, -1] <- asvs
 
 #left_join(troph_attr_table, troph_asv_table, by = "Sample") %>%
 #  write_csv('output/troph_total.csv')
 geodist <- read.table('../output/geodist.tsv')
-troph_attr_table <- left_join(troph_attr_table, geodist, by = "Site")
-troph_attr_table <- inner_join(troph_attr_table, asv_table, by = "Sample")
-write_csv(troph_attr_table, '../output/gab_troph_total.csv')
 
-gen_attr_table <- left_join(gen_attr_table, geodist, by = "Site")
-gen_attr_table <- inner_join(gen_attr_table, asv_table, by = "Sample")
-write_csv(gen_attr_table, '../output/gab_gen_total.csv')
+# Export
+troph_total_rare <- left_join(troph_attr_table, geodist, by = "Site")
+troph_total_rare <- inner_join(troph_total_rare, rare_asv_table, by = "Sample")
+write_csv(troph_total_rare, '../output/gab_all_rare_troph.csv')
+
+gen_total_rare <- left_join(gen_attr_table, geodist, by = "Site")
+gen_total_rare <- inner_join(gen_total_rare, rare_asv_table, by = "Sample")
+write_csv(gen_total_rare, '../output/gab_all_rare_gen.csv')
+
[…]

### 2019-06-14 R/brazil_cleaning.R
Update var comp

@@ -1,7 +1,7 @@
 library(tidyverse)
 
 ## Cleaning up metadata
-metadata <- read_tsv('data/brazil/Dimensions_meta_data.txt')
+metadata <- read_tsv('../data/brazil/Dimensions_meta_data.txt')
 metadata <- as.data.frame(metadata)
 ## Converting below detection limit to the detection limt
 ## Controversial decision - consider other options (zeros?)
@@ -14,7 +14,7 @@ metadata <- mutate_at(metadata, vars(pH:Error_N2O), as.numeric)
 as.data.frame(metadata)
 
 ## ASVs
-asvs <- read_tsv('data/brazil/16S_dada2_table_RDP_tax.txt')
+asvs <- read_tsv('../data/brazil/16S_dada2_table_RDP_tax.txt')
 colnames(asvs)
 taxa_abund <- 
   asvs %>% 
@@ -22,7 +22,7 @@ taxa_abund <-
   select(-ConsensusLineage) %>% 
   gather(Sample_ID, Abundance, 2:75) %>% 
   spread(OTU_ID, Abundance)
-gps <- read.table('data/brazil/Dimensions_GPS_dec_deg.txt', header = FALSE)
+gps <- read.table('../data/brazil/Dimensions_GPS_dec_deg.txt', header = FALSE)
 colnames(gps) <- c('Sample_ID', 'Lat', 'Long')
 
 all_data <-
@@ -31,4 +31,4 @@ all_data <-
   right_join(taxa_abund, by = 'Sample_ID')
 
 all_data$library_size <- rowSums(taxa_abund[, -1])
-write_tsv(all_data, 'output/dim_cleaned_data.tsv')
+write_tsv(all_data, '../output/dim_cleaned_data.tsv')

### 2019-06-14 R/gab_var_comp.R

@@ -4,8 +4,8 @@ library(vegan)
 library(broom)
 library(amorris)
 
-asv_table <- read.csv('../output/gab_asv_table.csv')
-
+all_data <- read.csv('../output/gab_all_rare_troph.csv')
+dim(all_data)
 ### Plot ASV abundances
 #i = 2
 #j = ncol(asv_table)
@@ -15,7 +15,7 @@ asv_table <- read.csv('../output/gab_asv_table.csv')
 #  geom_jitter()  
 
 ## Remove ASVs only present in 1 sample
-asvs <- asv_table[, -1] 
+asvs <- select(all_data, 
 asvs[asvs>0] <- 1
 asvs <- as.matrix(asvs)
 asv_table <- asv_table[rowSums(asvs) > 1, ]

### 2019-06-14 R/var_comp_brazil.R

@@ -8,7 +8,7 @@ library(amorris)
 
 ## Import and organize data
 
-all_data <- read.table('output/dim_cleaned_data.tsv', header = TRUE)
+all_data <- read.table('../output/dim_cleaned_data.tsv', header = TRUE)
 #all_data <- fread('output/dim_cleaned_data.tsv')
 #all_data[1:nrow(all_data), 1:40]
 #colnames(all_data[, 1:40])
@@ -73,7 +73,7 @@ env_sim <- 1/(1+env_dis)
 
 model <- varComp(CH4 ~ taxon, varcov = list(com = com_sim, geo = geo_sim, env = env_sim) )
 summary(model)
-
+CH4
 
 for (i in 11:20) {
   taxon <- asv_mat_ns[, i]

### 2019-06-15 R/analysis.Rmd
Add markdown document

---
title: "Royal Society paper"
subtitle: "Linking microbial communities to ecosystem functions: what we can learn from genotype-phenotype mapping in organisms"
author: "Andrew Morris"
date: "`r Sys.Date()`"
output: html_document
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=8, fig.height=6, fig.path='Figs/',
                      echo=FALSE, warning=FALSE, message=FALSE)
```

```{r load_packages, include=FALSE}

library(tidyverse)
library(lubridate)
library(broom)
library(broman)
library(morris)
```

Words

### 2019-06-15 R/gab_combine_data.R
update makefile

@@ -28,6 +28,7 @@ geodist <- read.table('../output/geodist.tsv')
 # Export
 troph_total_rare <- left_join(troph_attr_table, geodist, by = "Site")
 troph_total_rare <- inner_join(troph_total_rare, rare_asv_table, by = "Sample")
+troph_total_rare <- troph_total_rare[troph_total_rare$Wetland == 'Upland', ]
 write_csv(troph_total_rare, '../output/gab_all_rare_troph.csv')
 
 gen_total_rare <- left_join(gen_attr_table, geodist, by = "Site")

### 2019-06-20 R/var_stab_transform.R
Update variance stabilization script

@@ -1,9 +1,6 @@
-
 library(DESeq2)
 library(tidyverse)
-asv_data <- read.delim('data/16S_dada2_table_RDP_tax_edit.txt', header = TRUE)
-
-
+asv_data <- read.delim('../data/gabon/16S_dada2_table_RDP_tax_edit.txt', header = TRUE)
 asv_mat <- as.matrix(asv_data[,-c(1, ncol(asv_data))])
 rownames(asv_mat) <- asv_data$OTU_ID
 # colnames(asv_mat) <- colnames(asv_data[, -1])
@@ -24,10 +21,8 @@ vsd_data <-
 
 asv_table <- vsd_data %>%
   select(asv, starts_with('dna')) %>% 
-  select(asv, ends_with('MO')) %>% 
   gather(Sample, abund, 2:length(.)) %>% 
   spread(asv, abund) %>% 
   mutate(Sample = substring(Sample, 5))
-
-
-write_csv(asv_table, 'output/vst_asv_table.csv')
+asv_table[asv_table < 0] <- 0
+write_csv(asv_table, '../output/vst_asv_table.csv')

### 2019-06-20 R/gab_combine_data.R
Use VST data instead of rarefied data

@@ -1,5 +1,5 @@
 library(tidyverse)
-library(amorris)
+library(morris)
 
 gen_attr_table <- read_csv('../output/gab_gen_attr_table.csv')
 
@@ -9,10 +9,10 @@ troph_attr_table <- read.csv('../output/gab_troph_attr_table.csv', stringsAsFact
 troph_attr_table$Site <- paste0(troph_attr_table$Site, 
 				reverse_substr(troph_attr_table$Sample, 3, 3))
 ## Rarefied ASV Table
-rare_asv_table <- read.csv('../output/gab_rare_asv_table.csv')
+#asv_table <- read.csv('../output/gab_rare_asv_table.csv')
 
 ## DESeq2 variance stabilized asv table
-# troph_asv_table <- read_csv('../output/vst_asv_table.csv')
+asv_table <- read_csv('../output/vst_asv_table.csv')
 
 ## Raw sequence counts
 pa_asv_table <- read_csv('../output/gab_asv_table.csv')
@@ -23,16 +23,16 @@ pa_asv_table[, -1] <- asvs
 
 #left_join(troph_attr_table, troph_asv_table, by = "Sample") %>%
 #  write_csv('output/troph_total.csv')
-geodist <- read.table('../output/geodist.tsv')
+geodist <- read.table('../output/geodist.tsv', stringsAsFactors = FALSE)
 
 # Export
 troph_total_rare <- left_join(troph_attr_table, geodist, by = "Site")
-troph_total_rare <- inner_join(troph_total_rare, rare_asv_table, by = "Sample")
-troph_total_rare <- troph_total_rare[troph_total_rare$Wetland == 'Upland', ]
+troph_total_rare <- inner_join(troph_total_rare, asv_table, by = "Sample")
+#troph_total_rare <- troph_total_rare[troph_total_rare$Wetland == 'Upland', ]
 write_csv(troph_total_rare, '../output/gab_all_rare_troph.csv')
 
 gen_total_rare <- left_join(gen_attr_table, geodist, by = "Site")
-gen_total_rare <- inner_join(gen_total_rare, rare_asv_table, by = "Sample")
+gen_total_rare <- inner_join(gen_total_rare, asv_table, by = "Sample")
 write_csv(gen_total_rare, '../output/gab_all_rare_gen.csv')
 
 troph_total_pa <- left_join(troph_attr_table, geodist, by = "Site")

### 2019-06-20 R/gab_var_comp.R
Clean up varcomp script

@@ -2,138 +2,67 @@ library(tidyverse)
 library(varComp)
 library(vegan)
 library(broom)
-library(amorris)
+library(morris)
 
-all_data <- read.csv('../output/gab_all_rare_troph.csv')
-dim(all_data)
-### Plot ASV abundances
-#i = 2
-#j = ncol(asv_table)
-#  asv_table[, c(1, i:j)] %>% 
-#  gather(ASV, ABU, 2:ncol(asv_table[, c(1, i:j)])) %>% 
-#ggplot(aes(y = ABU, x = ASV)) +
-#  geom_jitter()  
+all_data <- read.csv('../output/gab_all_rare_troph.csv', stringsAsFactors = FALSE)
 
 ## Remove ASVs only present in 1 sample
-asvs <- select(all_data, 
+asvs <- as.matrix(all_data[, grep("^asv", colnames(all_data))])
 asvs[asvs>0] <- 1
-asvs <- as.matrix(asvs)
-asv_table <- asv_table[rowSums(asvs) > 1, ]
-asv_mat <- as.matrix(asv_table[, 2:ncol(asv_table)])
-
-### High variance ASVs
-#ColVar <- function(x, ...) {
-#  colSums((x - colMeans(x, ...))^2, ...)/(dim(x)[2] - 1)
-#}
-#asv_mat_var <- ColVar(asv_mat)
-#asv_mat_high_var<-asv_mat[,log10(asv_mat_var) > 0]
-#ncol(asv_mat_high_var)
-#plot(density(log10(asv_mat_var)))
-#rowSums(asv_mat)
-#colSums(asv_mat)
-#asv_mat_ns <- asv_mat[, colSums(asv_mat) > 1]
-#colSums(asv_mat_ns)
+all_data <- all_data[rowSums(asvs) > 1, ]
+asv_mat <- as.matrix(all_data[, grep("^asv", colnames(all_data))])
+dim(asvs); dim(asv_mat)
 
 # Community similarity
-taxon <- asv_mat[, 3]
-com_dissim <- vegdist(asv_mat[, -3])
-com_dissim <- as.matrix(com_dissim)
-com_sim <- 1-com_dissim
+asv <- 3
+taxon <- asv_mat[, asv]
+com_sim <- 1 - as.matrix(vegdist(asv_mat[, -asv]))
 
-
-troph_attr_table <- read.csv('../output/troph_attr_table.csv', stringsAsFactors = F)
-troph_attr_table$Site <- paste0(troph_attr_table$Site, 
[…]

### 2019-06-20 R/analysis.Rmd
Add varcomp to rmd, update lm and rarefy scripts

@@
 ---
+
+# Overview
+
+This is the primary analysis script for the analysis associated with the Royal Society
+paper.
 
@@
 library(morris)
+library(varComp)
+library(vegan)
+
 ```
 
-Words
+```{r import_data}
+
+all_data <- read.csv('../output/gab_all_rare_troph.csv', stringsAsFactors = FALSE)
+
+```
+
+# Variance Component Analysis
+
+```{r sim_mat}
+
+## Remove ASVs only present in 1 sample
+asvs <- as.matrix(all_data[, grep("^asv", colnames(all_data))])
+asvs[asvs>0] <- 1
+all_data <- all_data[rowSums(asvs) > 1, ]
+asv_mat <- as.matrix(all_data[, grep("^asv", colnames(all_data))])
+
+dim(asvs) - dim(asv_mat)
+rowSums(asvs)
+
+# Community similarity
+asv <- 3
+taxon <- asv_mat[, asv]
+#com_dis <- as.matrix(dist(as.matrix(scale(asv_mat[, -asv]))))
+#com_sim <- 1/(1+com_dis)
+com_sim <- 1 - as.matrix(vegdist(asv_mat[, -asv]))
+
+env_variables <- all_data[, 1:8]
+
+# Function
+lowk <- all_data$Low_final_k
+
+# Geographic similarity
+xy <- all_data[, c('X', 'Y')]
+geo_dis <- as.matrix(dist(as.matrix(scale(xy))))
+geo_sim <- 1/(1+geo_dis)
+
+# environmental dissimilarity
+wetland <- all_data[, "Wetland"]
+env <- all_data[, c('WFPS', 'N_percent', 'C_percent')]
+env_dis <- as.matrix(dist(as.matrix(scale(env))))
+env_sim <- 1/(1+env_dis)
+
+```
+
+```{r varcomp_diagnostics}
+
+mantel(com_sim, geo_sim)
+qplot(x = xy$X, y = xy$Y) + geom_jitter(aes(color = all_data$Wetland))
+
+ggplot(mapping = aes(lowk)) + 
+  geom_histogram(bins = 20) + 
+  labs(x = "Low Final K", y = "Number of Samples")
+ggplot(mapping = aes(taxon)) + 
+  geom_histogram(bins = 20) + 
+  labs(x = "Taxon Abundance", y = "Number of Samples")
+
+all_data[all_data$Y > 30000, 'geocode'] <- 'A'
+all_data[all_data$Y < 30000 & all_data$Y > 10000, 'geocode'] <- 'B'
+all_data[all_data$Y < 10000, 'geocode'] <- 'C'
+all_data$geocode <- factor(all_data$geocode)
+
+ggplot(mapping = aes(c(com_sim))) + 
+  geom_histogram(binwidth = 0.05) + 
+  labs(x = "Community Similarity (Bray-Curtis)", y = "Number of Samples")
+ggplot(mapping = aes(c(geo_sim))) + 
+  geom_histogram(binwidth = 0.05) + 
+  labs(x = "Spatial Similarity (Euclidean)", y = "Number of Samples")
+ggplot(mapping = aes(c(env_sim))) + 
+  geom_histogram(binwidth = 0.05) + 
+  labs(x = "Environment Similarity (Euclidean)", y = "Number of Samples")
+
+```
+
+Community and geography are correlated and sites cluster around three locations
+Also, geography uses up more of the similarity matrix space (see histograms)
+Therefore, I coded geography as three factors (one for each cluster of sites)
+Also, wetland/upland covaries with those three sites, but moisture content is
+included in the environmental similarity matrix and should capture that variation
+
+```{r fit_var_comp}
+
+model <- varComp(lowk ~ taxon, varcov = list(com = com_sim, geo = geo_sim, env = env_sim) )
+summary(model)
+
+model <- varComp(lowk ~ taxon + all_data$geocode, varcov = list(com = com_sim, env = env_sim) )
+summary(model)
+levels(all_data$geocode)
+all_data$geocode <- factor(all_data$geocode, levels(all_data$geocode)[c(2,3,1)])
+model1 <- varComp(lowk ~ taxon + all_data$geocode, varcov = list(com = com_sim, env = env_sim) )
+summary(model1)
+
+```

### 2019-06-20 R/gab_lm_model.R

@@ -2,7 +2,7 @@ library(tidyverse)
 library(purrr)
 library(broom)
 library(modelr)
-library(amorris)
+library(morris)
 source('functions.R')
 
 # Import all model output

### 2019-06-20 R/gab_rarefy_asv_table.R

@@ -1,8 +1,9 @@
 library(tidyverse)
-library(amorris)
+library(morris)
 
 # Import ASV table
 asv_table <- read.csv('../output/gab_asv_table.csv')
+dim(asv_table)
 
 # Check library size
 sort(rowSums(as.matrix(asv_table[, -1])))
@@ -19,4 +20,3 @@ rare_mat <- rarefy_mean(comm_mat, n = 1000)
 rare_asv <- cbind(samples, as_tibble(rare_mat))
 
 write_csv(rare_asv, '../output/gab_rare_asv_table.csv')
-

### 2019-06-21 R/gab_cleaning.R
Change seasonally inundated wetland to upland and remove community attributes

@@ -1,6 +1,6 @@
 
 library(tidyverse)
-library(amorris)
+library(morris)
 
 # Import ------------------------------------------
 raw_troph_data <- read_tsv('../data/gabon/Gabon_methanotrophy_table_scaled.txt')
@@ -17,7 +17,7 @@ troph_data <- raw_troph_data %>%
 
 # Add Wetland variable
 troph_data %<>% left_join(data_frame(Land_type = unique(troph_data$Land_type), 
-           Wetland = factor(c('Wetland', 'Upland', 'Upland', 'Wetland', 'Wetland',
+           Wetland = factor(c('Wetland', 'Upland', 'Upland', 'Wetland', 'Upland',
                               'Wetland', 'Upland', 'Upland', 'Upland'))), by = 'Land_type') %>% 
   select(Sample:Experiment, Wetland, Description:PC3_16S_rna)
 
@@ -121,6 +121,14 @@ num_predictors <- gen_num_data %>% dplyr::select(-CH4, -H_CH4_percent)
 # Attribute table
 gen_attr_table <- gen_data %>% 
   select(-Rep, -Experiment, -Description, -Land_type, -Bulk_dens, -Mois_cont)
+
+# Remove extraneous attributes
+colnames(gen_attr_table)
+gen_attr_table <- gen_attr_table %>% 
+  select(Site:Copies)
+colnames(troph_attr_table)
+troph_attr_table <- troph_attr_table %>% 
+  select(Sample:pmoa_transcript_num)
 
 write_csv(gen_attr_table, '../output/gab_gen_attr_table.csv')
 

### 2019-06-21 R/gab_combine_data.R
Replace rarefied with VST ASV table

@@ -1,21 +1,23 @@
 library(tidyverse)
 library(morris)
 
-gen_attr_table <- read_csv('../output/gab_gen_attr_table.csv')
-
+# Methanogen metadata
+gen_attr_table <- read.csv('../output/gab_gen_attr_table.csv', stringsAsFactors = FALSE)
 gen_attr_table$Site <- paste0(gen_attr_table$Site, 
 				reverse_substr(gen_attr_table$Sample, 3, 3))
+
+# Methanotrophy metadata
 troph_attr_table <- read.csv('../output/gab_troph_attr_table.csv', stringsAsFactors=F)
 troph_attr_table$Site <- paste0(troph_attr_table$Site, 
 				reverse_substr(troph_attr_table$Sample, 3, 3))
 ## Rarefied ASV Table
-#asv_table <- read.csv('../output/gab_rare_asv_table.csv')
+#rare_asv_table <- read.csv('../output/gab_rare_asv_table.csv')
 
 ## DESeq2 variance stabilized asv table
-asv_table <- read_csv('../output/vst_asv_table.csv')
+vst_asv_table <- read.csv('../output/vst_asv_table.csv', stringsAsFactors = FALSE)
 
 ## Raw sequence counts
-pa_asv_table <- read_csv('../output/gab_asv_table.csv')
+pa_asv_table <- vst_asv_table #read_csv('../output/gab_asv_table.csv')
 asvs <- pa_asv_table[, -1] 
 ## Presence Absence
 asvs[asvs>0] <- 1
@@ -26,14 +28,14 @@ pa_asv_table[, -1] <- asvs
 geodist <- read.table('../output/geodist.tsv', stringsAsFactors = FALSE)
 
 # Export
-troph_total_rare <- left_join(troph_attr_table, geodist, by = "Site")
-troph_total_rare <- inner_join(troph_total_rare, asv_table, by = "Sample")
+troph_total_vst <- left_join(troph_attr_table, geodist, by = "Site")
+troph_total_vst <- inner_join(troph_total_vst, vst_asv_table, by = "Sample")
 #troph_total_rare <- troph_total_rare[troph_total_rare$Wetland == 'Upland', ]
-write_csv(troph_total_rare, '../output/gab_all_rare_troph.csv')
+write_csv(troph_total_vst, '../output/gab_all_vst_troph.csv')
 
-gen_total_rare <- left_join(gen_attr_table, geodist, by = "Site")
[…]

### 2019-06-21 R/var_comp.R
Convert varComp script to R with Roxygen comments for markdown rendering

@@ -1,0 +1,142 @@
+#' ---
+#' title: "Royal Society paper"
+#' subtitle: "Linking microbial communities to ecosystem functions: what we can learn from genotype-phenotype mapping in organisms"
+#' author: "Andrew Morris"
+#' date: "`r Sys.Date()`"
+#' output: html_document
+#' ---
+#' 
+#' # Overview
+#' 
+#' This is the primary analysis script for the analysis associated with the Royal Society
+#' paper. 
+
+#+ r global_options, include=FALSE
+knitr::opts_chunk$set(fig.width=8, fig.height=6, fig.path='Figs/',
+                      echo=FALSE, warning=FALSE, message=FALSE)
+
+#+ r load_packages, include=FALSE
+
+library(tidyverse)
+library(lubridate)
+library(broom)
+library(broman)
+library(morris)
+library(varComp)
+library(vegan)
+
+#+ r import_data
+
+all_data <- read.csv('../output/gab_all_vst_troph.csv', stringsAsFactors = FALSE)
+
+#' # Variance Component Analysis
+
+#+ r sim_mat
+
+## Remove ASVs only present in 1 sample
+asvs <- as.matrix(all_data[, grepl("^[asv]", colnames(all_data))])
+vars <- all_data[, !grepl("^[asv]", colnames(all_data))]
+asvs <- asvs[, colSums(asvs) > 0]
+all_data <- cbind(vars, asvs)
+
+# Community similarity
+asv <- 3
+taxon <- asvs[, asv]
+#com_dis <- as.matrix(dist(as.matrix(scale(asv_mat[, -asv]))))
+#com_sim <- 1/(1+com_dis)
+com_sim <- 1 - as.matrix(vegdist(asvs[, -asv]))
+
+# Geographic similarity
+xy <- all_data[, c('X', 'Y')]
+geo_dis <- as.matrix(dist(as.matrix(scale(xy))))
+geo_sim <- 1/(1+geo_dis)
+
+# environmental dissimilarity
+env <- all_data[, c('WFPS', 'N_percent', 'C_percent')]
+env_dis <- as.matrix(dist(as.matrix(scale(env))))
+env_sim <- 1/(1+env_dis)
+
+all_data[all_data$Y > 30000, 'geocode'] <- 'A'
+all_data[all_data$Y < 30000 & all_data$Y > 10000, 'geocode'] <- 'B'
[…]

### 2019-06-21 R/var_stab_transform.R

@@ -1,5 +1,6 @@
 library(DESeq2)
 library(tidyverse)
+
 asv_data <- read.delim('../data/gabon/16S_dada2_table_RDP_tax_edit.txt', header = TRUE)
 asv_mat <- as.matrix(asv_data[,-c(1, ncol(asv_data))])
 rownames(asv_mat) <- asv_data$OTU_ID

### 2019-06-21 R/gab_var_comp.R
Clean up R/ to not include GWAS or Brazil

@@ -1,68 +1,0 @@
-library(tidyverse)
-library(varComp)
-library(vegan)
-library(broom)
-library(morris)
-
-all_data <- read.csv('../output/gab_all_rare_troph.csv', stringsAsFactors = FALSE)
-
-## Remove ASVs only present in 1 sample
-asvs <- as.matrix(all_data[, grep("^asv", colnames(all_data))])
-asvs[asvs>0] <- 1
-all_data <- all_data[rowSums(asvs) > 1, ]
-asv_mat <- as.matrix(all_data[, grep("^asv", colnames(all_data))])
-dim(asvs); dim(asv_mat)
-
-# Community similarity
-asv <- 3
-taxon <- asv_mat[, asv]
-com_sim <- 1 - as.matrix(vegdist(asv_mat[, -asv]))
-
-env_variables <- all_data[, 1:8]
-# Function
-lowk <- all_data$Low_final_k
-hist(lowk)
-hist(taxon)
-# Geographic similarity
-xy <- all_data[, c('X', 'Y')]
-geo_dis <- as.matrix(dist(as.matrix(scale(xy))))
-geo_sim <- 1/(1+geo_dis)
-
-# environmental dissimilarity
-wetland <- all_data[, "Wetland"]
-env <- all_data[, c('WFPS', 'N_percent', 'C_percent')]
-env_dis <- as.matrix(dist(as.matrix(scale(env))))
-env_sim <- 1/(1+env_dis)
-
-mantel(com_sim, geo_sim)
-qplot(x = xy$X, y = xy$Y)
-
-# Community and geography are correlated and sites cluster around three locations
-# Also, geography uses up more of the similarity matrix space (see histograms)
-# Therefore, I coded geography as three factors (one for each cluster of sites)
-# Also, wetland/upland covaries with those three sites, but moisture content is
-# included in the environmental similarity matrix and should capture that variation
-all_data[all_data$Y > 30000, 'geocode'] <- 'A'
-all_data[all_data$Y < 30000 & all_data$Y > 10000, 'geocode'] <- 'B'
-all_data[all_data$Y < 10000, 'geocode'] <- 'C'
-all_data$geocode <- factor(all_data$geocode)
-hist(com_sim)
-hist(env_sim)
-hist(geo_sim)
-qplot(x = xy$X, y = xy$Y) + geom_jitter(aes(color = all_data$Wetland))
-
[…]

### 2019-06-21 R/var_comp.R

@@ -111,8 +111,10 @@ summary(model1)
 
 anova(model, model2)
 anova(model)
-summary(model <- varComp(Low_final_k ~ taxon, data = all_data))
-summary(model2 <- varComp(Low_final_k ~ taxon + geocode, data = all_data))
+model <- varComp(Low_final_k ~ taxon, data = all_data)
+model2 <- varComp(Low_final_k ~ taxon + geocode, data = all_data)
+
+
 model <- varComp(Low_final_k ~ taxon, data = all_data,
                  varcov = list(env = env_sim))
 model <- varComp(Low_final_k ~ taxon, data = all_data,

### 2019-06-24 manuscript/main.tex
Add key question

@@
 
-What is novel about this? The first anyone has treated an ecosystem like an
+\section*{Tasks} 
+\begin{enumerate}
+\item What is novel about this? The first anyone has treated an ecosystem like an
 organism. The first time anyone has done this correlative approach well
 by considering environmental and spatial structure in the data. 
+\item Figure out variance partitioning (with Hannah)
+\item Maybe variance partitioning plus correlations (varPart)
+\item Do PC correction in all combinations
+\item Identify significant taxa and build a venn diagram figure (or
+something)
+\item 
+\end{enumerate}
+\section{Introduction}
+
+I want to know whether microbial community composition is important for determining ecosystem-scale functions independent of the underlying environmental variation.
+
+Microorganisms mediate many biogeochemical processes, but despite 
+much effort to link microbial communities
+to these functions the mechanism by which they regulate these
+processes remains unknown. Ecologists have investigated the effects of changing
+biodiversity on ecosystem function for decades, but many of their approaches
+are not applicable in microbial ecology due to the nature of microbial data. 
+This limits our ability to predict and manage globally relevant biogeochemical cycles. To
+address this challenge, we propose integrating current theory on community
+assembly and the functioning of ecosystems with lessons from genotype-phenotype
+mapping in quantitative genetics. This will address the
+question of how does variation in microbial community structure contribute to
+variation in the rate of ecosystem function?
+
+\subsection{Biodiversity increases ecosystem function in macrobial communities}
+
+Local species richness is correlated with ecosystem function
+\citep{hooper2005}. Most BEF research has focused on terrestrial plant 
+communities where species losses lead to decreases in community abundance or biomass
+\citep{naeem2003}. There are
+two major explanations for this phenomenon. The first is the sampling effect
+where a more diverse community is more likely to have one particularly
+productive species \citep{cardinale2006}. The second is niche complementarity where more species are
+more likely to occupy separate niches within the same ecosystem resulting in
+more efficient resource use and greater biomass.
+
+Regional species diversity may also be important for
+local ecosystem functioning when metacommunity dynamics regulate the
+diversity and composition of communities \citep{leibold2017}. 
+Typical BEF research focuses on local selection
+as a driver of ecosystem function. However, if regional
+diversity is important, then dispersal processes could alter local species diversity and composition resulting in
+a change in ecosystem function. This model is referred to as Community Assembly
+and the Functioning of Ecosystems (CAFE). Community assembly can alter ecosystem
+function if limiting dispersal reduces the likelihood of a highly productive
+species from dispersing to a local site or if high dispersal of an unproductive
+species overwhelms local selection.
+
+The CAFE approach has been demonstrated using the Price equation where a change in ecosystem function between two, matched
+observations can be attributed to species additions, species losses, and
+changes in abundance of shared species \citep{fox2006,
+bannar-martin2018}. To
+calculate these vectors, function must be attributed at the species level (e.g.
+biomass measurements of individual plant species). The CAFE framework demonstrates
+that not only local richness, 
+but also the composition of communities is important to the functioning of ecosystems.
+
+\subsection{Current approaches to microbial BEF reveal little about 
+structure-function mapping}
+
+The trajectory of microbial BEF research has differed from that of macrobial BEF
+for three reasons. First, microbial communities are
+immensely complex with orders of magnitude more taxa than macrobial communities
+and great intra-taxon variation for certain traits. Second, the way we measure
+microbes using marker gene sequences does not allow us to attribute portions of
+a function of interest to individual taxonomic groups . We can only
+measure function in aggregate across an entire community. Finally, we cannot be
+sure that the taxonomic groups that putatively perform a function are the ones
+contributing in this moment.
+
+As a result of these limitations, microbial ecologists have taken a different
+approach to assessing BEF relationships for microbial communities than others
+have for macrobial communities. This is generally a correlative approach where
+certain community attributes that can be derived from a community matrix are
+tested in terms of their relationship with function. These approaches make
+different assumptions about how microbial communities should map onto function.
+On one end of the spectrum, it is hypothesized that functional gene abundance - genes that
+code for an enzyme that performs one step of a process - should be correlated
+with the rate of the process associated with that gene. In this case, qPCR
+is used to quantify the abundance of a functional gene to test whether it is
+correlated with the rate of a process. This approach has had varied success and even the abundance of a
+functional gene that putatively performs a process is rarely correlated with
+that process \citep{rocca2015}. On the other end of the
+spectrum, investigators make few assumptions about which members of the community should
+alter function and so compute
+taxonomic richness or other diversity metrics from the entire Prokaryotic
+community inferred from 16S rRNA gene sequences. Taxonomic or functional
+diversity of the Prokaryote community generally adds little variance explained
+to models of ecosystem function \citep{graham2016}.
+
+\subsection{Biology through analogy: structure-function mapping is similar
+to genotype-phenotype mapping}
+
+The challenges described above for mapping microbial community structure to
+ecosystem function are analogous to the problem of identifying the
+genetic basis of complex traits in organismal populations. In that case,
+each individual organism has a level for a phenotype (e.g. presence
+or absence of a disease state) as well as a vector of potentially
+thousands of genetic loci with different alleles or single nucleotide
+polymorphisms (SNPs). To identify the genetic basis for this trait, we sample
+a population of organisms and try to correlate a matrix of SNPs with a
+vector of phenotypes. In most cases, there are many more loci than
+individuals, we cannot directly attribute a phenotypic state to a particular
+SNP, and we do not know the function of genes associated with these
+SNPs. This mirrors the problem of microbial structure-function mapping
+and presents a potential set of analyses that could be applied in
+microbial ecology.
+
+In addition to the nature of the data, population geneticists have other
+considerations when trying to make robust associations between phenotype
+and SNPs. For example, loci within the genome that are close together tend
+not to assort independently due to linkage disequilibrium. In addition,
+organisms exhibit population stratification due to a variety of historical
+and selective effects. For example, shared ancestry among European
+populations of humans results in an association between genotype and
+geography across Europe \citep{novembre2008}. This population structure
+can lead to false positives when trying to correlate human genotypes
+with genetic diseases or other phenotypes such as height (citation).
+
+In general, organisms exhibit population differentiation due to spatial
+and environmental patterns \citep{wright1943}. For geography, local genetic
+drift and reduced gene flow due to geographic separation can lead to population
+differentiation \citep{wright1943}. In addition, environmental distance can lead to both selection
+against migrants \citep{hendry2004} and local selection on endemic populations to
+generate variation. While
+these processes have primarily been described for macrobial populations, the
+same principles apply to microorganisms and should be considered
+in microbial structure-function mapping \citep{martiny2006}.
+
+\subsection{Reframing the question: Is variation in microbial community
+structure associated with differences in ecosystem function across space,
+time, and environments?}
+
+Typical microbial ecosystem function research focuses on whether microbial
+community composition is important at the ecosystem level
+\citep{schimel1998, cavigelli2000, balser2005, graham2016}. Given that
+microbial functional groups typically mediate biogeochemical processes
+and so the loss of a functional group will certainly alter function,
+the answer is "yes". However, this question does not address whether
+microbial community variation is actually useful for predicting or 
+managing ecosystem functions. To address that problem, we need to know
+whether variation within a functional group matters to variation in
+ecosystem function. If local selection always optimizes the available
+biodiversity to maximize process rates given the underlying environmental
+variation, then composition should not matter, because we should be able
+to predict ecosystem function from abiotic factors alone. Environmental
+treatments, common gardens, and reciprocal transplants have shown
+different functional potential for different community compositions.
+In addition, spatial patterns for microbial communities suggest that
+there should be observable diversification due to environmental and
+spatial processes \citep{wright1943, martiny2006}. Therefore, variation
+in microbial community composition seems to result in variation in 
+ecosystem function. However, in order for microbial community data to be
+useful for ecosystem modeling, the challenge is to identify specifically what variation in community composition
+is associated with function across space, time, and environmental gradients.
+To accomplish that, we need to separate the close association between environmental variation, community
+composition, and ecosystem function to identify the aspects of microbial
+biodiversity that are important for ecosystem function.
+
+\subsection*{Extra text for 1.3/1.4}
+
+\begin{enumerate}
+    \item Things exhibit population differentiation due to spatial and environmental 
+    patterns \citep{wright1943}
+    \item Microbes exhibit similar patterns \citep{martiny2006}
+    \item Pop gen tries to identify regions of the genome correlated with
+    traits
+    \item Considering ecosystems as individuals within a population presents
+    an analogous problem of determining community markers of ecosystem traits (functions)
+    \item Thus the goal of microbial BEF is to identify generalizable patterns
+    of genetic variation in microbes to feed into hypothesis testing about
+    potential mechanisms or to incorporate into ecosystem models
+\end{enumerate}
+
+Typically, microbial community structure-function research questions are framed
+as does microbial composition matter to ecosystem function? The answer to this
+question is obviously "yes" given that microorganisms mediate many of the
+processes of interest. The question then becomes does microbial community
+variation within a functional group or overall matter to ecosystem function? 
+Generally it is assumed that microbial composition is functionally irrelevant, 
+though studies of environmental treatments, common gardens, and reciprocal
+transplants have shown that different community compositions are correlated with
+different functions. The challenge with these studies is separating the close
+connection between environmental variation, community composition, and ecosystem
+function. Microorganisms cannot be useful to ecosystem modeling if there
+is no identifiable variation in composition that consistently alters function
+across space, time, and environmental gradients. The question "do we need to
+include microorganisms in ecosystem models?" depends on identifying variation in
+the community associated with function independent of the environment. If that's
+true, then the next step in understanding or modeling these processes is 
+identify what components of the community alter ecosystem function.
+
+There are a number of ways that environmental or spatial covariates could alter
+or obscure connections between microbial community structure and ecoystem
+function. For example, microbial communities exhibit strong biogeographic
+patterns \citep{martiny2006}. The historical effects of geographic proximity,
+such as dispersal limitation, could lead to spurious associations between taxa
+and function if taxa are unevenly distributed across the functional landscape.
+Environmental selection can also shape microbial community structure, so
+selection on the microbial community that is independent of the
+process could lead to erroneous conclusions about the
+underlying drivers of ecosystem function \citep{thompson2017}. For
+example, pH is known to regulate both microbial biodiversity patterns
+\citep{fierer2006} and methane cycling dynamics \citep{ye2012}, which could lead
+to microbial taxa correlated with methane cycling, but mechanistically
+unrelated. Environmental factors that directly influence function, such as
+moisture, could alter function with or without affecting the community. 
+Finally, community stratification where taxa co-occur in certain ecosystems due to
+selection or historical effects for reasons unrelated to the ecosystem function
+could lead to spurious associations.
+
+\subsection{A new framework for structure-function mapping in microbial
+communities}
+
+Microbial ecosystem function could be assessed using statistical techniques
+borrowed from quantitative genetics along with the conceptual framework for
+community assembly processes provided by metacommunity theory and CAFE. This
+would be a two-step process to first identify the portion of variation in
+ecosystem function attributable to different aspects of an ecosystem: the
+community, the underlying environmental variation, and to spatial processes.
+
+\begin{equation}
+V_F=V_E+V_S+V_C+\epsilon
+\end{equation}
+
+Ecosystem function is represented by $V_F$, $V_E$ is variation in the
+environment and covariation between the community and the environment, $V_S$ is
+variation in spatial distance between samples and covariation between the
+community and space, and $V_C$ is variation in the community independent of
+local selection and historical effects. The parameter $V_C$ also incorporates
+unmeasured environmental and spatial effects.
+
+Each of these variance components also correspond to alternative hypotheses for
+what drives ecosystem function. If most of the variation in $V_F$ is
+attributable to $V_E$, then the environment drives function either directly
+(abiotic processes) or indirectly by selecting for a particular microbial
+community that maximizes function. In this model, the community is simply a 
+conduit through which the environment acts on function. Alternatively, if $V_S$
+explains most of the variation in $V_F$ then this suggests that function is
+primarily a result of historical contingency and differences in dispersal traits.
+Finally, $V_C$ could explain most of the variation in $V_F$. This would be
+consistent with reciprocal transplant and common garden experiments showing that
+community composition drivers ecosystem function independent of the environment.
+
+This framework allows us to address the following questions: 1. Does variation
+in microbial community structure explain a measurable portion of the variation in
+ecosystem function or can all of that variation be attributed to the underlying
+effect of the environmental and spatial processes? If so, 2. Which members of
+the community are significantly correlated with function after controlling for
+the covariances of space and environment? Finally, 3. Do these taxa match our hypotheses about
+what controls ecosystem function (e.g. functional group abundance) or are there
+alternative mechanisms that need to be explored?
+
+\section{Methods}
+
+Sample 16S rRNA gene sequences, environmental covariates, geographic distance,
+ecosystem function. Use the DADA2 pipeline to assign ASVs. 
+phylogenetic and taxonomic assignment Null
+Models \citep{chase2011, stegen2015}
+
+Would it be valuable to use a null model (as in Chase or Stegen) where we
+generate random communities based on the distribution across samples and
+abundance within samples of each taxon. Then try to correlate those abundances
+with our observed rates of ecosystem function to see what distribution of
+structure-function correlations to expect given our community data. These sort
+of null community generating techniques can omit certain community assembly
+processes (such as homogenizing selection or stochastic processes)
+
+An alternative testing approach would be to use Hannah’s permutation test idea
+where we randomly permute the process rates ~1000 times and rebuild the model
+each time to see what the distribution of significant taxa is.
+
+VC - Correlations between taxa and process determined by accounting for spatial
+structure, environmental covariates, and community structure. Represents effect
+of taxonomy, unmeasured environmental variables VE - Account for community
+structure and correlate environmental variables cov(C, E) - The part of the venn
+diagram that disappears when controlling for environment and space VC:VE - Test
+the correlations between community and process rate at a variety of
+environmental conditions.  epsilon/error term - Effects of natural variation,
+unmeasured variables
+
+Another modeling options is HMSC from \cite{ovaskainen2017}
+
+\section{Results}
+
+For high-affinity methane oxidation in Gabon: We experimentally controlled some
+environmental variables and our measured environmental variables are not
+significantly correlated with process rate (C, N, WFPS) Space is correlated with
+function
+
+Be careful with attributing a certain \% of the variance to an individual taxa
+when performing partial regression on the residuals. If a taxa explains 60\% of
+the variance in the residuals after controlling for environment and space that
+doesn’t mean it explains 60\% of the overall variance.
+
+\section{Discussion}
+Combination of qPCR and pathway partinioing, qSIP \citep{hungate2015},
+and NanoSIMS \citep{mayali2012} may allow us to assign portions of
+function to particular taxa or functional groups.
+Problems that need to be addressed: We shouldn’t assume that functional groups
+that putatively perform a process are the rate-limiting step or the strongest
+correlate.  There may be unknown members of the community that perform or limit
+a process The final step of the process may not be the rate-limiting step: it
+may be substrate production or nutrient acquisition or something else (e.g.
+Qusheng Jin’s work on modeling methanogenesis) We need to consider the
+underlying biogeographic patterns shaping the distribution of microbial
+communities (à la GWAS).
+
+\section{Conclusion}
 

### 2019-06-24 R/clean_metadata.R
Update varcomp, change seasonal wetland to upland

@@ -21,15 +21,15 @@ troph_data %<>% left_join(data_frame(Land_type = unique(troph_data$Land_type),
                               'Wetland', 'Upland', 'Upland', 'Upland'))), by = 'Land_type') %>% 
   select(Sample:Experiment, Wetland, Description:PC3_16S_rna)
 
-# Add WFPS variable
-Site_Pd <- data_frame(
-  Site = factor(unique(troph_data$Site)),
-  Pd = c(2.65, 2.65, 2.65, 2.65, 1.55, 2.65, 1.55, 2.65, 2.65, 2.65, 2.65, 2.65, 2.65, 2.65, 2.65)
-)
+## Add WFPS variable
+#Site_Pd <- data_frame(
+#  Site = factor(unique(troph_data$Site)),
+#  Pd = c(2.65, 2.65, 2.65, 2.65, 1.55, 2.65, 1.55, 2.65, 2.65, 2.65, 2.65, 2.65, 2.65, 2.65, 2.65)
+#)
 
-troph_data %<>% left_join(Site_Pd, by = 'Site') %>% 
-  mutate(WFPS = calc_wfps(Bulk_dens, Mois_cont, Pd)) %>% 
-  select(Sample:Low_final_k, WFPS, Bulk_dens:PC3_16S_rna, -Pd)
+#troph_data %<>% left_join(Site_Pd, by = 'Site') %>% 
+#  mutate(WFPS = calc_wfps(Bulk_dens, Mois_cont, Pd)) %>% 
+#  select(Sample:Low_final_k, WFPS, Bulk_dens:PC3_16S_rna, -Pd)
 
 # Separate numeric data (remove factors)
 troph_num_data <- troph_data %>% 
@@ -40,7 +40,7 @@ num_predictors <- troph_num_data %>% dplyr::select(-Vmax, -Low_final_k)
 
 # Attribute table
 troph_attr_table <- troph_data %>% 
-  select(-Rep, -Experiment, -Description, -Land_type, -Bulk_dens, -Mois_cont)
+  select(-Rep, -Experiment, -Description, -Land_type)
 
 # Organizing attribute table for easy lmer
 atr_tbl_long <- troph_attr_table %>% 
@@ -102,14 +102,15 @@ gen_data <- troph_data %>%
   right_join(gen_data) %>% 
   mutate(Site = factor(Site))
 
-Site_Pd <- data_frame(
[…]

### 2019-06-24 R/var_comp.R

@@ -20,18 +20,25 @@ knitr::opts_chunk$set(fig.width=8, fig.height=6, fig.path='Figs/',
 library(tidyverse)
 library(lubridate)
 library(broom)
-library(broman)
-library(morris)
+#library(broman)
+#library(morris)
 library(varComp)
 library(vegan)
-
+library(gap)
+library(parallel)
+#library(qvalue)
+num.cores <- detectCores()
 #+ r import_data
 
-all_data <- read.csv('../output/gab_all_vst_troph.csv', stringsAsFactors = FALSE)
+all_data <- as.data.frame(data.table::fread('../output/gab_all_vst_troph.csv'))
 
 #' # Variance Component Analysis
 
 #+ r sim_mat
+
+calc_sim_matrix <- function(x) {
+1/(1+as.matrix(dist(as.matrix(scale(x)))))
+}
 
 ## Remove ASVs only present in 1 sample
 asvs <- as.matrix(all_data[, grepl("^[asv]", colnames(all_data))])
@@ -39,51 +46,47 @@ vars <- all_data[, !grepl("^[asv]", colnames(all_data))]
 asvs <- asvs[, colSums(asvs) > 0]
 all_data <- cbind(vars, asvs)
 
-# Community similarity
-asv <- 3
-taxon <- asvs[, asv]
-#com_dis <- as.matrix(dist(as.matrix(scale(asv_mat[, -asv]))))
-#com_sim <- 1/(1+com_dis)
-com_sim <- 1 - as.matrix(vegdist(asvs[, -asv]))
+## Community similarity
+#asv <- 3
+#taxon <- asvs[, asv]
+#colnames(asvs)[asv]
+#com_sim <- 1 - as.matrix(vegdist(asvs[, -c(asv)]))
+#
+## Geographic similarity
+#geo_sim <- calc_sim_matrix(all_data[, c('X', 'Y')])
+#
+## environmental similarity'
+#env_sim <- calc_sim_matrix(all_data[, c('Bulk_dens', 'Mois_cont', 'N_percent', 'C_percent')])
 
-# Geographic similarity
-xy <- all_data[, c('X', 'Y')]
-geo_dis <- as.matrix(dist(as.matrix(scale(xy))))
-geo_sim <- 1/(1+geo_dis)
-
-# environmental dissimilarity
-env <- all_data[, c('WFPS', 'N_percent', 'C_percent')]
[…]

### 2019-06-24 R/analysis.R
Move model fitting to its own script (var_comp.R) and put all analysis output into an Rmd (analysis.R)

@@ -3,13 +3,11 @@
 #' subtitle: "Linking microbial communities to ecosystem functions: what we can learn from genotype-phenotype mapping in organisms"
 #' author: "Andrew Morris"
 #' date: "`r Sys.Date()`"
-#' output: html_document
+#' output: 
+#'   html_document:
+#'     number_sections: TRUE
 #' ---
 #' 
-#' # Overview
-#' 
-#' This is the primary analysis script for the analysis associated with the Royal Society
-#' paper. 
 
 #+ r global_options, include=FALSE
 knitr::opts_chunk$set(fig.width=8, fig.height=6, fig.path='Figs/',
@@ -29,12 +27,48 @@ library(parallel)
 #library(qvalue)
 library(knitr)
 library(kableExtra)
+library(eulerr)
 num.cores <- detectCores()
 #+ r import_data
 
 all_data <- as.data.frame(data.table::fread('../output/gab_all_vst_troph.csv'))
+n_samp <- nrow(all_data)
+n_asvs <- ncol(all_data[, grepl("^[asv]", colnames(all_data))])
 
+#' # Overview
+#' 
+#' The goal of this project is to answer the question: is microbial community 
+#' composition important for determining ecosystem-scale function independent 
+#' of the underlying environmental variation? To address this, I partitioned
+#' variation between community and environment (while controlling for geographic
+#' location using site as a factor). I also identified taxa that are significantly
+#' correlated with function after accounting for the covariance of geography,
+#' environment, and community similarity. Currently, significant taxa are identified
+#' using the Bonferroni correction (with the number of non-independent tests equal
+#' to the number of taxa included in each model). Eventually, I would like to
+#' determine p.values using a permutation test.
+#'
+#' These data were collected from Gabon, Africa and include `r n_samp` samples
+#' and `r n_asvs` ASVs from the DADA2 pipeline. The community matrix was transformed
+#' using the `varianceStabilizingTransformation` from `DESeq2`.
+#' This analysis only includes high affinity
+#' methane oxidation measurements (Low Final K). The spatial data were converted
[…]

### 2019-06-24 R/var_comp.R

@@ -1,42 +1,20 @@
-#' ---
-#' title: "Royal Society paper"
-#' subtitle: "Linking microbial communities to ecosystem functions: what we can learn from genotype-phenotype mapping in organisms"
-#' author: "Andrew Morris"
-#' date: "`r Sys.Date()`"
-#' output: html_document
-#' ---
-#' 
-#' # Overview
-#' 
-#' This is the primary analysis script for the analysis associated with the Royal Society
-#' paper. 
-
-#+ r global_options, include=FALSE
-knitr::opts_chunk$set(fig.width=8, fig.height=6, fig.path='Figs/',
-                      echo=FALSE, warning=FALSE, message=FALSE)
-
-#+ r load_packages, include=FALSE
+# load_packages, include=FALSE
 
 library(tidyverse)
-library(lubridate)
 library(broom)
-#library(broman)
-#library(morris)
 library(varComp)
 library(vegan)
-#library(gap)
 library(parallel)
-#library(qvalue)
-library(knitr)
-library(kableExtra)
+
 num.cores <- detectCores()
-#+ r import_data
+
+# import_data
 
 all_data <- as.data.frame(data.table::fread('../output/gab_all_vst_troph.csv'))
 
-#' # Variance Component Analysis
+# Variance Component Analysis
 
-#+ r sim_mat
+# sim_mat
 
 calc_sim_matrix <- function(x) {
 1/(1+as.matrix(dist(as.matrix(scale(x)))))
@@ -90,14 +68,14 @@ ggplot(mapping = aes(c(env_sim))) +
   geom_histogram(binwidth = 0.05) + 
   labs(x = "Environment Similarity (Euclidean)", y = "Number of Samples")
 
-#'
-#'Community and geography are correlated and sites cluster around three locations
-#'Also, geography uses up more of the similarity matrix space (see histograms)
-#'Therefore, I coded geography as three factors (one for each cluster of sites)
-#'Also, wetland/upland covaries with those three sites, but moisture content is
-#'included in the environmental similarity matrix and should capture that variation
+#
+#Community and geography are correlated and sites cluster around three locations
+#Also, geography uses up more of the similarity matrix space (see histograms)
[…]

### 2019-06-25 R/analysis.R
remove old Rmd and update text in analysis.R

@@ -16,7 +16,7 @@ knitr::opts_chunk$set(fig.width=8, fig.height=6, fig.path='Figs/',
 #+ r load_packages, include=FALSE
 
 library(tidyverse)
-library(lubridate)
+#library(lubridate)
 library(broom)
 #library(broman)
 #library(morris)
@@ -38,7 +38,7 @@ n_asvs <- ncol(all_data[, grepl("^[asv]", colnames(all_data))])
 #' # Overview
 #' 
 #' The goal of this project is to answer the question: is microbial community 
-#' composition important for determining ecosystem-scale function independent 
+#' composition important for determining the rate of ecosystem-scale functions independent 
 #' of the underlying environmental variation? To address this, I partitioned
 #' variation between community and environment (while controlling for geographic
 #' location using site as a factor). I also identified taxa that are significantly
@@ -139,16 +139,16 @@ model <- varComp(Low_final_k ~ asv, data = data,
                  varcov = list(com = com_sim, env = env_sim, geo = geo_sim) )
 summary(model)
 }
-
 varcomp_diag(all_data)
 
-#' ## Geographic Correlated with Community Composition
+#' ## Geography Correlated with Community Composition
 #'
 #' Community and geography are correlated as demonstrated by the Mantel test for
 #' the community similarity matrix and the geographic similarity matrix (see Mantel test output). 
 #' Even after variance stabilization, the community similarity metric is skewed
 #' towards one end of the variance space while geography uses up more of the 
-#' similarity matrix space (see histograms)
+#' similarity matrix space (see histograms). This results in geography
+#' explaining virtually all of the variation in ecosystem function.
 #' Since sites are clustered around three locations,
 #' I coded geography as three factors (one for each cluster of sites)
 #'Also, wetland/upland covaries with those three sites, but moisture content is
@@ -348,8 +348,8 @@ my_model_output  %>%
   left_join(taxon_table) %>% 
 group_by(comps) %>% 
   filter(p.value <= 0.05 / (length(unique(asv)))) %>% 
-  select(Phylum:Genus) %>% 
-  arrange(comps) %>% 
+  select(comps, Phylum:Genus) %>% 
+  arrange(comps, Phylum, Class, Order, Family, Genus) %>% 
   kable() %>% 
[…]

### 2019-06-25 R/var_comp.R
Replace mclapply with foreach and perform permutation test

@@ -1,16 +1,20 @@
-# load_packages, include=FALSE
 
 library(tidyverse)
 library(broom)
 library(varComp)
 library(vegan)
 library(parallel)
+library(foreach)
+library(doParallel)
+library(data.table)
 
 num.cores <- detectCores()
 
+registerDoParallel(cores = num.cores)
+
 # import_data
 
-all_data <- as.data.frame(data.table::fread('../output/gab_all_vst_troph.csv'))
+all_data <- as.data.frame(fread('../output/gab_all_vst_troph.csv'))
 
 # Variance Component Analysis
 
@@ -44,121 +48,97 @@ all_data[all_data$Y < 30000 & all_data$Y > 10000, 'geocode'] <- 'B'
 all_data[all_data$Y < 10000, 'geocode'] <- 'C'
 all_data$geocode <- factor(all_data$geocode)
 
-#+ r varcomp_diagnostics
+## Permute response
+permute_no_replace <- function(perm_vec, nperm = 1000) {
+samples <- matrix(ncol = nperm, nrow = length(perm_vec))
+foreach(i = seq_len(nperm), .combine = "cbind") %do% {
+  sample_i <- sample(perm_vec)
+  while (any(apply(samples, 2, function(x) {all(sample_i == x)}), na.rm = T)) {
+    sample_i <- sample(perm_vec)
+  }
+  samples[, i] <- sample_i
+}
+}
+permute_no_replace_par <- function(perm_vec, nperm = 1000) {
+samples <- matrix(ncol = nperm, nrow = length(perm_vec))
+samples <- cbind(samples, truth = perm_vec)
+foreach(i = seq_len(nperm), .combine = "cbind") %dopar% {
+  sample_i <- sample(perm_vec)
+  while (any(apply(samples, 2, function(x) {all(sample_i == x)}), na.rm = T)) {
+    sample_i <- sample(perm_vec)
+  }
+  samples[, i] <- sample_i
+}
+}
 
-mantel(com_sim, geo_sim)
-ggplot(all_data, aes(x = X, y = Y)) + geom_point(aes(color = all_data$Wetland))
-ggplot(all_data, aes(x = X, y = Y)) + geom_point(aes(color = all_data$geocode))
+permute_no_replace_for <- function(perm_vec, nperm = 1000) {
+samples <- matrix(ncol = nperm, nrow = length(perm_vec))
+samples <- cbind(samples, truth = perm_vec)
[…]

### 2019-07-02 R/brazil_cleaning.R
Create brazil taxonomy table

@@ -1,5 +1,6 @@
 library(tidyverse)
 library(DESeq2)
+library(data.table)
 
 ## Cleaning up metadata
 metadata <- read_tsv('../data/brazil/Dimensions_meta_data.txt')
@@ -14,6 +15,18 @@ metadata <- as_tibble(metadata)
 metadata <- mutate_at(metadata, vars(pH:Error_N2O), as.numeric)
 ## ASVs
 asvs <- as.data.frame(data.table::fread('../data/brazil/16S_dada2_table_RDP_tax.txt'))
+
+# Pull out taxonomy
+taxon_table <- 
+  asvs %>% 
+  select(OTU_ID, ConsensusLineage) %>% 
+  mutate(OTU_ID = paste('asv', OTU_ID, sep = "_")) %>% 
+  select(asv = OTU_ID, ConsensusLineage) %>% 
+  separate(ConsensusLineage, c('Domain', 'Phylum', 'Class', 'Order', 'Family',
+                               'Genus', 'Species', 'Subspecies'), ";")
+fwrite(taxon_table, '../output/brazil_taxon_table.csv')
+
+# Pull out abundances and do variance stabilizing transformation
 asvs <-
   asvs %>% 
   select(-ConsensusLineage)

### 2019-07-02 R/analysis.R
Figure out h2GE

@@ -29,7 +29,14 @@ library(knitr)
 library(kableExtra)
 library(eulerr)
 library(GGally)
+library(phyloseq)
+library(doParallel)
+library(gap)
 num.cores <- detectCores()
+registerDoParallel(cores = 3)
+
+#+ r tree_analysis
+tree <- read_tree('../data/gabon/16S_rep_seqs_tree_Gabon.nwk')
 #+ r import_data
 
 taxon_table <- as.data.frame(data.table::fread('../output/taxon_table.csv'))
@@ -524,5 +531,45 @@ dry_model_output  %>%
 
 
 plot_asv_euler(dry_model_output)
+
+#+ phylo_tree
+colnames(asvs) <- substr(colnames(asvs), 5, nchar(colnames(asvs)[1]))
+otu_table <- otu_table(asvs, taxa_are_rows = FALSE)
+phylo_object <- phyloseq(otu_table, tree)
+unifrac <- UniFrac(phylo_object, TRUE)
+uni_sim <- 1/(1+as.matrix(unifrac))
+ggplot(mapping = aes(c(uni_sim))) +
+  geom_histogram(bins = 20)
+
+model <- varComp(Low_final_k ~ 1, data = all_data,
+                 varcov = list(uni = uni_sim, com = com_sim) )
+summary(model)
+h2 <- h2GE(c(model$varComps, err = model$sigma2), vcov(model, what = 'varComp'))
+var_comps_se <- data.frame(comp = c('uni', 'com'), h2 = c(h2$h2G, h2$h2GE), 
+           SE = c(sqrt(h2$Varh2G), sqrt(h2$Varh2GE)))
+ggplot(var_comps_se, aes(x = comp, y = h2, ymax = h2 + SE, ymin = h2 - SE)) +
+  geom_pointrange() 
+
+ggplot(mapping = aes(x = c(com_sim), y = c(uni_sim))) +
+  geom_point() +
+  geom_abline(intercept = 0, slope = 1) +
+  coord_cartesian(xlim = c(0, 1), ylim = c(0, 1))
+model <- varComp(Low_final_k ~ 1, data = all_data,
+                 varcov = list(com = uni_sim) )
+summary(model)
+
+h2G(c(model$varComps['com'], err = model$sigma2), vcov(model, what = 'varComp'))
+
+model <- varComp(Low_final_k ~ asv, data = all_data,
+                 varcov = list(com = uni_sim) )
+summary(model)
+
[…]

### 2019-07-02 R/analysis_brazil.R
Analyze brazil dataset

@@ -1,0 +1,436 @@
+#' ---
+#' title: "Royal Society paper"
+#' subtitle: "Linking microbial communities to ecosystem functions: what we can learn from genotype-phenotype mapping in organisms"
+#' author: "Andrew Morris"
+#' date: "`r Sys.Date()`"
+#' output: 
+#'   html_document:
+#'     number_sections: TRUE
+#' ---
+#' 
+
+#+ r global_options, include=FALSE
+knitr::opts_chunk$set(fig.width=8, fig.height=6, fig.path='Figs/',
+                      echo=FALSE, warning=FALSE, message=FALSE)
+
+#+ r load_packages, include=FALSE
+
+library(tidyverse)
+#library(lubridate)
+library(broom)
+#library(broman)
+#library(morris)
+library(varComp)
+library(vegan)
+#library(gap)
+library(parallel)
+#library(qvalue)
+library(knitr)
+library(kableExtra)
+library(eulerr)
+library(GGally)
+num.cores <- detectCores() - 1
+#+ r import_data
+
+taxon_table <- as.data.frame(data.table::fread('../output/brazil_taxon_table.csv'))
+all_data <- as.data.frame(data.table::fread('../output/brazil_cleaned_data.tsv'))
+all_data <- all_data[!is.na(all_data$Long), ]
+n_samp <- nrow(all_data)
+n_asvs <- ncol(all_data[, grepl("^[asv]", colnames(all_data))])
+
+## Remove ASVs only present in 1 sample
+asvs <- as.matrix(all_data[, grepl("^[asv]", colnames(all_data))])
+vars <- all_data[, !grepl("^[asv]", colnames(all_data))]
+asvs <- asvs[, colSums(asvs) > 0]
+all_data <- cbind(vars, asvs)
+
+#' # Overview
+#' 
+#' The goal of this project is to answer the question: is microbial community 
+#' composition important for determining the rate of ecosystem-scale functions independent 
+#' of the underlying environmental variation? To address this, I partitioned
+#' variation between community and environment (while controlling for geographic
+#' location using site as a factor). I also identified taxa that are significantly
+#' correlated with function after accounting for the covariance of geography,
+#' environment, and community similarity. Currently, significant taxa are identified
+#' using the Bonferroni correction (with the number of non-independent tests equal
[…]

### 2019-07-02 R/var_comp_brazil.R

@@ -1,0 +1,236 @@
+library(tidyverse)
+library(broom)
+library(varComp)
+library(vegan)
+library(parallel)
+library(foreach)
+library(doParallel)
+library(data.table)
+library(microbenchmark)
+
+registerDoParallel(cores = 3)
+num.cores <- detectCores() - 1
+
+all_data <- as.data.frame(fread('../output/brazil_cleaned_data.tsv'))
+all_data <- all_data[!is.na(all_data$Long), ]
+
+# Variance Component Analysis
+
+# sim_mat
+calc_sim_matrix <- function(x) {
+1/(1+as.matrix(dist(as.matrix(scale(x)))))
+}
+
+## Remove ASVs only present in 1 sample
+asvs <- as.matrix(all_data[, grepl("^[asv]", colnames(all_data))])
+vars <- all_data[, !grepl("^[asv]", colnames(all_data))]
+asvs <- asvs[, colSums(asvs) > 0]
+all_data <- cbind(vars, asvs)
+
+## Functions
+#hist(all_data[, 'CH4'])
+#hist(log(all_data$CH4 + abs(min(all_data$CH4)) + 1))
+#hist(all_data[, 'CO2'])
+#hist(log(all_data[, 'CO2']))
+#hist(all_data[, 'N2O'])
+all_data$log_CH4_c <- log(all_data$CH4 + abs(min(all_data$CH4)) + 1)
+all_data$log_CO2 <- log(all_data$CO2)
+functions <- all_data[, c('log_CH4_c', 'log_CO2', 'N2O')]
+# Community similarity
+asv <- 3
+taxon <- asvs[, asv]
+asv_tax <- colnames(asvs)[asv]
+com_sim <- 1 - as.matrix(vegdist(asvs[, -c(asv)]))
+
+# Geographic similarity
+geo_sim <- calc_sim_matrix(all_data[, c('Lat', 'Long')])
+
+# environmental similarity'
+env_sim <- calc_sim_matrix(select(all_data, pH:Total_moisture))
+
+#plot(all_data$Lat, all_data$Long)
+
+
+## Permute response
+#permute_no_replace_for <- function(perm_vec, nperm = 1000) {
+#samples <- matrix(ncol = nperm, nrow = length(perm_vec))
+#samples <- cbind(samples, truth = perm_vec)
+#for (i in seq_len(nperm)) {
+#  sample_i <- sample(perm_vec)
[…]

### 2019-07-02 R/var_comp_test.R
Try out different parallelization approaches

@@ -1,0 +1,240 @@
+library(tidyverse)
+library(broom)
+library(varComp)
+library(vegan)
+library(parallel)
+library(foreach)
+library(doParallel)
+library(data.table)
+library(microbenchmark)
+
+registerDoParallel(cores = 4)
+num.cores <- 4
+all_data <- as.data.frame(fread('../output/brazil_cleaned_data.tsv'))
+all_data <- all_data[!is.na(all_data$Long), ]
+
+# Variance Component Analysis
+
+# sim_mat
+calc_sim_matrix <- function(x) {
+1/(1+as.matrix(dist(as.matrix(scale(x)))))
+}
+
+## Remove ASVs only present in 1 sample
+asvs <- as.matrix(all_data[, grepl("^[asv]", colnames(all_data))])
+vars <- all_data[, !grepl("^[asv]", colnames(all_data))]
+asvs <- asvs[, colSums(asvs) > 0]
+all_data <- cbind(vars, asvs)
+
+## Functions
+#hist(all_data[, 'CH4'])
+#hist(log(all_data$CH4 + abs(min(all_data$CH4)) + 1))
+#hist(all_data[, 'CO2'])
+#hist(log(all_data[, 'CO2']))
+#hist(all_data[, 'N2O'])
+all_data$log_CH4_c <- log(all_data$CH4 + abs(min(all_data$CH4)) + 1)
+all_data$log_CO2 <- log(all_data$CO2)
+functions <- all_data[, c('log_CH4_c', 'log_CO2', 'N2O')]
+# Community similarity
+asv <- 3
+taxon <- asvs[, asv]
+asv_tax <- colnames(asvs)[asv]
+com_sim <- 1 - as.matrix(vegdist(asvs[, -c(asv)]))
+
+# Geographic similarity
+geo_sim <- calc_sim_matrix(all_data[, c('Lat', 'Long')])
+
+# environmental similarity'
+env_sim <- calc_sim_matrix(select(all_data, pH:Total_moisture))
+
+#plot(all_data$Lat, all_data$Long)
+
+
+## Permute response
+#permute_no_replace_for <- function(perm_vec, nperm = 1000) {
+#samples <- matrix(ncol = nperm, nrow = length(perm_vec))
+#samples <- cbind(samples, truth = perm_vec)
+#for (i in seq_len(nperm)) {
+#  sample_i <- sample(perm_vec)
[…]

### 2019-07-02 manuscript/main.tex
Update text

@@
 by considering environmental and spatial structure in the data. 
-\item Figure out variance partitioning (with Hannah)
-\item Maybe variance partitioning plus correlations (varPart)
-\item Do PC correction in all combinations
-\item Identify significant taxa and build a venn diagram figure (or
-something)
-\item 
+\item Interpret model output
+\item Decide whether to include Gabon data, Brazil data, or both
 \end{enumerate}
+
+\section{Outline}
+\begin{enumerate}
+  \item First paragraph laying out the problem
+  \item BEF research on macroorganisms has demonstrated that productivity increases with
+    taxonomic and functional diversity.
+  \item This research has been less definitive for microbial communities. There
+    are two main hypotheses that aren't very well supported: function should
+    scale with community size or abundance of marker genes.
+  \item We propose that this BEF approach to linking microbial communities
+    to ecosystem function is inappropriate and instead this problem is more
+    analagous to the population genetics challenge of linking many genetic
+    markers to an organismal phenotype.
+  \item To incorporate this idea in to microbial community function research,
+    we should consider the 
+\end{enumerate}
+
 \section{Introduction}
 
-I want to know whether microbial community composition is important for determining ecosystem-scale functions independent of the underlying environmental variation.
-
-Microorganisms mediate many biogeochemical processes, but despite 
-much effort to link microbial communities
-to these functions the mechanism by which they regulate these
-processes remains unknown. Ecologists have investigated the effects of changing
-biodiversity on ecosystem function for decades, but many of their approaches
-are not applicable in microbial ecology due to the nature of microbial data. 
-This limits our ability to predict and manage globally relevant biogeochemical cycles. To
+Microorganisms mediate many biogeochemical processes and yet the link between
+microbial biodiversity and ecosystem function remains unknown. 
+Ecologists have investigated the effects of changing
+biodiversity on ecosystem function for decades. This research has primarily
+focused on linking terrestrial plant community richness with productivity
+\citep{hooper2005}. More recently, there has been an interest in linking
+microbial community composition to host disease, soil greenhouse gas emissions,
+and plant productivity. However, it remains unclear the extent to which 
+microbial community composition is important for understanding ecosystem-scale
+functions.  This uncertainty limits our ability to predict and manage 
+globally relevant biogeochemical cycles and host traits. To
 address this challenge, we propose integrating current theory on community
 assembly and the functioning of ecosystems with lessons from genotype-phenotype
-mapping in quantitative genetics. This will address the
-question of how does variation in microbial community structure contribute to
-variation in the rate of ecosystem function?
+mapping in quantitative genetics. This connection will address
+whether microbial community
+composition is important for determining ecosystem-scale functions independent
+of the underlying environmental variation.
 

### 2019-07-04 R/analysis.R
Clean up h2GE analysis

@@ -533,6 +533,19 @@ dry_model_output  %>%
 plot_asv_euler(dry_model_output)
 
 #+ phylo_tree
+
+x <- 3
+asv <- asvs[, x]
+asv_tax <- colnames(asvs)[x]
+com_sim <- 1 - as.matrix(vegdist(asvs[, -c(x)]))
+
+# Geographic similarity
+geo_sim <- calc_sim_matrix(all_data[, c('X', 'Y')])
+
+# environmental similarity'
+env_sim <- calc_sim_matrix(all_data[, c('Bulk_dens', 'Mois_cont', 'N_percent', 'C_percent')])
+asvs <- as.matrix(all_data[, grepl("^[asv]", colnames(all_data))])
+asv <- asvs[, x]
 colnames(asvs) <- substr(colnames(asvs), 5, nchar(colnames(asvs)[1]))
 otu_table <- otu_table(asvs, taxa_are_rows = FALSE)
 phylo_object <- phyloseq(otu_table, tree)
@@ -541,14 +554,6 @@ uni_sim <- 1/(1+as.matrix(unifrac))
 ggplot(mapping = aes(c(uni_sim))) +
   geom_histogram(bins = 20)
 
-model <- varComp(Low_final_k ~ 1, data = all_data,
-                 varcov = list(uni = uni_sim, com = com_sim) )
-summary(model)
-h2 <- h2GE(c(model$varComps, err = model$sigma2), vcov(model, what = 'varComp'))
-var_comps_se <- data.frame(comp = c('uni', 'com'), h2 = c(h2$h2G, h2$h2GE), 
-           SE = c(sqrt(h2$Varh2G), sqrt(h2$Varh2GE)))
-ggplot(var_comps_se, aes(x = comp, y = h2, ymax = h2 + SE, ymin = h2 - SE)) +
-  geom_pointrange() 
 
 ggplot(mapping = aes(x = c(com_sim), y = c(uni_sim))) +
   geom_point() +
@@ -568,8 +573,14 @@ h2G(c(model$varComps['com'], err = model$sigma2), vcov(model, what = 'varComp'))
 
 model <- varComp(Low_final_k ~ asv + geocode, data = all_data,
                  varcov = list(com = uni_sim, env = env_sim) )
+all_data$geocode
+class(uni_sim)
 summary(model)
 vcov(model, what = 'varComp')
[…]

### 2019-07-04 R/analysis_brazil.R

@@ -432,5 +432,15 @@ group_by(comps) %>%
 
 #plot_asv_euler(my_model_output)
 
+model <- varComp(log_CH4_c ~ asv + Region + Land_type, data = all_data,
+                 varcov = list(com = com_sim, env = env_sim) )
+summary(model)
+vcov(model, what = 'varComp')
+h2 <- h2GE(c(model$varComps, err = model$sigma2), vcov(model, what = 'varComp'))
+
+var_comps_se <- data.frame(comp = c('com', 'env'), h2 = c(h2$h2G, h2$h2GE), 
+           SE = c(sqrt(h2$Varh2G), sqrt(h2$Varh2GE)))
+ggplot(var_comps_se, aes(x = comp, y = h2, ymax = h2 + SE, ymin = h2 - SE)) +
+  geom_pointrange() 
 
 sessionInfo()

### 2019-07-04 manuscript/main.tex
Add updated manuscript text

--- 2019-06-24 manuscript/main.tex
@@
 by considering environmental and spatial structure in the data. 
-\item Figure out variance partitioning (with Hannah)
-\item Maybe variance partitioning plus correlations (varPart)
-\item Do PC correction in all combinations
-\item Identify significant taxa and build a venn diagram figure (or
-something)
-\item 
+\item Interpret model output
+\item Decide whether to include Gabon data, Brazil data, or both
 \end{enumerate}
+
 \section{Introduction}
 
-I want to know whether microbial community composition is important for determining ecosystem-scale functions independent of the underlying environmental variation.
+Microorganisms mediate many biogeochemical processes and yet the link between
+microbial biodiversity and ecosystem function remains unknown. 
+Ecologists have investigated the effects of changing
+biodiversity on ecosystem function for decades, but this research has primarily
+focused on linking terrestrial plant community richness with productivity
+\citep{hooper2005}. More recently, there has been an interest in linking
+microbial community composition to ecosystem functions such as host disease 
+and soil greenhouse gas emissions \citep{schimel1998, singh2010}. 
+However, it remains unclear to what extent 
+microbial community composition is important for determining the rates of ecosystem-scale
+functions.  This uncertainty limits our ability to predict and manage host
+health and global biogeochemical cycles. To
+address this challenge, we propose integrating traditional
+biodiversity-ecosystem function research with ideas from genotype-phenotype
+mappingin population genetics. This connection will help determine
+whether microbial community
+composition alters ecosystem-scale functions independent
+of the underlying environmental variation.
 
-Microorganisms mediate many biogeochemical processes, but despite 
-much effort to link microbial communities
-to these functions the mechanism by which they regulate these
-processes remains unknown. Ecologists have investigated the effects of changing
-biodiversity on ecosystem function for decades, but many of their approaches
-are not applicable in microbial ecology due to the nature of microbial data. 
-This limits our ability to predict and manage globally relevant biogeochemical cycles. To
-address this challenge, we propose integrating current theory on community
-assembly and the functioning of ecosystems with lessons from genotype-phenotype
-mapping in quantitative genetics. This will address the
-question of how does variation in microbial community structure contribute to
-variation in the rate of ecosystem function?
+If local selection always optimizes the available biodiversity to maximize
+ecosystem function, them microbial community composition should not matter for
+pedicting the rates of microbially mediated functions. Instead, ecosystem function
+should be predicted from the underlying environmental variation. In this case,
+the microbial community simply acts as a conduit through which the abiotic
+environment alters ecosystem function. However, experiments that manipulate the
+connection between environmental factors and microbial community composition
+such as environmental treatments, common gardens, and reciprocal transplants
+have shown different rates of ecosystem functions for different microbial
+communities. This has been observed for plant decomposition
+\citep{strickland2009}, plant phenology \citep{panke-buisse2015}, soil nitrogen
+cycling \citep{balser2005}, and soil greenhouse gas emissions
+\citep{cavigelli2000}. Therefore, variation in microbial community composition
+seems to generate variation in ecosystem function. However, in order to
+incorporate microorganisms into ecosystem models, we need to identify what
+aspects of microbial biodiversity are associated with variation in ecosystem
+function.
+
+Most studies of microbial community function focus on one of two aspects of
+microbial community structure that are hypothesized to be predictive of
+ecosystem function. The first aspect is functional gene or transcript abundance.
+In this case, qPCR or shotgun metagonemic sequencing is used to target a gene
+or transcript that is a putative marker for a microbial process. For example,
+the gene \textit{mcrA}, which encodes the enzyme that performs the final step in
+methanogenesis, is hypothesized to be correlated with the rate of
+methane production. Other examples include \textit{pmoA} and methanotrophy,
+\textit{nifH} and nitrification, and \textit{nosZ} and denitrification. The
+rational for this hypothesis is that the abundance of a functional group (as
+inferred from marker gene abundance) should correspond to the rate of the
+process performed by that group. There are examples where this relationship
+holds \citep{freitag2009, freitag2010, schnyder2018}. However, a review of these
+studies found that the abundance of a functional gene or transcript is rarely
+correlated with the rate of the corresponding process, with most effects either
+negative or not significant \citep{rocca2015}.
+
+The second aspect of microbial community structure hypothesized to predict
+ecosystem function is taxonomic or functional
+diversity. Diversity is either estimated using a measure of community
+composition such as 16S rRNA gene ribotyping or manipulated through some proxy
+of diversity such as sequential dilution or varying filter sizes. Studies that
+experimentally manipulate diversity generally find a relationship between
+the applied diversity treatment and ecosystem functions including methanotrophy
+\citep{schnyder2018}, phosphorus leaching, greenhouse gas emissions
+\citep{wagg2014}, and decomposition \citep{maron2018}. However, in each of these
+cases, richness is confounded with other factors such as presence or absence of
+major phylogenetic groups (nematodes, fungi) or abundance of microbial cells. In
+addition, assembled microbial communities pose the same challenges as
+macroorganismal diversity experiments, for example no replication at the highest
+diversity treatment. A review of microbial diversity studies shows that
+microbial community taxonomic and functional diversity add little variance
+explained to models of ecosystem function. Overall, functional gene and
+abundance and community doversity improve models of ecosystem function less than
+one third of the time and increase variance explained by only 8 percentage
+points \citep{graham2016}.
+
+These approaches follow on the tradition of macroorganismal
+biodiversity-ecosystem function research by measuring or manipulating diversity
+and comparing it to function or by looking at the association between richness
+or abundance of a functional group (e.g. plants) and the function performed by
+that group (e.g. productivity). However, in the case of microbes, our knowledge
+of functional groups is quite limited. We cannot be sure that we're accurately
+estimating the abundance or diversity of soil microbial functional groups the
+way we can, say, plants in a grassland.
+
+Organisms exhibit population differentiation due to geographic and ecological
+processes \citep{wright1943}. Neutral processes such as local genetic drift and
+reduced gene flow due to geographic separation can lead to population
+differentiation. In addition, environmental distance can lead to selection
+against migrants \citep{hendry2004} and local selection on endemic populations
+to generate variation. Microorganisms exhibit biogeographic patterns due to
+ecological selection and geographic separation \citep{martiny2006}. If dispersal
+is non-limiting then the most productive species should always be present and
+composition should not matter for local ecosystem function. However, if important
+functional groups are dispersal-limited, lost to drift, or are unable to
+colonize a site, then there could be important functional consequences of
+community differentiation to geographic or ecological isolation.
+
+There are a numbber of ways that environmental or spatial distance could obscure
+microbial BEF relationships. For example, historical effects of geographic
+isolation such as dispersal limitation could lead to spurious associations
+between the abundance of taxa and the rate of ecosystem function. Community
+stratification where taxa co-occur due to historical effects or local selection.
+Environmental selection can also shape microbial community
+structure. For example, pH is known to regulate both broad biodiversity patterns for
+microbial communities \citep{fierer2006} and methane cycling dynamics
+\citep{ye2012}. This could like to microbial taxa that are correlated with
+methane cycling, but are mechanisticalyl unrelated. In fact, if the rate of
+ecosystem function is high in one set of ecosystems, function will be
+associated with any taxa that are abundant in that set of
+ecosystems.
+
+In order to understand the biological basis of ecosystem function, we need to
+separate the close association between environmental variation, community
+composition, and ecosystem function in order to identify what aspects of
+microbial biodiversity determine the rate of ecosystem function.
+
+The challenge of mapping microbial community structure to ecosystem function is
+analogous to the challenge of identifying the genetic basis of complex traits in
+organismal populations. In the latter case, a population exhibits variation in a
+phenotype (e.g. hight or disease state) as well as variation in potentially
+thousands of genetic loci (usually identified as single nucleotide polymorphisms
+(SNPs)). To identify the genetic basis for a trait, investigators sample from
+this population and correlate phenotype with genotype. In most cases, there are
+many more loci than individuals and we do not know the function of genes
+associated with these SNPs. Genetic markers identified in these genome-wide association
+studies can be the target for future research to dissect the function of those
+genes, can be added or removed to test their association with the trait, or can
+be used as marker for that disease or phenotype even if it has no known
+mechanistic association.
+
+This mirrors the problem of microbial
+structure-function mapping. In this case, a population of hosts or ecosystems exhibit
+variation in a particular trait or ecosystem function putatively mediated by
+microorganisms as well as variation in metagenomic composition. Investigators
+could sample from this population of ecosystems and correlate the rate of an
+ecosystem function with microbial community composition or gene composition.
+
+Population geneticists trying to make robust associations between genetic loci
+and phenotype must take into account population stratification. Population
+stratification due to shared ancestry or local selection can lead to spurious
+associations between genes and traits. For example, human height varies
+across Europe in a latitudinal gradient. Positive associations between genes and
+phenotypes can arise because of true associations, through linkage
+disequilibrium with true genes, or through admixture \citep{lander1994}.
+Microbial structure-function connections could pose these same problems. There
+could be true associations between a microorganism and the ecosystem function
+performed by that taxon, or there could be there could be systematic
+co-occurrence patterns between one taxon and a causal taxon like metabolic
+networks, or there could be admixture due to shared history or selection within
+a sub-set of ecosystems. Robust structure-function associations rely on
+controlling for these factors.
 

### 2019-07-07 R/clean_metadata.R
Include land type in Gabon dataset

@@ -40,7 +40,7 @@ num_predictors <- troph_num_data %>% dplyr::select(-Vmax, -Low_final_k)
 
 # Attribute table
 troph_attr_table <- troph_data %>% 
-  select(-Rep, -Experiment, -Description, -Land_type)
+  select(-Rep, -Experiment, -Description)
 
 # Organizing attribute table for easy lmer
 atr_tbl_long <- troph_attr_table %>% 
@@ -121,7 +121,7 @@ num_predictors <- gen_num_data %>% dplyr::select(-CH4, -H_CH4_percent)
 
 # Attribute table
 gen_attr_table <- gen_data %>% 
-  select(-Rep, -Experiment, -Description, -Land_type)
+  select(-Rep, -Experiment, -Description)
 
 # Remove extraneous attributes
 colnames(gen_attr_table)

### 2019-07-07 R/analysis.R
Clean up varComp analyses and start implementing new approach

@@ -22,16 +22,17 @@ library(broom)
 #library(morris)
 library(varComp)
 library(vegan)
-#library(gap)
 library(parallel)
-#library(qvalue)
+library(qvalue)
 library(knitr)
 library(kableExtra)
-library(eulerr)
-library(GGally)
+#library(eulerr)
+#library(GGally)
 library(phyloseq)
 library(doParallel)
 library(gap)
+library(lmtest)
+
 num.cores <- detectCores()
 registerDoParallel(cores = 3)
 
@@ -39,8 +40,8 @@ registerDoParallel(cores = 3)
 tree <- read_tree('../data/gabon/16S_rep_seqs_tree_Gabon.nwk')
 #+ r import_data
 
-taxon_table <- as.data.frame(data.table::fread('../output/taxon_table.csv'))
-all_data <- as.data.frame(data.table::fread('../output/gab_all_vst_troph.csv'))
+taxon_table <- data.table::fread('../output/taxon_table.csv', data.table = FALSE)
+all_data <- data.table::fread('../output/gab_all_vst_troph.csv', data.table = FALSE)
 n_samp <- nrow(all_data)
 n_asvs <- ncol(all_data[, grepl("^[asv]", colnames(all_data))])
 
@@ -49,6 +50,12 @@ asvs <- as.matrix(all_data[, grepl("^[asv]", colnames(all_data))])
 vars <- all_data[, !grepl("^[asv]", colnames(all_data))]
 asvs <- asvs[, colSums(asvs) > 0]
 all_data <- cbind(vars, asvs)
+
+# Add location factor
+all_data[all_data$Y > 30000, 'geocode'] <- 'A'
+all_data[all_data$Y < 30000 & all_data$Y > 10000, 'geocode'] <- 'B'
+all_data[all_data$Y < 10000, 'geocode'] <- 'C'
+all_data$geocode <- factor(all_data$geocode)
 
 #' # Overview
 #' 
@@ -109,7 +116,10 @@ all_data <- cbind(vars, asvs)
 #+ trad_correlations
 
 colnames(all_data[, 1:13])
-ggpairs(all_data[, 3:13])
+
+print(ggplot(all_data, aes(x = X, y = Y)) + 
+  geom_point(aes(color = all_data$Wetland)) +
+  labs(title = "Distribution of sites highlighted by wetland or upland"))
 
[…]

### 2019-07-07 R/analysis_brazil.R

@@ -27,13 +27,16 @@ library(parallel)
 #library(qvalue)
 library(knitr)
 library(kableExtra)
-library(eulerr)
-library(GGally)
+#library(eulerr)
+#library(GGally)
+library(gap)
+library(lmtest)
+
 num.cores <- detectCores() - 1
 #+ r import_data
 
-taxon_table <- as.data.frame(data.table::fread('../output/brazil_taxon_table.csv'))
-all_data <- as.data.frame(data.table::fread('../output/brazil_cleaned_data.tsv'))
+taxon_table <- data.table::fread('../output/brazil_taxon_table.csv', data.table = FALSE)
+all_data <- data.table::fread('../output/brazil_cleaned_data.tsv', data.table = FALSE)
 all_data <- all_data[!is.na(all_data$Long), ]
 n_samp <- nrow(all_data)
 n_asvs <- ncol(all_data[, grepl("^[asv]", colnames(all_data))])
@@ -109,16 +112,16 @@ all_data <- cbind(vars, asvs)
 #+ trad_correlations, fig.width=10, fig.height=10
 
 colnames(all_data[, 1:40])
-ggcorr(all_data[, c(2:3, 8:34, 36, 38)], nbreaks = 4, label = TRUE)
+#ggcorr(all_data[, c(2:3, 8:34, 36, 38)], nbreaks = 4, label = TRUE)
 
 #+ other_correlations
-
-log_CH4_c <- with(all_data, log(CH4 + abs(min(CH4)) + 1))
+all_data$log_CH4_c <- with(all_data, log(CH4 + abs(min(CH4)) + 1))
 pca <- rda(as.matrix(all_data[, grepl("^[asv]", colnames(all_data))]))
 screeplot(pca)
-#summary(pca)
-pc1 <- scores(pca)$sites[, 'PC1']
-pc2 <- scores(pca)$sites[, 'PC2']
+
+all_data$pc1 <- scores(pca, choices = 1, display = 'sites', scaling = 3)
+all_data$pc2 <- scores(pca, choices = 2, display = 'sites', scaling = 3)
+all_data$pc3 <- scores(pca, choices = 3, display = 'sites', scaling = 3)
 
 ggplot(all_data, aes(x = pc1, y = CH4)) +
        geom_point() +
@@ -160,21 +163,42 @@ ggplot(all_data, aes(x = pc2, y = N2O)) +
        geom_smooth(method = 'lm')
[…]

### 2019-07-09 manuscript/main.tex
Update text with latest revision

@@
 \title{Linking microbial communities to ecosystem functions: what we can learn
-from genotype-phenotype mapping in organisms} \author{Andrew H. Morris, Brendan
-J. M. Bohannan}
+from genotype-phenotype mapping in organisms}
+
+\author{Andrew H. Morris, Brendan J. M. Bohannan}
 
@@
 
+\section*{Abstract}
+
 \section{Introduction}
@@
 biodiversity-ecosystem function research with ideas from genotype-phenotype
-mappingin population genetics. This connection will help determine
+mapping in population genetics. This connection will help determine
 whether microbial community
@@
 ecosystem function, them microbial community composition should not matter for
-pedicting the rates of microbially mediated functions. Instead, ecosystem function
-should be predicted from the underlying environmental variation. In this case,
-the microbial community simply acts as a conduit through which the abiotic
+pedicting the rates of microbially-mediated functions. Instead, 
+the rates of these functions should be determined by the
+underlying environmental variation. In this case, the microbial community simply
+acts as a conduit through which the abiotic
 environment alters ecosystem function. However, experiments that manipulate the
@@
 \textit{nifH} and nitrification, and \textit{nosZ} and denitrification. The
-rational for this hypothesis is that the abundance of a functional group (as
+rationale for this hypothesis is that the abundance of a functional group (as
 inferred from marker gene abundance) should correspond to the rate of the
@@
 microbial community taxonomic and functional diversity add little variance
-explained to models of ecosystem function. Overall, functional gene and
-abundance and community doversity improve models of ecosystem function less than
-one third of the time and increase variance explained by only 8 percentage
+explained to models of ecosystem function. Overall, functional gene
+abundance and community diversity improve models of ecosystem function less than
+one third of the time and increase variance explained by an average of only 8 percentage
 points \citep{graham2016}.
@@
 estimating the abundance or diversity of soil microbial functional groups the
-way we can, say, plants in a grassland.
+way we can plants in a grassland.
 
@@
 against migrants \citep{hendry2004} and local selection on endemic populations
-to generate variation. Microorganisms exhibit biogeographic patterns due to
+to generate variation. Microorganisms also exhibit biogeographic patterns due to
 ecological selection and geographic separation \citep{martiny2006}. If dispersal
@@
 
-There are a numbber of ways that environmental or spatial distance could obscure
+There are a number of ways that environmental or spatial distance could obscure
 microbial BEF relationships. For example, historical effects of geographic
@@
 microbial communities \citep{fierer2006} and methane cycling dynamics
-\citep{ye2012}. This could like to microbial taxa that are correlated with
-methane cycling, but are mechanisticalyl unrelated. In fact, if the rate of
+\citep{ye2012}. This could lead to microbial taxa that are correlated with
+methane cycling, but are mechanistically unrelated. In fact, if the rate of
 ecosystem function is high in one set of ecosystems, function will be
-associated with any taxa that are abundant in that set of
-ecosystems.
-
+associated with any taxa that are abundant in those ecosystems. 
 In order to understand the biological basis of ecosystem function, we need to
@@
 organismal populations. In the latter case, a population exhibits variation in a
-phenotype (e.g. hight or disease state) as well as variation in potentially
+phenotype (e.g. height or disease state) as well as variation in potentially
 thousands of genetic loci (usually identified as single nucleotide polymorphisms
@@
 be used as marker for that disease or phenotype even if it has no known
-mechanistic association.
-
-This mirrors the problem of microbial
+mechanistic association. This mirrors the problem of microbial
 structure-function mapping. In this case, a population of hosts or ecosystems exhibit
@@
 could be true associations between a microorganism and the ecosystem function
-performed by that taxon, or there could be there could be systematic
+performed by that taxon, or there could be systematic
 co-occurrence patterns between one taxon and a causal taxon like metabolic
 networks, or there could be admixture due to shared history or selection within
-a sub-set of ecosystems. Robust structure-function associations rely on
+a subset of ecosystems. This is true for ASVs in ribotyping 16S datasets and for
+genes in metagenomic (or other -omic) datsets. 
+Robust structure-function associations rely on
 controlling for these factors.
+
+\section{Methods}
+
+\subsection{Sample collection}
+
+\subsection{Determination of the rates of ecosystem functions}
+
+\subsection{Measurement of environmental variables}
+
+\subsection{Sequencing and bioinformatic processing}
+
+DADA2 pipeline for amplicon sequence variant (ASV) assignment.
+
+We used the variance stabilizing transformation from the \texttt{DESeq2}
+package to adjust for differences in sequencing depth between samples.
+
+\subsection{Statistical analysis}
+
+All statistical analyses were performed using the \texttt{R} statistical
+environment. 
+We first applied traditional tests of microbial biodiversity-ecosystem function
+relationships. To do this, we fit ordinary least squares regression models with
+the rate of ecosystem function as the response with one predictor for each
+model. These predictors include functional gene and transcript abundance,
+diversity measures including richness, Shannon's diversity index, and Simpson's
+diversity index, measures of microbial community composition using the first
+several principal component axes from a principal component
+analysis (PCA) chosen based on visual inspection of a screeplot incpluding a PCA for
+abundance data and a PCA for presence/absence data, and several 
+environmental variables.
+
+Next, we used a statistical approach common in population genetics genome-wide
+association studies. We fit variance component mixed models from the
+\texttt{varComp} package to test whether variation in microbial
+community composition expains a significant portion of the variation in
+ecosystem function \citep{qu2013}. First, we fit an intercept-only model with one random effect
+for community similarity calculated from Bray-Curtis dissimilarity for abundance
+data and the Jaccard index for presence/absance data. We tested the significance
+of these components using a linear score test by comparing nested models with and
+without the microbial community variance component \citep{qu2013}. We then estimated
+the community heritability of this function by calculating the degree of
+variation in the rate of ecosystem function due to variation in community
+composition between ecosystems including standard errors on that estimate 
+using the \texttt{h2G} function from the
+\texttt{gap} package. We then tested whether microbial community composition explains a
+significant portion of the variation in ecosystem function after accounting for
+variation in the environment and accounting for the spatial distance between
+samples. We did this by fitting an intercept-only model with variance components
+for environmental similarity from euclidean distance of a matrix of
+environmental covariates and spatial similarity
+also using euclidean distance from the latitude and longitude coordinates where
+the samples were collected. We then used a linear score test to compare
+nested models with and without the microbial community variance component to
+determine whether community composition adds any variance explained after
+accounting for spatial and environmental covariates. We then estimated the
+variance explained by each component (environment, geography, and community
+composition) with standard errors of the estimates using the \texttt{h2GE}
+function.
+
+After determining which variance components were significant, we tested the
+effect of each ASV on the rate of ecosystem function. We fit a variance
+component model for each ASV with ASV abundance as the fixed effect and random
+effects for each of the variance components. This tests the relationship between
+taxon abundance and the rate of ecosystem function after controlling for
+population stratification and data structure. Significant ASVs were identified
+by controlling the false discovery rate
+
+\section{Results}
+
+\begin{itemize}
+  \item For low-affinity and high-affinity methane oxidation in Gabon and
+    methane production in Brazil, community composition explains a significant
+    portion of the variation in ecosystem function. (linear score test or
+    likelihood ratio test)
+  \item On its own, the portion of `community heritability' varies by function.
+    \textasciitilde99\% (SE = 20\%) for high affinity, \textasciitilde61\% (SE =
+    39\%)for low affinity, etc. 
+  \item Geographic proximity: high affinity 0.85 (0.10), low affinity n.s. 0.34
+    (0.31)
+  \item Environmental similarity: high affinity 0.61 (0.23), low affinity 0.61 (0.36) 
+  \item Some of these become n.s. when combined suggesting they explain an
+    overlapping portion of variation. Perhaps this means we should fit the
+    variance component model with just the community similarity matrix knowing
+    that that will capture the important components of historical effects and
+    selection.
+  \item Function varies strongly by ecosystem type: in Brazil, pasture has
+    higher methane fluxes than primary or secondary forest; in Gabon, peatlands
+    and wetlands have higher low affinity while plantations and forests have
+    higher high affinity.
+  \item Possible next step: fit a variance component model with community
+    similarity as the random effect and taxon abundance as the fixed effect and
+    collect p.values for each taxon (though this may return none). Or include
+    all three variance components. Adjust for false discovery rate to get
+    significant taxa.
+\end{itemize}
+
+\section{Discussion}
+
+In this paper, we argue that the traditional approach to microbial BEF research
+ignores important understandings about the nature of ecological processes in
+determining the distribution of microorganisms. The marker gene hypothesis has
+proven unsuccesful in predicting the rates of ecosystem functions from the
+abundance of functional genes and in our example study was not significant in one case and
+was significant in another, though explained a small amount of the variation.
+Similarly, diversity metrics from the entire Prokrayotic community were not
+explanatory. These results agree with recent reviews of this literature and
+demonstrate that these approaches are not a fruitful avenue for elucidating
+microbial structure-function connections.
+
+Instead, microbial ecologists should treat microbial communities as analogous to
+organisms with the metagenome representing the community genome. In this case,
+ecosystem function represents a phenotype at the community level. As in
+organismal biology, this ecosystem phenotype will be a mix of genotypic effects and
+environmental effects. Using this conceptual framework, population genetic
+approaches can be applied to microbial community datasets to identify
+
+Treating ecosystems like organisms. Calculating community heritability. Possible
+approaches: variance component modeling, experimental ecosystem
+selection/experimental evolution, CAFE and the Price equation. Need to collect
+data in such a way to optimally use these techniques that already exist.
+
+\section{Conclusions}
+Microbial community ecosystem function research has demonstrated that
+functional gene abundance is not a good predictor of ecosystem function and
+that diversity metrics do not add much to ecosystem models. Therefore, we
+argue that microbial ecologists should look more agnostically at microbial
+community structure to identify whether microbial community composition is
+important for determing the rate of ecosystem function across space, time,
+and ecological conditions. In this way we can identify which taxa or aspects of
+microbial community structure determine the rates of ecosystem functions and so
+will be useful for predicting and managing those rates.
+
+The problem of
+connecting a complex microbial community matrix with an ecosystem-scale function
+is not unique to microbial ecology. Population geneticists attempting to like
+organismal phenotypes to its genotypic and environmental basis have dealt with
+this problem for decades. In order to make robust associations between the
+presence or abundance of specific microbial taxa and the rate of ecosystem-scale
+functions, we should take lessons from population genetics in identifying
+correlations between community structure and ecosystem function. We suggest
+three approaches to addressing this question:
+
+\begin{enumerate}
+  \item Metagenome-wide association studies/Community-wide association
+    studies/Ecosystem-wide association studies to survey natural ecosystems
+    for potential biomarkers or drivers of ecosystem function. Appropriately
+    control for community structure patterns to make robust
+    structure-function associations. Combine this
+    with experimental manipulations such as reciprocal transplants or
+    environmental treatments to identify community-environment (CxE)
+    interactions.
+  \item Use artificial ecosystem selection to enrich for groups of taxa that
+    respond to selection on a function. Use this to identify organisms that
+    might play a critical role in ecosystem function. Use a response to
+    selection to calculate ``community heritability''or the prortion of
+    variation in function due to community composition.
+  \item Measure rates of ecosystem functions on the individual cell or
+    individual taxon level in order to partition changes in ecosystem
+    function to gains, losses, or changes in abundance of
+    specific taxa using the Price equation. Accomplish this by combining
+    isotope source partitioning with functional group assessments, using
+    labeling approaches like stable-isotope probing, ...
+\end{enumerate}
+
+\section*{Old Text}
 

### 2019-07-16 manuscript/main.tex
Update manuscript text

@@
 
-\section*{Tasks} 
-\begin{enumerate}
-\item What is novel about this? The first anyone has treated an ecosystem like an
-organism. The first time anyone has done this correlative approach well
-by considering environmental and spatial structure in the data. 
-\item Interpret model output
-\item Decide whether to include Gabon data, Brazil data, or both
-\end{enumerate}
-
 \section*{Abstract}
@@
 \citep{hooper2005}. More recently, there has been an interest in linking
-microbial community composition to ecosystem functions such as host disease 
+microbial community structure to the functioning of ecosystems both inside and
+outside of a host, such as disease states, plant productivity,
 and soil greenhouse gas emissions \citep{schimel1998, singh2010}. 
@@
 whether microbial community
-composition alters ecosystem-scale functions independent
+composition alters ecosystem-scale function independent
 of the underlying environmental variation.
@@
 If local selection always optimizes the available biodiversity to maximize
-ecosystem function, them microbial community composition should not matter for
-pedicting the rates of microbially-mediated functions. Instead, 
+ecosystem function, then microbial community composition should not matter for
+predicting the rates of microbially-mediated functions. Instead, 
 the rates of these functions should be determined by the
@@
 have shown different rates of ecosystem functions for different microbial
-communities. This has been observed for plant decomposition
+communities under the same conditions \citep{reed2007}. 
+This has been observed for plant decomposition
 \citep{strickland2009}, plant phenology \citep{panke-buisse2015}, soil nitrogen
@@
 seems to generate variation in ecosystem function. However, in order to
-incorporate microorganisms into ecosystem models, we need to identify what
-aspects of microbial biodiversity are associated with variation in ecosystem
-function.
+incorporate microorganisms into ecosystem models, 
+we should distinguish between the portion
+of microbial biodiversity that is contributing to ecosystem function as a
+consequence of the environmental conditions and that contributing uniquely due to
+variation in the community independent of the environment.
 
 Most studies of microbial community function focus on one of two aspects of
-microbial community structure that are hypothesized to be predictive of
+microbial community structure that are hypothesized to predict
 ecosystem function. The first aspect is functional gene or transcript abundance.
-In this case, qPCR or shotgun metagonemic sequencing is used to target a gene
+In this case, qPCR or shotgun metagonemic sequencing is used to estimate the
+abundance of a gene
 or transcript that is a putative marker for a microbial process. For example,
 the gene \textit{mcrA}, which encodes the enzyme that performs the final step in
-methanogenesis, is hypothesized to be correlated with the rate of
+methanogenesis, is hypothesized to predict the rate of
 methane production. Other examples include \textit{pmoA} and methanotrophy,
@@
 ecosystem function is taxonomic or functional
-diversity. Diversity is either estimated using a measure of community
-composition such as 16S rRNA gene ribotyping or manipulated through some proxy
+diversity. Diversity is either estimated from sequence variants of a barcode
+gene such as the 16S rRNA gene or manipulated through some proxy
 of diversity such as sequential dilution or varying filter sizes. Studies that
@@
 addition, assembled microbial communities pose the same challenges as
-macroorganismal diversity experiments, for example no replication at the highest
-diversity treatment. A review of microbial diversity studies shows that
+macroorganismal diversity experiments, for example lacking replication at the highest
+diversity treatment such that that treatment will always have the most
+productive taxon. A review of microbial diversity studies shows that
 microbial community taxonomic and functional diversity add little variance
@@
 that group (e.g. productivity). However, in the case of microbes, our knowledge
-of functional groups is quite limited. We cannot be sure that we're accurately
+of functional groups is quite limited. We cannot be sure that we are accurately
 estimating the abundance or diversity of soil microbial functional groups the
-way we can plants in a grassland.
+way we can plants in a grassland. In addition, we do not know the function of
+most microbial taxa and metagenomic sequences recovered from environmental 
+samples. Because of this, we should look more
+agnostically at the microbial community to identify taxa or groups of taxa that
+are important for predicting the rate of ecosystem function. We can accomplish
+this by drawing on ideas from population genetics.
+
+The challenge of mapping microbial community structure to ecosystem function is
+analogous to the challenge of identifying the genetic basis of complex traits in
+organismal populations. In the latter case, a population exhibits variation in a
+phenotype (e.g. height or disease state) as well as variation in potentially
+thousands of single nucleotide polymorphisms
+(SNPs). To identify the genetic basis for a trait, investigators sample from
+this population and correlate phenotype with genotype. In most cases, there are
+many more loci than individuals and we do not know whether the SNPs are causally
+linked or are simply in linkage disequilibrium with a causal mutation. 
+Genetic loci identified in these genome-wide association
+studies can be the target for future research to dissect the function of those
+genes, can be added or removed to test their association with the trait, or can
+be used as markers for that disease or phenotype even if it has no known
+mechanistic association. This mirrors the problem of microbial
+structure-function mapping. In this case, a population of ecosystems exhibit
+variation in a particular trait or ecosystem function putatively mediated by
+microorganisms as well as variation in metagenomic composition. Investigators
+could sample from this population of ecosystems and correlate the rate of an
+ecosystem function with microbial community composition or gene composition. 
+In order to do this, microbial ecologists should incorporate similar
+methodological and statistical approaches as population geneticists in order to
+make robust associations between microbes and ecosystem function.
+
+Population geneticists trying to make robust associations between genetic loci
+and phenotype must consider several points when designing a genome-wide
+association study. First is
+to identify the precise trait expected to be influenced by genetic variation. 
+Then, to choose a population that exhibits
+substantial variation in the trait of interest. This population should also be
+otherwise relatively homogeneous to avoid other factors that might influence the
+trait. After sampling, statistical methods should take into account population
+stratification due to shared ancestry or local selection that can lead to spurious
+associations between genes and traits. For example, both human height and
+genotype vary systematically across Europe in a latitudinal gradient, which can
+lead to spurious genotype-phenotype associations \citep{novembre2008}.
+
+Positive associations between genes and
+phenotypes can arise because of true associations, through linkage
+disequilibrium with true genes, or through admixture \citep{lander1994}.
+Microbial structure-function connections could pose these same problems. There
+could be true associations between a microorganism and the ecosystem function
+performed by that taxon, or there could be systematic
+co-occurrence patterns between one taxon and a causal taxon,
+or there could be admixture due to shared history or selection within
+a subset of ecosystems. This is true for 16S amplicons and for
+genes in metagenomic (or other -omic) datsets. 
+Robust structure-function associations rely on
+controlling for the underlying geographic and ecological covariance.
 
@@
 colonize a site, then there could be important functional consequences of
-community differentiation to geographic or ecological isolation.
+community differentiation due to geographic or ecological isolation.
 
@@
 
-The challenge of mapping microbial community structure to ecosystem function is
-analogous to the challenge of identifying the genetic basis of complex traits in
-organismal populations. In the latter case, a population exhibits variation in a
-phenotype (e.g. height or disease state) as well as variation in potentially
-thousands of genetic loci (usually identified as single nucleotide polymorphisms
-(SNPs)). To identify the genetic basis for a trait, investigators sample from
-this population and correlate phenotype with genotype. In most cases, there are
-many more loci than individuals and we do not know the function of genes
-associated with these SNPs. Genetic markers identified in these genome-wide association
-studies can be the target for future research to dissect the function of those
-genes, can be added or removed to test their association with the trait, or can
-be used as marker for that disease or phenotype even if it has no known
-mechanistic association. This mirrors the problem of microbial
-structure-function mapping. In this case, a population of hosts or ecosystems exhibit
-variation in a particular trait or ecosystem function putatively mediated by
-microorganisms as well as variation in metagenomic composition. Investigators
-could sample from this population of ecosystems and correlate the rate of an
-ecosystem function with microbial community composition or gene composition.
+Successfully connecting microbial community structure to ecosystem function
+depends on being precise about the variation we want to explain. The first and
+often overlooked step is choosing a
+clearly defined function that exhibits considerable variation across ecosystems. 
+Next, be intentional about choosing the
+population to be sampled to maximize the amount of variation. For example, how
+might ecosystems that are otherwise quite similar vary in function?
+Alternatively, how might one ecosystem change in response to a single
+environmental change? This could be achieved by sampling across similar ecosystems 
+or by imposing experimental conditions that minimize environmental variation or
+vary only abiotic factors of interest. 
+It is important to take an agnostic approach towards identifying taxa that are
+indicators of ecosystem function. It is also important to remember that any
+relationships identified in this manner are provisional until further tests are
+performed. The work is not done until the organism has been manipulated through
+addition, knockout, or variation in abundance.
 
-Population geneticists trying to make robust associations between genetic loci
-and phenotype must take into account population stratification. Population
-stratification due to shared ancestry or local selection can lead to spurious
-associations between genes and traits. For example, human height varies
-across Europe in a latitudinal gradient. Positive associations between genes and
-phenotypes can arise because of true associations, through linkage
-disequilibrium with true genes, or through admixture \citep{lander1994}.
-Microbial structure-function connections could pose these same problems. There
-could be true associations between a microorganism and the ecosystem function
-performed by that taxon, or there could be systematic
-co-occurrence patterns between one taxon and a causal taxon like metabolic
-networks, or there could be admixture due to shared history or selection within
-a subset of ecosystems. This is true for ASVs in ribotyping 16S datasets and for
-genes in metagenomic (or other -omic) datsets. 
-Robust structure-function associations rely on
-controlling for these factors.
+In this paper we propose integrating microbial biodiversity-ecosystem function
+research with ideas and techniques from population genetics to advance our
+understanding of the role of microbial commuities in the functioning of
+ecosystems. Perhaps the most important part of that discussion is to define the
+question carefully. In microbial ecology
+structure-function research is focused on answering the question: what is the
+unique contribution of the microbial community to ecosystem function? This is of
+great interest for two reasons. First, because it addresses what is the singular
+contribution of biodiversity to the functioning of ecosystems. Second, on a
+practical level, because answering that question will be essential to
+incorporate a microbial component into ecosystem models.
 
+Is variation in microbial community
+structure associated with differences in ecosystem function across space,
+time, and environments?
 \section{Methods}
@@
 taxon abundance and the rate of ecosystem function after controlling for
-population stratification and data structure. Significant ASVs were identified
-by controlling the false discovery rate
+population stratification and data structure. The community similarity matrix
+was reconstructed for each model to exclude the predictor ASV. 
+Significant ASVs were identified by controlling the false discovery rate using
+the \texttt{qvalue} package.
 
 \section{Results}
+
+To demonstrate the ideas presented in the Introduction, we re-analyzed data from
+a forthcoming publication \citep{meyer2019}. With these data, we demonstrate the
+importance of selecting the appropriate population to sample from. We are
+precise about the variation we are measuring. Specifically, we measured the rate
+of a specific microbial pathway. In order to minimize environmenal variation, we
+brought intact soil cores into the lab to incubate them under similar
+temperature and methane concentration. We also statistically controlled for
+environmental covariates that we couldn't isolate in the lab including soil
+carbon and nitrogen, soil bulk density, and soil moisture content. Finally, we
+surveyed the microbial community agnostically for microbial markers of ecosystem
+function.
+
+We found that for two pathways of methane oxidation in
 
@@
 abundance of functional genes and in our example study was not significant in one case and
-was significant in another, though explained a small amount of the variation.
+was significant in another, though explained only a small portion of the variation.
 Similarly, diversity metrics from the entire Prokrayotic community were not
@@
 demonstrate that these approaches are not a fruitful avenue for elucidating
-microbial structure-function connections.
+microbial structure-function connections. After identifying significant taxa,
+it is notable that none of the significant taxa fall into expected functional
+groups. For example, none of the taxa identified in this analysis were
+methanotrophs.
 
@@
 environmental effects. Using this conceptual framework, population genetic
-approaches can be applied to microbial community datasets to identify
+approaches can be applied to microbial community datasets to identify  links
+between the composition of microbial communities and the rates of ecosystem
+functions.
 
-Treating ecosystems like organisms. Calculating community heritability. Possible
-approaches: variance component modeling, experimental ecosystem
-selection/experimental evolution, CAFE and the Price equation. Need to collect
-data in such a way to optimally use these techniques that already exist.
+While this analysis focuses on using genome-wide association studies as a model
+for mapping microbial community structure to ecosystem-function, other designs
+are possible. For example, artificial ecosystem selection is a viable approach
+to enrich for communities with high or low rates of ecosystem function
+\citep{swenson2000, panke-buisse2015}. While community-wide association or
+ecosystem-wide association or metagenome-wide association or microbiome-wide
+association studies have sufficient power to identify taxa that are individually
+associated with ecosystem function, whole-community selection could identify
+assemblages of organisms with high rates of function that often co-occur in
+productive ecosystems.
+
+Another possibility that has not to our knowledge been demonstrated with any
+microbial datasets is to apply the ecological form of the Price equation to
+estimate the effects of microbial biodiversity on ecosystem function
+\citep{fox2006, bannar-martin2018}. Using this equation, a change in ecosystem function between two, matched
+observations can be attributed to species additions, species losses, and
+changes in abundance of shared species \citep{fox2006,
+bannar-martin2018}. To
+calculate these vectors, function must be attributed at the species level (e.g.
+biomass measurements of individual plant species). To apply this method to
+microbial communities we will need to measure ecosystem function on a
+per-species basis. There are several ways to do that including quantitative stable-isotope
+probing to measure the accumulation of a substrate in the DNA of a portion of
+the microbial community \citep{hungate2015}, NanoSIMS to measure the incorporation of nutrients at
+the single-cell level \citep{mayali2012}, and isotope source partitioning coupled with measurements
+of microbial pathways through quantification of marker genes through qPCR or
+metagenomics/metatranscriptomics.
+
+This is novel, because it represent the first time that anyone has treated an
+ecosystem like an organism. Further, it is the first time that anyone has
+used this GWAS-esque approach while accounting for both environmental and
+spatial structure in the data.
 
@@
 
-\section*{Old Text}
-
-\subsection{Biodiversity increases ecosystem function in macrobial communities}
-
-Local species richness is correlated with ecosystem function
-\citep{hooper2005}. Most BEF research has focused on terrestrial plant 
-communities where species losses lead to decreases in community abundance or biomass
-\citep{naeem2003}. There are
-two major explanations for this phenomenon. The first is the sampling effect
-where a more diverse community is more likely to have one particularly
-productive species \citep{cardinale2006}. The second is niche complementarity where more species are
-more likely to occupy separate niches within the same ecosystem resulting in
-more efficient resource use and greater biomass.
-
-Regional species diversity may also be important for
-local ecosystem functioning when metacommunity dynamics regulate the
-diversity and composition of communities \citep{leibold2017}. 
-Typical BEF research focuses on local selection
-as a driver of ecosystem function. However, if regional
-diversity is important, then dispersal processes could alter local species diversity and composition resulting in
-a change in ecosystem function. This model is referred to as Community Assembly
-and the Functioning of Ecosystems (CAFE). Community assembly can alter ecosystem
-function if limiting dispersal reduces the likelihood of a highly productive
-species from dispersing to a local site or if high dispersal of an unproductive
-species overwhelms local selection.
-
-The CAFE approach has been demonstrated using the Price equation where a change in ecosystem function between two, matched
-observations can be attributed to species additions, species losses, and
-changes in abundance of shared species \citep{fox2006,
-bannar-martin2018}. To
-calculate these vectors, function must be attributed at the species level (e.g.
-biomass measurements of individual plant species). The CAFE framework demonstrates
-that not only local richness, 
-but also the composition of communities is important to the functioning of ecosystems.
-
-\subsection{Current approaches to microbial BEF reveal little about 
-structure-function mapping}
-
-The trajectory of microbial BEF research has differed from that of macrobial BEF
-for three reasons. First, microbial communities are
-immensely complex with orders of magnitude more taxa than macrobial communities
-and great intra-taxon variation for certain traits. Second, the way we measure
-microbes using marker gene sequences does not allow us to attribute portions of
-a function of interest to individual taxonomic groups . We can only
-measure function in aggregate across an entire community. Finally, we cannot be
-sure that the taxonomic groups that putatively perform a function are the ones
-contributing in this moment.
-
-As a result of these limitations, microbial ecologists have taken a different
-approach to assessing BEF relationships for microbial communities than others
-have for macrobial communities. This is generally a correlative approach where
-certain community attributes that can be derived from a community matrix are
-tested in terms of their relationship with function. These approaches make
-different assumptions about how microbial communities should map onto function.
-On one end of the spectrum, it is hypothesized that functional gene abundance - genes that
-code for an enzyme that performs one step of a process - should be correlated
-with the rate of the process associated with that gene. In this case, qPCR
-is used to quantify the abundance of a functional gene to test whether it is
-correlated with the rate of a process. This approach has had varied success and even the abundance of a
-functional gene that putatively performs a process is rarely correlated with
-that process \citep{rocca2015}. On the other end of the
-spectrum, investigators make few assumptions about which members of the community should
-alter function and so compute
-taxonomic richness or other diversity metrics from the entire Prokaryotic
-community inferred from 16S rRNA gene sequences. Taxonomic or functional
-diversity of the Prokaryote community generally adds little variance explained
-to models of ecosystem function \citep{graham2016}.
-
-\subsection{Biology through analogy: structure-function mapping is similar
-to genotype-phenotype mapping}
-
-The challenges described above for mapping microbial community structure to
-ecosystem function are analogous to the problem of identifying the
-genetic basis of complex traits in organismal populations. In that case,
-each individual organism has a level for a phenotype (e.g. presence
-or absence of a disease state) as well as a vector of potentially
-thousands of genetic loci with different alleles or single nucleotide
-polymorphisms (SNPs). To identify the genetic basis for this trait, we sample
-a population of organisms and try to correlate a matrix of SNPs with a
-vector of phenotypes. In most cases, there are many more loci than
-individuals, we cannot directly attribute a phenotypic state to a particular
-SNP, and we do not know the function of genes associated with these
-SNPs. This mirrors the problem of microbial structure-function mapping
-and presents a potential set of analyses that could be applied in
-microbial ecology.
-
-In addition to the nature of the data, population geneticists have other
-considerations when trying to make robust associations between phenotype
-and SNPs. For example, loci within the genome that are close together tend
-not to assort independently due to linkage disequilibrium. In addition,
-organisms exhibit population stratification due to a variety of historical
-and selective effects. For example, shared ancestry among European
-populations of humans results in an association between genotype and
-geography across Europe \citep{novembre2008}. This population structure
-can lead to false positives when trying to correlate human genotypes
-with genetic diseases or other phenotypes such as height (citation).
-
-In general, organisms exhibit population differentiation due to spatial
-and environmental patterns \citep{wright1943}. For geography, local genetic
-drift and reduced gene flow due to geographic separation can lead to population
-differentiation \citep{wright1943}. In addition, environmental distance can lead to both selection
-against migrants \citep{hendry2004} and local selection on endemic populations to
-generate variation. While
-these processes have primarily been described for macrobial populations, the
-same principles apply to microorganisms and should be considered
-in microbial structure-function mapping \citep{martiny2006}.
-
-\subsection{Reframing the question: Is variation in microbial community
-structure associated with differences in ecosystem function across space,
-time, and environments?}
-
-Typical microbial ecosystem function research focuses on whether microbial
-community composition is important at the ecosystem level
-\citep{schimel1998, cavigelli2000, balser2005, graham2016}. Given that
-microbial functional groups typically mediate biogeochemical processes
-and so the loss of a functional group will certainly alter function,
-the answer is "yes". However, this question does not address whether
-microbial community variation is actually useful for predicting or 
-managing ecosystem functions. To address that problem, we need to know
-whether variation within a functional group matters to variation in
-ecosystem function. If local selection always optimizes the available
-biodiversity to maximize process rates given the underlying environmental
-variation, then composition should not matter, because we should be able
-to predict ecosystem function from abiotic factors alone. Environmental
-treatments, common gardens, and reciprocal transplants have shown
-different functional potential for different community compositions.
-In addition, spatial patterns for microbial communities suggest that
-there should be observable diversification due to environmental and
-spatial processes \citep{wright1943, martiny2006}. Therefore, variation
-in microbial community composition seems to result in variation in 
-ecosystem function. However, in order for microbial community data to be
-useful for ecosystem modeling, the challenge is to identify specifically what variation in community composition
-is associated with function across space, time, and environmental gradients.
-To accomplish that, we need to separate the close association between environmental variation, community
-composition, and ecosystem function to identify the aspects of microbial
-biodiversity that are important for ecosystem function.
-
-\subsection*{Extra text for 1.3/1.4}
-
-\begin{enumerate}
-    \item Things exhibit population differentiation due to spatial and environmental 
-    patterns \citep{wright1943}
-    \item Microbes exhibit similar patterns \citep{martiny2006}
-    \item Pop gen tries to identify regions of the genome correlated with
-    traits
-    \item Considering ecosystems as individuals within a population presents
-    an analogous problem of determining community markers of ecosystem traits (functions)
-    \item Thus the goal of microbial BEF is to identify generalizable patterns
-    of genetic variation in microbes to feed into hypothesis testing about
-    potential mechanisms or to incorporate into ecosystem models
-\end{enumerate}
-
-Typically, microbial community structure-function research questions are framed
-as does microbial composition matter to ecosystem function? The answer to this
-question is obviously "yes" given that microorganisms mediate many of the
-processes of interest. The question then becomes does microbial community
-variation within a functional group or overall matter to ecosystem function? 
-Generally it is assumed that microbial composition is functionally irrelevant, 
-though studies of environmental treatments, common gardens, and reciprocal
-transplants have shown that different community compositions are correlated with
-different functions. The challenge with these studies is separating the close
-connection between environmental variation, community composition, and ecosystem
-function. Microorganisms cannot be useful to ecosystem modeling if there
-is no identifiable variation in composition that consistently alters function
-across space, time, and environmental gradients. The question "do we need to
-include microorganisms in ecosystem models?" depends on identifying variation in
-the community associated with function independent of the environment. If that's
-true, then the next step in understanding or modeling these processes is 
-identify what components of the community alter ecosystem function.
-
-There are a number of ways that environmental or spatial covariates could alter
-or obscure connections between microbial community structure and ecoystem
-function. For example, microbial communities exhibit strong biogeographic
-patterns \citep{martiny2006}. The historical effects of geographic proximity,
-such as dispersal limitation, could lead to spurious associations between taxa
-and function if taxa are unevenly distributed across the functional landscape.
-Environmental selection can also shape microbial community structure, so
-selection on the microbial community that is independent of the
-process could lead to erroneous conclusions about the
-underlying drivers of ecosystem function \citep{thompson2017}. For
-example, pH is known to regulate both microbial biodiversity patterns
-\citep{fierer2006} and methane cycling dynamics \citep{ye2012}, which could lead
-to microbial taxa correlated with methane cycling, but mechanistically
-unrelated. Environmental factors that directly influence function, such as
-moisture, could alter function with or without affecting the community. 
-Finally, community stratification where taxa co-occur in certain ecosystems due to
-selection or historical effects for reasons unrelated to the ecosystem function
-could lead to spurious associations.
-
-\subsection{A new framework for structure-function mapping in microbial
-communities}
-
-Microbial ecosystem function could be assessed using statistical techniques
-borrowed from quantitative genetics along with the conceptual framework for
-community assembly processes provided by metacommunity theory and CAFE. This
-would be a two-step process to first identify the portion of variation in
-ecosystem function attributable to different aspects of an ecosystem: the
-community, the underlying environmental variation, and to spatial processes.
-
-\begin{equation}
-V_F=V_E+V_S+V_C+\epsilon
-\end{equation}
-
-Ecosystem function is represented by $V_F$, $V_E$ is variation in the
-environment and covariation between the community and the environment, $V_S$ is
-variation in spatial distance between samples and covariation between the
-community and space, and $V_C$ is variation in the community independent of
-local selection and historical effects. The parameter $V_C$ also incorporates
-unmeasured environmental and spatial effects.
-
-Each of these variance components also correspond to alternative hypotheses for
-what drives ecosystem function. If most of the variation in $V_F$ is
-attributable to $V_E$, then the environment drives function either directly
-(abiotic processes) or indirectly by selecting for a particular microbial
-community that maximizes function. In this model, the community is simply a 
-conduit through which the environment acts on function. Alternatively, if $V_S$
-explains most of the variation in $V_F$ then this suggests that function is
-primarily a result of historical contingency and differences in dispersal traits.
-Finally, $V_C$ could explain most of the variation in $V_F$. This would be
-consistent with reciprocal transplant and common garden experiments showing that
-community composition drivers ecosystem function independent of the environment.
-
-This framework allows us to address the following questions: 1. Does variation
-in microbial community structure explain a measurable portion of the variation in
-ecosystem function or can all of that variation be attributed to the underlying
-effect of the environmental and spatial processes? If so, 2. Which members of
-the community are significantly correlated with function after controlling for
-the covariances of space and environment? Finally, 3. Do these taxa match our hypotheses about
-what controls ecosystem function (e.g. functional group abundance) or are there
-alternative mechanisms that need to be explored?
-
-\section{Methods}
-
-Sample 16S rRNA gene sequences, environmental covariates, geographic distance,
-ecosystem function. Use the DADA2 pipeline to assign ASVs. 
-phylogenetic and taxonomic assignment Null
-Models \citep{chase2011, stegen2015}
-
-Would it be valuable to use a null model (as in Chase or Stegen) where we
-generate random communities based on the distribution across samples and
-abundance within samples of each taxon. Then try to correlate those abundances
-with our observed rates of ecosystem function to see what distribution of
-structure-function correlations to expect given our community data. These sort
-of null community generating techniques can omit certain community assembly
-processes (such as homogenizing selection or stochastic processes)
-
-An alternative testing approach would be to use Hannah’s permutation test idea
-where we randomly permute the process rates ~1000 times and rebuild the model
-each time to see what the distribution of significant taxa is.
-
-VC - Correlations between taxa and process determined by accounting for spatial
-structure, environmental covariates, and community structure. Represents effect
-of taxonomy, unmeasured environmental variables VE - Account for community
-structure and correlate environmental variables cov(C, E) - The part of the venn
-diagram that disappears when controlling for environment and space VC:VE - Test
-the correlations between community and process rate at a variety of
-environmental conditions.  epsilon/error term - Effects of natural variation,
-unmeasured variables
-
-Another modeling options is HMSC from \cite{ovaskainen2017}
-
-\section{Results}
-
-For high-affinity methane oxidation in Gabon: We experimentally controlled some
-environmental variables and our measured environmental variables are not
-significantly correlated with process rate (C, N, WFPS) Space is correlated with
-function
-
-Be careful with attributing a certain \% of the variance to an individual taxa
-when performing partial regression on the residuals. If a taxa explains 60\% of
-the variance in the residuals after controlling for environment and space that
-doesn’t mean it explains 60\% of the overall variance.
-
-\section{Discussion}
-Combination of qPCR and pathway partinioing, qSIP \citep{hungate2015},
-and NanoSIMS \citep{mayali2012} may allow us to assign portions of
-function to particular taxa or functional groups.
-Problems that need to be addressed: We shouldn’t assume that functional groups
-that putatively perform a process are the rate-limiting step or the strongest
-correlate.  There may be unknown members of the community that perform or limit
-a process The final step of the process may not be the rate-limiting step: it
-may be substrate production or nutrient acquisition or something else (e.g.
-Qusheng Jin’s work on modeling methanogenesis) We need to consider the
-underlying biogeographic patterns shaping the distribution of microbial
-communities (à la GWAS).
-
-\section{Conclusion}
-
-\section*{Outline} 
-\begin{enumerate} 
-  \item Problem statement: Microorganisms putatively mediate biogeochemical
-    cycles and yet we have been unable to demonstrate consistent BEF or CAFE
-  relationships.   
-\item  Opening: Biodiversity ecosystem function research has demonstrated the
-  importance of protecting biodiversity to maintain ecosystem function.
-  However, more recently the importance of community assembly processes such as
-  dispersal have shown that regional diversity is potentially as important as
-  local processes. Therefore, metacommunity dynamics should play a role in local
-  ecosystem functioning.  
-  \item Challenge: Applying CAFE to microbes would be
-  convenient, however microbes pose several unique challenges. 1. Microbial
-  communities are orders of magnitude more complex than macrobial communities.
-  2. We don’t know who performs each function of interest. 3. We cannot
-  attribute a portion of the rate of an ecosystem function to a particular
-  individual or species like we can in plant communities.  
-  \item Solution: The
-  microbial community assembly - ecosystem function relationship is more
-  analogous to the problem of mapping complex phenotypic traits to genetic
-  variation in organisms. In this case we have thousands of genetic loci with
-  unknown functions that may contribute positively or negatively (or not at all)
-  to a particular trait.  
-  \item Studies linking genotype to phenotype have to
-  consider similar processes as metacommunities within the
-  metapopulation such as confounding geography with genotypic similarity.
-  Applying this to microbial metacommunities, associations between microbes and
-  function could be due to a number of factors other than the “genetic” effect
-  of microbial community structure on ecosystem function. By accounting for the
-  covariance structure (space, environment, community interactions) we can
-  identify true structure-function connections.  
-  \item Approach: Demonstrate the
-  previous approaches to microbial structure-function (e.g. abundance-function,
-  diversity-function, PCA or indicator species analysis) and demonstrate our new
-  approach, which accounts for metacommunity dynamics in local ecosystem
-  function. Step 1. Variance partitioning between community, environment, and
-  space. Step 2. GWAS-esque association mapping to identify microbes.  
-  \item Take home: The way we study the functioning of microbial communities is
-  inherently limiting in our ability to apply BEF or CAFE approaches. Instead,
-  we need to learn from GWAS to identify robust structure-function
-  relationships. What we learn from this is that 1. Contrary to past results,
-  microbes do explain a substantial portion of the variation in ecosystem
-  function when we account for other sources of variation. 2. Stochastic
-  processes can alter the relationship between environment/community and
-  function 3.  we cannot assume that we know what part of the community controls
-  ecosystem function (e.g. functional genes) i.e. non-intuitive features can
-  explain a lot of residual variance.  
-  \item Conclusion: This work presents a
-  framework for testing the relationship between microbial community structure
-  and ecosystem function and demonstrates the importance of understanding
-  microbial genetic variation in predicting ecosystem function.  
-  \end{enumerate}
-
-\section*{Possible things to include:}
-Why do we think that communities will be important for understanding ecosystem
-function?
-
-\begin{equation}
-V_F=V_E+V_C+V_EV_C+cov(C,E)+\epsilon
-\end{equation}
-\begin{equation}
-F_{ij}=C_i+E_j+CE_{ij}+\epsilon_{ijk}
-\end{equation}
-
-In this equation, VF represents variation in the rate of an ecosystem function.
-VE is variation in the environment that independently contributes to the rate of
-the ecosystem function. VC is variation in microbial community structure (i.e.
-composition, abundance, or diversity) that independently contributes to
-ecosystem function. The interaction term VE:VC represents situations in which
-the difference in function between two communities varies by environment. For
-example, community A could outperform community B in environment 1, but not in
-environment 2. Alternatively, community A could outperform B in environment 1
-and environment 2, but the magnitude of the difference varies by environment.
-The term covC,E captures covariance between community and environment where a
-change in the environment produces an associated change in the community.
-Finally,  is an error term for natural variation in these relationships.
-
-In the context of deterministic versus stochastic/neutral community assembly,
-each portion of the variance represents different assembly processes. The cov(C,
-E) represents selection whereas VC represents neutral processes of drift,
-homogenizing dispersal, or dispersal-limitation. The interaction term VC:VE
-represents differences in the rate constants/Q10s for different communities or
-neutral processes of community assembly under a particular environmental
-condition. VE is independent of the community and so doesn’t represent a
-community assembly process, but is simply a physicochemical driver of VF such as
-gas diffusion through water, tortuosity of soil pores, redox conditions, … I’m
-not sure this paragraph is correct
-
-VE = Direct physicochemical effects on process rates including abiotic
-reactions, diffusivity, temperature limitation/Q10 (that doesn’t affect taxon
-abundance or activity but does affect e.g. reaction rates), soil stabilization
-of chemicals (sorption), tortuosity of soil pores, wind speed (affecting
-differential pressure between soil air and atmosphere), concentration gradient
-between soil and atmosphere.
-
-cov(C, E) = Environmental selection on soil microorganisms that directly
-influence function. Environmental variables that directly affect function and
-also select for a microbe unrelated to function. An environmental variable that
-both directly affects function and indirectly affects function via a microbe.
-
-VC:VE = Context-dependent organismal traits including different temperature
-optima, substrate affinities, enzyme specificity, redox status/oxygen tolerance,
-pH optima.
-
-Epsilon = Stochastic processes. Taxa occurrences or
-environmental variables completely unrelated to function.
-
-I can imagine several mechanisms by which a significant correlation between a
-taxon and a process rate might be detected.  An environmental variable directly
-affects function and selects for a microbe unrelated to function (false
-positive). (VE) This would be removed by correcting for environment, but would
-remain if it was due to an unmeasured environmental variable (and so would
-appear in VC).  An environmental variable selects for a microbe that directly
-affects function (cov(C, E)). This would be removed by correcting for the right
-environmental variable, but would be revealed by controlling for everything
-except the environment.  A taxon directly affects function (true positive)  and
-co-occurs with another microbe unrelated to function (false positive) (This
-could be VC or cov(C,E)). False positive would be removed by correcting for
-community similarity. True positive would be lost if perfect covariance between
-function, true positive, and false positive.  Function exhibits a spatial signal
-(due to underlying spatial VE or VC) and a taxon exhibits a similar spatial
-signal, but for different reasons (selection on different VE or VC or historical
-effects (e.g. dispersal limitation)). Would be removed by correcting for space,
-but would not be removed if there was a perfect covariance.
-
-Why is it important to understand VC?  At this point in the development of
-microbial ecology, we know that both selection and dispersal limitation shape
-microbial community assembly. At the extremes, a single bacterial phylotype
-cannot instantly be present in all environments. Likewise, oxygen will kill
-obligate anaerobes. It also appears, to the best of our knowledge, that
-microbial communities of different origins exhibit different functional rates
-when introduced to identical environmental conditions. Therefore, environmental
-selection on microbial communities cannot fully explain variation in ecosystem
-functions if they are mediated by microorganisms. Put another way, even if we
-could perfectly measure VE, we cannot fully understand VF.
-
-The question that follows from this is: Does VC matter on the scales relevant to
-humans? To answer that question, we have to understand the magnitude of
-
-Null models of community assembly
-
-VC or VE:VC could also represent unmeasured environmental variables.
-
 \bibliographystyle{ecology} 

### 2019-07-16 R/analysis.R
Playing around with varComp, qvalues, permutations test, etc.

@@ -37,25 +37,120 @@ num.cores <- detectCores()
 registerDoParallel(cores = 3)
 
 #+ r tree_analysis
-tree <- read_tree('../data/gabon/16S_rep_seqs_tree_Gabon.nwk')
+#tree <- read_tree('../data/gabon/16S_rep_seqs_tree_Gabon.nwk')
 #+ r import_data
 
 taxon_table <- data.table::fread('../output/taxon_table.csv', data.table = FALSE)
 all_data <- data.table::fread('../output/gab_all_vst_troph.csv', data.table = FALSE)
+gen_data <- data.table::fread('../output/gab_all_vst_gen.csv', data.table = FALSE)
 n_samp <- nrow(all_data)
 n_asvs <- ncol(all_data[, grepl("^[asv]", colnames(all_data))])
 
+all_data <- all_data[all_data$Wetland == 'Upland', ]
+
 ## Remove ASVs only present in 1 sample
-asvs <- as.matrix(all_data[, grepl("^[asv]", colnames(all_data))])
-vars <- all_data[, !grepl("^[asv]", colnames(all_data))]
+remove_asvs_in_1_sample <- function(dataset) {
+asvs <- as.matrix(dataset[, grepl("^[asv]", colnames(dataset))])
+vars <- dataset[, !grepl("^[asv]", colnames(dataset))]
 asvs <- asvs[, colSums(asvs) > 0]
-all_data <- cbind(vars, asvs)
+cbind(vars, asvs)
+}
+all_data <- remove_asvs_in_1_sample(all_data)
+gen_data <- remove_asvs_in_1_sample(gen_data)
 
 # Add location factor
 all_data[all_data$Y > 30000, 'geocode'] <- 'A'
 all_data[all_data$Y < 30000 & all_data$Y > 10000, 'geocode'] <- 'B'
 all_data[all_data$Y < 10000, 'geocode'] <- 'C'
 all_data$geocode <- factor(all_data$geocode)
+
+pca <- rda(as.matrix(all_data[, grepl("^[asv]", colnames(all_data))]))
+screeplot(pca)
+head(summary(pca))
+all_data$pc1 <- scores(pca, choices = 1, display = 'sites', scaling = 3)
+all_data$pc2 <- scores(pca, choices = 2, display = 'sites', scaling = 3)
[…]

### 2019-07-16 R/analysis_brazil.R

@@ -47,6 +47,8 @@ vars <- all_data[, !grepl("^[asv]", colnames(all_data))]
 asvs <- asvs[, colSums(asvs) > 0]
 all_data <- cbind(vars, asvs)
 
+all_data$log_CH4_c <- with(all_data, log(CH4 + abs(min(CH4)) + 1))
+
 #' # Overview
 #' 
 #' The goal of this project is to answer the question: is microbial community 
@@ -115,7 +117,6 @@ colnames(all_data[, 1:40])
 #ggcorr(all_data[, c(2:3, 8:34, 36, 38)], nbreaks = 4, label = TRUE)
 
 #+ other_correlations
-all_data$log_CH4_c <- with(all_data, log(CH4 + abs(min(CH4)) + 1))
 pca <- rda(as.matrix(all_data[, grepl("^[asv]", colnames(all_data))]))
 screeplot(pca)
 
@@ -299,6 +300,15 @@ geo_sim <- calc_sim_matrix(all_data[, c('Lat', 'Long')])
 # environmental similarity'
 env_sim <- calc_sim_matrix(select(all_data, pH:Total_moisture))
 
+
+#+ screwing_around_with_models
+all_data$region_land_type <- factor(paste(all_data$Region, all_data$Land_type, sep = '_'))
+levels(all_data$region_land_type)
+fit <- varComp(log_CH4_c ~ region_land_type, all_data,
+               ~ cholRoot(com_sim):region_land_type)
+summary(fit)
+h2GE(c(fit$varComps[-c(1, 5)], err = fit$sigma2), vcov(fit, what = 'varComp', drop = T ))
+
 #' ## Model diagnostics for all data
 
 # Diagnostics
@@ -330,7 +340,16 @@ print(ggplot(mapping = aes(c(geo_sim))) +
 print(ggplot(mapping = aes(c(env_sim))) + 
   geom_histogram(binwidth = 0.05) + 
   labs(x = "Environment Similarity (Euclidean)", y = "Number of Samples"))
+model <- varComp(log_CH4_c ~ asv, all_data,
+                 ~ ibs(com_sim))
+summary(model)
 
+h2G(c(model$varComps, err = model$sigma2), vcov(model, what = 'varComp', drop = FALSE))
+model <- varComp(log_CH4_c ~ asv, all_data,
+                 varcov = list(com = com_sim))
+summary(model)
[…]

### 2019-07-16 R/combine_data.R

@@ -1,47 +1,41 @@
 library(tidyverse)
 library(morris)
+library(data.table)
 
 # Methanogen metadata
-gen_attr_table <- read.csv('../output/gab_gen_attr_table.csv', stringsAsFactors = FALSE)
+gen_attr_table <- fread('../output/gab_gen_attr_table.csv', data.table = FALSE)
 gen_attr_table$Site <- paste0(gen_attr_table$Site, 
 				reverse_substr(gen_attr_table$Sample, 3, 3))
 
 # Methanotrophy metadata
-troph_attr_table <- read.csv('../output/gab_troph_attr_table.csv', stringsAsFactors=F)
+troph_attr_table <- fread('../output/gab_troph_attr_table.csv', data.table = FALSE)
 troph_attr_table$Site <- paste0(troph_attr_table$Site, 
 				reverse_substr(troph_attr_table$Sample, 3, 3))
 ## Rarefied ASV Table
 #rare_asv_table <- read.csv('../output/gab_rare_asv_table.csv')
 
 ## DESeq2 variance stabilized asv table
-vst_asv_table <- read.csv('../output/vst_asv_table.csv', stringsAsFactors = FALSE)
+vst_asv_table <- fread('../output/vst_asv_table.csv', data.table = FALSE)
 
-## Raw sequence counts
-pa_asv_table <- vst_asv_table #read_csv('../output/gab_asv_table.csv')
-asvs <- pa_asv_table[, -1] 
-## Presence Absence
-asvs[asvs>0] <- 1
-pa_asv_table[, -1] <- asvs
+### Raw sequence counts
+#pa_asv_table <- vst_asv_table #read_csv('../output/gab_asv_table.csv')
+#asvs <- pa_asv_table[, -1] 
+### Presence Absence
+#asvs[asvs>0] <- 1
+#pa_asv_table[, -1] <- asvs
 
 #left_join(troph_attr_table, troph_asv_table, by = "Sample") %>%
 #  write_csv('output/troph_total.csv')
 geodist <- read.table('../output/geodist.tsv', stringsAsFactors = FALSE)
+#geodist <- read.table('../data/gabon/Geodist_gabon_methane_samples_lat_long.txt', stringsAsFactors = FALSE)
+#colnames(geodist) <- c('Site', 'X', 'Y')
 
 # Export
 troph_total_vst <- left_join(troph_attr_table, geodist, by = "Site")
[…]

### 2019-07-16 R/analysis.R
Clean up scripts and rewrite for new analysis

@@ -474,130 +474,6 @@ qplot(c(com_sim), c(env_sim))
 
 
 
-### Genetics example	
-trt=gl(2, 15)
-set.seed(2340)
-dat=data.frame(trt=trt)
-dat$SNP=matrix(sample(0:2, 120, replace=TRUE), 30)
-dat$Y  = as.numeric(trt)+rnorm(30) + dat$SNP%*%rnorm(4)
-str(dat)
-(vcf0 = varComp(Y~trt, dat, ~ibs(SNP)))
-(vcf00 = varComp(Y~trt, dat, varcov = list(`ibs(SNP)`=IBS(dat$SNP)))) ## same as above
-(vcf1 = varComp(Y~trt, dat, ~ibs(SNP):trt)) ## two variance components
-h2GE(c(vcf1$varComps, err = vcf1$sigma2), vcov(vcf1, what = 'varComp', drop = FALSE))
-range(tcrossprod(ibs(dat$SNP)) - IBS(dat$SNP))
-range(tcrossprod(cholRoot(com_sim)) - com_sim)
-1 - as.matrix(dist(x, method='manhattan') * .5
-     /max(1, ncol(x)) )
-VEG <- 1 - as.matrix(vegdist(asvs) * .5 / max(1, ncol(asvs)))
-
-dat$trt[1]=NA
-varComp(Y~trt, dat, ~ibs(SNP))  ## 29 observations compared to 30 in vcf0
-dat$SNP[2,1]=NA
-varComp(Y~trt, dat, ~ibs(SNP))  ## still, 29 observations, as ibs handles sporadic NA
-dat$SNP[3, ]=NA
-varComp(Y~trt, dat, ~ibs(SNP))  ## 28 observations, as no genotype for the 3rd obs
-dat
-fun_sim <- calc_sim_matrix(all_data$Vmax)
-qplot(c(geo_sim), c(fun_sim)) + stat_smooth(method = 'lm')
-mantel(fun_sim, geo_sim)
-mantel(fun_sim, com_sim)
-mantel(fun_sim, env_sim)
-
-summary(aov(Low_final_k ~ geocode, all_data))
-
-
-all_data$Wetland <- factor(all_data$Wetland)
-all_data$geocode <- factor(all_data$geocode, levels = c("A", "C", "B"))
-ggplot(all_data, aes(x = X, y = Y, color = Low_final_k)) + 
-  geom_jitter()
-
[…]

### 2019-07-16 R/var_comp.R

@@ -14,7 +14,6 @@ library(data.table)
 registerDoParallel(cores = 28)
 
 all_data <- as.data.frame(fread('../output/gab_all_vst_troph.csv'))
-
 # Variance Component Analysis
 
 # sim_mat
@@ -48,42 +47,20 @@ all_data[all_data$Y < 10000, 'geocode'] <- 'C'
 all_data$geocode <- factor(all_data$geocode)
 
 ## Permute response
-permute_no_replace <- function(perm_vec, nperm = 1000) {
-samples <- matrix(ncol = nperm, nrow = length(perm_vec))
-foreach(i = seq_len(nperm), .combine = "cbind") %do% {
-  sample_i <- sample(perm_vec)
-  while (any(apply(samples, 2, function(x) {all(sample_i == x)}), na.rm = T)) {
-    sample_i <- sample(perm_vec)
-  }
-  samples[, i] <- sample_i
-}
-}
-permute_no_replace_par <- function(perm_vec, nperm = 1000) {
-samples <- matrix(ncol = nperm, nrow = length(perm_vec))
-samples <- cbind(samples, truth = perm_vec)
-foreach(i = seq_len(nperm), .combine = "cbind") %dopar% {
-  sample_i <- sample(perm_vec)
-  while (any(apply(samples, 2, function(x) {all(sample_i == x)}), na.rm = T)) {
-    sample_i <- sample(perm_vec)
-  }
-  samples[, i] <- sample_i
-}
-}
-
-permute_no_replace_for <- function(perm_vec, nperm = 1000) {
-samples <- matrix(ncol = nperm, nrow = length(perm_vec))
-samples <- cbind(samples, truth = perm_vec)
-for (i in seq_len(nperm)) {
-  sample_i <- sample(perm_vec)
-  while (any(apply(samples, 2, function(x) {all(sample_i == x)}), na.rm = T)) {
-    sample_i <- sample(perm_vec)
-  }
-  samples[, i] <- sample_i
-  colnames(samples)[i] <- paste0('perm.', i)
-}
-return(samples)
-}
-all_perm_lowk <- permute_no_replace_for(all_data$Low_final_k, nperm = 1000)
+#permute_no_replace_for <- function(perm_vec, nperm = 1000) {
+#samples <- matrix(ncol = nperm, nrow = length(perm_vec))
+#samples <- cbind(samples, truth = perm_vec)
+#for (i in seq_len(nperm)) {
+#  sample_i <- sample(perm_vec)
[…]

### 2019-07-19 R/analysis.R
Change to using all data and create figures and tables for publication

@@ -18,7 +18,7 @@ knitr::opts_chunk$set(fig.width=8, fig.height=6, fig.path='Figs/',
 library(tidyverse)
 #library(lubridate)
 library(broom)
-#library(broman)
+library(broman)
 #library(morris)
 library(varComp)
 library(vegan)
@@ -35,6 +35,7 @@ library(lmtest)
 
 num.cores <- detectCores()
 registerDoParallel(cores = 3)
+options(knitr.kable.NA = '')
 
 #+ r tree_analysis
 #tree <- read_tree('../data/gabon/16S_rep_seqs_tree_Gabon.nwk')
@@ -45,8 +46,6 @@ all_data <- data.table::fread('../output/gab_all_vst_troph.csv', data.table = FA
 gen_data <- data.table::fread('../output/gab_all_vst_gen.csv', data.table = FALSE)
 n_samp <- nrow(all_data)
 n_asvs <- ncol(all_data[, grepl("^[asv]", colnames(all_data))])
-
-all_data <- all_data[all_data$Wetland == 'Upland', ]
 
 ## Remove ASVs only present in 1 sample
 remove_asvs_in_1_sample <- function(dataset) {
@@ -82,23 +81,23 @@ screeplot(geo_pca)
 summary(geo_pca)
 all_data$geo_pc1 <- scores(geo_pca, choices = 1, display = 'sites', scaling = 3)
 
-pca <- rda(as.matrix(gen_data[, grepl("^[asv]", colnames(gen_data))]))
-screeplot(pca)
-head(summary(pca))
-gen_data$pc1 <- scores(pca, choices = 1, display = 'sites', scaling = 3)
-gen_data$pc2 <- scores(pca, choices = 2, display = 'sites', scaling = 3)
-gen_data$pc3 <- scores(pca, choices = 3, display = 'sites', scaling = 3)
-
-env_pca <- rda(as.matrix(gen_data[, c('Bulk_dens', 'Mois_cont', 'pH', 'N_percent', 'C_percent')]))
-screeplot(env_pca)
-head(summary(env_pca))
-gen_data$env_pc1 <- scores(env_pca, choices = 1, display = 'sites', scaling = 3)
-gen_data$env_pc2 <- scores(env_pca, choices = 2, display = 'sites', scaling = 3)
-
-geo_pca <- rda(as.matrix(gen_data[, c('X', 'Y')]))
-screeplot(geo_pca)
[…]

### 2019-07-19 R/var_comp.R

@@ -40,7 +40,7 @@ geo_sim <- calc_sim_matrix(all_data[, c('X', 'Y')])
 # environmental similarity'
 env_sim <- calc_sim_matrix(all_data[, c('Bulk_dens', 'Mois_cont', 'N_percent', 'C_percent')])
 
-# 
+ 
 all_data[all_data$Y > 30000, 'geocode'] <- 'A'
 all_data[all_data$Y < 30000 & all_data$Y > 10000, 'geocode'] <- 'B'
 all_data[all_data$Y < 10000, 'geocode'] <- 'C'
@@ -69,46 +69,52 @@ calc_varcomp_per <- function(model) {
 c(model$varComps, err = model$sigma2)/sum(c(model$varComps, model$sigma2)))))
 }
 
-fit_varComps <- function(x, y, all_data) {
+fit_varComps <- function(x, y, all_data, func = "Low_final_k") {
   #print(paste0(round(x/ncol(asvs)*100, 2), ' % of ASVs and ', round(y/ncol(perm_lowk)*100, 2), ' % of permutations'))
   print(paste0(round(x/ncol(asvs)*100, 2), ' % of ASVs'))
 asv <- asvs[, x]
 asv_tax <- colnames(asvs)[x]
-lowk <- all_data$Low_final_k
-#lowk <- perm_lowk[, y]
+func <- all_data[, func]
+#func <- perm_lowk[, y]
+n_perm <- 1
 #n_perm <- colnames(perm_lowk)[y]
-com_sim <- 1 - as.matrix(vegdist(asvs[, -x]))
+com_sim <- 1 - as.matrix(vegdist(asvs[, -x], method = 'jaccard'))
 parts <- NULL
-model <- varComp(lowk ~ asv ,
+
+geo <- cholRoot(geo_sim)
+com <- cholRoot(com_sim)
+env <- cholRoot(env_sim)
+
+model <- varComp(func ~ asv ,
                  data = all_data,
-                 varcov = list(geo = geocode, com = com_sim, env = env_sim) )
+                 ~ geo + com + env)
 parts <- bind_cols(tibble(asv = asv_tax, perm = n_perm, comps = 'gce'), calc_varcomp_per(model))
 
-model <- varComp(lowk ~ asv, data = all_data,
-                 varcov = list(com = com_sim, env = env_sim) )
+model <- varComp(func ~ asv, data = all_data,
+                 ~ com + env)
[…]

### 2019-07-19 R/geodist_processing.R
update geodist

@@ -1,7 +1,7 @@
 library(sp)
 library(rgdal)
 
-geodist <- read.table('data/Geodist_gabon_methane_samples_lat_long.txt')
+geodist <- read.table('../data/gabon/Geodist_gabon_methane_samples_lat_long.txt')
 
 
 colnames(geodist) <- c('Site', 'X', 'Y')

### 2019-07-19 manuscript/main.tex

@@
 \documentclass{article}
-%\usepackage[utf8]{inputenc}
-%\usepackage{amsmath}
-%\usepackage{amsfonts}
-%\usepackage{amssymb}
-%\usepackage{textalpha}
 \usepackage{natbib}
+\usepackage{subfiles}
+\usepackage{booktabs}
+\usepackage{multirow}
+\usepackage{makecell}
+\usepackage{graphicx}
+\graphicspath{{../Figures/}}
 
@@
 time, and environments?
+
 \section{Methods}
@@
 community composition expains a significant portion of the variation in
-ecosystem function \citep{qu2013}. First, we fit an intercept-only model with one random effect
-for community similarity calculated from Bray-Curtis dissimilarity for abundance
+ecosystem function \citep{qu2013}. First, we fit an intercept-only model with
+each random effect in all combinations. Variance components included deographic location was coded as a factor, environmental
+similarity from euclidean distance of four edaphic variables including total
+nitrogen, total carbon, moisture content, and bulk density, and
+community similarity calculated from Bray-Curtis dissimilarity for abundance
 data and the Jaccard index for presence/absance data. We tested the significance
 of these components using a linear score test by comparing nested models with and
-without the microbial community variance component \citep{qu2013}. We then estimated
-the community heritability of this function by calculating the degree of
-variation in the rate of ecosystem function due to variation in community
-composition between ecosystems including standard errors on that estimate 
+without each variance component \citep{qu2013}. We then estimated
+the percent of variation in ecosystem function explained by each component 
+including standard errors on that estimate 
 using the \texttt{h2G} function from the
@@
 
-To demonstrate the ideas presented in the Introduction, we re-analyzed data from
-a forthcoming publication \citep{meyer2019}. With these data, we demonstrate the
+\includegraphics[width=0.5\textwidth]{lowk_pmoa}
+\includegraphics[width=0.5\textwidth]{lowk_pc1}
+\includegraphics[width=0.5\textwidth]{lowk_rich}
+\includegraphics[width=0.5\textwidth]{lowk_shan}
+
+All three variance components are significant individually. Community
+significant in the community + environment model and geography
+significant in the geography + environment model. Table \ref{tab:pvalues}
+
+\subfile{../tables/var_comp_pvalues}
+
+Including different covariates removes different numbers of
+significant taxa. Some covariates add taxa that were not previously
+identified as significant. Table \ref{tab:n_taxa}
+
+\subfile{../tables/all_taxa_n}
+
+After controlling for geographic proximity, environmental similarity, and
+similarity in community composition, four taxa were significantly correlated
+with the rate of methane oxidation. None of these four taxa are known methane
+oxidizers or methane producers. Table  \ref{tab:sig_taxa}
+
+\subfile{../tables/all_taxa_sig}
+
+\section{Discussion}
+
+In this paper, we argue that the traditional approach to microbial BEF research
+ignores important understandings about the nature of ecological processes in
+determining the distribution of microorganisms. The marker gene hypothesis has
+proven unsuccesful in predicting the rates of ecosystem functions from the
+abundance of functional genes and in our example study was not significant.
+Similarly, diversity metrics from the entire Prokaryotic community were not
+explanatory. These results agree with recent reviews of this literature and
+demonstrate that these approaches are not a fruitful avenue for elucidating
+microbial structure-function connections.
+
+Instead, microbial ecologists should treat microbial communities as analogous to
+organisms with the metagenome representing the community genome. In this case,
+ecosystem function represents a phenotype at the community level. As in
+organismal biology, this ecosystem phenotype will be a mix of genotypic effects and
+environmental effects. Using this conceptual framework, population genetic
+approaches can be applied to microbial community datasets to identify  links
+between the composition of microbial communities and the rates of ecosystem
+functions.
+
+There are two ways to interpret an association between taxon abundance and rate
+of function. One is a statistical co-occurrence and the other is a biological
+interpetation, for example pH affects function and pH affects taxon or pH
+affects taxon, which affects function.
+A significant association between a taxon and the rate of ecosystem function
+could be explained a number of ways. One possibility is that the taxon is
+statistically related, because it is causally connected to the function. This
+could be direct, for example an organism that consumes methane, or indirect, for
+example an organism that regulates substrates necessary for consumers of
+methane. In either of these cases, this would represent a taxon that could be
+useful as a biomarker of function or an organism to investigate to understand
+the mechanistic relationship between communities and function. Alternativelly, a significant
+association could occur due to admixture within a community. In a community of
+microorganisms, any organism that tends to be in high abundance under the same
+conditions as methane oxidizers will be correlated with methane oxidation even
+if it has no mechanistic relationship with the rate of methane flux. In fact,
+any taxa in high abundance in highly methanotrophic communities will be
+statistically related to function.
+
+In this variance component modeling framework, many taxa that are significantly
+associated with function become non-significant after accounting for different
+covariates. This could be explained by different relationships between the
+environment, community assembly processes, and the rate of ecosystem function.
+For example, a taxon significantly correlated with function that becomes
+non-significant after controlling for the underlying environmental variation
+could be acting as a conduit through which the environment alters function.
+Alternatively, if a taxon becomes non-significant after controlling for
+geographic proximity, this taxon could be related to function due to historical
+effects such as past environmental conditions or through dispersal from a nearby
+patch. Alternatively, geography could represent unmeasured environmental
+variables that change over distance. Finally, if incorporating community
+similarity makes a taxon non-significant, then
+
+None of the significant taxa identified by the full variance component model 
+fall into expected functional groups. For example, none of the taxa identified
+in this analysis were methanotrophs or methanogens.
+
+Broad vs narrow processes: Some genetic diseases such as Parkinson's 
+are controlled by a single locus. However, in most GWAS studies the phenotype of
+interest is a result of many loci of small effect (as well as environmental
+factors) such as human height and LDL cholesterol. For ecosystem functions that
+consist of physiologically or phylogenetically "narrow" process (sensu
+\cite{schimel1995b}), it's possible that a single marker gene could control the
+rate of a process. For example, methane flux in permafrost in Sweden may be
+controlled by a single taxon \citep{mccalley2014}. However, most microbial ecosystem functions are likely the result
+of many taxa interacting directly or indirectly with function.
+Trying to identify many taxa of small effect is made easier by constraining that
+variation by defining your population in a way tha makes them similar: 
+similar ecosystem, soil type, and abiotic conditions.
+
+With these data, we demonstrate the
 importance of selecting the appropriate population to sample from. We are
@@
 function.
-
-We found that for two pathways of methane oxidation in
-
-\begin{itemize}
-  \item For low-affinity and high-affinity methane oxidation in Gabon and
-    methane production in Brazil, community composition explains a significant
-    portion of the variation in ecosystem function. (linear score test or
-    likelihood ratio test)
-  \item On its own, the portion of `community heritability' varies by function.
-    \textasciitilde99\% (SE = 20\%) for high affinity, \textasciitilde61\% (SE =
-    39\%)for low affinity, etc. 
-  \item Geographic proximity: high affinity 0.85 (0.10), low affinity n.s. 0.34
-    (0.31)
-  \item Environmental similarity: high affinity 0.61 (0.23), low affinity 0.61 (0.36) 
-  \item Some of these become n.s. when combined suggesting they explain an
-    overlapping portion of variation. Perhaps this means we should fit the
-    variance component model with just the community similarity matrix knowing
-    that that will capture the important components of historical effects and
-    selection.
-  \item Function varies strongly by ecosystem type: in Brazil, pasture has
-    higher methane fluxes than primary or secondary forest; in Gabon, peatlands
-    and wetlands have higher low affinity while plantations and forests have
-    higher high affinity.
-  \item Possible next step: fit a variance component model with community
-    similarity as the random effect and taxon abundance as the fixed effect and
-    collect p.values for each taxon (though this may return none). Or include
-    all three variance components. Adjust for false discovery rate to get
-    significant taxa.
-\end{itemize}
-
-\section{Discussion}
-
-In this paper, we argue that the traditional approach to microbial BEF research
-ignores important understandings about the nature of ecological processes in
-determining the distribution of microorganisms. The marker gene hypothesis has
-proven unsuccesful in predicting the rates of ecosystem functions from the
-abundance of functional genes and in our example study was not significant in one case and
-was significant in another, though explained only a small portion of the variation.
-Similarly, diversity metrics from the entire Prokrayotic community were not
-explanatory. These results agree with recent reviews of this literature and
-demonstrate that these approaches are not a fruitful avenue for elucidating
-microbial structure-function connections. After identifying significant taxa,
-it is notable that none of the significant taxa fall into expected functional
-groups. For example, none of the taxa identified in this analysis were
-methanotrophs.
-
-Instead, microbial ecologists should treat microbial communities as analogous to
-organisms with the metagenome representing the community genome. In this case,
-ecosystem function represents a phenotype at the community level. As in
-organismal biology, this ecosystem phenotype will be a mix of genotypic effects and
-environmental effects. Using this conceptual framework, population genetic
-approaches can be applied to microbial community datasets to identify  links
-between the composition of microbial communities and the rates of ecosystem
-functions.
 
@@
 metagenomics/metatranscriptomics.
+
+\subsection{Caveats}
+
+Future studies could improve on the approach illustrated here in a number of
+ways. First, GWAS studies rely on high power to detect significant associations
+between phenotype and SNPs of small effect. This study consistent of 44 samples,
+which is typical for microbial ecology sudies, but would benefit from a large
+sampling of ecosystems and variation in both community composition and function. 
+Second, this exploratory study sampled across
+a diversity of wetland and upland ecosystems in the Congo basin. However, upland
+ecosystems such as forests are known to have the greatest rates of high-affinity
+methane oxidation. Identifying connections between variation in community
+composition and variation in ecosystem function would benefit from sampling a
+population of ecosystems that is known to exhibit variation in the rate of
+the target function. Third, it is important to define a narrow pathway
+definition. Functions such as low-affinity methane oxidation and bulk methane
+production are relatively broad pathways and as such exhibit no significant
+correlation with individual taxa \citep{meyer2019}. Here we have measured
+high-affinity methane oxidation, which is performed by relatively few taxa.
+Functions such as this are more phylogenetically conserved and as such are more
+likely to have significant connections between community and function. Fourth,
+these environmental variables measured may or may not be the correct ones to
+measure to accurately account for the environmental effect on function.
+Moisture, structure, total nitrogen, and total carbon are relatively course
+measurements of environmental variation. Including more environmental variables
+and variables known to influence the rate of the function of interest would more
+accurately estimate the effect of the environment of communities and function.
+This could also be applied to the geographic location of samples. Our study
+sites were clustered around three locations that are far apart and so didn't
+fully capturing the geographic variation underlying these ecosystems. Sampling
+in a grid or logarithmic pattern might more suitable capture variation across
+the landscape.
 

### 2019-07-20 manuscript/main.tex
Revise Introduction

@@
 
-Population geneticists trying to make robust associations between genetic loci
-and phenotype must consider several points when designing a genome-wide
-association study. First is
-to identify the precise trait expected to be influenced by genetic variation. 
-Then, to choose a population that exhibits
-substantial variation in the trait of interest. This population should also be
-otherwise relatively homogeneous to avoid other factors that might influence the
-trait. After sampling, statistical methods should take into account population
-stratification due to shared ancestry or local selection that can lead to spurious
-associations between genes and traits. For example, both human height and
+Population geneticists make robust associations between genetic loci
+and phenotypes by carefully considering study design when performing a genome-wide
+association study. Studies should consider that positive associations between genes and
+phenotypes can arise because of true associations, through linkage
+disequilibrium with true genes, or through admixture where traits at high
+frequency in a sub-population will be associated with any allele more common in
+that group \citep{lander1994}. A common example of admixture is that human height and
 genotype vary systematically across Europe in a latitudinal gradient, which can
-lead to spurious genotype-phenotype associations \citep{novembre2008}.
+lead to spurious genotype-phenotype associations \citep{novembre2008}. To
+correct for this, it is
+imporant to sample from a population that exhibits variation in the trait of
+interest, but is otherwise relatively homogeneous
+to avoid spurious associations due to admixture. In addition, it is important to standardize phenotyping
+criteria to provide a narrowly defined trait. While certain genetic diseases are
+controlled by only one or a few genetic loci, common diseases tend to be
+controlled by many common variants of small effect. To detect these
+associations requires large sample sizes for association tests. After
+sampling the genotype and phenotype of this population, association tests should
+adjust for population stratification using covariates, which are commonly
+variables like sex and age as well as shared ancestry based on
+genotypic similarity. Finally, after comparing thousands or millions of genetic
+loci, investigators should correct for multiple comparisons.
 
-Positive associations between genes and
-phenotypes can arise because of true associations, through linkage
-disequilibrium with true genes, or through admixture \citep{lander1994}.
-Microbial structure-function connections could pose these same problems. There
-could be true associations between a microorganism and the ecosystem function
-performed by that taxon, or there could be systematic
-co-occurrence patterns between one taxon and a causal taxon,
-or there could be admixture due to shared history or selection within
-a subset of ecosystems. This is true for 16S amplicons and for
-genes in metagenomic (or other -omic) datsets. 
-Robust structure-function associations rely on
-controlling for the underlying geographic and ecological covariance.
+Studies attempting to link microbial community structure to ecosystem function
+could pose similar challenges. In this case, individual taxa are analogous to
+genetic loci and their abundance or presence can be association with the rate of 
+ecosystem function. There
+could be true associations between a microorganism and the ecosystem function if
+that organism is involved in the metabolic pathway producing that function.
+Alternatively, there 
+could be systematic co-occurrence patterns between a taxon and ecosystem
+function due to admixture if that organism tends to reside in ecosystems with
+high rates of ecosystem function.
 
-Organisms exhibit population differentiation due to geographic and ecological
-processes \citep{wright1943}. Neutral processes such as local genetic drift and
-reduced gene flow due to geographic separation can lead to population
-differentiation. In addition, environmental distance can lead to selection
-against migrants \citep{hendry2004} and local selection on endemic populations
-to generate variation. Microorganisms also exhibit biogeographic patterns due to
-ecological selection and geographic separation \citep{martiny2006}. If dispersal
-is non-limiting then the most productive species should always be present and
-composition should not matter for local ecosystem function. However, if important
-functional groups are dispersal-limited, lost to drift, or are unable to
-colonize a site, then there could be important functional consequences of
-community differentiation due to geographic or ecological isolation.
-
-There are a number of ways that environmental or spatial distance could obscure
-microbial BEF relationships. For example, historical effects of geographic
-isolation such as dispersal limitation could lead to spurious associations
-between the abundance of taxa and the rate of ecosystem function. Community
-stratification where taxa co-occur due to historical effects or local selection.
-Environmental selection can also shape microbial community
-structure. For example, pH is known to regulate both broad biodiversity patterns for
-microbial communities \citep{fierer2006} and methane cycling dynamics
-\citep{ye2012}. This could lead to microbial taxa that are correlated with
-methane cycling, but are mechanistically unrelated. In fact, if the rate of
+Microbial admixture could arise through a
+number of ecological mechanisms illustrated in conceptual diagram 1. For
+example, an environmental condition such as pH could contribute to a high rate
+of ecosystem function and select for a taxon even if that taxon is unrelated to
+function. An association could also arise if an organism disperses from a
+low-rate ecosystem to an ecosystem with a high rate of ecosystem function
+through source-sink dynamics. In fact, if the rate of
 ecosystem function is high in one set of ecosystems, function will be
@@
 
-Successfully connecting microbial community structure to ecosystem function
-depends on being precise about the variation we want to explain. The first and
-often overlooked step is choosing a
-clearly defined function that exhibits considerable variation across ecosystems. 
-Next, be intentional about choosing the
-population to be sampled to maximize the amount of variation. For example, how
-might ecosystems that are otherwise quite similar vary in function?
-Alternatively, how might one ecosystem change in response to a single
-environmental change? This could be achieved by sampling across similar ecosystems 
-or by imposing experimental conditions that minimize environmental variation or
-vary only abiotic factors of interest. 
-It is important to take an agnostic approach towards identifying taxa that are
-indicators of ecosystem function. It is also important to remember that any
-relationships identified in this manner are provisional until further tests are
-performed. The work is not done until the organism has been manipulated through
-addition, knockout, or variation in abundance.
+%Microorganisms exhibit biogeographic patterns due to
+%ecological selection and geographic separation \citep{martiny2006}. 
+%Organisms exhibit population differentiation due to geographic and ecological
+%processes \citep{wright1943}. Neutral processes such as local genetic drift and
+%reduced gene flow due to geographic separation can lead to population
+%differentiation. In addition, environmental distance can lead to selection
+%against migrants \citep{hendry2004} and local selection on endemic populations
+%to generate variation.
+
+The same study design decisions used in GWAS could be applied to microbial
+studies of structure-function connections. First, it is important to choose a 
+subset of ecosystems that are relatively homogeneous in their conditions other
+than that function of interest. This could be achieved by sampling across
+similar types of ecosystems or by imposing conditions that minimize
+environmental varition or vary only abiotic factors of interest. Next, it is important to provide a narrow
+definition of function and a consistant measurement approach across studies.
+Also, some functions are putatively controlled by one or a few taxa
+\citep{mccalley2014}, but most functions are likely controlled by many
+interacting microorganisms. To detect these associations depends on a large
+sample size. After collecting samples, the underlying sample structure should be
+controlled for by including environmental and spatial covariates and by
+accounting for population structure. Finally, statistical approaches to adjust
+for multiple comparisons should be applied. This approach could be applied to 16S amplicons or
+genes in metagenomic (or other -omic) datasets.
+
+%clearly defined function that exhibits considerable variation across ecosystems. 
+%Next, be intentional about choosing the
+%population to be sampled to maximize the amount of variation. For example, how
+%might ecosystems that are otherwise quite similar vary in function?
+%Alternatively, how might one ecosystem change in response to a single
+%environmental change?
+%It is important to take an agnostic approach towards identifying taxa that are
+%indicators of ecosystem function. It is also important to remember that any
+%relationships identified in this manner are provisional until further tests are
+%performed. The work is not done until the organism has been manipulated through
+%addition, knockout, or variation in abundance.
 
@@
 research with ideas and techniques from population genetics to advance our
-understanding of the role of microbial commuities in the functioning of
+understanding of the role of microbial communities in the functioning of
 ecosystems. Perhaps the most important part of that discussion is to define the
-question carefully. In microbial ecology
-structure-function research is focused on answering the question: what is the
-unique contribution of the microbial community to ecosystem function? This is of
+question carefully. Microbial structure-function research should focus on answering the question: what is the
+unique contribution of the microbial community to ecosystem function independent
+of the environment? This is of
 great interest for two reasons. First, because it addresses what is the singular
@@
 practical level, because answering that question will be essential to
-incorporate a microbial component into ecosystem models.
-
-Is variation in microbial community
-structure associated with differences in ecosystem function across space,
-time, and environments?
+incorporate a microbial component into ecosystem models. To address that
+question, we demonstrate that one ecosystem function, high-affinity methane
+oxidation, is significantly association with variation in microbial community
+composition. We also show that three components of population stratification
+contribute significantly to variation in ecosystem function. Finally, we
+demonstrate one approach to identifying the component of microbial
+community that contributes to ecosystem function use an association test that
+controls for these three
+covariates and find four microbial taxa that are significantly associated with
+ecosytsem function.
 

### 2019-07-20 R/analysis.R

@@ -856,6 +856,10 @@ all_model_output  %>%
   ungroup()
 
 colnames(sig_taxa_all)
+taxa <- 
+sig_taxa_all %>% 
+  select(comps, Phylum:Genus)
+
 sig_taxa_all %>% 
   filter(comps == 'gce')  %>% 
   select(Phylum:Genus) %>% 

### 2019-07-26 manuscript/main.tex
Update text and figures

@@
 \usepackage{graphicx}
-\graphicspath{{../Figures/}}
+\graphicspath{{../figures/}}
 
@@
 Microbial admixture could arise through a
-number of ecological mechanisms illustrated in conceptual diagram 1. For
+number of ecological mechanisms illustrated in the conceptual diagram in Figure 1. For
 example, an environmental condition such as pH could contribute to a high rate
@@
 demonstrate one approach to identifying the component of microbial
-community that contributes to ecosystem function use an association test that
+community that contributes to ecosystem function using an association test that
 controls for these three
@@
 the rate of ecosystem function as the response with one predictor for each
-model. These predictors include functional gene and transcript abundance,
+model. These predictors include functional gene abundance,
 diversity measures including richness, Shannon's diversity index, and Simpson's
@@
 ecosystem function \citep{qu2013}. First, we fit an intercept-only model with
-each random effect in all combinations. Variance components included deographic location was coded as a factor, environmental
+each random effect in all combinations. Variance components included geographic location was coded as a factor, environmental
 similarity from euclidean distance of four edaphic variables including total
@@
 
+Methane oxidation rate was significantly correlated with the first principal
+component from a PCA of the
+community matrix of ASVs inferred from 16S rRNA genes (Figure
+\ref{fig:corrs}, Table 1). Other typical
+measurements of community structure including functional gene abundance for the
+gene encoding the final step in the methanotrophy pathway, richness, and Shannon
+diversity were not significantly correlated with methane oxidation rate.
+
+\subfile{../tables/trad_cor_table}
+
+\begin{figure}
 \includegraphics[width=0.5\textwidth]{lowk_pmoa}
@@
 \includegraphics[width=0.5\textwidth]{lowk_shan}
+\caption{\label{fig:corrs}
+  \textbf{Methane oxidation rate is correlated with community composition, but not
+functional gene abundance or diversity}.
+  Correlations between community attributes and ecosystem function. A) Abundance
+  of the functional gene \textit{pmoA}, B) community composition based on
+  principal component 1 of a PCA, C) richness, and D) Shannon's diversity index. n = 44 except
+for A where n = 42, because two samples failed to amplify during quantitative
+PCR.}
+\end{figure}
 
-All three variance components are significant individually. Community
-significant in the community + environment model and geography
-significant in the geography + environment model. Table \ref{tab:pvalues}
+Three variance components explained significant portion of the variation in methane
+oxidation when tested individually including community similaity from
+Bray-Curtis distance, environmental similarity, and geographic similarity. When
+combined, most of these terms became not significant with two exceptions. In the
+community + environment model, community was still significant and in the
+geography + environment model, geography was still significant (Table 2).
 
@@
 
-Including different covariates removes different numbers of
-significant taxa. Some covariates add taxa that were not previously
-identified as significant. Table \ref{tab:n_taxa}
+There were 460 taxa that were significantly correlated with methane oxidation
+rate. After including each covariate, fewer taxa were significant. All variance
+components removed some significant taxa and several of the models added taxa 
+that were not previously
+identified as significant. Controlling for community similarity removed all
+significant taxa and controlling for geography and community removed all
+significant taxa. After controlling for all three covariates, only four taxa
+remained significant (Table 3).
 
@@
 with the rate of methane oxidation. None of these four taxa are known methane
-oxidizers or methane producers. Table  \ref{tab:sig_taxa}
+oxidizers or methane producers (needs citations) (Table 4).
 
@@
 proven unsuccesful in predicting the rates of ecosystem functions from the
-abundance of functional genes and in our example study was not significant.
+abundance of functional genes and in our example study was not a good predictor
+of ecosystem function.
 Similarly, diversity metrics from the entire Prokaryotic community were not

### 2019-07-26 R/analysis.R
Playing around

@@ -40,13 +40,16 @@ options(knitr.kable.NA = '')
 #+ r tree_analysis
 #tree <- read_tree('../data/gabon/16S_rep_seqs_tree_Gabon.nwk')
 #+ r import_data
-
 taxon_table <- data.table::fread('../output/taxon_table.csv', data.table = FALSE)
 all_data <- data.table::fread('../output/gab_all_vst_troph.csv', data.table = FALSE)
 gen_data <- data.table::fread('../output/gab_all_vst_gen.csv', data.table = FALSE)
 n_samp <- nrow(all_data)
 n_asvs <- ncol(all_data[, grepl("^[asv]", colnames(all_data))])
 
+asvs <- as.matrix(all_data[, grepl("^[asv]", colnames(all_data))])
+vars <- all_data[, !grepl("^[asv]", colnames(all_data))]
+asvs <- asvs[, colSums(asvs) > 0]
+all_data <- cbind(vars, asvs)
 ## Remove ASVs only present in 1 sample
 remove_asvs_in_1_sample <- function(dataset) {
 asvs <- as.matrix(dataset[, grepl("^[asv]", colnames(dataset))])
@@ -238,18 +241,20 @@ taxon_table$taxon <- apply(taxon_table, 1, function(r) { r[
 
 #+ trad_correlations
 
+
 colnames(all_data[, 1:13])
 
 print(ggplot(all_data, aes(x = X, y = Y)) + 
   geom_point(aes(color = all_data$Wetland)) +
   labs(title = "Distribution of sites highlighted by wetland or upland"))
+
 ggplot(all_data, aes(x = pmoa_copy_num, y = Low_final_k)) +
   geom_point() +
   geom_smooth(method = 'lm') +
   theme_classic() +
-  labs(x = expression(italic('pmoA') ~ 'copy number'), y = expression('Methane oxidation')) + 
-  ggsave(file = '../Figures/lowk_pmoa.pdf', width = 4, height = 4)
-summary(lm(Low_final_k ~ pmoa_copy_num, data = all_data))
+  labs(x = expression(italic('pmoA') ~ 'copy number'), y = expression('Methane oxidation (k)')) + 
+  annotate(geom = 'text', x = 0, y = 2.5, label = 'A', size = 6) +
+  ggsave(file = '../figures/lowk_pmoa.pdf', width = 4, height = 4)
 
 ggplot(all_data, aes(x = pmoa_copy_num, y = Vmax)) +
[…]

### 2019-07-26 R/var_comp.R

@@ -12,7 +12,7 @@ library(data.table)
 #registerDoSNOW(cl)
 #import_data
 registerDoParallel(cores = 28)
-
+taxon_table <- fread('../output/taxon_table.csv', data.table = FALSE, na.strings = "")
 all_data <- as.data.frame(fread('../output/gab_all_vst_troph.csv'))
 # Variance Component Analysis
 
@@ -22,11 +22,70 @@ calc_sim_matrix <- function(x) {
 1/(1+as.matrix(dist(as.matrix(scale(x)))))
 }
 
-## Remove ASVs only present in 1 sample
 asvs <- as.matrix(all_data[, grepl("^[asv]", colnames(all_data))])
-vars <- all_data[, !grepl("^[asv]", colnames(all_data))]
+# Total dataset n asvs
+dim(asvs)
 asvs <- asvs[, colSums(asvs) > 0]
+# n asvs present in at least one sample
+dim(asvs)
+vars <- all_data[, !grepl("^[asv]", colnames(all_data))]
+
+
+#### Cluster ASVs at the genus level
+## n Unique ASVS
+#length(unique(taxon_table$asv))
+## n ASVs with no Genus assignment
+#sum(is.na(taxon_table$Genus))
+#dim(asvs)
+#asvs <- cbind(Sample = all_data$Sample, all_data[, grepl("^[asv]", colnames(all_data))])
+#asvs <- 
+#  asvs %>% 
+#  gather(asv, abundance, 2:ncol(asvs)) %>% 
+#  left_join(taxon_table) %>% 
+#  select(Sample, Genus, abundance) %>% 
+#  group_by(Sample, Genus) %>% 
+#  summarize(abundance = sum(abundance)) %>% 
+#  spread(Genus, abundance)
+#dim(asvs)
+#asvs[1:5, 1:5]
+#asvs <- as.matrix(asvs[, -1])
+#dim(asvs)
+#asvs <- asvs[, colSums(asvs) > 0]
+#
+###
+
+PA <- asvs
+PA[PA<2] = 0
+PA[PA>1] = 1
+PA[1:10, 1:10]
+sum(colSums(PA) == 44)
+sum(colSums(PA) == 0)
+sum(apply(PA,2,sum) == 44)
+sum(apply(PA,2,sum) == 0)
+PA <- PA[, !colSums(PA) %in% c(0, 44)]
+colnames(PA)
+asvs <- PA
+
+#dim(asvs)
+#sum(asvs > 0)
[…]

### 2019-07-26 R/analysis.R
Add processing genus-level models

@@ -84,6 +84,12 @@ screeplot(geo_pca)
 summary(geo_pca)
 all_data$geo_pc1 <- scores(geo_pca, choices = 1, display = 'sites', scaling = 3)
 
+all_data %>% 
+  ggplot(aes(Low_final_k)) +
+  facet_grid(geocode ~ Wetland) +
+  geom_density()
+colnames(all_data[1:20])
+
 #pca <- rda(as.matrix(gen_data[, grepl("^[asv]", colnames(gen_data))]))
 #screeplot(pca)
 #head(summary(pca))
@@ -730,6 +736,7 @@ wet_model_output <- as.data.frame(data.table::fread('../talapas-output/var_comp_
 dry_model_output <- as.data.frame(data.table::fread('../talapas-output/var_comp_out_dry.csv'))
 dry_model_output_bin <- as.data.frame(data.table::fread('../talapas-output/var_comp_out_dry_bin.csv'))
 dry_model_output_geo <- as.data.frame(data.table::fread('../talapas-output/var_comp_out_dry_geo.csv'))
+model_output_genus <- as.data.frame(data.table::fread('../talapas-output/var_comp_genus.csv'))
 
 #' ## Variance components
 
@@ -766,6 +773,10 @@ dry_model_output_bin %>%
         col.names = c('Components', 'Geography', 'Community', 'Environment', 'Error'),
         digits = 2, format = 'latex') %>% 
   kable_styling()
+
+model_output_genus %>% 
+  group_by(comps) %>% 
+  summarize_at(vars(geo, com, env, err), mean)
 
 #' ## Number of taxa identified with each covariate
 
@@ -810,6 +821,14 @@ my_qs <- qvalue(dry_model_output_geo$p.value[dry_model_output_geo$comps == i])
 summary(my_qs)
 dry_model_output_geo$qvalues[dry_model_output_geo$comps == i] <- my_qs$qvalues
 }
+
+for (i in unique(model_output_genus$comps)) {
+  print(i)
+my_qs <- qvalue(model_output_genus$p.value[model_output_genus$comps == i])
+summary(my_qs)
+model_output_genus$qvalues[model_output_genus$comps == i] <- my_qs$qvalues
+}
+
 #+ r n_taxa_ided
 
 n_taxa <- 
@@ -867,6 +886,12 @@ dry_model_output_geo %>%
   filter(qvalues < 0.05)  %>% 
[…]

### 2019-07-27 R/combine_data.R
Switch data combining to phyloseq

@@ -1,41 +1,56 @@
 library(tidyverse)
 library(morris)
-library(data.table)
-
-# Methanogen metadata
-gen_attr_table <- fread('../output/gab_gen_attr_table.csv', data.table = FALSE)
-gen_attr_table$Site <- paste0(gen_attr_table$Site, 
-				reverse_substr(gen_attr_table$Sample, 3, 3))
+library(phyloseq)
 
 # Methanotrophy metadata
-troph_attr_table <- fread('../output/gab_troph_attr_table.csv', data.table = FALSE)
+troph_attr_table <- data.table::fread('../output/gab_troph_attr_table.csv', data.table = FALSE)
 troph_attr_table$Site <- paste0(troph_attr_table$Site, 
 				reverse_substr(troph_attr_table$Sample, 3, 3))
-## Rarefied ASV Table
-#rare_asv_table <- read.csv('../output/gab_rare_asv_table.csv')
+# Methanogenesis metadata
+gen_attr_table <- data.table::fread('../output/gab_gen_attr_table.csv', data.table = FALSE)
+gen_attr_table$Site <- paste0(gen_attr_table$Site, 
+				reverse_substr(gen_attr_table$Sample, 3, 3))
+geodist <- read.table('../output/geodist.tsv', stringsAsFactors = FALSE)
+
+sample_data <- full_join(troph_attr_table, gen_attr_table)
+sample_data <- left_join(sample_data, geodist, by = "Site")
+rownames(sample_data) <- sample_data$Sample
+sample_data <- sample_data[, -1]
+
+sample_data[sample_data$Y > 30000, 'geocode'] <- 'A'
+sample_data[sample_data$Y < 30000 & sample_data$Y > 10000, 'geocode'] <- 'B'
+sample_data[sample_data$Y < 10000, 'geocode'] <- 'C'
+sample_data$geocode <- factor(sample_data$geocode)
 
 ## DESeq2 variance stabilized asv table
-vst_asv_table <- fread('../output/vst_asv_table.csv', data.table = FALSE)
+otu_table <- data.table::fread('../output/vst_asv_table.csv', data.table = FALSE)
+rownames(otu_table) <- otu_table[, 1]
+otu_table <- otu_table[, -1]
+colnames(otu_table) <- substr(colnames(otu_table), 5, nchar(colnames(otu_table)[1]))
+otu_table <- as.matrix(otu_table)
[…]

### 2019-07-27 R/analysis.R
Cleaning up analyses for publication

@@ -16,147 +16,30 @@ knitr::opts_chunk$set(fig.width=8, fig.height=6, fig.path='Figs/',
 #+ r load_packages, include=FALSE
 
 library(tidyverse)
-#library(lubridate)
 library(broom)
 library(broman)
-#library(morris)
 library(varComp)
 library(vegan)
-library(parallel)
 library(qvalue)
 library(knitr)
 library(kableExtra)
-#library(eulerr)
-#library(GGally)
 library(phyloseq)
-library(doParallel)
 library(gap)
 library(lmtest)
-
-num.cores <- detectCores()
-registerDoParallel(cores = 3)
+source('functions.R')
 options(knitr.kable.NA = '')
 
-#+ r tree_analysis
-#tree <- read_tree('../data/gabon/16S_rep_seqs_tree_Gabon.nwk')
 #+ r import_data
-taxon_table <- data.table::fread('../output/taxon_table.csv', data.table = FALSE)
-all_data <- data.table::fread('../output/gab_all_vst_troph.csv', data.table = FALSE)
-gen_data <- data.table::fread('../output/gab_all_vst_gen.csv', data.table = FALSE)
-n_samp <- nrow(all_data)
-n_asvs <- ncol(all_data[, grepl("^[asv]", colnames(all_data))])
+physeq <- readRDS('../output/physeq.rds')
 
-asvs <- as.matrix(all_data[, grepl("^[asv]", colnames(all_data))])
-vars <- all_data[, !grepl("^[asv]", colnames(all_data))]
-asvs <- asvs[, colSums(asvs) > 0]
-all_data <- cbind(vars, asvs)
-## Remove ASVs only present in 1 sample
-remove_asvs_in_1_sample <- function(dataset) {
-asvs <- as.matrix(dataset[, grepl("^[asv]", colnames(dataset))])
-vars <- dataset[, !grepl("^[asv]", colnames(dataset))]
-asvs <- asvs[, colSums(asvs) > 0]
-cbind(vars, asvs)
-}
-all_data <- remove_asvs_in_1_sample(all_data)
-gen_data <- remove_asvs_in_1_sample(gen_data)
+# Remove 0 abundance taxa
+otu_table(physeq) <- otu_table(physeq)[, colSums(otu_table(physeq)) > 0]
 
-# Add location factor
[…]

### 2019-07-27 R/combine_data.R

@@ -10,9 +10,11 @@ troph_attr_table$Site <- paste0(troph_attr_table$Site,
 gen_attr_table <- data.table::fread('../output/gab_gen_attr_table.csv', data.table = FALSE)
 gen_attr_table$Site <- paste0(gen_attr_table$Site, 
 				reverse_substr(gen_attr_table$Sample, 3, 3))
+gen_attr_table$Wetland <- "Wetland"
 geodist <- read.table('../output/geodist.tsv', stringsAsFactors = FALSE)
 
 sample_data <- full_join(troph_attr_table, gen_attr_table)
+sample_data$Experiment <- reverse_substr(sample_data$Sample, 1, 2)
 sample_data <- left_join(sample_data, geodist, by = "Site")
 rownames(sample_data) <- sample_data$Sample
 sample_data <- sample_data[, -1]
@@ -28,29 +30,21 @@ rownames(otu_table) <- otu_table[, 1]
 otu_table <- otu_table[, -1]
 colnames(otu_table) <- substr(colnames(otu_table), 5, nchar(colnames(otu_table)[1]))
 otu_table <- as.matrix(otu_table)
-class(otu_table)
-colnames(otu_table)
 
 ## Taxonomy
 taxon_table <- data.table::fread('../output/taxon_table.csv', data.table = FALSE)
 taxon_table$Species <- ""
 rownames(taxon_table) <- substr(taxon_table$asv, 5, nchar(taxon_table$asv[1]))
-taxon_table[1:10, ]
 taxon_table <- taxon_table[, -1]
 taxon_table <- as.matrix(taxon_table)
 
 ## Read tree
 tree <- read_tree('../data/gabon/16S_rep_seqs_tree_Gabon.nwk')
-class(tree)
 
 ## Make phyloseq object
 ASV <- otu_table(otu_table, taxa_are_rows = FALSE)
-sample_names(ASV)
 TAX <- tax_table(taxon_table)
-sample_names(TAX)
 sample_data <- sample_data(sample_data)
-sample_names(sample_data)
 
 physeq <- phyloseq(ASV, TAX, sample_data, tree)
 saveRDS(physeq, '../output/physeq.rds')
-

### 2019-07-27 R/functions.R

@@ -1,4 +1,33 @@
 
+varCompGE <- function(model, drop = FALSE) {
+  if (length(model$varComps) <= 1) {
+h2G(c(model$varComps, err = model$sigma2), vcov(model, what = 'varComp', drop = drop))
+  } else {
+h2GE(c(model$varComps, err = model$sigma2), vcov(model, what = 'varComp', drop = drop))
+  }
+}
+
+permute_no_replace_for <- function(perm_vec, nperm = 999) {
+samples <- matrix(ncol = nperm, nrow = length(perm_vec))
+samples <- cbind(samples, truth = perm_vec)
+for (i in seq_len(nperm)) {
+  sample_i <- sample(perm_vec)
+  while (any(apply(samples, 2, function(x) {all(sample_i == x)}), na.rm = T)) {
+    sample_i <- sample(perm_vec)
+  }
+  samples[, i] <- sample_i
+  colnames(samples)[i] <- paste0('perm.', i)
+}
+return(samples)
+}
+
+#taxon_table$taxon <- apply(taxon_table, 1, function(r) { r[ 
+#                      if( any(is.na(r))){ 
+#                          (which(is.na(r))[1]-1) 
+#                       }else{
+#                          (length(r)-0)}
+#                                      ] }
+#       )
 ##### Functions
 
 lowk_model <- function(df) {

### 2019-07-27 R/var_comp.R

@@ -12,7 +12,7 @@ library(data.table)
 #registerDoSNOW(cl)
 #import_data
 registerDoParallel(cores = 28)
-taxon_table <- fread('../output/taxon_table.csv', data.table = FALSE, na.strings = "")
+#taxon_table <- fread('../output/taxon_table.csv', data.table = FALSE, na.strings = "")
 all_data <- as.data.frame(fread('../output/gab_all_vst_troph.csv'))
 # Variance Component Analysis
 
@@ -54,17 +54,17 @@ vars <- all_data[, !grepl("^[asv]", colnames(all_data))]
 #
 ###
 
-PA <- asvs
-PA[PA<2] = 0
-PA[PA>1] = 1
-PA[1:10, 1:10]
-sum(colSums(PA) == 44)
-sum(colSums(PA) == 0)
-sum(apply(PA,2,sum) == 44)
-sum(apply(PA,2,sum) == 0)
-PA <- PA[, !colSums(PA) %in% c(0, 44)]
-colnames(PA)
-asvs <- PA
+#PA <- asvs
+#PA[PA<2] = 0
+#PA[PA>1] = 1
+#PA[1:10, 1:10]
+#sum(colSums(PA) == 44)
+#sum(colSums(PA) == 0)
+#sum(apply(PA,2,sum) == 44)
+#sum(apply(PA,2,sum) == 0)
+#PA <- PA[, !colSums(PA) %in% c(0, 44)]
+#colnames(PA)
+#asvs <- PA
 
 #dim(asvs)
 #sum(asvs > 0)
@@ -188,7 +188,7 @@ system.time(results <- foreach(i = seq_len(ncol(asvs)), .combine = 'rbind',
                                .packages = c('vegan', 'varComp', 'tidyverse')) %dopar% { 
             suppressMessages(suppressWarnings(fit_varComps(x = i, all_data = all_data)))
                  })
-fwrite(results, "../output/var_comp_genus.csv")
+fwrite(results, "../output/var_comp_abund.csv")
 ##+ fit_var_comp_wet
 #
 #wet_data <- filter(all_data, Wetland == 'Wetland')

### 2019-07-27 R/analysis.R
Continuing to straighten up scripts

@@ -35,10 +35,31 @@ physeq <- readRDS('../output/physeq.rds')
 # Remove 0 abundance taxa
 otu_table(physeq) <- otu_table(physeq)[, colSums(otu_table(physeq)) > 0]
 
+# Add variables
+sample_data(physeq)$richness <- rowSums(com_mat > 0)
+sample_data(physeq)$shannon <- diversity(com_mat, "shannon")
+sample_data(physeq)$simpson <- diversity(com_mat, "simpson")
+
+pca <- rda(otu_table(physeq))
+screeplot(pca)
+head(summary(pca))
+sample_data(physeq)$pc1 <- scores(pca, choices = 1, display = 'sites', scaling = 3)
+sample_data(physeq)$pc2 <- scores(pca, choices = 2, display = 'sites', scaling = 3)
+sample_data(physeq)$pc3 <- scores(pca, choices = 3, display = 'sites', scaling = 3)
+sample_data(physeq)$pc4 <- scores(pca, choices = 4, display = 'sites', scaling = 3)
+sample_data(physeq)$pc5 <- scores(pca, choices = 5, display = 'sites', scaling = 3)
+sample_data(physeq)$pc6 <- scores(pca, choices = 6, display = 'sites', scaling = 3)
+sample_data(physeq)$pc7 <- scores(pca, choices = 7, display = 'sites', scaling = 3)
+
 sample_data(physeq) %>% 
   filter(Experiment == 'MO') %>% 
   ggplot(aes(Low_final_k)) +
-  facet_grid(geocode ~ Wetland) +
+  geom_density()
+
+sample_data(physeq) %>% 
+  filter(Experiment == 'MO') %>% 
+  ggplot(aes(Low_final_k)) +
+  facet_grid(Wetland ~ geocode) +
   geom_density()
 
 #' # Overview
@@ -99,7 +120,6 @@ sample_data(physeq) %>%
 
 #+ trad_correlations
 sample_data <- as(sample_data(physeq), 'data.frame')
-
 sample_data(physeq) %>% 
   ggplot(mapping = aes(x = pmoa_copy_num, y = Low_final_k)) +
   geom_point() +
@@ -111,10 +131,7 @@ sample_data(physeq) %>%
 ggsave(file = '../figures/lowk_pmoa.pdf', width = 4, height = 4)
 
 
-pca <- rda(otu_table(physeq))
-screeplot(pca)
-head(summary(pca))
[…]

### 2019-07-27 R/clean_sample_data.R

@@ -1,4 +1,3 @@
-
 library(tidyverse)
 library(morris)
 
@@ -130,6 +129,10 @@ gen_attr_table <- gen_attr_table %>%
 colnames(troph_attr_table)
 troph_attr_table <- troph_attr_table %>% 
   select(Sample:pmoa_transcript_num)
+
+# Rename Copies
+gen_attr_table$mcra_copy_num <- gen_attr_table$Copies
+gen_attr_table <- subset(gen_attr_table, select = -c(Copies))
 
 write_csv(gen_attr_table, '../output/gab_gen_attr_table.csv')
 

### 2019-10-08 manuscript/manuscript_rev1.Rmd
Create revision and reviewer comments

---
title: "Linking microbial communities to ecosystem functions: what we can learn from genotype-phenotype mapping in organisms"
output: 
    word_document:
        reference_docx: reference.docx
bibliography: library.bib
csl: philosophical-transactions-of-the-royal-society-b.csl 
---

```{r setup, include=FALSE}

library(tidyverse)
library(knitr)
library(flextable)
opts_chunk$set(echo = FALSE)

```

# Abstract

Microorganisms mediate many important ecosystem functions, yet it remains
unclear to what extent microbial diversity or community composition is
important for determining the rates of ecosystem-scale functions. This
uncertainty limits our ability to predict and manage crucial
microbially mediated processes, such as nutrient loss and greenhouse gas
emissions. Our lack of understanding stems from the relatively large
diversity of microorganisms, the difficulty in directly identifying
functional groups, and our limited ability to manipulate microbial community
attributes. For this reason, we propose that integrating traditional
biodiversity-ecosystem function research with ideas from genotype-phenotype
mapping could provide the new perspective our discipline needs. We identify
three insights from genotype-phenotype mapping that could be useful for
microbial biodiversity-ecosystem function studies: the concept of “agnostic”
mapping, the use of more powerful ways to account for multiple comparisons,
and the incorporation of covariates into models of ecosystem function. We
illustrate the potential for these approaches to elucidate microbial
biodiversity-ecosystem function relationships by analyzing a subset of
published data measuring methane oxidation rates from incubations of tropical
soil. We assert that combining the approaches of traditional
biodiversity-ecosystem function research with ideas from genotype-phenotype
mapping will not only generate novel hypotheses about how complex microbial
communities drive ecosystem function, but also help scientists predict and
manage changes to ecosystem functions resulting from human activities.

## Introduction

### A new perspective on microbial community-ecosystem function relationships is needed

Ecologists have long investigated the effects of changing biodiversity on
ecosystem function, documenting, for example, relationships between terrestrial
plant community richness and primary productivity [@hooper2005]. The study of
the relationship between microbial biodiversity and ecosystem functions is much
more recent and has generated mixed results [@schimel1998, @singh2010]. Despite
the fact that microbes mediate many important ecosystem functions, it remains
maddeningly unclear to what extent microbial diversity or community composition
is important for determining the rates of ecosystem-scale functions.  For
example, numerous studies have attempted to correlate microbial functional
group abundance or diversity with the rate of various ecosystem functions and
most are unsuccessful [@rocca2015, @graham2016]. This uncertainty limits our
ability to predict and manage crucial microbially mediated ecosystem functions,
such as nutrient loss and greenhouse gas emissions. As a discipline, we need a
new perspective on these questions.

Our lack of understanding stems from many sources. Microbial communities are
more diverse and contain greater numbers of individuals than plant communities and are thus more
difficult to sample comprehensively. Unlike plants, functional groups of
microbes are rarely determined through observation or direct measurement, but
rather must be indirectly inferred based on taxonomic assignment of marker
genes (such as 16S rRNA genes), by sequencing functional genes
from metagenomes, or by quantifying specific functional genes using qPCR. This adds uncertainty in our ability to estimate the size of functional groups.
The ability to experimentally manipulate microbial diversity or community
composition is much more limited for microbes. Given all of these differences
(and many others), the challenge of linking microbial communities to ecosystem
functions may be less like that of plant communities, and more analogous to the
complex task of linking genomic variation to organismal phenotypes. For this
reason, we propose that integrating traditional biodiversity-ecosystem function
research with ideas from genotype-phenotype mapping could provide the new
perspective our discipline needs.

### There is evidence that microbial diversity can matter

Perhaps our uncertainty arises because there is no general relationship between
microbial biodiversity and any particular ecosystem function.  For example, if
local selection always optimizes the available microbial biodiversity to
maximize ecosystem function, then microbial community composition should not
matter for predicting the rates of microbially mediated functions. Instead, the
rates of these functions should be determined by the underlying environmental
variation. In this case, the microbial community would simply act as a conduit
through which the abiotic environment alters ecosystem function. Attempts to
document a general relationship between microbial community attributes and
function would then fail (since the primary drivers are environmental).

However, there is evidence that this is not always the case, A limited number
of studies have manipulated the connection between environmental factors and
microbial community composition through “common garden” or reciprocal
transplant experiments, and they frequently report different rates of ecosystem
functions for different microbial communities under the same environmental
conditions [@reed2007].  This has been observed for plant decomposition
[@glassman2018, @strickland2009], plant phenology [@panke-buisse2015], soil
nitrogen cycling [@balser2005], and soil greenhouse gas emissions
[@cavigelli2000], among others. Therefore, variation in microbial community
composition can be associated with variation in ecosystem function, independent
of environmental variation.  So why has this been difficult to consistently
document in the field?

### Traditional approaches to quantifying this relationship have provided only minor improvements

Most comparative studies of microbial community function in the field focus on
one of two aspects of microbial community structure that are hypothesized to
predict ecosystem function. The first aspect is “functional” gene or transcript
abundance.  In this case, qPCR or shotgun metagenomic sequencing is used to
estimate the abundance of a gene or transcript that is a putative marker for a
microbial process (and thus a marker for the functional group that performs
this process). For example, the gene mcrA, which encodes a subunit of the
enzyme that performs the final step in methanogenesis, is commonly used as a
marker for methanogenesis and for the methanogen functional group. Other
examples include pmoA and methanotrophy, nifH and nitrification, and nosZ and
denitrification. The abundance of these markers is often hypothesized to be
predictive of the rate of their associated processes (for example, the
abundance of mcrA is often hypothesized to be related to the rate of
methanogenesis). There are examples where this relationship is present
[@freitag2009, @freitag2010, @schnyder2018].  However, a review of such studies
found that the abundance of a functional gene or transcript is rarely
correlated with the rate of the corresponding process and depends on the
function of interest, with most effects either negative or not significant
[@rocca2015].

The second aspect of microbial community structure hypothesized to predict
ecosystem function is taxonomic or functional diversity. Diversity is either
estimated from sequence variants of a barcode gene such as the 16S rRNA gene or
manipulated through some proxy of diversity such as sequential dilution or
varying filter sizes. Studies that experimentally manipulate diversity
generally find a relationship between the applied diversity treatment and
ecosystem functions, including for methanotrophy [@schnyder2018], phosphorus
leaching [@wagg2014], greenhouse gas emissions [@wagg2014], decomposition
[@maron2018], and nitrogen cycling [@philippot2013]. However, in many cases,
richness is confounded with other factors such as the presence or absence of
major phylogenetic groups (nematodes, fungi) or abundance of microbial cells.
In addition, assembled microbial communities pose the same challenges as
macroorganismal diversity experiments, for example the highest diversity
treatment will more likely contain the most productive taxon. A review of
microbial diversity studies shows that microbial community taxonomic and
functional diversity add little explanatory power to models of ecosystem
function [@graham2016]. Overall, functional gene abundance and community
diversity improve models of ecosystem function less than one third of the time
and increase variance explained by an average of only 8 percentage points
[@graham2016].

### Genotype-phenotype mapping as a source of inspiration

Given the relative lack of success to date, it is time to rethink how we
approach the challenge of mapping microbial community structure to ecosystem
function.  Microbial ecologists are not the only biologists who are attempting
to determine the relationship between a complex set of highly-variable data and
an aggregate function.  This kind of “many-to-one” mapping is analogous to the
challenge of identifying the genetic basis of complex traits in organismal
populations. In such “genotype-phenotype” mapping studies, a population
exhibits variation in a phenotype (e.g., height or disease state) as well as
variation in potentially thousands of single nucleotide polymorphisms (SNPs).
To identify the genetic basis for a trait, investigators sample from this
population and correlate phenotype with genotype. In most cases, there are many
more loci than individuals and we do not know whether the SNPs are causally
linked or are simply in linkage disequilibrium with a causal mutation.

There are a number of parallels between this challenge faced by organismal
biologists and that facing microbial community ecologists.  They are both
“many-to-one” mapping challenges, which involve large numbers of statistical
comparisons.  Both are attempting to identify causal relationships that are
potentially confounded by very complex patterns of covariation.  There is often
no strong expectation about which entities (i.e., which genomic regions or which
microbial taxa) are most likely to be causally-related to phenotype or
function, and thus “agnostic” approaches are needed.  Both are sensitive to
exactly how the “mapping” question is asked, and ultimately require
manipulation (of genes or taxa) to establish causation.  We describe each of
these parallels below in more detail, and provide an example of how these ideas
could be applied to microbial data.

### The importance of a taxonomically “agnostic” approach

The traditional approach to microorganismal biodiversity-ecosystem function
research is to measure or manipulate the diversity of a taxonomic group (e.g.,
plants) and look for an association with the function performed by that group
(e.g., primary productivity).  In the broadest sense, we can think of plants as
a “functional group,” i.e., a group of taxa united by their ability to perform
a particular ecosystem function. Ecologists may further divide a functional
group (e.g., plants) into smaller functional groupings (e.g., forbs), defined
by the details of how they perform their particular ecosystem function.

However, for microbes our knowledge of functional groups is much more limited.
From a very limited number of cultured isolates we have a provisional
understanding of which microbes might be involved in some ecosystem functions.
And by sequencing the genomes of these isolates, we have identified genetic
markers for certain functions. But most microbial taxa remain uncultured and we
do not know the function of most microbial taxa detected in environmental
samples [@hug2016, @steen2019]. In addition, there have been recent discoveries
of functions in unexpected taxonomic groups, for example methanogenesis by
fungi and cyanobacteria, a function previously considered restricted to the
Archaea [@bizic-ionescu2019, @lenhart2012]. Because of this, it would be
prudent to look more agnostically at microbial communities to identify taxa or
groups of taxa that are important for predicting the rate of ecosystem
functions rather than assuming that the genetic markers we have provisionally
identified for a given function represent the most likely taxa involved.  This
agnostic approach is analogous to the approach of many genotype-phenotype
mapping studies (e.g.,  genome-wide association studies or GWAS), which often
look for associations between a phenotype and loci anywhere in a genome.

### Dealing with multiple comparisons and covariation

Microbial biodiversity-ecosystem function studies and genotype-phenotype
mapping studies are both “many-to-one” mapping challenges, which involve large
numbers of statistical tests.  Hundreds to thousands of tests are routinely
made per study, greatly inflating the number of false positives identified
using statistical hypothesis testing approaches.  Some microbial
biodiversity-ecosystem function studies do not correct for this, while others
use approaches that may unnecessarily inflate the false negative rate, such as
the Bonferroni correction. The Bonferroni correction has been widely considered
to be too conservative, particular for exploratory studies designed to generate
hypotheses [@perneger1998, @rothman1990, @thomas1985]. Statisticians have
developed a number of less-conservative approaches to correct for multiple
comparisons by controlling the false discovery rate in order to balance the
tradeoff between Type I and Type II errors [@benjamini1995, @storey2002].
These approaches are commonly used in genotype-phenotype mapping studies.

It is widely accepted that organisms, including microorganisms, exhibit
population stratification due to geographic and environmental separation
[@martiny2006, @wright1943]. Genome-wide association studies generally account
for population structure due to shared ancestry among cases and controls when
modeling the connection between genotype and phenotype. The classic example is
the latitudinal gradient of both height and genotypic similarity in Europe,
which results in spurious associations between human height and genetic
variation [@berg2019, @novembre2008, @sohail2019]. To correct for this
covariance structure, GWAS models incorporate genotypic similarity to correct
for shared ancestry using a variety of methods, such as principal component
correction or variance component modeling [@kang2010, @price2006]. Typically,
microbial biodiversity-ecosystem function studies do not account for population
stratification (i.e., community similarity across samples), although there are
some exceptions [@meyer2019, @qin2012]. Genome-wide association studies
generally ignores the underlying environmental and spatial distance between
samples and instead uses shared ancestry as a proxy for these variables.
However, community similarity (the community analogue of shared ancestry among
organisms) is not as tightly linked to geography or environment as is shared
ancestry among organisms, and it could be very useful to account for these
separately in microbial studies, especially if one is particularly interested
in how composition alters function independent of the underlying environmental
variation.

## An example: high-affinity methane oxidation

To illustrate the ideas outlined above, we reanalyzed a subset previously
published data [@meyer2019] using a modified procedure from the original
version.  A full description of the study design, samples, and data generation
can be found in that manuscript. Briefly, these data were gathered from intact
soil cores taken from diverse ecosystems of the Congo Basin in Gabon, Africa.
Cores were incubated in the laboratory under different concentrations of
methane to identify the rates of specific methane cycling pathways. For this
example, we will analyze data from just one of these pathways, high-affinity
methane oxidation (the oxidation of atmospheric concentrations of methane),
which we will refer to below as “methane oxidation”. In addition, for
simplicity we only include amplicon sequences from the DNA-inferred community
and not the RNA-inferred community, both of which are presented in the original
paper [@meyer2019].  The data we analyzed include methane oxidation
measurements, amplicon sequence variants (ASVs; @callahan2017) generated using
DADA2 and inferred from 16S rRNA gene sequences, pmoA abundance estimates (via
qPCR), latitude and longitude, and four environmental covariates (soil
moisture, bulk density, carbon, and nitrogen).

Analyses were conducted in the R statistical environment using the phyloseq
package [@mcmurdie2013, @rcoreteam2019].  The relative abundances of ASVs were
corrected using the variance stabilizing transformation from DESeq2 [@love2014,
@mcmurdie2014]. We first test typical measures of microbial structure including
functional gene abundance and community richness, which was estimated using the
breakaway package [@willis2015]. We then demonstrate significant covariation
between community structure (estimated as Bray-Curtis distance using vegan),
environmental variation (euclidean distance), and geographic distance
(euclidean distance) using Mantel tests [@bray1957, @oksanen2019].  Finally, we
present one approach to identifying taxa which are significantly associated
with function independent of the environment by fitting variance component
models using varComp to test the relationship between relative abundance of
each ASV and methane oxidation rate [@qu2013]. To illustrate how including
different covariates (environmental, geographic, and community) can result in
different conclusions about which taxa are associated with function, we fit
this model with and without random effects variance components for
environmental similarity, geographic site ID, and Bray-Curtis similarity.
Significant taxa were determined by controlling the false discovery rate at
q-value < 0.05 [@storey2002]. Figures were created using ggplot2
[@wickham2019].

Methane oxidation rate was not significantly correlated with pmoA gene
abundance or richness (Table 1, Figure 1). We tested collinearity between each
pair of distance matrices for community, environment, and geography using
Mantel tests and estimated p-values by permutation. We found a moderate and
significant correlation between community composition and environmental
variation, geography and community composition, and geography and environmental
variation (Table 2, Figure 2). Principal coordinate plots show that beta
diversity of samples separated by site ID and by ecosystem type (wetland or
upland; Figure 2). Finally, we tested the effect of the relative abundance of
each ASV on methane oxidation rate. We found different numbers of taxa
significantly associated with methane oxidation depending on which covariates
were included in the model (Table 3).  In particular, 460 taxa were identified
with no covariates whereas 6 were identified after including all covariates.
Though we cannot infer function from 16S sequences, these 6 taxa fall into
three genera and one class with cultured representatives that are not known to
consume methane [@belova2009, @chang2011, @domeignoz-horta2019, @fritz2004].
Their effect on methane oxidation rate ranges from 0.5 to 1.5 (Figure 3).  An
effect of 1.5 means that for a one unit increase in relative abundance there is
a 1.5 increase in k (the rate of exponential decrease in methane concentration
over time).

## Discussion

In this paper, we argue that the traditional approach to microbial
biodiversity-ecosystem function research often ignores the complexity of
microbial community dynamics, in particular the complex patterns of covariation
that often arise among microbial community similarity, environmental
similarity, and spatial proximity.  We propose that a better approach would be
to learn from other complex many-to-one mapping problems in biology,
particularly genotype-phenotype mapping.  These studies often explicitly
account for covariation, they correct for multiple comparisons in powerful
ways, and they frequently take an “agnostic approach” that does not assume a
particular relationship between structure and function.

We illustrate the potential for these approaches to elucidate microbial
biodiversity-ecosystem function relationships by analyzing a subset of
published data from incubations of tropical soil.  In this example, soil cores
from different ecosystems exhibited different rates of methane oxidation, but
simple correlations between these rates and the diversity or abundance of
putative methane oxidizing bacteria were not informative.  These results agree
with recent reviews of the literature that demonstrate that such simple
approaches are often not a fruitful avenue for elucidating microbial
structure-function connections.

We then asked if the relative abundance of particular taxa were related to
ecosystem function, and we identified 460 taxa whose relative abundances were
associated with rates of methane oxidation after controlling the false
discovery rate (q-value < 0.05).  These taxa could be related to ecosystem
function in multiple ways.  The most interesting possibility is that each of
these taxa is statistically related because it is causally connected to the
function. This could be direct, for example an organism that consumes methane,
or indirect, for example an organism that regulates substrates necessary for
consumers of methane. In either of these cases, the taxon could be useful as a
biomarker of function or as an organism to investigate in order to better
understand the biological drivers of variation in methane oxidation.

Alternatively, a significant association could occur for non-causal reasons.
For example, any organism that tends to be in high abundance where methane
oxidation rates are high would be correlated with methane oxidation, even if it
has no causal relationship. This could be because such an organism is favored
under the same environmental conditions that favor methane oxidizing bacteria
(or that favor methane oxidation in general).  Such covariation can drive
associations that are not causal, but the effects of such covariation can be
reduced through a number of approaches (many developed by biologists interested
in genotype-phenotype mapping).

In our example, we showed that the abundances of microbial taxa exhibit complex
patterns of covariation with each other and with environmental conditions and
spatial location. These biogeographic patterns might be even stronger if we
could sample the populations more intensively [@meyer2018]. Once we account for this
covariation, our list of taxa associated with function was reduced to 6.  These
6 taxa include taxa that are not known to directly contribute to methane
oxidation, suggesting that the drivers of methane oxidation may be indirect,
perhaps mediated by ecological interactions among taxa from multiple functional
groups.

This approach is powerful because it accounts for the covariances that arise
from the fundamental ecological processes that drive community assembly and
underlie biogeographical patterns [@martiny2006, @vellend2010]. It is also powerful because it is
focused on a much more specific version of the “biodiversity-ecosystem
function” question, i.e., it asks “which taxa are uniquely associated with
function?” By “uniquely associated,” we mean those taxa associated with
function irrespective of environmental conditions, local community structure,
or spatial proximity.  This is not only a more specific question than is
usually asked in microbial biodiversity-ecosystem function studies, but it is
also a more appropriate one, especially if one is interested in how to
incorporate microbial community data into ecosystem models. For modeling what
is usually important is to identify those taxa that add explanatory power
beyond that provided by other factors (such as environmental conditions).

We identified three insights from genotype-phenotype mapping that could be
useful for microbial biodiversity-ecosystem function studies: the concept of
“agnostic” mapping, the use of more powerful ways to account for multiple
comparisons, and the incorporation of covariates.  But there are other insights
to be gained as well.  For example, some phenotypes (e.g., the propensity for
diseases such as Parkinson's) are controlled by a single genetic locus [@kerem1989, @macdonald1992].
However, most phenotypes studied to date arise from the influence of many loci
of small effect (as well as environmental factors; [@hindorff2009, reich2001]). Organismal
biologists have developed approaches tailored to identifying such loci; for
example, by sampling organismal populations in a way that constrains genetic
variation unrelated to phenotype. Similarly, for some ecosystem functions, it
is possible that a single microbial taxon could substantially influence its
rate. For example, methane flux from permafrost in Sweden may be controlled by
a single taxon [@mccalley2014]. However, most microbially mediated ecosystem functions are
likely the result of interactions (direct and indirect) among many taxa.  We
could maximize our ability to identify taxa of small effect by constraining
variation, e.g., by sampling microbial communities in a way that makes them
similar in structure (similar ecosystem, soil type, and abiotic conditions)
while still varying in function.

Ultimately, the relationships identified in genotype-phenotype mapping studies
must be verified.  There are multiple ways that this verification is
accomplished. In some cases, organisms can be artificially selected for a
particular phenotype (e.g., through experimental evolution in an environment
that favors the phenotype of interest) and the genetic changes that occur in
response to selection can be compared to those identified via mapping studies
(such as GWAS).  An analogous approach for microbial biodiversity-ecosystem
function studies would be to apply artificial ecosystem selection (sensu @swenson2000)
on a given function and compare the taxa that change in response to selection
with those identified via a comparative approach (such as the one illustrated
in our example).

The most common way that loci identified in a mapping study are verified is
through manipulative genetics.  The identified loci can be knocked out or
over-expressed, and the effect on phenotype compared to that predicted from
mapping studies.  There is no direct analogue for this in microbial
biodiversity-ecosystem function studies.  In some cases it may be possible to
inhibit a particular functional group through the use of a specific
antimicrobial or a chemical inhibitor [@maxson2016], but this is not generally true.  It
may be possible in some cases to isolate a microorganism of interest in pure
culture and add it back to a particular ecosystem, transiently increasing its
abundance (roughly analogous to “overexpressing” a gene).  Synthetic
communities (contrived assemblages of microorganisms) may ultimately be the
most powerful way to test hypotheses about microbial biodiversity-ecosystem
function relationships, but currently these approaches are limited by the small
number of taxa that can be routinely cultured from most environments.

### Conclusion

Microbial biodiversity-ecosystem function research has demonstrated that
functional group abundance (measured by qPCR of functional genes) is often not
a good predictor of ecosystem function and that microbial
diversity metrics on average do not add much power to ecosystem models.  A new
perspective on how to determine the relationship between microbial communities
and ecosystem functions is sorely needed.  Organismal biologists have over a
hundred years of experience identifying relationships between complex sets of
highly-variable data (genotypes or genome sequences) and aggregate functions
(organismal phenotypes).  We assert that combining the approaches of
traditional biodiversity-ecosystem function research with ideas from
genotype-phenotype mapping could provide this new perspective.  This
integration could not only make underutilized approaches such as covariate
modeling and artificial selection more available to microbial ecologists, but
also provide instructive examples of how best to conceive of microbial
biodiversity-ecosystem function questions.  If this integration is successful,
it is possible that in the not-so-distant future our field will be able to
robustly identify taxa, genes, or even molecules that will allow us to
accurately predict the response of ecosystems to environmental change. Doing so
will not only generate novel hypotheses about how complex microbial communities
drive ecosystem function, but also help scientists predict and manage changes
to ecosystem functions resulting from human activities.

# Acknowledgments

This project was supported by the National Science Foundation – Dimensions of
Biodiversity program (DEB 14422214), the National Science Foundation Graduate
Research Fellowship Program (DGE 1255832), and the ARCS Foundation Florence and
Mike Nudelman Scholarship. The data re-analyzed here were generated with
permission of the Government of Gabon, Centre National de la Recherche
Scientifique et Technologique (Permit No AR0035/14/MESRS/CENAREST/CG/CST/CSAR)
and the assistance of F. Bivigou, H. Memiaghe, L. Tchignoumba, E.  Tobi, and I.
Akendengue. The University of Oregon, the Gabon-Oregon Transnational Center on
Environment and Development, the Smithsonian Conservation Biology Institute,
and Shell Gabon provided financial and logistical support for the original
study re-analyzed here. We also thank W. Cresko for insightful discussions on
population genetics and H.  Tavalire for contributions to the statistical
analyses.

# References

::: {#refs} 
:::

# Tables and Figures

```{r copy_num}
copy_number <- data.frame(Term = c('pmoA copy number', 'Richness'),
                          Estimate = c(0.019, 0.001),
                          SE = c(0.011, 0.001),
                          t_statistic = c(1.700, 0.694),
                          p_value = c(0.096, 0.491),
                          n = c(42, 44))
flextable(copy_number) %>% 
  compose(i = 1, j = 1, 
    value = as_paragraph(as_i("pmoA"), " copy number")) %>% 
  compose(part = "header", j = 4, 
    value = as_paragraph("t-statistic")) %>% 
  compose(part = "header", j = 5, 
    value = as_paragraph("p-value")) %>% 
  colformat_int(col_keys = 'n', big.mark = ",")  %>% 
  autofit()
```

Table 1. Functional gene abundance and ASV richness are not good predictors of
methane oxidation rate. Linear models predicting methane oxidation rate from
measures of microbial community structure.

```{r mantel}
mantel <- data.frame(Terms = c('Community ~ Environment', 'Community ~ Geography', 'Environment ~ Geography'),
                     mantel = c(0.479, 0.360, 0.241),
                     quantiles = c(0.096, 0.055, 0.060),
                     p_value = c(0.001, 0.001, 0.001))
flextable(mantel) %>% 
  compose(part = "header", j = 2, 
    value = as_paragraph("Mantel statistics (r)")) %>% 
  compose(part = "header", j = 3, 
    value = as_paragraph("95% upper quantile of permutations")) %>% 
  compose(part = "header", j = 4, 
    value = as_paragraph("p-value")) %>% 
  autofit()
```

Table 2. Mantel tests for each pair of dissimilarity matrices. Community
distance matrix was based on Bray-Curtis distance while both environment and
geography distance matrices were based on Euclidean distance. P-values
determined by permutation test with 999 permutations.

```{r n_taxa, eval=FALSE}
mantel <- data.frame(Terms = c('None', 'Geo', 'Com', 'Env', 'Geo + Env', 'Geo + Env', 'Com + Env', 'Geo + Com + Env')
flextable(mantel) %>% 
  compose(part = "header", j = 2, 
    value = as_paragraph("Mantel statistics (r)")) %>% 
  compose(part = "header", j = 3, 
    value = as_paragraph("95% upper quantile of permutations")) %>% 
  compose(part = "header", j = 4, 
    value = as_paragraph("p-value")) %>% 
  autofit()
```

Table 3. Number of significant taxa after including each set of covariates in a
variance component model. Removed and Added columns are relative to the
no-covariate model. Significance is determined by controlling the false
discovery rate at q-values < 0.05.

```{r methane, out.width="768px", out.height="364px"}
knitr::include_graphics('../figures/trad_cors.pdf')
```

Figure 1. Methane oxidation rate is not correlated with functional gene
abundance or ASV richness. Correlations between community attributes and
ecosystem function. A) Abundance of the functional gene pmoA (n = 42) and B)
ASV richness (n = 44). Lines represent the ordinary least squares regression
line with standard errors.

```{r structure, out.width="960px", out.height="1152px"}
knitr::include_graphics('../figures/pcoa_multi.pdf')
```

Figure 2. Microbial community composition is spatially and environmentally
structured. Principle coordinate plots of Bray-Curtis distance representing the
first three axes of community composition. In A, B, and C, points are
identified by Site ID and in D, E, and F, points are identified by wetland or
upland ecosystem. All four environmental covariates separate strongly by
wetland/upland. Axis length is proportional to variance explained as indicated
in parentheses. PC = principal coordinate.

```{r taxa, out.width="768px", out.height="384px"}
knitr::include_graphics('../figures/effect_sizes.pdf')
```

Figure 3. Taxa associated with methane oxidation rates after controlling for
geographic location, environmental similarity, and community composition.
Points are estimates for the linear relationship between the relative abundance
of a single ASV and methane oxidation rate with standard errors from variance
component models including similarity matrices as covariates for community and
environment and site ID for geographic location. Amplicon sequence variants are
labelled at the finest resolution available: genus for all except the Group 1
Acidobacterium. Points are identified by Phylum. Significant taxa were
determined by controlling the false discovery rate at q-value < 0.05.

### 2019-10-08 manuscript/manuscript_rev1.Rmd
Update tables in revision 1

@@
 ```{r copy_num}
-copy_number <- data.frame(Term = c('pmoA copy number', 'Richness'),
-                          Estimate = c(0.019, 0.001),
-                          SE = c(0.011, 0.001),
-                          t_statistic = c(1.700, 0.694),
-                          p_value = c(0.096, 0.491),
-                          n = c(42, 44))
-flextable(copy_number) %>% 
+model_out <- read.csv('../output/model_out.csv')
+flextable(model_out) %>% 
   compose(i = 1, j = 1, 
@@
     value = as_paragraph("p-value")) %>% 
-  colformat_int(col_keys = 'n', big.mark = ",")  %>% 
   autofit()
@@
 ```{r mantel}
-mantel <- data.frame(Terms = c('Community ~ Environment', 'Community ~ Geography', 'Environment ~ Geography'),
-                     mantel = c(0.479, 0.360, 0.241),
-                     quantiles = c(0.096, 0.055, 0.060),
-                     p_value = c(0.001, 0.001, 0.001))
-flextable(mantel) %>% 
+mantel_table <- read.table('../output/mantel_table.tsv')
+flextable(mantel_table) %>% 
   compose(part = "header", j = 2, 
@@
 
-```{r n_taxa, eval=FALSE}
-mantel <- data.frame(Terms = c('None', 'Geo', 'Com', 'Env', 'Geo + Env', 'Geo + Env', 'Com + Env', 'Geo + Com + Env')
-flextable(mantel) %>% 
+```{r n_taxa}
+n_taxa <- read.table('../output/comp_adds_removes.tsv')
+flextable(n_taxa) %>% 
+  compose(part = "header", j = 1, 
+    value = as_paragraph("Terms")) %>% 
   compose(part = "header", j = 2, 
-    value = as_paragraph("Mantel statistics (r)")) %>% 
+    value = as_paragraph("Removed")) %>% 
   compose(part = "header", j = 3, 
-    value = as_paragraph("95% upper quantile of permutations")) %>% 
+    value = as_paragraph("Added")) %>% 
   compose(part = "header", j = 4, 
-    value = as_paragraph("p-value")) %>% 
+    value = as_paragraph("Significant")) %>% 
   autofit()

### 2019-10-18 R/7_analysis.R
Update figures for ARCS poster

@@ -27,6 +27,7 @@ library(gap)
 library(breakaway)
 library(ape)
 library(ggpubr)
+library(flextable)
 
 source('functions.R')
 
@@ -86,6 +87,24 @@ ggarrange(pmoa_plot, rich_plot,
 
 ggsave(file = '../figures/trad_cors.pdf', width = 4*2, height = 4)
 
+(pmoa_plot_arcs <- sample_data(physeq) %>% 
+  ggplot(aes(x = pmoa_copy_num, y = pos_lowk)) +
+  geom_point() +
+  geom_smooth(method = 'lm') +
+  labs(x = 'Abundance of methane consumers', y = expression('Methane oxidation rate (-k)')))
+
+(rich_plot_arcs <- sample_data(physeq) %>% 
+  ggplot(mapping = aes(x = rich_breakaway, y = pos_lowk)) +
+  geom_point() +
+  geom_smooth(method = 'lm') +
+  labs(x = 'Number of bacterial species', y = expression('Methane oxidation rate (-k)')))
+
+ggarrange(pmoa_plot_arcs, rich_plot_arcs,
+          labels = "AUTO",
+          ncol = 2, nrow = 1)
+
+ggsave(file = '../figures/trad_cors_arcs.pdf', width = 6, height = 3)
+
 # Table 1
 
 pmoa_fit <- lm(Low_final_k ~ pmoa_copy_num, data = as(sample_data(physeq), 'data.frame'))
@@ -96,7 +115,10 @@ rich_fit <- lm(Low_final_k ~ rich_breakaway, data = as(sample_data(physeq), 'dat
 summary(rich_fit)
 glance(rich_fit)
 
-rbind(tidy(pmoa_fit)[2, ], tidy(rich_fit)[2, ])
+model_out <- rbind(tidy(pmoa_fit)[2, ], tidy(rich_fit)[2, ])
+colnames(model_out) <- c('Term', 'Estimate', 'SE', 't-statistic', 'p-value')
+write.table(model_out, '../output/model_out.csv', sep = ",")
+
 
 # Pull out important axes an plot pairwise with axes proportional to var expl
 
@@ -181,6 +203,7 @@ ggarrange(W1, W2, W3,
           #align = 'v', 
           common.legend = TRUE))
 ggsave('../figures/pcoa_multi.pdf', width = 5*2, height = 4*3)
+ggsave('../figures/pcoa_multi_arcs.pdf', width = 10*3/4, height = 12*3/4)
 
 #' ## Collinearity
 
[…]

### 2019-10-18 manuscript/manuscript_rev1.Rmd
Incorporate reviewer comments

@@
 
+# To Do
+
+ - [ ] Soften the language surrounding what progress we have made in microbial
+   BEF - currently it is too harsh. Reference some of the newer
+   multi-functionality literature.
+ - [ ] Add an explanation of what figure two represents.
+ - [ ] Reframe the paper about how there is ample evidence that underground
+   biodiversity might play a role in ecosystem functioning. Add the caveat that
+   current approaches to incorporating microbes into earth system models don't
+   provide much increased variance explained. Use this to pitch an alternative
+   approach: learning from geno-pheno mapping to angostically search for taxa
+   that may regulate function upstream of the actual process being studied.
+
 # Abstract
@@
 important for determining the rates of ecosystem-scale functions. This
-uncertainty limits our ability to predict and manage crucial
-microbially mediated processes, such as nutrient loss and greenhouse gas
-emissions. Our lack of understanding stems from the relatively large
-diversity of microorganisms, the difficulty in directly identifying
-functional groups, and our limited ability to manipulate microbial community
-attributes. For this reason, we propose that integrating traditional
-biodiversity-ecosystem function research with ideas from genotype-phenotype
-mapping could provide the new perspective our discipline needs. We identify
-three insights from genotype-phenotype mapping that could be useful for
-microbial biodiversity-ecosystem function studies: the concept of “agnostic”
-mapping, the use of more powerful ways to account for multiple comparisons,
-and the incorporation of covariates into models of ecosystem function. We
-illustrate the potential for these approaches to elucidate microbial
+uncertainty limits our ability to predict and manage crucial microbially
+mediated processes, such as nutrient loss and greenhouse gas emissions. Our
+lack of understanding stems from the relatively large diversity of
+microorganisms, the difficulty in directly identifying functional groups, and
+our limited ability to manipulate microbial community attributes. For this
+reason, we propose that integrating traditional biodiversity-ecosystem function
+research with ideas from genotype-phenotype mapping could provide the new
+perspective our discipline needs. We identify two insights from
+genotype-phenotype mapping that could be useful for microbial
+biodiversity-ecosystem function studies: the concept of “agnostic” mapping and
+the incorporation of covariates into models of ecosystem function to identify
+microorganisms that uniquely contribute to ecosystem function. We illustrate
+the potential for these approaches to elucidate microbial
 biodiversity-ecosystem function relationships by analyzing a subset of
@@
 
-### A new perspective on microbial community-ecosystem function relationships is needed
+Ecology is broadly focused on understanding what creates the biodiversity we
+observe and how that biodiversity shapes the ecosystems that humans depend
+on. Earth's microbiome represents an important microcosm for that general
+pursuit as most of the major biogeochemical cycles are mediated by
+microorganisms. We have shown that certain ecosystem functions are positively
+correlated with microbial biodiversity. However, attempts to understand the
+mechanism underlying that relationship have borne little fruit. We argue that
+to overcome this challenge we should look for inspiration from population
+genetics and the process of genotype-phenotype mapping. In particular, we should look
+more agnostically at microbial
+communities to identify taxa that regulate ecosystem-scale functions. In
+addition, we should attempt to understand the complex covariance that exists
+between microbial communities, biogeographic patterns, and the underlying abiotic
+environment. Finally, we reframe the question "Is microbial biodiversity
+related to ecosystem function?" as "What is the unique contribution of the
+microbiome to ecosystem function independent of the environment?"
+
+### There is evidence that microbial diversity matters
+
+There is evidence that microbial communities should matter for regulating
+ecosystem function. Some studies have manipulated the connection between
+environmental factors and microbial community composition through common garden
+or reciprocal transplant experiments. These studies frequently report different
+rates of ecosystem functions for different microbial communities under the same
+environmental conditions [@reed2007]. This has been observed for plant
+decomposition [@glassman2018, @strickland2009], plant phenology
+[@panke-buisse2015], soil nitrogen cycling [@balser2005], and soil greenhouse
+gas emissions [@cavigelli2000], among other functions. Therefore, variation in microbial
+community composition can be associated with variation in ecosystem function.
+
+Studies using a comparative approach rather than a manipulative one have also
+demonstrated a link between microbial biodiversity and ecosystem function.  In
+general, there is a significant correlation between belowground biodiversity
+and multiple ecosystem functions [@jing2015, @delgado-baquerizo2016,
+@delgado-baquerizo2017, @wagg2014]. It is unclear whether the positive
+diversity-function relationship is due to the independent effect of variation
+in community composition on ecosystem function or due to environmental
+conditions that select for both diverse communities and high rates of ecosystem
+function. Determining why we see this diversity-function relationship depends
+on identifying the components of the community that are related to ecosystem
+function.
+
+### We need a new perspective on microbial biodiversity-ecosystem function relationships
+
+Most sudies of microbial community function focus on either functional gene
+abundance or community diversity.
+In the case of functional genes, qPCR or shotgun metagenomic sequencing is used to
+estimate the abundance of a gene or transcript that is a putative marker for a
+microbial process (and thus a marker for the functional group that performs
+that process). For example, the gene _mcrA_, which encodes a subunit of the
+enzyme that performs the final step in methanogenesis, is commonly used as a
+marker for methanogenesis and for the methanogen functional group. Other
+examples include _pmoA_ for methanotrophy, _nifH_ for nitrification, and _nosZ_ for
+denitrification. The abundance of these markers is often hypothesized to
+predict the rate of their associated processes (for example, the
+abundance of _mcrA_ is expected to relate to the rate of
+methanogenesis). There are examples where this relationship is present
+[@freitag2009, @freitag2010, @schnyder2018].  However, a review of such studies
+found that the abundance of a functional gene or transcript is rarely
+correlated with the rate of the corresponding process and depends on the
+function of interest, with most effects either negative or not significant
+[@rocca2015].
+
+The second aspect of microbial community structure hypothesized to predict
+ecosystem function is taxonomic or functional diversity. Diversity is either
+estimated from sequence variants of a barcode gene such as the 16S rRNA gene or
+manipulated through some proxy of diversity such as sequential dilution or
+varying filter sizes. Studies that experimentally manipulate diversity
+generally find a relationship between the applied diversity treatment and
+ecosystem function, including methanotrophy [@schnyder2018], phosphorus
+leaching [@wagg2014], greenhouse gas emissions [@wagg2014], decomposition
+[@maron2018], and nitrogen cycling [@philippot2013]. However, in many cases
+richness is confounded with other factors such as the presence or absence of
+major phylogenetic groups (nematodes, fungi) or abundance of microbial cells.
+In addition, assembled microbial communities risk confounding diversity with
+composition, for example by more likely including the most productive taxa in
+the highest diversity treatment. In general, the relationship between microbial
+diversity is positive and significant, but adds little variance over abiotic
+conditions or macroorganismal diversity [@graham2016, @jing2015].
+
+Despite the positive diversity-function relationship for microbes, models that
+incorporate microbial features only marginally improve predictions over
+environmental variables [@graham2016]. In addition, we've made little progress
+in identifyng the components of microbial communities that drive these
+diversity-function relationships [@rocca2015].
+Overall, functional gene abundance and community diversity improve models of
+ecosystem function less than one third of the time and increase variance
+explained by an average of only 8 percentage points [@graham2016].
+
+Up to this point, microbial biodiversity-ecosystem function research has
+followed a similar path to macrobial biodiversity-ecosystem function research,
+particularly compared to plant communities. However, there are important
+differences between plants and microbes that might inform how we pursue this
+research going forward. For example, microbial communities are more diverse and
+contain greater numbers of individuals than plant communities and are thus more
+difficult to sample comprehensively. Unlike plants, functional groups of
+microbes are rarely determined through observation or direct measurement, but
+rather must be indirectly inferred based on taxonomic assignment of marker
+genes (such as 16S rRNA genes), by sequencing functional genes from
+metagenomes, or by quantifying specific functional genes using qPCR.  This adds
+uncertainty in our ability to estimate the size of functional groups.  The
+ability to experimentally manipulate diversity or community composition is much
+more limited for microbes. Given all of these differences (and many others),
+the challenge of linking microbial communities to ecosystem functions may be
+less like that of plant communities and more analogous to the complex task of
+linking genomic variation to organismal phenotypes. For this reason, we propose
+that integrating traditional biodiversity-ecosystem function research with
+ideas from genotype-phenotype mapping could provide the new perspective our
+discipline needs.
+
+### Genotype-phenotype as a source of inspiration
+
+Given the relative lack of success to date, it is time to rethink how we
+approach the challenge of mapping microbial community structure to ecosystem
+function. Microbial ecologists are not the only biologists who are attempting
+to determine the relationship between a complex set of highly-variable data and
+an aggregate function. This kind of “many-to-one” mapping is analogous to the
+challenge of identifying the genetic basis of complex traits in organismal
+populations. In such genotype-phenotype mapping studies, a population
+exhibits variation in a phenotype (e.g., height or disease state) as well as
+variation in thousands of genetic loci or single nucleotide polymorphisms (SNPs).
+To identify the genetic basis for a trait, investigators sample from this
+population and correlate phenotype with genotype. In most cases, there are many
+more loci than individuals and we do not know whether the SNPs are causally
+linked or are simply in linkage disequilibrium with a causal mutation.
+
+There are a number of parallels between the challenge faced by organismal
+biologists and that facing microbial community ecologists. They are both
+“many-to-one” mapping challenges which involve large numbers of statistical
+comparisons. Both are attempting to identify causal relationships that are
+potentially confounded by complex patterns of covariation. For example,
+with local selection or distance-decay. Further, a positive association may
+either indicate a causal relationship or simply an association between a
+significant taxon and a causal taxon (i.e., microbial linkage
+disequilibrium). There is often no strong expectation about which entities
+(i.e., which genomic regions or which microbial taxa) are most likely to be
+causally-related to phenotype or function, and thus “agnostic” approaches are
+needed.  Both are sensitive to exactly how the “mapping” question is asked, and
+both ultimately require manipulation (of genes or taxa) to establish causation. We
+describe each of these parallels below in more detail and provide an example
+of how these ideas could be applied to microbial data.
+
+### The importance of a taxonomically “agnostic” approach
+
+[This is super redundant with the previous section.]
+
+The traditional approach to macroorganismal biodiversity-ecosystem function
+research is to measure or manipulate the diversity of a taxonomic group (e.g.,
+plants) and look for an association with the function performed by that group
+(e.g., primary productivity). In the broadest sense, we can think of plants as
+a “functional group,” i.e., a group of taxa united by their ability to perform
+a particular ecosystem function. Ecologists may further divide a functional
+group (e.g., plants) into smaller functional groupings (e.g., forbs), defined
+by the details of how they perform their particular ecosystem function.
+
+However, for microbes our knowledge of functional groups is much more limited.
+From a very limited number of cultured isolates we have a provisional
+understanding of which microbes might be involved in some ecosystem functions.
+And by sequencing the genomes of these isolates, we have identified genetic
+markers for certain functions. But most microbial taxa remain uncultured and we
+do not know the function of most microbial taxa detected in environmental
+samples [@hug2016, @martiny2019, @steen2019]. In addition, there have been
+recent discoveries of functions in unexpected taxonomic groups, for example
+methanogenesis by fungi and cyanobacteria, a function previously considered
+restricted to the Archaea [@bizic-ionescu2019, @lenhart2012]. Therefore,
+estimating microbial functional group abundance can be challenging.
+
+Beyond, for example, identifying previously undocumented methanotrophs that are
+related to methane oxidation, it's possible that more complex ecological
+regulatory patterns exist in microbial communities. In other words, it may
+be that methanotroph abundance does not limit methanotrophy in a given ecosystem.
+[Something about metabolic networks...]
+Instead, the rate
+of methane flux to the atmosphere could be controlled by methanotrophs,
+methanogens, bacteria that produce the fermentative byproducts that methanogens
+depend on, nutrient cyclers that regulate the availability of nitrogen or
+phosporus, or even grazers and other predators or pathogens. In other words, the regulation of
+ecosystem function by microbial communities could be direct or indirect, much
+like genetic regulation within an organism.
+
+Because of this, it would be prudent to look more agnostically at microbial
+communities to identify taxa or groups of taxa that are important for
+predicting the rate of ecosystem functions rather than assuming that the
+genetic markers we have provisionally identified for a given function represent
+the most likely taxa involved. This agnostic approach is analogous to the
+approach of many genotype-phenotype mapping studies (e.g., genome-wide
+association studies or GWAS), which often look for associations between a
+phenotype and loci anywhere in a genome.
+
+### Controlling for population stratification
+
+It is widely accepted that organisms, including microorganisms, exhibit
+population stratification due to geographic and environmental separation
+[@martiny2006, @wright1943]. Genome-wide association studies generally control
+for population stratification using shared ancestry among cases and controls when
+modeling the connection between genotype and phenotype. The classic example is
+the latitudinal gradient of both height and genotypic similarity in Europe,
+which results in spurious associations between human height and genetic
+variation [@berg2019, @novembre2008, @sohail2019]. Perhaps a more illustrative
+example is the correlation between chopstick usage and certain genetic markers
+in Asian populations, which is unlikely to be due to an actual genetic basis
+for chopstick skills. Spurious associations arise
+because of heterogeneity in population structure where any genetic variant in a
+population with an extreme phenotype will tend to be associated with that
+phenotype whether it is related or not.
+
+Genome-wide association studies correct for stratification using a variety of methods.
+Generally, they ignore the underlying environmental and spatial
+distance between samples and instead use shared ancestry as a proxy for local
+selection and assortative mating.  A common approach is to perform a regression
+of phenotype and shared ancestry (computed as the first one or more principal
+components of the genotype matrix) and then use the residuals from this model
+as the values for phenotype in a
+subsequent regression using the genotypes directly [@price2006]. This principal
+component analysis is designed to test the effect of individual genes after
+removing shared ancestry among individuals. Another approach, employed in our
+example, is variance component modeling (or mixed modeling, heirarchical
+modeling, etc.) where genotypic similarity is included as a covariate in the
+model [@kang2010].
+
+While these represent two common statistical approaches to
+control for population stratification, it is also considered good practice to
+control for stratification experimentally. This can be accomplished by 
+selectng a study population that is relatively homogeneous. Otherwise, even
+large, well-replicated GWAS studies can result in
+false positives. For example, it was recently revealed that GWAS studies on
+height of Europeans failed to adequately control for stratification. When their
+results were confirmed using a similarly sized, but more homogeneous study
+population, many of the genetic markers for height were no longer significant [@berg2019,
+@sohail2019].
+
+[This paragraph needs work]
+
+Typically, microbial biodiversity-ecosystem function studies do not account for
+population stratification (i.e., community similarity across samples), although
+there are some exceptions [@meyer2019, @qin2012, @lloyd-price2019].  However,
+community similarity (the community analogue of shared ancestry among
+organisms) is not as tightly linked to geography or environment as is shared
+ancestry among organisms, and it could be very useful to account for these
+separately in microbial studies, especially if one is particularly interested
+in how composition alters function independent of the underlying environmental
+variation.
+
+### What might this look like/More about the analogy
+
+The relationship between structure and function could be complex and indirect.
+i.e. not just methanogens making methane, but metabolic and regulatory networks
+
+A significant correlation between a taxon and a function could mean a number of
+things.
+1. That microbe is performing that function right now and its abundance is
+   related to function
+2.
+
+This whole compositional data/relative abundance thing also seems problematic
+for correlating abundance with function. Really what we're saying is "this read is differentially abundant in low vs. high flux ecosystems." Not abundance is _correlated_ with flux.
+1. It's not **abundance**.
+2. An increase in relative abundance in one taxon implies nothing about the
+   abundance of that taxon
+
+   wootincrease in relative abundance in one taxon implies nothing about the
+abun
+3. It is perhaps somewhat more likely that
+
+### We need a new perspective on microbial community-ecosystem function relationships
+
+Many of the topics in this first section are covered later (e.g. in diversity
+matters). What is the purpose of this section - how does it move the story
+along? It should motivate this work and put it in a larger context. This is the
+hook. Maybe the best way to do this is introduce what we know, then create a
+conflict by showing what we don't know, and then give a brief read map of how
+we're going to fix that. Possibly send these 2-3 paragraphs to Bebo. Maybe also
+read some writing advice on hooks (maybe in Schimel's book).
 
@@
 
+Numerous studies have attempted to link microbial diversity or functional group
+abundance  to ecosystem function. In general, there is a significant correlation between
+belowground biodiversity and multiple ecosystem functions [@jing2015,
+@delgado-baquerizo2016, @delgado-baquerizo2017, @wagg2014]. However, models that incorporate microbial features
+only marginally improve predictions overall environmental variables [@graham2016]. In addition, we've made
+little progress in identifyng the components of microbial communities that
+drive these diversity-function relationships [@rocca2015]. It is unclear
+whether the positive diversity-function relationship is due to the independent
+effect of variation in community composition on ecosystem function or due to
+environmental conditions that select for both diverse communities and high
+rates of ecosystem function. Determining why we see this diversity-function
+relationship depends on identifying the components of the community that are
+related to ecosystem function independent of the environmental variation.
+
+I don't love the below paragraph
+
 Our lack of understanding stems from many sources. Microbial communities are
-more diverse and contain greater numbers of individuals than plant communities and are thus more
+more diverse and contain greater numbers of individuals than plant communities
+and are thus more
 difficult to sample comprehensively. Unlike plants, functional groups of
@@
 
-### There is evidence that microbial diversity can matter
+### There is evidence that microbial diversity matters
+
+Perhaps remove this paragraph. Maybe add an explanation of "independent of the
+environment" slash "uniquely attributed"
 
@@
 
-However, there is evidence that this is not always the case, A limited number
-of studies have manipulated the connection between environmental factors and
+However, there is evidence that this is not always the case [super awkward]. Some studies have
+manipulated the connection between environmental factors and
 microbial community composition through “common garden” or reciprocal
@@
 do not know the function of most microbial taxa detected in environmental
-samples [@hug2016, @steen2019]. In addition, there have been recent discoveries
+samples [@hug2016] (@martiny2019 or @steen2019). In addition, there have been
+recent discoveries
 of functions in unexpected taxonomic groups, for example methanogenesis by
 fungi and cyanobacteria, a function previously considered restricted to the
-Archaea [@bizic-ionescu2019, @lenhart2012]. Because of this, it would be
+Archaea [@bizic-ionescu2019, @lenhart2012].
+
+Beyond, for example, identifying previously undocumented methanotrophs that are
+related to methane oxidation, it's possible that more complex ecological
+regulatory patterns exist in microbial communities. In other words, it may not
+be methanotrophs limiting methanotrophy in a given ecosystem. Instead, the rate
+of methane flux to the atmosphere could be controlled by methanotrophs,
+methanogens, bacteria that produce the fermentative byproducts that methanogens
+depend on, nutrient cyclers that regulate the availability of nitrogen or
+phosporus, or even grazers and other predators or pathogens. The regulation of ecosystem function by microbial communities could be direct or indirect.
+
+Because of this, it would be
 prudent to look more agnostically at microbial communities to identify taxa or
@@
 
-### Dealing with multiple comparisons and covariation
+### Dealing with multiple comparisons
 
@@
 
+### Controlling for population stratification
+
+Maybe introduce "uniquely associated" here.
+
 It is widely accepted that organisms, including microorganisms, exhibit
@@
 which results in spurious associations between human height and genetic
-variation [@berg2019, @novembre2008, @sohail2019]. To correct for this
+variation [@berg2019, @novembre2008, @sohail2019]. Spurious associations arise
+because of heterogeneity in population structure where any genetic variant in a
+population with an extreme phenotype will tend to be associated with that
+phenotype whether it is related or not. To correct for this
 covariance structure, GWAS models incorporate genotypic similarity to correct
-for shared ancestry using a variety of methods, such as principal component
-correction or variance component modeling [@kang2010, @price2006]. Typically,
-microbial biodiversity-ecosystem function studies do not account for population
-stratification (i.e., community similarity across samples), although there are
-some exceptions [@meyer2019, @qin2012]. Genome-wide association studies
-generally ignores the underlying environmental and spatial distance between
-samples and instead uses shared ancestry as a proxy for these variables.
-However, community similarity (the community analogue of shared ancestry among
-organisms) is not as tightly linked to geography or environment as is shared
-ancestry among organisms, and it could be very useful to account for these
-separately in microbial studies, especially if one is particularly interested
-in how composition alters function independent of the underlying environmental
-variation.
+for shared ancestry using a variety of methods. A common example is to correct
+for population structure by performing regression of the first one or more
+principal components of genotype on both genotype and phenotype. Then, the
+residuals from this model are used as the new genotype and phenotype with
+structure removed [@price2006]. Another approach, employed in our example, is
+variance component
+modeling (or mixed modeling, heirarchical modeling, etc.) where genotypic
+similarity is included as a covariate in the model as a random
+effect [@kang2010]. While these represent two common statistical approaches to
+control for population stratification, it is also considered good practice to
+select a study population that is relatively homogeneous when designing an
+experiment. Otherwise, even large, well-replicated GWAS studies can result in
+false positives, as recently revealed by two studies reversing conclusions from
+previous studies of the relationship between genetics and human height
+[@berg2019, @sohail2019].
+
+Typically, microbial biodiversity-ecosystem
+function studies do not account for population stratification (i.e., community
+similarity across samples), although there are some exceptions [@meyer2019,
+@qin2012]. Genome-wide association studies generally ignore the underlying
+environmental and spatial distance between samples and instead uses shared
+ancestry as a proxy for these variables.  However, community similarity (the
+community analogue of shared ancestry among organisms) is not as tightly linked
+to geography or environment as is shared ancestry among organisms, and it could
+be very useful to account for these separately in microbial studies, especially
+if one is particularly interested in how composition alters function
+independent of the underlying environmental variation.
 
@@
 
-Methane oxidation rate was not significantly correlated with pmoA gene
-abundance or richness (Table 1, Figure 1). We tested collinearity between each
+## Results & Discussion
+
+Methane oxidation rate was not significantly correlated with the abundance of
+methanotrophs (inferred from _pmoA_ gene
+abundance) or 16S read richness (Table 1, Figure 1). We tested collinearity between each
 pair of distance matrices for community, environment, and geography using
@@
 diversity of samples separated by site ID and by ecosystem type (wetland or
-upland; Figure 2). Finally, we tested the effect of the relative abundance of
+upland; Figure 2). This indicates substantial spatial and environmental
+structure for microbial populations which should be controlled for to
+rigorously identify associations between taxa and function. Finally, we tested
+the effect of the relative abundance of
 each ASV on methane oxidation rate. We found different numbers of taxa
 significantly associated with methane oxidation depending on which covariates
-were included in the model (Table 3).  In particular, 460 taxa were identified
-with no covariates whereas 6 were identified after including all covariates.
+were included in the model (Table 3). After correcting for the false discovery
+rate, 460 unique 16S reads were identified as significantly correlated with function.
+In the model that controlled for community structure using all three covariates
+(community, environment, and geography), only 6 of these reads were found to be
+significiantly correlated with methane oxidation rate.
 Though we cannot infer function from 16S sequences, these 6 taxa fall into
@@
 consume methane [@belova2009, @chang2011, @domeignoz-horta2019, @fritz2004].
-Their effect on methane oxidation rate ranges from 0.5 to 1.5 (Figure 3).  An
+Their effect on methane oxidation rate ranges from 0.5 to 1.5 (Figure 3). (The
+slope of the relationship between relative abundance and methane oxidation
+rate.) An
 effect of 1.5 means that for a one unit increase in relative abundance there is
@@
 over time).
-
-## Discussion
 
@@
 structure-function connections.
+
+[redundant with the former results section]
 
@@
 also a more appropriate one, especially if one is interested in how to
-incorporate microbial community data into ecosystem models. For modeling what
-is usually important is to identify those taxa that add explanatory power
-beyond that provided by other factors (such as environmental conditions).
+incorporate microbial community data into ecosystem models. These models depend
+on identifying taxa that add explanatory power
+beyond that provided by other factors, such as environmental conditions.
 
@@
 
-### Conclusion
+## Conclusion
 

### 2019-10-18 manuscript/manuscript_rev1b.Rmd
Incorporate reviewer comments

--- 2019-10-18 manuscript/manuscript_rev1.Rmd
@@
----
-title: "Linking microbial communities to ecosystem functions: what we can learn from genotype-phenotype mapping in organisms"
-output: 
-    word_document:
-        reference_docx: reference.docx
-bibliography: library.bib
-csl: philosophical-transactions-of-the-royal-society-b.csl 
----
+## Outline
 
-```{r setup, include=FALSE}
+### We need a new perspective on microbial community-ecosystem function relationships
 
-library(tidyverse)
-library(knitr)
-library(flextable)
-opts_chunk$set(echo = FALSE)
+### There is evidence the microbial diversity matters
 
-```
+### Current approaches to quantify this relationship have provided only minor improvements
 
-# To Do
+### Genotype-phenotype as a source of inspiration
 
- - [ ] Soften the language surrounding what progress we have made in microbial
-   BEF - currently it is too harsh. Reference some of the newer
-   multi-functionality literature.
- - [ ] Add an explanation of what figure two represents.
- - [ ] Reframe the paper about how there is ample evidence that underground
-   biodiversity might play a role in ecosystem functioning. Add the caveat that
-   current approaches to incorporating microbes into earth system models don't
-   provide much increased variance explained. Use this to pitch an alternative
-   approach: learning from geno-pheno mapping to angostically search for taxa
-   that may regulate function upstream of the actual process being studied.
+### A taxonomically "agnostic" approach
 
-# Abstract
+### Controlling for population stratification/the effect of the microbiome independent of the environment
 
-Microorganisms mediate many important ecosystem functions, yet it remains
-unclear to what extent microbial diversity or community composition is
-important for determining the rates of ecosystem-scale functions. This
-uncertainty limits our ability to predict and manage crucial microbially
-mediated processes, such as nutrient loss and greenhouse gas emissions. Our
-lack of understanding stems from the relatively large diversity of
-microorganisms, the difficulty in directly identifying functional groups, and
-our limited ability to manipulate microbial community attributes. For this
-reason, we propose that integrating traditional biodiversity-ecosystem function
-research with ideas from genotype-phenotype mapping could provide the new
-perspective our discipline needs. We identify two insights from
-genotype-phenotype mapping that could be useful for microbial
-biodiversity-ecosystem function studies: the concept of “agnostic” mapping and
-the incorporation of covariates into models of ecosystem function to identify
-microorganisms that uniquely contribute to ecosystem function. We illustrate
-the potential for these approaches to elucidate microbial
-biodiversity-ecosystem function relationships by analyzing a subset of
-published data measuring methane oxidation rates from incubations of tropical
-soil. We assert that combining the approaches of traditional
-biodiversity-ecosystem function research with ideas from genotype-phenotype
-mapping will not only generate novel hypotheses about how complex microbial
-communities drive ecosystem function, but also help scientists predict and
-manage changes to ecosystem functions resulting from human activities.
-
-## Introduction
+## Intro
 
@@
 3. It is perhaps somewhat more likely that
-
-### We need a new perspective on microbial community-ecosystem function relationships
-
-Many of the topics in this first section are covered later (e.g. in diversity
-matters). What is the purpose of this section - how does it move the story
-along? It should motivate this work and put it in a larger context. This is the
-hook. Maybe the best way to do this is introduce what we know, then create a
-conflict by showing what we don't know, and then give a brief read map of how
-we're going to fix that. Possibly send these 2-3 paragraphs to Bebo. Maybe also
-read some writing advice on hooks (maybe in Schimel's book).
-
-Ecologists have long investigated the effects of changing biodiversity on
-ecosystem function, documenting, for example, relationships between terrestrial
-plant community richness and primary productivity [@hooper2005]. The study of
-the relationship between microbial biodiversity and ecosystem functions is much
-more recent and has generated mixed results [@schimel1998, @singh2010]. Despite
-the fact that microbes mediate many important ecosystem functions, it remains
-maddeningly unclear to what extent microbial diversity or community composition
-is important for determining the rates of ecosystem-scale functions.  For
-example, numerous studies have attempted to correlate microbial functional
-group abundance or diversity with the rate of various ecosystem functions and
-most are unsuccessful [@rocca2015, @graham2016]. This uncertainty limits our
-ability to predict and manage crucial microbially mediated ecosystem functions,
-such as nutrient loss and greenhouse gas emissions. As a discipline, we need a
-new perspective on these questions.
-
-Numerous studies have attempted to link microbial diversity or functional group
-abundance  to ecosystem function. In general, there is a significant correlation between
-belowground biodiversity and multiple ecosystem functions [@jing2015,
-@delgado-baquerizo2016, @delgado-baquerizo2017, @wagg2014]. However, models that incorporate microbial features
-only marginally improve predictions overall environmental variables [@graham2016]. In addition, we've made
-little progress in identifyng the components of microbial communities that
-drive these diversity-function relationships [@rocca2015]. It is unclear
-whether the positive diversity-function relationship is due to the independent
-effect of variation in community composition on ecosystem function or due to
-environmental conditions that select for both diverse communities and high
-rates of ecosystem function. Determining why we see this diversity-function
-relationship depends on identifying the components of the community that are
-related to ecosystem function independent of the environmental variation.
-
-I don't love the below paragraph
-
-Our lack of understanding stems from many sources. Microbial communities are
-more diverse and contain greater numbers of individuals than plant communities
-and are thus more
-difficult to sample comprehensively. Unlike plants, functional groups of
-microbes are rarely determined through observation or direct measurement, but
-rather must be indirectly inferred based on taxonomic assignment of marker
-genes (such as 16S rRNA genes), by sequencing functional genes
-from metagenomes, or by quantifying specific functional genes using qPCR. This adds uncertainty in our ability to estimate the size of functional groups.
-The ability to experimentally manipulate microbial diversity or community
-composition is much more limited for microbes. Given all of these differences
-(and many others), the challenge of linking microbial communities to ecosystem
-functions may be less like that of plant communities, and more analogous to the
-complex task of linking genomic variation to organismal phenotypes. For this
-reason, we propose that integrating traditional biodiversity-ecosystem function
-research with ideas from genotype-phenotype mapping could provide the new
-perspective our discipline needs.
-
-### There is evidence that microbial diversity matters
-
-Perhaps remove this paragraph. Maybe add an explanation of "independent of the
-environment" slash "uniquely attributed"
-
-Perhaps our uncertainty arises because there is no general relationship between
-microbial biodiversity and any particular ecosystem function. For example, if
-local selection always optimizes the available microbial biodiversity to
-maximize ecosystem function, then microbial community composition should not
-matter for predicting the rates of microbially mediated functions. Instead, the
-rates of these functions should be determined by the underlying environmental
-variation. In this case, the microbial community would simply act as a conduit
-through which the abiotic environment alters ecosystem function. Attempts to
-document a general relationship between microbial community attributes and
-function would then fail (since the primary drivers are environmental).
-
-However, there is evidence that this is not always the case [super awkward]. Some studies have
-manipulated the connection between environmental factors and
-microbial community composition through “common garden” or reciprocal
-transplant experiments, and they frequently report different rates of ecosystem
-functions for different microbial communities under the same environmental
-conditions [@reed2007].  This has been observed for plant decomposition
-[@glassman2018, @strickland2009], plant phenology [@panke-buisse2015], soil
-nitrogen cycling [@balser2005], and soil greenhouse gas emissions
-[@cavigelli2000], among others. Therefore, variation in microbial community
-composition can be associated with variation in ecosystem function, independent
-of environmental variation. So why has this been difficult to consistently
-document in the field?
-
-### Traditional approaches to quantifying this relationship have provided only minor improvements
-
-Most comparative studies of microbial community function in the field focus on
-one of two aspects of microbial community structure that are hypothesized to
-predict ecosystem function. The first aspect is “functional” gene or transcript
-abundance.  In this case, qPCR or shotgun metagenomic sequencing is used to
-estimate the abundance of a gene or transcript that is a putative marker for a
-microbial process (and thus a marker for the functional group that performs
-this process). For example, the gene mcrA, which encodes a subunit of the
-enzyme that performs the final step in methanogenesis, is commonly used as a
-marker for methanogenesis and for the methanogen functional group. Other
-examples include pmoA and methanotrophy, nifH and nitrification, and nosZ and
-denitrification. The abundance of these markers is often hypothesized to be
-predictive of the rate of their associated processes (for example, the
-abundance of mcrA is often hypothesized to be related to the rate of
-methanogenesis). There are examples where this relationship is present
-[@freitag2009, @freitag2010, @schnyder2018].  However, a review of such studies
-found that the abundance of a functional gene or transcript is rarely
-correlated with the rate of the corresponding process and depends on the
-function of interest, with most effects either negative or not significant
-[@rocca2015].
-
-The second aspect of microbial community structure hypothesized to predict
-ecosystem function is taxonomic or functional diversity. Diversity is either
-estimated from sequence variants of a barcode gene such as the 16S rRNA gene or
-manipulated through some proxy of diversity such as sequential dilution or
-varying filter sizes. Studies that experimentally manipulate diversity
-generally find a relationship between the applied diversity treatment and
-ecosystem functions, including for methanotrophy [@schnyder2018], phosphorus
-leaching [@wagg2014], greenhouse gas emissions [@wagg2014], decomposition
-[@maron2018], and nitrogen cycling [@philippot2013]. However, in many cases,
-richness is confounded with other factors such as the presence or absence of
-major phylogenetic groups (nematodes, fungi) or abundance of microbial cells.
-In addition, assembled microbial communities pose the same challenges as
-macroorganismal diversity experiments, for example the highest diversity
-treatment will more likely contain the most productive taxon. A review of
-microbial diversity studies shows that microbial community taxonomic and
-functional diversity add little explanatory power to models of ecosystem
-function [@graham2016]. Overall, functional gene abundance and community
-diversity improve models of ecosystem function less than one third of the time
-and increase variance explained by an average of only 8 percentage points
-[@graham2016].
-
-### Genotype-phenotype mapping as a source of inspiration
-
-Given the relative lack of success to date, it is time to rethink how we
-approach the challenge of mapping microbial community structure to ecosystem
-function.  Microbial ecologists are not the only biologists who are attempting
-to determine the relationship between a complex set of highly-variable data and
-an aggregate function.  This kind of “many-to-one” mapping is analogous to the
-challenge of identifying the genetic basis of complex traits in organismal
-populations. In such “genotype-phenotype” mapping studies, a population
-exhibits variation in a phenotype (e.g., height or disease state) as well as
-variation in potentially thousands of single nucleotide polymorphisms (SNPs).
-To identify the genetic basis for a trait, investigators sample from this
-population and correlate phenotype with genotype. In most cases, there are many
-more loci than individuals and we do not know whether the SNPs are causally
-linked or are simply in linkage disequilibrium with a causal mutation.
-
-There are a number of parallels between this challenge faced by organismal
-biologists and that facing microbial community ecologists.  They are both
-“many-to-one” mapping challenges, which involve large numbers of statistical
-comparisons.  Both are attempting to identify causal relationships that are
-potentially confounded by very complex patterns of covariation.  There is often
-no strong expectation about which entities (i.e., which genomic regions or which
-microbial taxa) are most likely to be causally-related to phenotype or
-function, and thus “agnostic” approaches are needed.  Both are sensitive to
-exactly how the “mapping” question is asked, and ultimately require
-manipulation (of genes or taxa) to establish causation.  We describe each of
-these parallels below in more detail, and provide an example of how these ideas
-could be applied to microbial data.
-
-### The importance of a taxonomically “agnostic” approach
-
-The traditional approach to microorganismal biodiversity-ecosystem function
-research is to measure or manipulate the diversity of a taxonomic group (e.g.,
-plants) and look for an association with the function performed by that group
-(e.g., primary productivity).  In the broadest sense, we can think of plants as
-a “functional group,” i.e., a group of taxa united by their ability to perform
-a particular ecosystem function. Ecologists may further divide a functional
-group (e.g., plants) into smaller functional groupings (e.g., forbs), defined
-by the details of how they perform their particular ecosystem function.
-
-However, for microbes our knowledge of functional groups is much more limited.
-From a very limited number of cultured isolates we have a provisional
-understanding of which microbes might be involved in some ecosystem functions.
-And by sequencing the genomes of these isolates, we have identified genetic
-markers for certain functions. But most microbial taxa remain uncultured and we
-do not know the function of most microbial taxa detected in environmental
-samples [@hug2016] (@martiny2019 or @steen2019). In addition, there have been
-recent discoveries
-of functions in unexpected taxonomic groups, for example methanogenesis by
-fungi and cyanobacteria, a function previously considered restricted to the
-Archaea [@bizic-ionescu2019, @lenhart2012].
-
-Beyond, for example, identifying previously undocumented methanotrophs that are
-related to methane oxidation, it's possible that more complex ecological
-regulatory patterns exist in microbial communities. In other words, it may not
-be methanotrophs limiting methanotrophy in a given ecosystem. Instead, the rate
-of methane flux to the atmosphere could be controlled by methanotrophs,
-methanogens, bacteria that produce the fermentative byproducts that methanogens
-depend on, nutrient cyclers that regulate the availability of nitrogen or
-phosporus, or even grazers and other predators or pathogens. The regulation of ecosystem function by microbial communities could be direct or indirect.
-
-Because of this, it would be
-prudent to look more agnostically at microbial communities to identify taxa or
-groups of taxa that are important for predicting the rate of ecosystem
-functions rather than assuming that the genetic markers we have provisionally
-identified for a given function represent the most likely taxa involved.  This
-agnostic approach is analogous to the approach of many genotype-phenotype
-mapping studies (e.g., genome-wide association studies or GWAS), which often
-look for associations between a phenotype and loci anywhere in a genome.
-
-### Dealing with multiple comparisons
-
-Microbial biodiversity-ecosystem function studies and genotype-phenotype
-mapping studies are both “many-to-one” mapping challenges, which involve large
-numbers of statistical tests.  Hundreds to thousands of tests are routinely
-made per study, greatly inflating the number of false positives identified
-using statistical hypothesis testing approaches.  Some microbial
-biodiversity-ecosystem function studies do not correct for this, while others
-use approaches that may unnecessarily inflate the false negative rate, such as
-the Bonferroni correction. The Bonferroni correction has been widely considered
-to be too conservative, particular for exploratory studies designed to generate
-hypotheses [@perneger1998, @rothman1990, @thomas1985]. Statisticians have
-developed a number of less-conservative approaches to correct for multiple
-comparisons by controlling the false discovery rate in order to balance the
-tradeoff between Type I and Type II errors [@benjamini1995, @storey2002].
-These approaches are commonly used in genotype-phenotype mapping studies.
-
-### Controlling for population stratification
-
-Maybe introduce "uniquely associated" here.
-
-It is widely accepted that organisms, including microorganisms, exhibit
-population stratification due to geographic and environmental separation
-[@martiny2006, @wright1943]. Genome-wide association studies generally account
-for population structure due to shared ancestry among cases and controls when
-modeling the connection between genotype and phenotype. The classic example is
-the latitudinal gradient of both height and genotypic similarity in Europe,
-which results in spurious associations between human height and genetic
-variation [@berg2019, @novembre2008, @sohail2019]. Spurious associations arise
-because of heterogeneity in population structure where any genetic variant in a
-population with an extreme phenotype will tend to be associated with that
-phenotype whether it is related or not. To correct for this
-covariance structure, GWAS models incorporate genotypic similarity to correct
-for shared ancestry using a variety of methods. A common example is to correct
-for population structure by performing regression of the first one or more
-principal components of genotype on both genotype and phenotype. Then, the
-residuals from this model are used as the new genotype and phenotype with
-structure removed [@price2006]. Another approach, employed in our example, is
-variance component
-modeling (or mixed modeling, heirarchical modeling, etc.) where genotypic
-similarity is included as a covariate in the model as a random
-effect [@kang2010]. While these represent two common statistical approaches to
-control for population stratification, it is also considered good practice to
-select a study population that is relatively homogeneous when designing an
-experiment. Otherwise, even large, well-replicated GWAS studies can result in
-false positives, as recently revealed by two studies reversing conclusions from
-previous studies of the relationship between genetics and human height
-[@berg2019, @sohail2019].
-
-Typically, microbial biodiversity-ecosystem
-function studies do not account for population stratification (i.e., community
-similarity across samples), although there are some exceptions [@meyer2019,
-@qin2012]. Genome-wide association studies generally ignore the underlying
-environmental and spatial distance between samples and instead uses shared
-ancestry as a proxy for these variables.  However, community similarity (the
-community analogue of shared ancestry among organisms) is not as tightly linked
-to geography or environment as is shared ancestry among organisms, and it could
-be very useful to account for these separately in microbial studies, especially
-if one is particularly interested in how composition alters function
-independent of the underlying environmental variation.
-
-## An example: high-affinity methane oxidation
-
-To illustrate the ideas outlined above, we reanalyzed a subset previously
-published data [@meyer2019] using a modified procedure from the original
-version.  A full description of the study design, samples, and data generation
-can be found in that manuscript. Briefly, these data were gathered from intact
-soil cores taken from diverse ecosystems of the Congo Basin in Gabon, Africa.
-Cores were incubated in the laboratory under different concentrations of
-methane to identify the rates of specific methane cycling pathways. For this
-example, we will analyze data from just one of these pathways, high-affinity
-methane oxidation (the oxidation of atmospheric concentrations of methane),
-which we will refer to below as “methane oxidation”. In addition, for
-simplicity we only include amplicon sequences from the DNA-inferred community
-and not the RNA-inferred community, both of which are presented in the original
-paper [@meyer2019].  The data we analyzed include methane oxidation
-measurements, amplicon sequence variants (ASVs; @callahan2017) generated using
-DADA2 and inferred from 16S rRNA gene sequences, pmoA abundance estimates (via
-qPCR), latitude and longitude, and four environmental covariates (soil
-moisture, bulk density, carbon, and nitrogen).
-
-Analyses were conducted in the R statistical environment using the phyloseq
-package [@mcmurdie2013, @rcoreteam2019].  The relative abundances of ASVs were
-corrected using the variance stabilizing transformation from DESeq2 [@love2014,
-@mcmurdie2014]. We first test typical measures of microbial structure including
-functional gene abundance and community richness, which was estimated using the
-breakaway package [@willis2015]. We then demonstrate significant covariation
-between community structure (estimated as Bray-Curtis distance using vegan),
-environmental variation (euclidean distance), and geographic distance
-(euclidean distance) using Mantel tests [@bray1957, @oksanen2019].  Finally, we
-present one approach to identifying taxa which are significantly associated
-with function independent of the environment by fitting variance component
-models using varComp to test the relationship between relative abundance of
-each ASV and methane oxidation rate [@qu2013]. To illustrate how including
-different covariates (environmental, geographic, and community) can result in
-different conclusions about which taxa are associated with function, we fit
-this model with and without random effects variance components for
-environmental similarity, geographic site ID, and Bray-Curtis similarity.
-Significant taxa were determined by controlling the false discovery rate at
-q-value < 0.05 [@storey2002]. Figures were created using ggplot2
-[@wickham2019].
-
-## Results & Discussion
-
-Methane oxidation rate was not significantly correlated with the abundance of
-methanotrophs (inferred from _pmoA_ gene
-abundance) or 16S read richness (Table 1, Figure 1). We tested collinearity between each
-pair of distance matrices for community, environment, and geography using
-Mantel tests and estimated p-values by permutation. We found a moderate and
-significant correlation between community composition and environmental
-variation, geography and community composition, and geography and environmental
-variation (Table 2, Figure 2). Principal coordinate plots show that beta
-diversity of samples separated by site ID and by ecosystem type (wetland or
-upland; Figure 2). This indicates substantial spatial and environmental
-structure for microbial populations which should be controlled for to
-rigorously identify associations between taxa and function. Finally, we tested
-the effect of the relative abundance of
-each ASV on methane oxidation rate. We found different numbers of taxa
-significantly associated with methane oxidation depending on which covariates
-were included in the model (Table 3). After correcting for the false discovery
-rate, 460 unique 16S reads were identified as significantly correlated with function.
-In the model that controlled for community structure using all three covariates
-(community, environment, and geography), only 6 of these reads were found to be
-significiantly correlated with methane oxidation rate.
-Though we cannot infer function from 16S sequences, these 6 taxa fall into
-three genera and one class with cultured representatives that are not known to
-consume methane [@belova2009, @chang2011, @domeignoz-horta2019, @fritz2004].
-Their effect on methane oxidation rate ranges from 0.5 to 1.5 (Figure 3). (The
-slope of the relationship between relative abundance and methane oxidation
-rate.) An
-effect of 1.5 means that for a one unit increase in relative abundance there is
-a 1.5 increase in k (the rate of exponential decrease in methane concentration
-over time).
-
-In this paper, we argue that the traditional approach to microbial
-biodiversity-ecosystem function research often ignores the complexity of
-microbial community dynamics, in particular the complex patterns of covariation
-that often arise among microbial community similarity, environmental
-similarity, and spatial proximity.  We propose that a better approach would be
-to learn from other complex many-to-one mapping problems in biology,
-particularly genotype-phenotype mapping.  These studies often explicitly
-account for covariation, they correct for multiple comparisons in powerful
-ways, and they frequently take an “agnostic approach” that does not assume a
-particular relationship between structure and function.
-
-We illustrate the potential for these approaches to elucidate microbial
-biodiversity-ecosystem function relationships by analyzing a subset of
-published data from incubations of tropical soil.  In this example, soil cores
-from different ecosystems exhibited different rates of methane oxidation, but
-simple correlations between these rates and the diversity or abundance of
-putative methane oxidizing bacteria were not informative.  These results agree
-with recent reviews of the literature that demonstrate that such simple
-approaches are often not a fruitful avenue for elucidating microbial
-structure-function connections.
-
-[redundant with the former results section]
-
-We then asked if the relative abundance of particular taxa were related to
-ecosystem function, and we identified 460 taxa whose relative abundances were
-associated with rates of methane oxidation after controlling the false
-discovery rate (q-value < 0.05).  These taxa could be related to ecosystem
-function in multiple ways.  The most interesting possibility is that each of
-these taxa is statistically related because it is causally connected to the
-function. This could be direct, for example an organism that consumes methane,
-or indirect, for example an organism that regulates substrates necessary for
-consumers of methane. In either of these cases, the taxon could be useful as a
-biomarker of function or as an organism to investigate in order to better
-understand the biological drivers of variation in methane oxidation.
-
-Alternatively, a significant association could occur for non-causal reasons.
-For example, any organism that tends to be in high abundance where methane
-oxidation rates are high would be correlated with methane oxidation, even if it
-has no causal relationship. This could be because such an organism is favored
-under the same environmental conditions that favor methane oxidizing bacteria
-(or that favor methane oxidation in general).  Such covariation can drive
-associations that are not causal, but the effects of such covariation can be
-reduced through a number of approaches (many developed by biologists interested
-in genotype-phenotype mapping).
-
-In our example, we showed that the abundances of microbial taxa exhibit complex
-patterns of covariation with each other and with environmental conditions and
-spatial location. These biogeographic patterns might be even stronger if we
-could sample the populations more intensively [@meyer2018]. Once we account for this
-covariation, our list of taxa associated with function was reduced to 6.  These
-6 taxa include taxa that are not known to directly contribute to methane
-oxidation, suggesting that the drivers of methane oxidation may be indirect,
-perhaps mediated by ecological interactions among taxa from multiple functional
-groups.
-
-This approach is powerful because it accounts for the covariances that arise
-from the fundamental ecological processes that drive community assembly and
-underlie biogeographical patterns [@martiny2006, @vellend2010]. It is also powerful because it is
-focused on a much more specific version of the “biodiversity-ecosystem
-function” question, i.e., it asks “which taxa are uniquely associated with
-function?” By “uniquely associated,” we mean those taxa associated with
-function irrespective of environmental conditions, local community structure,
-or spatial proximity.  This is not only a more specific question than is
-usually asked in microbial biodiversity-ecosystem function studies, but it is
-also a more appropriate one, especially if one is interested in how to
-incorporate microbial community data into ecosystem models. These models depend
-on identifying taxa that add explanatory power
-beyond that provided by other factors, such as environmental conditions.
-
-We identified three insights from genotype-phenotype mapping that could be
-useful for microbial biodiversity-ecosystem function studies: the concept of
-“agnostic” mapping, the use of more powerful ways to account for multiple
-comparisons, and the incorporation of covariates.  But there are other insights
-to be gained as well.  For example, some phenotypes (e.g., the propensity for
-diseases such as Parkinson's) are controlled by a single genetic locus [@kerem1989, @macdonald1992].
-However, most phenotypes studied to date arise from the influence of many loci
-of small effect (as well as environmental factors; [@hindorff2009, reich2001]). Organismal
-biologists have developed approaches tailored to identifying such loci; for
-example, by sampling organismal populations in a way that constrains genetic
-variation unrelated to phenotype. Similarly, for some ecosystem functions, it
-is possible that a single microbial taxon could substantially influence its
-rate. For example, methane flux from permafrost in Sweden may be controlled by
-a single taxon [@mccalley2014]. However, most microbially mediated ecosystem functions are
-likely the result of interactions (direct and indirect) among many taxa.  We
-could maximize our ability to identify taxa of small effect by constraining
-variation, e.g., by sampling microbial communities in a way that makes them
-similar in structure (similar ecosystem, soil type, and abiotic conditions)
-while still varying in function.
-
-Ultimately, the relationships identified in genotype-phenotype mapping studies
-must be verified.  There are multiple ways that this verification is
-accomplished. In some cases, organisms can be artificially selected for a
-particular phenotype (e.g., through experimental evolution in an environment
-that favors the phenotype of interest) and the genetic changes that occur in
-response to selection can be compared to those identified via mapping studies
-(such as GWAS).  An analogous approach for microbial biodiversity-ecosystem
-function studies would be to apply artificial ecosystem selection (sensu @swenson2000)
-on a given function and compare the taxa that change in response to selection
-with those identified via a comparative approach (such as the one illustrated
-in our example).
-
-The most common way that loci identified in a mapping study are verified is
-through manipulative genetics.  The identified loci can be knocked out or
-over-expressed, and the effect on phenotype compared to that predicted from
-mapping studies.  There is no direct analogue for this in microbial
-biodiversity-ecosystem function studies.  In some cases it may be possible to
-inhibit a particular functional group through the use of a specific
-antimicrobial or a chemical inhibitor [@maxson2016], but this is not generally true.  It
-may be possible in some cases to isolate a microorganism of interest in pure
-culture and add it back to a particular ecosystem, transiently increasing its
-abundance (roughly analogous to “overexpressing” a gene).  Synthetic
-communities (contrived assemblages of microorganisms) may ultimately be the
-most powerful way to test hypotheses about microbial biodiversity-ecosystem
-function relationships, but currently these approaches are limited by the small
-number of taxa that can be routinely cultured from most environments.
-
-## Conclusion
-
-Microbial biodiversity-ecosystem function research has demonstrated that
-functional group abundance (measured by qPCR of functional genes) is often not
-a good predictor of ecosystem function and that microbial
-diversity metrics on average do not add much power to ecosystem models.  A new
-perspective on how to determine the relationship between microbial communities
-and ecosystem functions is sorely needed.  Organismal biologists have over a
-hundred years of experience identifying relationships between complex sets of
-highly-variable data (genotypes or genome sequences) and aggregate functions
-(organismal phenotypes).  We assert that combining the approaches of
-traditional biodiversity-ecosystem function research with ideas from
-genotype-phenotype mapping could provide this new perspective.  This
-integration could not only make underutilized approaches such as covariate
-modeling and artificial selection more available to microbial ecologists, but
-also provide instructive examples of how best to conceive of microbial
-biodiversity-ecosystem function questions.  If this integration is successful,
-it is possible that in the not-so-distant future our field will be able to
-robustly identify taxa, genes, or even molecules that will allow us to
-accurately predict the response of ecosystems to environmental change. Doing so
-will not only generate novel hypotheses about how complex microbial communities
-drive ecosystem function, but also help scientists predict and manage changes
-to ecosystem functions resulting from human activities.
-
-# Acknowledgments
-
-This project was supported by the National Science Foundation – Dimensions of
-Biodiversity program (DEB 14422214), the National Science Foundation Graduate
-Research Fellowship Program (DGE 1255832), and the ARCS Foundation Florence and
-Mike Nudelman Scholarship. The data re-analyzed here were generated with
-permission of the Government of Gabon, Centre National de la Recherche
-Scientifique et Technologique (Permit No AR0035/14/MESRS/CENAREST/CG/CST/CSAR)
-and the assistance of F. Bivigou, H. Memiaghe, L. Tchignoumba, E.  Tobi, and I.
-Akendengue. The University of Oregon, the Gabon-Oregon Transnational Center on
-Environment and Development, the Smithsonian Conservation Biology Institute,
-and Shell Gabon provided financial and logistical support for the original
-study re-analyzed here. We also thank W. Cresko for insightful discussions on
-population genetics and H.  Tavalire for contributions to the statistical
-analyses.
-
-# References
-
-::: {#refs} 
-:::
-
-# Tables and Figures
-
-```{r copy_num}
-model_out <- read.csv('../output/model_out.csv')
-flextable(model_out) %>% 
-  compose(i = 1, j = 1, 
-    value = as_paragraph(as_i("pmoA"), " copy number")) %>% 
-  compose(part = "header", j = 4, 
-    value = as_paragraph("t-statistic")) %>% 
-  compose(part = "header", j = 5, 
-    value = as_paragraph("p-value")) %>% 
-  autofit()
-```
-
-Table 1. Functional gene abundance and ASV richness are not good predictors of
-methane oxidation rate. Linear models predicting methane oxidation rate from
-measures of microbial community structure.
-
-```{r mantel}
-mantel_table <- read.table('../output/mantel_table.tsv')
-flextable(mantel_table) %>% 
-  compose(part = "header", j = 2, 
-    value = as_paragraph("Mantel statistics (r)")) %>% 
-  compose(part = "header", j = 3, 
-    value = as_paragraph("95% upper quantile of permutations")) %>% 
-  compose(part = "header", j = 4, 
-    value = as_paragraph("p-value")) %>% 
-  autofit()
-```
-
-Table 2. Mantel tests for each pair of dissimilarity matrices. Community
-distance matrix was based on Bray-Curtis distance while both environment and
-geography distance matrices were based on Euclidean distance. P-values
-determined by permutation test with 999 permutations.
-
-```{r n_taxa}
-n_taxa <- read.table('../output/comp_adds_removes.tsv')
-flextable(n_taxa) %>% 
-  compose(part = "header", j = 1, 
-    value = as_paragraph("Terms")) %>% 
-  compose(part = "header", j = 2, 
-    value = as_paragraph("Removed")) %>% 
-  compose(part = "header", j = 3, 
-    value = as_paragraph("Added")) %>% 
-  compose(part = "header", j = 4, 
-    value = as_paragraph("Significant")) %>% 
-  autofit()
-```
-
-Table 3. Number of significant taxa after including each set of covariates in a
-variance component model. Removed and Added columns are relative to the
-no-covariate model. Significance is determined by controlling the false
-discovery rate at q-values < 0.05.
-
-```{r methane, out.width="768px", out.height="364px"}
-knitr::include_graphics('../figures/trad_cors.pdf')
-```
-
-Figure 1. Methane oxidation rate is not correlated with functional gene
-abundance or ASV richness. Correlations between community attributes and
-ecosystem function. A) Abundance of the functional gene pmoA (n = 42) and B)
-ASV richness (n = 44). Lines represent the ordinary least squares regression
-line with standard errors.
-
-```{r structure, out.width="960px", out.height="1152px"}
-knitr::include_graphics('../figures/pcoa_multi.pdf')
-```
-
-Figure 2. Microbial community composition is spatially and environmentally
-structured. Principle coordinate plots of Bray-Curtis distance representing the
-first three axes of community composition. In A, B, and C, points are
-identified by Site ID and in D, E, and F, points are identified by wetland or
-upland ecosystem. All four environmental covariates separate strongly by
-wetland/upland. Axis length is proportional to variance explained as indicated
-in parentheses. PC = principal coordinate.
-
-```{r taxa, out.width="768px", out.height="384px"}
-knitr::include_graphics('../figures/effect_sizes.pdf')
-```
-
-Figure 3. Taxa associated with methane oxidation rates after controlling for
-geographic location, environmental similarity, and community composition.
-Points are estimates for the linear relationship between the relative abundance
-of a single ASV and methane oxidation rate with standard errors from variance
-component models including similarity matrices as covariates for community and
-environment and site ID for geographic location. Amplicon sequence variants are
-labelled at the finest resolution available: genus for all except the Group 1
-Acidobacterium. Points are identified by Phylum. Significant taxa were
-determined by controlling the false discovery rate at q-value < 0.05.

### 2019-11-12 manuscript/manuscript_rev1.Rmd
Incorporating Brendan's edits

--- 2019-10-08 manuscript/manuscript_rev1.Rmd
@@
 
-Microorganisms mediate many important ecosystem functions, yet it remains
-unclear to what extent microbial diversity or community composition is
-important for determining the rates of ecosystem-scale functions. This
-uncertainty limits our ability to predict and manage crucial
-microbially mediated processes, such as nutrient loss and greenhouse gas
-emissions. Our lack of understanding stems from the relatively large
-diversity of microorganisms, the difficulty in directly identifying
-functional groups, and our limited ability to manipulate microbial community
-attributes. For this reason, we propose that integrating traditional
-biodiversity-ecosystem function research with ideas from genotype-phenotype
-mapping could provide the new perspective our discipline needs. We identify
-three insights from genotype-phenotype mapping that could be useful for
-microbial biodiversity-ecosystem function studies: the concept of “agnostic”
-mapping, the use of more powerful ways to account for multiple comparisons,
-and the incorporation of covariates into models of ecosystem function. We
-illustrate the potential for these approaches to elucidate microbial
+Microbial physiological processes are intimately involved in nutrient cycling.
+However, it remains unclear to what extent microbial diversity or community
+composition is important for determining the rates of ecosystem-scale
+functions. There are numerous examples where microbial diversity is correlated
+with the rate of an ecosystem function, but exactly how microbial communities
+“map” onto ecosystem functions remains unresolved.  This uncertainty limits our
+ability to predict and manage crucial microbially mediated processes such as
+nutrient loss and greenhouse gas emissions. Our lack of understanding stems
+from the relatively large diversity of microorganisms, the difficulty in
+directly identifying functional groups, and our limited ability to manipulate
+microbial community attributes. For this reason, we propose that integrating
+traditional biodiversity-ecosystem function research with ideas from
+genotype-phenotype mapping could provide the new perspective our discipline
+needs. We identify two insights from genotype-phenotype mapping that could be
+useful for microbial biodiversity-ecosystem function studies: the concept of
+“agnostic” mapping and controlling for population stratification to identify
+microorganisms that uniquely contribute to ecosystem function. We illustrate
+the potential for these approaches to elucidate microbial
 biodiversity-ecosystem function relationships by analyzing a subset of
-published data measuring methane oxidation rates from incubations of tropical
-soil. We assert that combining the approaches of traditional
-biodiversity-ecosystem function research with ideas from genotype-phenotype
-mapping will not only generate novel hypotheses about how complex microbial
-communities drive ecosystem function, but also help scientists predict and
-manage changes to ecosystem functions resulting from human activities.
+published data measuring methane oxidation rates from tropical soils. We assert
+that combining the approaches of traditional biodiversity-ecosystem function
+research with ideas from genotype-phenotype mapping will not only generate
+novel hypotheses about how complex microbial communities drive ecosystem
+function, but also help scientists predict and manage changes to ecosystem
+functions resulting from human activities.
 
@@
 
-### A new perspective on microbial community-ecosystem function relationships is needed
+Ecology is broadly focused on understanding biodiversity and how that
+biodiversity shapes the ecosystems that humans depend on. Many ecosystem
+processes essential to all of life are mediated by microorganisms and therefore
+understanding the relationship between microbial biodiversity and ecosystem
+function is important [@schimel1998, @singh2010]. Certain ecosystem functions
+are correlated with microbial diversity, indicating that we should be able to
+determine what aspects of microbial biodiversity influence ecosystem function.
+However, attempts to describe that mapping have borne little fruit [@rocca2015,
+@graham2016]. We argue that to overcome this challenge we should look to other
+successful attempts at mapping biological variation onto higher order
+processes. In particular, population genetics and the process of
+genotype-phenotype mapping provide a number of potentially useful insights. For
+example, genotype-phenotype mapping often makes few assumptions about the
+nature of the map, i.e., it is “agnostic.” In addition, population geneticists
+have developed rigorous methods for reducing potentially confounding
+relationships such as geographic structuring of populations. Finally, the
+ultimate goal of genotype-phenotype mapping is to identify the unique
+contribution of genotype to phenotype separate from other drivers of phenotypic
+variation such as the environment. Inspired by this, we suggest a reframe of
+the question, "is microbial biodiversity related to ecosystem function?" to, "what
+is the unique contribution of the microbiome to ecosystem function independent
+of the environment?"
 
-Ecologists have long investigated the effects of changing biodiversity on
-ecosystem function, documenting, for example, relationships between terrestrial
-plant community richness and primary productivity [@hooper2005]. The study of
-the relationship between microbial biodiversity and ecosystem functions is much
-more recent and has generated mixed results [@schimel1998, @singh2010]. Despite
-the fact that microbes mediate many important ecosystem functions, it remains
-maddeningly unclear to what extent microbial diversity or community composition
-is important for determining the rates of ecosystem-scale functions.  For
-example, numerous studies have attempted to correlate microbial functional
-group abundance or diversity with the rate of various ecosystem functions and
-most are unsuccessful [@rocca2015, @graham2016]. This uncertainty limits our
-ability to predict and manage crucial microbially mediated ecosystem functions,
-such as nutrient loss and greenhouse gas emissions. As a discipline, we need a
-new perspective on these questions.
+### There is evidence that microbial biodiversity matters for ecosystem function
 
-Our lack of understanding stems from many sources. Microbial communities are
-more diverse and contain greater numbers of individuals than plant communities and are thus more
-difficult to sample comprehensively. Unlike plants, functional groups of
-microbes are rarely determined through observation or direct measurement, but
-rather must be indirectly inferred based on taxonomic assignment of marker
-genes (such as 16S rRNA genes), by sequencing functional genes
-from metagenomes, or by quantifying specific functional genes using qPCR. This adds uncertainty in our ability to estimate the size of functional groups.
-The ability to experimentally manipulate microbial diversity or community
-composition is much more limited for microbes. Given all of these differences
-(and many others), the challenge of linking microbial communities to ecosystem
-functions may be less like that of plant communities, and more analogous to the
-complex task of linking genomic variation to organismal phenotypes. For this
-reason, we propose that integrating traditional biodiversity-ecosystem function
-research with ideas from genotype-phenotype mapping could provide the new
-perspective our discipline needs.
+There is increasing evidence for microbial biodiversity-ecosystem function
+relationships. For example, there are positive correlations between microbial
+diversity and ecosystem multifunctionality for a variety of ecosystems and most
+major lineages of microorganisms [@jing2015, @delgado-baquerizo2016,
+@delgado-baquerizo2017]. Adding microbial diversity or abundance to ecosystem
+models can in some cases improve model accuracy [@graham2016]. Contrived
+communities that vary in richness and communities created through sequential
+dilution or varying filter size to generate differences in diversity can also
+exhibit differences in ecosystem function [@philippot2013, @wagg2014, @schnyder2018,
+@maron2018].  Finally, reciprocal transplant and common garden experiments that
+manipulate the connection between community composition and environment reveal
+differences in ecosystem function for communities of distinct
+origins [@cavigelli2000, @strickland2009, @glassman2018]. Given these
+relationships, many investigators have now moved on to the challenge of
+identifying the aspects of microbial biodiversity (e.g., specific taxa, genes,
+functional groups, etc.) that influence a given ecosystem function; however,
+this has proven especially challenging.
 
-### There is evidence that microbial diversity can matter
+### The mapping between microbial biodiversity and ecosystem function has been elusive
 
-Perhaps our uncertainty arises because there is no general relationship between
-microbial biodiversity and any particular ecosystem function.  For example, if
-local selection always optimizes the available microbial biodiversity to
-maximize ecosystem function, then microbial community composition should not
-matter for predicting the rates of microbially mediated functions. Instead, the
-rates of these functions should be determined by the underlying environmental
-variation. In this case, the microbial community would simply act as a conduit
-through which the abiotic environment alters ecosystem function. Attempts to
-document a general relationship between microbial community attributes and
-function would then fail (since the primary drivers are environmental).
-
-However, there is evidence that this is not always the case, A limited number
-of studies have manipulated the connection between environmental factors and
-microbial community composition through “common garden” or reciprocal
-transplant experiments, and they frequently report different rates of ecosystem
-functions for different microbial communities under the same environmental
-conditions [@reed2007].  This has been observed for plant decomposition
-[@glassman2018, @strickland2009], plant phenology [@panke-buisse2015], soil
-nitrogen cycling [@balser2005], and soil greenhouse gas emissions
-[@cavigelli2000], among others. Therefore, variation in microbial community
-composition can be associated with variation in ecosystem function, independent
-of environmental variation.  So why has this been difficult to consistently
-document in the field?
-
-### Traditional approaches to quantifying this relationship have provided only minor improvements
-
-Most comparative studies of microbial community function in the field focus on
-one of two aspects of microbial community structure that are hypothesized to
-predict ecosystem function. The first aspect is “functional” gene or transcript
-abundance.  In this case, qPCR or shotgun metagenomic sequencing is used to
-estimate the abundance of a gene or transcript that is a putative marker for a
-microbial process (and thus a marker for the functional group that performs
-this process). For example, the gene mcrA, which encodes a subunit of the
-enzyme that performs the final step in methanogenesis, is commonly used as a
-marker for methanogenesis and for the methanogen functional group. Other
-examples include pmoA and methanotrophy, nifH and nitrification, and nosZ and
+Most studies that attempt to identify the aspects of microbial biodiversity
+that influence a given ecosystem function focus on “functional” gene or
+transcript abundance.  In this case, qPCR or shotgun metagenomic sequencing is
+used to estimate the abundance of a gene or transcript that is a putative
+marker for a microbial process (and thus a marker for the functional group that
+performs this process). For example, the gene _mcrA_, which encodes a subunit of
+the enzyme that performs the final step in methanogenesis, is commonly used as
+a marker for methanogenesis and for the methanogen functional group. Other
+examples include _pmoA_ and methanotrophy, _nifH_ and nitrification, and _nosZ_ and
 denitrification. The abundance of these markers is often hypothesized to be
 predictive of the rate of their associated processes (for example, the
-abundance of mcrA is often hypothesized to be related to the rate of
-methanogenesis). There are examples where this relationship is present
-[@freitag2009, @freitag2010, @schnyder2018].  However, a review of such studies
-found that the abundance of a functional gene or transcript is rarely
-correlated with the rate of the corresponding process and depends on the
-function of interest, with most effects either negative or not significant
-[@rocca2015].
+abundance of _mcrA_ is often hypothesized to be related to the rate of
+methanogenesis).
 
-The second aspect of microbial community structure hypothesized to predict
-ecosystem function is taxonomic or functional diversity. Diversity is either
-estimated from sequence variants of a barcode gene such as the 16S rRNA gene or
-manipulated through some proxy of diversity such as sequential dilution or
-varying filter sizes. Studies that experimentally manipulate diversity
-generally find a relationship between the applied diversity treatment and
-ecosystem functions, including for methanotrophy [@schnyder2018], phosphorus
-leaching [@wagg2014], greenhouse gas emissions [@wagg2014], decomposition
-[@maron2018], and nitrogen cycling [@philippot2013]. However, in many cases,
-richness is confounded with other factors such as the presence or absence of
-major phylogenetic groups (nematodes, fungi) or abundance of microbial cells.
-In addition, assembled microbial communities pose the same challenges as
-macroorganismal diversity experiments, for example the highest diversity
-treatment will more likely contain the most productive taxon. A review of
-microbial diversity studies shows that microbial community taxonomic and
-functional diversity add little explanatory power to models of ecosystem
-function [@graham2016]. Overall, functional gene abundance and community
-diversity improve models of ecosystem function less than one third of the time
-and increase variance explained by an average of only 8 percentage points
-[@graham2016].
+Some ecosystem functions in certain ecosystems can be predicted from the
+abundance or transcriptional activity of genetic markers for those functions.
+For example, soil methane production and consumption can under some
+circumstances be predicted from the genetic markers _mcrA_ and _pmoA_
+[@freitag2009, @freitag2010, @mccalley2014]. However, for most ecosystem
+functions the abundance of a functional gene or transcript is rarely positively
+correlated with the rate of the corresponding process [@rocca2015]. The
+cases where there is a positive correlation tend to be restricted to
+agricultural ecosystems and for certain functions within the nitrogen cycle
+[@rocca2015]. In general, including aspects of microbial biodiversity
+(e.g., functional gene abundance or diversity) improve models of ecosystem
+function less than one third of the time, and increase variance explained by an
+average of only 8 percentage points over environmental variables [@graham2016].
 
-### Genotype-phenotype mapping as a source of inspiration
+## Genotype-phenotype mapping as a source of inspiration
 
-Given the relative lack of success to date, it is time to rethink how we
-approach the challenge of mapping microbial community structure to ecosystem
-function.  Microbial ecologists are not the only biologists who are attempting
-to determine the relationship between a complex set of highly-variable data and
-an aggregate function.  This kind of “many-to-one” mapping is analogous to the
-challenge of identifying the genetic basis of complex traits in organismal
-populations. In such “genotype-phenotype” mapping studies, a population
-exhibits variation in a phenotype (e.g., height or disease state) as well as
-variation in potentially thousands of single nucleotide polymorphisms (SNPs).
-To identify the genetic basis for a trait, investigators sample from this
-population and correlate phenotype with genotype. In most cases, there are many
-more loci than individuals and we do not know whether the SNPs are causally
-linked or are simply in linkage disequilibrium with a causal mutation.
+In the approaches described above, microbial ecologists often use microbiome
+data to infer taxonomic composition, essentially creating species lists from
+data such as 16S rRNA marker genes or shotgun metagenomes. Interpreting
+microbiome data in this way has allowed us to use approaches from
+biodiversity-ecosystem function research (which are often focused on taxonomic
+or functional groups), but it has generally not been useful for creating more
+detailed descriptions of the relationship between microbial biodiversity and
+ecosystem function. But this approach is not the only way one could determine
+the relationship between a complex set of highly-variable data and an aggregate
+function.
+
+This kind of “many-to-one” mapping is analogous to the challenge of identifying
+the genetic basis of complex traits in organismal populations. In such
+“genotype-phenotype” mapping studies, a population exhibits variation in a
+phenotype (e.g., height or disease state) as well as variation in potentially
+thousands of single nucleotide polymorphisms (SNPs). To identify the genetic
+basis for a trait, investigators sample from this population and correlate
+phenotype with genotype. While some phenotypes (e.g., the propensity for
+diseases such as Parkinson’s) are controlled by a single genetic locus
+[@kerem1989, @macdonald1992], most traits depend on a large number of genes
+that control variation in that phenotype [@hindorff2009, reich2001]. In
+addition, there is often no a priori expectation about which regions of the
+genome control that trait so we must search for genetic markers
+throughout the genome. If a marker is significantly correlated with the
+phenotype of interest, this either indicates it is inside of a gene with a
+direct or indirect effect on phenotype or that it is in linkage disequilibrium
+(i.e., non-random association between two alleles) with a causal gene.
 
 There are a number of parallels between this challenge faced by organismal
-biologists and that facing microbial community ecologists.  They are both
-“many-to-one” mapping challenges, which involve large numbers of statistical
-comparisons.  Both are attempting to identify causal relationships that are
-potentially confounded by very complex patterns of covariation.  There is often
-no strong expectation about which entities (i.e., which genomic regions or which
-microbial taxa) are most likely to be causally-related to phenotype or
-function, and thus “agnostic” approaches are needed.  Both are sensitive to
-exactly how the “mapping” question is asked, and ultimately require
-manipulation (of genes or taxa) to establish causation.  We describe each of
-these parallels below in more detail, and provide an example of how these ideas
-could be applied to microbial data.
+biologists and that facing microbial community ecologists. They both involve
+large numbers of statistical comparisons. Both are attempting to identify
+causal relationships that are potentially confounded by complex patterns of
+covariation. There is often no strong expectations about which entities (i.e.,
+which genomic regions or which microbial genes or lineages) are most likely to
+be causally-related to phenotype or function, and thus “agnostic” approaches
+are needed. For some ecosystem functions, it is possible that a single
+microbial taxon could substantially influence its rate. For example, methane
+flux from permafrost in Sweden may be controlled by a single taxon [@mccalley2014]. But
+for most ecosystem functions, there could be many taxa of small effect that
+contribute to the rate of ecosystem function. Finally, both ultimately require
+manipulation (of genes or taxa) to establish causation.
 
@@
 
-The traditional approach to microorganismal biodiversity-ecosystem function
-research is to measure or manipulate the diversity of a taxonomic group (e.g.,
-plants) and look for an association with the function performed by that group
-(e.g., primary productivity).  In the broadest sense, we can think of plants as
-a “functional group,” i.e., a group of taxa united by their ability to perform
-a particular ecosystem function. Ecologists may further divide a functional
-group (e.g., plants) into smaller functional groupings (e.g., forbs), defined
-by the details of how they perform their particular ecosystem function.
+Most microbial biodiversity-ecosystem function research up to this point has
+used an approach analogous to that used by plant ecologists studying
+biodiversity-ecosystem function relationships. This approach is to measure or
+manipulate the diversity of a taxonomic group (e.g., plants) and look for an
+association with the function performed by that group (e.g., primary
+productivity). We can think of plants as a “functional group,” i.e., a group of
+taxa united by their ability to perform a particular ecosystem function. For
+microbes, estimating functional group abundance can be much more challenging.
+From a small number of cultured isolates, we have a provisional understanding
+of which microbes might be involved in some ecosystem functions. By sequencing
+the genomes of these isolates, we have identified genetic markers for certain
+functions, which we call “functional genes.” But most microbial taxa remain
+uncultured and we do not know the function of most microbial taxa detected in
+environmental samples [@hug2016, @martiny2019]. In addition, there have
+been recent discoveries of functions in unexpected taxonomic groups, for
+example methanogenesis by fungi and cyanobacteria, a function previously
+considered restricted to the Archaea [@bizic-ionescu2019, @lenhart2012].
 
-However, for microbes our knowledge of functional groups is much more limited.
-From a very limited number of cultured isolates we have a provisional
-understanding of which microbes might be involved in some ecosystem functions.
-And by sequencing the genomes of these isolates, we have identified genetic
-markers for certain functions. But most microbial taxa remain uncultured and we
-do not know the function of most microbial taxa detected in environmental
-samples [@hug2016, @steen2019]. In addition, there have been recent discoveries
-of functions in unexpected taxonomic groups, for example methanogenesis by
-fungi and cyanobacteria, a function previously considered restricted to the
-Archaea [@bizic-ionescu2019, @lenhart2012]. Because of this, it would be
-prudent to look more agnostically at microbial communities to identify taxa or
-groups of taxa that are important for predicting the rate of ecosystem
-functions rather than assuming that the genetic markers we have provisionally
-identified for a given function represent the most likely taxa involved.  This
-agnostic approach is analogous to the approach of many genotype-phenotype
-mapping studies (e.g.,  genome-wide association studies or GWAS), which often
-look for associations between a phenotype and loci anywhere in a genome.
+As illustrated earlier, these functional markers are not correlated with their
+corresponding ecosystem function in most ecosystems and for most processes. In
+addition, they provide little explanatory power to models of ecosystem
+function. Because of this, it might be prudent to look more agnostically at
+microbial communities to identify taxa, groups of taxa or genes that are
+important for predicting the rates of ecosystem functions rather than assuming
+that the genetic markers we have provisionally identified for a given function
+represent the most likely taxa or genes involved. This agnostic approach is
+analogous to the approach of many genotype-phenotype mapping studies (e.g.,
+genome-wide association studies), which often look for associations between a
+phenotype and loci anywhere in the genome.
 
-### Dealing with multiple comparisons and covariation
+Beyond finding new physiologies in unexpected places, there are other reasons
+for looking agnostically. In the case of microbial functions, it may be that
+the organism that performs a function is not the limiting factor for the rate
+of that function. For example, the rate of soil-to-atmosphere methane flux
+could be limited by methanogens or methanotrophs or the balance of the two. It
+could also be limited by the bacteria that produce the fermentative byproducts
+that methanogens depend on. Or there could be indirect limitation by organisms
+that liberate nitrogen or phosphorus into mineral forms. In other words, the
+influence of microbial communities on the rate ecosystem function could
+represent a complex metabolic network much like the regulation of gene
+expression in organisms that partially determine their phenotype. These broader
+patterns of biodiversity-ecosystem function relationships would be invisible to
+any study that focuses on putative functional groups without considering the
+reads that underlie those functional groups.
 
-Microbial biodiversity-ecosystem function studies and genotype-phenotype
-mapping studies are both “many-to-one” mapping challenges, which involve large
-numbers of statistical tests.  Hundreds to thousands of tests are routinely
-made per study, greatly inflating the number of false positives identified
-using statistical hypothesis testing approaches.  Some microbial
-biodiversity-ecosystem function studies do not correct for this, while others
-use approaches that may unnecessarily inflate the false negative rate, such as
-the Bonferroni correction. The Bonferroni correction has been widely considered
-to be too conservative, particular for exploratory studies designed to generate
-hypotheses [@perneger1998, @rothman1990, @thomas1985]. Statisticians have
-developed a number of less-conservative approaches to correct for multiple
-comparisons by controlling the false discovery rate in order to balance the
-tradeoff between Type I and Type II errors [@benjamini1995, @storey2002].
-These approaches are commonly used in genotype-phenotype mapping studies.
+### Controlling for population stratification
 
@@
 population stratification due to geographic and environmental separation
-[@martiny2006, @wright1943]. Genome-wide association studies generally account
-for population structure due to shared ancestry among cases and controls when
-modeling the connection between genotype and phenotype. The classic example is
-the latitudinal gradient of both height and genotypic similarity in Europe,
-which results in spurious associations between human height and genetic
-variation [@berg2019, @novembre2008, @sohail2019]. To correct for this
-covariance structure, GWAS models incorporate genotypic similarity to correct
-for shared ancestry using a variety of methods, such as principal component
-correction or variance component modeling [@kang2010, @price2006]. Typically,
-microbial biodiversity-ecosystem function studies do not account for population
-stratification (i.e., community similarity across samples), although there are
-some exceptions [@meyer2019, @qin2012]. Genome-wide association studies
-generally ignores the underlying environmental and spatial distance between
-samples and instead uses shared ancestry as a proxy for these variables.
-However, community similarity (the community analogue of shared ancestry among
+[@wright1943, @martiny2006]. This can lead to spurious associations between
+phenotypes and genetic markers that are at high frequency in isolated
+sub-populations. Association studies generally control for population
+stratification by accounting for shared ancestry among organisms in a
+population when modeling the connection between genotype and phenotype.
+Typically, microbial biodiversity-ecosystem function studies do not account for
+population stratification (i.e., community similarity among ecosystems),
+although there are some exceptions [@meyer2019, @qin2012, @lloyd-price2019].
+Community similarity (the community analogue of shared ancestry among
 organisms) is not as tightly linked to geography or environment as is shared
-ancestry among organisms, and it could be very useful to account for these
-separately in microbial studies, especially if one is particularly interested
-in how composition alters function independent of the underlying environmental
+ancestry, and it could be very useful to account for these separately in
+microbial studies, especially if one is particularly interested in how
+composition alters function independent of the underlying environmental
 variation.
+
+Genome-wide association studies correct for stratification using a variety of
+methods. Generally, they ignore the underlying environmental and spatial
+distance between samples and instead use shared ancestry as a proxy for local
+selection and assortative mating. A common approach is to perform a regression
+of phenotype and shared ancestry (computed as the first one or more principal
+components of the genotype matrix) and then use the residuals from this model
+as the values for phenotype in a subsequent regression using the genotypes
+directly [@price2006]. This principal component correction is designed to test the
+effect of individual genes after removing the effect of shared ancestry among
+individuals. Another approach, employed in our example, is variance component
+modeling (or mixed modeling, hierarchical modeling, etc.) where genotypic
+similarity is included as a covariate in the model to control for
+stratification while testing the genotype-phenotype connection [@kang2010].
+
+If we control for covariates such as community similarity, geographic distance,
+or environmental similarity, it changes the nature of our question. For
+example, if we test the correlation between the relative abundance of a taxon
+and the rate of methane flux we are asking “is this taxon correlated with
+methane flux?” If we find a significant result, that may be because variation
+in the abundance of that organism directly or indirectly contributes to methane
+flux. However, it might also be that that organism lives only in ecosystems
+that happen to have a high rate of ecosystem function. We would not be able to
+distinguish these outcomes with this simple correlation. If we then include
+environmental variables or environmental similarity as a covariate, we could
+start to ask, “Is the taxon uniquely associated with function independent of
+the environment?” By “uniquely associated”, we mean those taxa associated with
+function irrespective of environmental conditions, local community structure,
+or spatial proximity. This is not only a more specific question than is usually
+asked in microbial biodiversity-ecosystem function studies, but it is also a
+more appropriate one, especially if one is interested in how to incorporate
+microbial community data into ecosystem models. Finally, it is interesting on
+its own to understand whether microbial communities are selected by the
+underlying environmental conditions to produce a particular rate of ecosystem
+function (the conduit) or whether there is variation in community composition
+independent of the environment that explains some variation in ecosystem
+function.
 
@@
 
-To illustrate the ideas outlined above, we reanalyzed a subset previously
-published data [@meyer2019] using a modified procedure from the original
-version.  A full description of the study design, samples, and data generation
-can be found in that manuscript. Briefly, these data were gathered from intact
-soil cores taken from diverse ecosystems of the Congo Basin in Gabon, Africa.
-Cores were incubated in the laboratory under different concentrations of
-methane to identify the rates of specific methane cycling pathways. For this
-example, we will analyze data from just one of these pathways, high-affinity
-methane oxidation (the oxidation of atmospheric concentrations of methane),
-which we will refer to below as “methane oxidation”. In addition, for
-simplicity we only include amplicon sequences from the DNA-inferred community
-and not the RNA-inferred community, both of which are presented in the original
-paper [@meyer2019].  The data we analyzed include methane oxidation
-measurements, amplicon sequence variants (ASVs; @callahan2017) generated using
-DADA2 and inferred from 16S rRNA gene sequences, pmoA abundance estimates (via
-qPCR), latitude and longitude, and four environmental covariates (soil
-moisture, bulk density, carbon, and nitrogen).
+To illustrate the ideas outlined above, we reanalyzed a subset of previously
+published data from a paper that has demonstrated one successful approach
+for applying genotype-phenotype mapping to microbial communities [@meyer2019]. In our
+reanalysis, we do not intend to challenge the conclusions from that paper, but
+instead we want to demonstrate how to perform this type of study for microbial
+ecologists unfamiliar with association studies. A full description of the study
+design, samples, and data generation can be found in that manuscript. Briefly,
+these data were gathered from intact soil cores taken from diverse ecosystems
+of the Congo Basin in Gabon, Africa. Cores were incubated in the laboratory
+under different concentrations of methane to identify the rates of specific
+methane cycling pathways. For this example, we will analyze data from just one
+of these pathways, high-affinity methane oxidation (the oxidation of
+atmospheric concentrations of methane), which we will refer to simply as
+“methane oxidation”. In addition, we only include amplicon sequences from the
+DNA-inferred community and not the RNA-inferred community, both of which are
+presented in the original paper (@meyer2019). The data we analyzed include methane
+oxidation rates, amplicon sequence variants (ASVs; @callahan2017) generated using DADA2
+and inferred from 16S rRNA gene sequences, _pmoA_ abundance estimates (via qPCR),
+latitude and longitude, and four environmental covariates (soil moisture, bulk
+density, carbon, and nitrogen).
 
 Analyses were conducted in the R statistical environment using the phyloseq
-package [@mcmurdie2013, @rcoreteam2019].  The relative abundances of ASVs were
-corrected using the variance stabilizing transformation from DESeq2 [@love2014,
-@mcmurdie2014]. We first test typical measures of microbial structure including
-functional gene abundance and community richness, which was estimated using the
-breakaway package [@willis2015]. We then demonstrate significant covariation
-between community structure (estimated as Bray-Curtis distance using vegan),
-environmental variation (euclidean distance), and geographic distance
-(euclidean distance) using Mantel tests [@bray1957, @oksanen2019].  Finally, we
+package [@mcmurdie2013, @rcoreteam2019]. The relative abundances of ASVs were corrected using the variance
+stabilizing transformation from DESeq2 [@love2014,
+@mcmurdie2014]. We first test typical measures of
+microbial structure including functional gene abundance and community richness,
+which was estimated using the breakaway package [@willis2015]. We then demonstrate
+significant covariation between community structure (estimated as Bray-Curtis
+distance using vegan), environmental variation (euclidean distance), and
+geographic distance (euclidean distance) using Mantel tests [@bray1957, @oksanen2019]. Finally, we
 present one approach to identifying taxa which are significantly associated
@@
 models using varComp to test the relationship between relative abundance of
-each ASV and methane oxidation rate [@qu2013]. To illustrate how including
-different covariates (environmental, geographic, and community) can result in
-different conclusions about which taxa are associated with function, we fit
-this model with and without random effects variance components for
-environmental similarity, geographic site ID, and Bray-Curtis similarity.
-Significant taxa were determined by controlling the false discovery rate at
-q-value < 0.05 [@storey2002]. Figures were created using ggplot2
-[@wickham2019].
+each ASV and methane oxidation rate [@kang2010, @qu2013]. To illustrate how
+including different
+covariates (environmental, geographic, and community) can result in different
+conclusions about which taxa are associated with function, we fit this model
+with and without random effects variance components for environmental
+similarity, geographic site ID, and Bray-Curtis similarity. Significant taxa
+were determined by controlling the false discovery rate at q-value < 0.05 [@storey2002].
+Figures were created using ggplot2 [@wickham2019].
 
-Methane oxidation rate was not significantly correlated with pmoA gene
-abundance or richness (Table 1, Figure 1). We tested collinearity between each
-pair of distance matrices for community, environment, and geography using
-Mantel tests and estimated p-values by permutation. We found a moderate and
-significant correlation between community composition and environmental
-variation, geography and community composition, and geography and environmental
-variation (Table 2, Figure 2). Principal coordinate plots show that beta
-diversity of samples separated by site ID and by ecosystem type (wetland or
-upland; Figure 2). Finally, we tested the effect of the relative abundance of
-each ASV on methane oxidation rate. We found different numbers of taxa
-significantly associated with methane oxidation depending on which covariates
-were included in the model (Table 3).  In particular, 460 taxa were identified
-with no covariates whereas 6 were identified after including all covariates.
-Though we cannot infer function from 16S sequences, these 6 taxa fall into
-three genera and one class with cultured representatives that are not known to
-consume methane [@belova2009, @chang2011, @domeignoz-horta2019, @fritz2004].
-Their effect on methane oxidation rate ranges from 0.5 to 1.5 (Figure 3).  An
-effect of 1.5 means that for a one unit increase in relative abundance there is
-a 1.5 increase in k (the rate of exponential decrease in methane concentration
-over time).
+### Results & Discussion
 
-## Discussion
+Microbial biodiversity-ecosystem function studies typically test functional
+group abundance or community richness as it relates to ecosystem function. In
+our case, methane oxidation rate was not significantly correlated with the pmoA
+gene abundance or 16S read richness (Table 1, Figure 1). To demonstrate that
+the covariance structure of the data might alter our conclusions about which
+taxa regulate ecosystem function, we tested collinearity between each pair of
+distance matrices for community, environment, and geography. We found a
+moderate and significant correlation between community composition and
+environmental variation, geography and community composition, and geography and
+environmental variation (Table 2, Figure 2). To visualize this population
+stratification, principal coordinate plots show that beta diversity of samples
+separated by site ID and by ecosystem type (wetland or upland; Figure 2), which
+indicates substantial spatial and environmental structuring of microbial
+populations. This means that the presence or abundance of certain genetic
+markers (taxa) will be elevated in specific ecosystems. In this case,
+high-affinity methane oxidation is typically greater in upland ecosystems than
+in wetland ecosystems and so any taxa differentially abundant in uplands will
+be correlated with methane oxidation regardless of their involvement in that
+process. It is necessary to control for this structuring to rigorously identify
+associations between taxa and function.
 
-In this paper, we argue that the traditional approach to microbial
-biodiversity-ecosystem function research often ignores the complexity of
-microbial community dynamics, in particular the complex patterns of covariation
-that often arise among microbial community similarity, environmental
-similarity, and spatial proximity.  We propose that a better approach would be
-to learn from other complex many-to-one mapping problems in biology,
-particularly genotype-phenotype mapping.  These studies often explicitly
-account for covariation, they correct for multiple comparisons in powerful
-ways, and they frequently take an “agnostic approach” that does not assume a
-particular relationship between structure and function.
-
-We illustrate the potential for these approaches to elucidate microbial
-biodiversity-ecosystem function relationships by analyzing a subset of
-published data from incubations of tropical soil.  In this example, soil cores
-from different ecosystems exhibited different rates of methane oxidation, but
-simple correlations between these rates and the diversity or abundance of
-putative methane oxidizing bacteria were not informative.  These results agree
-with recent reviews of the literature that demonstrate that such simple
-approaches are often not a fruitful avenue for elucidating microbial
-structure-function connections.
-
-We then asked if the relative abundance of particular taxa were related to
-ecosystem function, and we identified 460 taxa whose relative abundances were
-associated with rates of methane oxidation after controlling the false
-discovery rate (q-value < 0.05).  These taxa could be related to ecosystem
-function in multiple ways.  The most interesting possibility is that each of
-these taxa is statistically related because it is causally connected to the
-function. This could be direct, for example an organism that consumes methane,
-or indirect, for example an organism that regulates substrates necessary for
-consumers of methane. In either of these cases, the taxon could be useful as a
-biomarker of function or as an organism to investigate in order to better
+To demonstrate this approach, we tested the effect of the relative abundance of
+each ASV on methane oxidation rate while controlling for different sets of
+covariates including environment, geography, and community. After controlling
+the false discovery rate, 460 unique 16S reads were identified as significantly
+correlated with function when no covariates were included in the model. We
+found different numbers of taxa significantly associated with methane oxidation
+depending on which covariates were included in the model (Table 3). Each of
+these sets of taxa represent different versions of the biodiversity-ecosystem
+function mapping question. For example, controlling for environmental variation
+tests for taxa that are related to function independent of environmental
+conditions. Similarly, controlling for geographic distance tests for taxa that
+are related to function independent of geographic proximity; geographic
+proximity could influence function via unmeasured environmental variation that
+is spatially structured or differences in dispersal history among sites due to
+physical distance. In the model that controlled for all three covariates
+(community, environment, and geography), only six ASVs were significantly
+correlated with methane oxidation rate (Figure 3). These taxa could be useful
+indicators for methane oxidation rate across space and different ecosystems. In
+addition, researchers could manipulate their abundance in order to better
 understand the biological drivers of variation in methane oxidation.
 
-Alternatively, a significant association could occur for non-causal reasons.
-For example, any organism that tends to be in high abundance where methane
-oxidation rates are high would be correlated with methane oxidation, even if it
-has no causal relationship. This could be because such an organism is favored
-under the same environmental conditions that favor methane oxidizing bacteria
-(or that favor methane oxidation in general).  Such covariation can drive
-associations that are not causal, but the effects of such covariation can be
-reduced through a number of approaches (many developed by biologists interested
-in genotype-phenotype mapping).
-
-In our example, we showed that the abundances of microbial taxa exhibit complex
-patterns of covariation with each other and with environmental conditions and
-spatial location. These biogeographic patterns might be even stronger if we
-could sample the populations more intensively [@meyer2018]. Once we account for this
-covariation, our list of taxa associated with function was reduced to 6.  These
-6 taxa include taxa that are not known to directly contribute to methane
-oxidation, suggesting that the drivers of methane oxidation may be indirect,
-perhaps mediated by ecological interactions among taxa from multiple functional
-groups.
+Notably, these six taxa fall into three genera and one class with cultured
+representatives that are not known to consume methane [@belova2009, @chang2011,
+@domeignoz-horta2019, @fritz2004]. These taxa could be
+related to ecosystem function in multiple ways. The most interesting
+possibility is that each of these taxa is statistically related because it is
+causally connected to the function. This could be direct, for example an
+organism that consumes methane, or indirect, for example an organism that
+regulates substrates necessary for methane oxidizers. Alternatively, a
+significant association could occur for non-causal reasons. For example, any
+organism that tends to be in high abundance where methane oxidation rates are
+high would be correlated with methane oxidation, even if it has no causal
+relationship. This could be because such an organism is favored under the same
+conditions that favor methane oxidation. Such covariation can
+drive associations that are not causal, but the effects of covariation would
+have been reduced by controlling for covariates in our tests.
 
@@
 from the fundamental ecological processes that drive community assembly and
-underlie biogeographical patterns [@martiny2006, @vellend2010]. It is also powerful because it is
+underlie biogeographical patterns [@martiny2006, @vellend2010]. It is also
+powerful because it is
 focused on a much more specific version of the “biodiversity-ecosystem
@@
 function irrespective of environmental conditions, local community structure,
-or spatial proximity.  This is not only a more specific question than is
-usually asked in microbial biodiversity-ecosystem function studies, but it is
-also a more appropriate one, especially if one is interested in how to
-incorporate microbial community data into ecosystem models. For modeling what
-is usually important is to identify those taxa that add explanatory power
-beyond that provided by other factors (such as environmental conditions).
+or spatial proximity. This is not only a more specific question than is usually
+asked in microbial biodiversity-ecosystem function studies, but it is also a
+more appropriate one, especially if one is interested in how to incorporate
+microbial community data into ecosystem models. These models depend on
+identifying taxa that add explanatory power beyond that provided by other
+factors, such as environmental conditions. Earth system models would benefit
+most from microbial bioindicators that capture variation unexplained by other,
+more easily measured variables (such as certain environmental conditions). The
+approach we describe is an important step toward identifying such biomarkers.
 
-We identified three insights from genotype-phenotype mapping that could be
-useful for microbial biodiversity-ecosystem function studies: the concept of
-“agnostic” mapping, the use of more powerful ways to account for multiple
-comparisons, and the incorporation of covariates.  But there are other insights
-to be gained as well.  For example, some phenotypes (e.g., the propensity for
-diseases such as Parkinson's) are controlled by a single genetic locus [@kerem1989, @macdonald1992].
-However, most phenotypes studied to date arise from the influence of many loci
-of small effect (as well as environmental factors; [@hindorff2009, reich2001]). Organismal
-biologists have developed approaches tailored to identifying such loci; for
-example, by sampling organismal populations in a way that constrains genetic
-variation unrelated to phenotype. Similarly, for some ecosystem functions, it
-is possible that a single microbial taxon could substantially influence its
-rate. For example, methane flux from permafrost in Sweden may be controlled by
-a single taxon [@mccalley2014]. However, most microbially mediated ecosystem functions are
-likely the result of interactions (direct and indirect) among many taxa.  We
-could maximize our ability to identify taxa of small effect by constraining
-variation, e.g., by sampling microbial communities in a way that makes them
-similar in structure (similar ecosystem, soil type, and abiotic conditions)
-while still varying in function.
+## Caveats and future directions
 
-Ultimately, the relationships identified in genotype-phenotype mapping studies
-must be verified.  There are multiple ways that this verification is
+Once taxa have been identified with an association test (such as the one we
+outline above), there are multiple ways they could be used for future study.
+One approach common in genetics, particularly for markers of genetic disorders,
+is to generate a polygenic score based on the summed effect of many genes on a
+phenotype of interest, such as the probability of developing a disorder. A
+similar aggregate bioindicator could be generated for ecosystems that would
+summarize the probability of having a particular ecosystem function or rate.
+This would be accomplished by measuring the abundance of the taxa identified in
+an association study and determining their association with the rate of an
+ecosystem function. Alternatively, the identified taxa could be incorporated
+into a structural equation model in an attempt to better understand the
+individual effects and interactions among taxa as they contribute to the rate
+of ecosystem function. This might give an indication of the relative importance
+of different taxa as compared to other factors, such as environmental
+variables.
+
+Ultimately, the relationships identified in any comparative mapping study
+must be verified. For genotype-phenotype studies in organisms there are
+multiple ways that this verification is
 accomplished. In some cases, organisms can be artificially selected for a
@@
 
-The most common way that loci identified in a mapping study are verified is
+The most common way that loci identified in a genotype-phenotype mapping study are verified is
 through manipulative genetics. The identified loci can be knocked out or
-over-expressed, and the effect on phenotype compared to that predicted from
-mapping studies.  There is no direct analogue for this in microbial
-biodiversity-ecosystem function studies.  In some cases it may be possible to
-inhibit a particular functional group through the use of a specific
-antimicrobial or a chemical inhibitor [@maxson2016], but this is not generally true.  It
-may be possible in some cases to isolate a microorganism of interest in pure
-culture and add it back to a particular ecosystem, transiently increasing its
-abundance (roughly analogous to “overexpressing” a gene).  Synthetic
-communities (contrived assemblages of microorganisms) may ultimately be the
-most powerful way to test hypotheses about microbial biodiversity-ecosystem
-function relationships, but currently these approaches are limited by the small
-number of taxa that can be routinely cultured from most environments.
+over-expressed and the effect on phenotype compared to that predicted from
+mapping studies. In the case of microbial biodiversity studies, it may be
+possible to inhibit a particular functional group through the use of a specific
+antimicrobial or a chemical inhibitor or by using phages with high
+host-specificity [@maxson2016], but this is not generally possible. In some cases, we
+may be able to isolate a microorganism of interest in pure culture and add it
+back to an ecosystem, transiently increasing its abundance (roughly analogous
+to “overexpressing” a gene). A greater focus on culture-based approaches could
+increase the success of these kinds of microbial enrichments. Finally,
+synthetic communities (contrived assemblages of microorganisms) may be the most
+powerful way to test hypotheses about microbial biodiversity-ecosystem function
+relationships, but currently these approaches are limited by the small number
+of taxa that can be routinely cultured from most environments.
 
-### Conclusion
+There are a number of limitations to the biodiversity-ecosystem function
+mapping approach we describe, some in common with organismal mapping studies
+and others unique. For example,
+simple linear models such as the variance component model used in this study
+are typical for genetics studies, but may not be the best way to identify
+correlations for microbiomes due to the unique challenges of microbial data.
+These are zero-inflated compositional data with differences in sequencing
+depth across samples. We have addressed these challenges by using a variance
+stabilizing transformation, but other models that test differential
+abundance and differential variance that can control for differences in
+sequencing depth and are robust to zero-inflation might be more appropriate
+(e.g., CORNCOB @martin2019).  Alternatively, clustering reads at higher
+taxonomic levels could provide a more continuous variation in taxon abundances
+across ecosystems by reducing the number of “bins.” The tradeoff for this is
+that it introduces bias based on the completeness of our taxonomic databases
+and the accuracy of 16S-based taxonomic assignment and it removes reads that
+lack a taxonomic assignment. Here, we chose to test ASVs at the level of the
+individual read so as not to bias our results by losing reads that lack a
+taxonomic assignment at a particular level.
 
-Microbial biodiversity-ecosystem function research has demonstrated that
-functional group abundance (measured by qPCR of functional genes) is often not
-a good predictor of ecosystem function and that microbial
-diversity metrics on average do not add much power to ecosystem models.  A new
-perspective on how to determine the relationship between microbial communities
-and ecosystem functions is sorely needed.  Organismal biologists have over a
-hundred years of experience identifying relationships between complex sets of
-highly-variable data (genotypes or genome sequences) and aggregate functions
-(organismal phenotypes).  We assert that combining the approaches of
-traditional biodiversity-ecosystem function research with ideas from
-genotype-phenotype mapping could provide this new perspective.  This
-integration could not only make underutilized approaches such as covariate
-modeling and artificial selection more available to microbial ecologists, but
-also provide instructive examples of how best to conceive of microbial
-biodiversity-ecosystem function questions.  If this integration is successful,
-it is possible that in the not-so-distant future our field will be able to
-robustly identify taxa, genes, or even molecules that will allow us to
-accurately predict the response of ecosystems to environmental change. Doing so
-will not only generate novel hypotheses about how complex microbial communities
-drive ecosystem function, but also help scientists predict and manage changes
-to ecosystem functions resulting from human activities.
+Experimentally, future studies could improve on our example by sampling a more
+homogeneous set of ecosystems. Our survey includes grasslands, pastures,
+forests, peatlands, and mineral soil wetlands among others. These ecosystems
+represent a range of moisture conditions that could regulate the abundance and
+activity of methane oxidizers and access to methane and oxygen, which methane
+oxidizers rely on. While this captured substantial variation in methane
+oxidation rates, sampling from such diverse ecosystems could result in spurious
+associations between taxa and function. For example, taxa differentially
+abundant in upland ecosystems that are unrelated to methane oxidation might
+appear correlated simply due to their presence in those ecosystems with high
+oxidation rates. Future studies could try restricting their search to a more
+homogeneous population of ecosystems specific to the question at hand.
+
+## Conclusion
+
+Microbial biodiversity-ecosystem function research has demonstrated positive
+correlations between diversity and ecosystem function. However, the abundances
+of microbial functional groups (as currently defined) are often not good
+predictors of ecosystem function and commonly do not add substantial
+explanatory power to ecosystem models. Therefore, a new perspective on how to
+determine the relationship between microbial communities and ecosystem
+functions is sorely needed. Organismal biologists have over a hundred years of
+experience identifying relationships between complex sets of highly-variable
+data (genotypes or genome sequences) and aggregate functions (organismal
+phenotypes). We assert that combining the approaches of traditional
+biodiversity-ecosystem function research with ideas from genotype-phenotype
+mapping could provide this new perspective. This integration could not only
+make underutilized approaches such as covariate modeling and artificial
+selection more available to microbial ecologists, but also provide instructive
+examples of how best to conceive of microbial biodiversity-ecosystem function
+questions. If this integration is successful, it is possible that in the
+not-so-distant future our field will be able to robustly identify taxa, genes,
+or even molecules that will allow us to accurately predict the response of
+ecosystems to environmental change. Doing so will not only generate novel
+hypotheses about how complex microbial communities drive ecosystem function,
+but also help scientists predict and manage changes to ecosystem functions
+resulting from human activities.
 
@@
 and Shell Gabon provided financial and logistical support for the original
-study re-analyzed here. We also thank W. Cresko for insightful discussions on
+study re-analyzed here. We also thank W. Cresko for insightful discussions regarding
 population genetics and H. Tavalire for contributions to the statistical

### 2019-12-10 manuscript/response_rev1.Rmd
Update revision and reviewer response

---
output: pdf_document
---

Response to Reviewers

Dear Dr. Martiny and Reviewers,

Thank you so much for your thoughtful and helpful reviews.

We have revised the manuscript to reflect your comments and have provided
specific responses to your comments below.

>Referee: 1
>
>Comments to Author(s) The manuscript ``Linking microbial communities to
>ecosystem functions: what we can learn from genotype-phenotype mapping in
>organisms'' by Morris et al proposes that integrating traditional
>biodiversity-ecosystem function research with genotype-phenotype mapping
>approach and suggests this  could provide the new perspective microbial BEF
>studies.
>
>I believe paper has a value in terms of expanding our tools to validate data
>from microbial BEF -particularly, identifying potential key taxa.  I have two
>major concerns;
>
>1. The paper set out with the projection that current approaches of testing
>   microbial BEF are not good enough. I do not see evidence how this approach
>   is better than on any existing approach (which link microbial diversity,
>   community composition with ecosystem functions). The manuscript does provide
>   an additional tool to potential identity of key taxa which are involved in
>   specific functions. Introduction should reflect that.

The intention of this paper is not to show that previous approaches are bad or
that this approach is “better.” Instead, the goal of this paper is to lay out
how this question has been addressed in the past, present a new way of framing
the question, and then demonstrate one example of how that new framing might be
used. We have revised throughout the paper to try to reflect that goal more
accurately. In particular, we have toned down the critique of past BEF work and
made the introduction clearer about the fact that we have learned a lot from
BEF research and that this is a new way of thinking about the problem.

>2. Paper has not done justice to recent progress (since 2015), we made in
>   demonstrating microbial BEF from field to global scales. Few examples
>   provided below 
>
>   Jing et al (2015). Nature Communications 
>
>   Delgado-Baquerizo et al., (2016) Nature Communication 
>
>   Delgado-Baquerizo et al., (2017) New Phytologist 
>
>   There are several more!

We have incorporated these references into the brief review of BEF research and
our review spans the literature from 2000 to 2018.

>Other comments
>1. L26-36. There have been number of recent papers published on this topic
>   which provide evidence of microbial BEF. Please see above.  
We have edited the text in this section to include these papers.
>2. L79-81. But there are other papers which provide support for microbial BEF. 
>3. L91-95. Here and several times later on- authors used one paper to suggest
>   that either there is no or extremely limited evidence of microbial BEF.
>   However, authors should  balance that there are multiple papers which
>   provided strong evidence in support of microbial BEF in field conditions
Here and elsewhere we have made it clear what aspect of the BEF literature
that we are critiquing and we have separated the critique from the review of
literature in support of microbial BEF.
4. L290-292. Can phage treatment be useful here?  Phage treatment is a possible
   approach and we have added it with a citation.
5. Reference list needs be carefully formatted to meet the criteria of the
   journalisms.

We have added the papers above and emphasized the support for microbial BEF,
particularly in the first paragraph of the Intro and in section 1.1. We have
also softened the language surrounding what we don’t know to more accurately
capture the state of the literature. We have also been more specific about what
questions remain open, particularly in section 1.2. We have updated the
citation format to match the Phil Trans B style.

Referee: 2

Comments to Author(s) Overview: For decades, ecologists have been developing
theoretical and empirical approaches for understanding the contribution of
species to ecosystem function. While advances have been made, relationships can
still be hard to detect, especially in microbial communities where there are
many taxa that are interacting and contributing in complex ways to different
functions. The premise of the Morris et al. manuscript is that ideas and
approaches borrowed from population genomics can be used to better understand
contributions of taxa to ecosystem function. The authors argue that that
biodiversity ecosystem functioning (BEF) relationships in microbial systems
have shown promise, but that they are often clouded by covariation between
environmental filters and ecosystem function as well as correction levels that
do not match the data. Specifically, methods from genotype-phenotype mapping
provide three insights, “agnostic” mapping, correction for multiple comparison,
and the ability to account for covariation between variables such as
environmental factors, geography, and community composition. The authors use an
example of methane oxidation in soil cores where previous analysis showed no
link between microbial community structure and function. After correcting for
multiple comparisons and accounting for covariation, their approach identified
six uniquely associated taxa. The authors conclude that genotype-phenotype
mapping can improve our ability to link microbial taxa with ecosystem functions
independent of environmental filters and other factors that covary with
community structure.

Major concerns:

-- The premise of using genotype-phenotype mapping to understand BEF is
intriguing. So, I was excited to read this manuscript. A major concern is that
I thought that authors spent too much space in the relatively short manuscript
focusing on the shortcomings of BEF research (sections in Abstracts and 1.1 –
1.3) relative to what I would have liked to have learned about
genotype-phenotype mapping (1.4). The latter needs to be better developed in
order for the manuscript to have its desired effect, which I assume is to help
researchers working on BEF to incorporate genotype-phenotype mapping approaches
to their research. I’m afraid that in the current version, the manuscript falls
short of providing the necessary background and justification.

We have shortened the review of the BEF literature and have tried to be more
clear and concise about what aspect of BEF we are critiquing, insofar as it
helps develop the mapping framework. We have also tried to unpack some of the
analogies that we use and provide examples of what they might like like in
microbial communities. See, for example, the new text in sections 2.1 and 2.2.

-- I sort of like the section advocating for an “agnostic” approach given that
we often don’t know a priori which taxa are contributing directly or indirectly
to a function. However, I was more skeptical about attributing multiple
comparisons to a genotype-phenotype approach. For example, it is well known
that Bonferroni corrections are severe and alternatives have been in place for
decades.  For example, without correction for false discovery rate (FDR), the
author report that 460 taxa contributed to rates of methane oxidation. In my
experience, all FDR-correction is doing is removing variables (in this case,
taxa) that have p-values above some adjusted threshold that is determined by
the total number of variables in a model.

The 460 taxa were significant after FDR correction, not before. we have updated
the text to make this clearer. Indeed, FDR correction sets a threshold for
significance that is less conservative than Bonferroni correction, but
ultimately it is just a significance threshold. It is true that correcting for
multiple comparisons does not come from population genetics. We have removed
the idea of correcting for multiple comparisons from the three “take-home”
points and have focused our text on describing the other two points. We have
also tried to make it clear that the main take home lesson from
genotype-phenotype mapping is not any particular statistical approach, but
rather a general way of thinking about these statistical relationships.

-- The section on population stratification starting on line 152 is
interesting. This doesn’t seem to be something that microbial ecologists often
do. I wonder if it would be helpful to expand upon this some more and perhaps
even show more explicitly how such spurious correlations could contribute
inferences either using the case study (methane oxidation) or another cartoon
type example, perhaps as a separate Box.

We have provided some examples of spurious correlations (using the methane
oxidation example) and have expanded on the conceptual ideas behind
stratification.

-- The section on high affinity methane oxidation needs work. I have no problem
with the case study per se, but I think it needs to be reworked if it is mean
to serve as a general example of a pipeline for how one might use
genotype-phenotype mapping to explore BEF. The second and third paragraphs of
this section don’t particularly seem aligned with the objective of the
manuscript.

Much of this section has been reworked and shaped into a Results & Discussion
section to interpret each step in the process and to better illustrate how and
why to perform each part of the analysis.

-- In the end, as someone who has never done genotype-phenotype mapping (but as
someone who has thought about BEF relationships from a statistical perspective)
I didn’t really come away feeling like a I had a deeper understanding or
appreciation for how I might use the authors’ recommended approach, but I think
with some more effort and revision, this could be achieved. Since this seems
like a fairly novel idea and one that spans disciplines, I would encourage the
authors to think about how the paper might serve in part as a primer that would
help readers understand the conceptual and statistical analogies between
genotype-phenotype mapping and BEF research.

This has been addressed by many of the other comments, particularly in
expanding on the analogy and how some of the conceptual components might apply
to microbial communities. We have also tried to make the purpose of the example
clearer and interpret the results more thoroughly. Finally, we have added a
caveats/future directions section that provides some suggestions of what other
changes could be made to improve on this approach in future studies.

-- Thought based on and around text on lines 104. Examples of phenotypes like
height are classic example of a quantitative trait that is influenced by many
genes. Could statistical approaches from quantitative genetics be used in a BEF
context?

The way we think about this is that genotype-phenotype mapping is the purview
of population genetics. Quantitative genetics is simply a subset of population
genenetics that deals with continuous rather than discrete traits. Ecosystem
function is more in line with quantitative genetics, because we’re dealing with
quantitative traits of ecosystems (e.g, rates of fluxes of molecules), but
ultimately the statistical approaches are the same. We have used “population
genetics” throughout in order to consistently use one term and so the ideas
apply more generally.

Minor Comments

-- Line 40: Environmental DNA is introduced here without a description of the
type of sequencing data being discussed, which may be critical to understanding
the information gathered from the DNA, be it phylogenetic or functional-gene
markers. Later, in line 208, 16S sequences are discussed as a way to identify
organisms that does not provide functional information. Even later, in line
302, it is clarified that functional group abundance can be measured via
genetic markers from environmental DNA.

We revised line 40 to specifically describe the type of sequencing data (16S
sequences or metagenomic sequences or qPCR of functional genes). We revised
line 302 to specifically refer to qPCR of functional genes. In all cases, the
phrase “environmental DNA” has been replaced with the specific type of
sequencing data being discussed. Sections 1.2 and 2.1 expand on the idea behind
functional genes.

-- Lines 140-167: Consider splitting Section 1.6 into two sections, one on
multiple comparisons and one on covariation, to stay consistent with the
expected discussion of three insights as identified in the abstract.

We removed the multiple comparisons section as per above comments and created a
separate section regarding covariation.

-- Lines 150-151: Consider expanding upon commonly used corrections in GWAS
studies. An additional sentence of two about a couple common corrections may
make this section stand on its own and allow the methods to be clearer to the
reader.

We added a few sentences elaborating on PC correction and variance component
modeling. Also, we added text regarding the design of the experiment around a
homogeneous population, which is considered a good practice in GWAS.

-- Lines 203-204: An explanation of the importance of site ID and ecosystem
type in structuring communities would aid the reader in understanding what
Figure 2 adds to the example. Is Figure 2 included to show the reader that the
environmental variation, community structure and geographic distance structure
communities? In general, it wasn’t clear to me that Figures 1 and 2 were
critical or necessary to include.

We added more interpretation of Figure 2.

-- Lines 264-267: The proposed method for identifying taxa of small effect
could benefit from a clearer set-up. The idea is overshadowed by the discussion
of detecting loci of small effect in genotype-phenotype studies and needs a
stronger connection to microbial biodiversity ecosystem function relationships
to stand on its own.

We moved this up to section 2 and made it clearer how this would work in BEF
studies.

-- Lines 256-259: The idea of “uniquely associated” taxa should be introduced
earlier in the text, perhaps section 1.4. Without explaining the change in
question from associated to uniquely associated, the example loses some of its
power and the importance of genotype-phenotype mapping methods is undersold.

We added discussion of uniquely associated to the section on controlling for
population structure.

-- Lines 50-52: consider revising for clarity

This text has been rewritten and moved to the “Population Stratification”
section

-- Line 57. Typo at end of topic sentence.

The comma has been corrected to a period.

-- Line 262-263: Reword sentence for clarity

This text has been revised.

-- Line 235: Other statistical approaches used for understanding causal
relationships include structural equation modeling (SEM) and path analysis.
These techniques also provide opportunity to identify latent variables, i.e.,
variables that are likely important but are difficult to measure. Might such
approaches be mentioned, and could they be used in conjunction with findings
from genotype-phenotype mapping? For example, it would be impossible to build
an SEM with 406 taxa and make any sense of what was going on, but with a
reduced set of taxa (e.g., 6), a SEM becomes very tractable.

We have added some discussion of how to use the identified taxa including a
form of microbial polygenic risk scores and including the taxa in a structural
equation model (Section 3.3).

-- To what degree to microbial BEF studies suffer from the fact that we are all
using relative abundance data? This is an obstacle that genotype-phenotype
mapping doesn’t have to deal with, correct?

We added text describing this to the discussion.

-- No hyphen needed in “microbially-mediated”

We have changed all instances of “microbially-mediated” to “microbially
mediated.”

### 2019-12-10 manuscript/manuscript_rev1.Rmd
Update revision and reviewer response

@@
 opts_chunk$set(echo = FALSE) 
-
 ```
@@
 composition is important for determining the rates of ecosystem-scale
-functions. There are numerous examples where microbial diversity is correlated
-with the rate of an ecosystem function, but exactly how microbial communities
-“map” onto ecosystem functions remains unresolved.  This uncertainty limits our
+functions. There are many examples of positive correlations between microbial
+diversity and ecosystem function, but how microbial communities “map” onto
+ecosystem functions remains unresolved. This uncertainty limits our
 ability to predict and manage crucial microbially mediated processes such as
-nutrient loss and greenhouse gas emissions. Our lack of understanding stems
-from the relatively large diversity of microorganisms, the difficulty in
-directly identifying functional groups, and our limited ability to manipulate
-microbial community attributes. For this reason, we propose that integrating
-traditional biodiversity-ecosystem function research with ideas from
-genotype-phenotype mapping could provide the new perspective our discipline
-needs. We identify two insights from genotype-phenotype mapping that could be
-useful for microbial biodiversity-ecosystem function studies: the concept of
-“agnostic” mapping and controlling for population stratification to identify
-microorganisms that uniquely contribute to ecosystem function. We illustrate
-the potential for these approaches to elucidate microbial
-biodiversity-ecosystem function relationships by analyzing a subset of
-published data measuring methane oxidation rates from tropical soils. We assert
-that combining the approaches of traditional biodiversity-ecosystem function
-research with ideas from genotype-phenotype mapping will not only generate
+nutrient losses and greenhouse gas emissions. To overcome this challenge, we
+propose integrating traditional biodiversity-ecosystem function research with
+ideas from genotype-phenotype mapping in organisms. We identify two insights
+from genotype-phenotype mapping that could be useful for microbial
+biodiversity-ecosystem function studies: the concept of searching
+“agnostically” for markers of ecosystem function and controlling for population
+stratification to identify microorganisms uniquely associated with ecosystem
+function. We illustrate the potential for these approaches to elucidate
+microbial biodiversity-ecosystem function relationships by analyzing a subset
+of published data measuring methane oxidation rates from tropical soils. We
+assert that combining the approaches of traditional biodiversity-ecosystem
+function research with ideas from genotype-phenotype mapping will generate
 novel hypotheses about how complex microbial communities drive ecosystem
-function, but also help scientists predict and manage changes to ecosystem
-functions resulting from human activities.
+function and help scientists predict and manage changes to ecosystem functions
+resulting from human activities.
+
+# Keywords
+
+Microbial ecology; Biodiversity-ecosystem function; Structure-function relationships;
+Microbial processes; Microbial community composition; Biogeochemical cycles
 
@@
 understanding the relationship between microbial biodiversity and ecosystem
-function is important [@schimel1998, @singh2010]. Certain ecosystem functions
+function is important [@schimel1998; @singh2010]. Certain ecosystem functions
 are correlated with microbial diversity, indicating that we should be able to
 determine what aspects of microbial biodiversity influence ecosystem function.
-However, attempts to describe that mapping have borne little fruit [@rocca2015,
+However, attempts to describe that mapping have borne little fruit [@rocca2015;
 @graham2016]. We argue that to overcome this challenge we should look to other
@@
 diversity and ecosystem multifunctionality for a variety of ecosystems and most
-major lineages of microorganisms [@jing2015, @delgado-baquerizo2016,
+major lineages of microorganisms [@jing2015; @delgado-baquerizo2016;
 @delgado-baquerizo2017]. Adding microbial diversity or abundance to ecosystem
@@
 dilution or varying filter size to generate differences in diversity can also
-exhibit differences in ecosystem function [@philippot2013, @wagg2014, @schnyder2018,
+exhibit differences in ecosystem function [@philippot2013; @wagg2014; @schnyder2018;
 @maron2018].  Finally, reciprocal transplant and common garden experiments that
@@
 differences in ecosystem function for communities of distinct
-origins [@cavigelli2000, @strickland2009, @glassman2018]. Given these
+origins [@cavigelli2000; @strickland2009; @glassman2018]. Given these
 relationships, many investigators have now moved on to the challenge of
@@
 marker for a microbial process (and thus a marker for the functional group that
-performs this process). For example, the gene _mcrA_, which encodes a subunit of
+performs that process). For example, the gene _mcrA_, which encodes a subunit of
 the enzyme that performs the final step in methanogenesis, is commonly used as
 a marker for methanogenesis and for the methanogen functional group. Other
-examples include _pmoA_ and methanotrophy, _nifH_ and nitrification, and _nosZ_ and
-denitrification. The abundance of these markers is often hypothesized to be
-predictive of the rate of their associated processes (for example, the
-abundance of _mcrA_ is often hypothesized to be related to the rate of
+examples include _pmoA_ and methanotrophy, _nifH_ and nitrification, and _nosZ_
+and denitrification. It is often hypothesized that the abundance of these
+markers is predictive of the rate of the associated processes (for example,
+it is hypothesized that the abundance of _mcrA_ is related to the rate of
 methanogenesis).
@@
 circumstances be predicted from the genetic markers _mcrA_ and _pmoA_
-[@freitag2009, @freitag2010, @mccalley2014]. However, for most ecosystem
+[@freitag2009; @freitag2010; @mccalley2014]. However, for most ecosystem
 functions the abundance of a functional gene or transcript is rarely positively
@@
 cases where there is a positive correlation tend to be restricted to
-agricultural ecosystems and for certain functions within the nitrogen cycle
+agricultural ecosystems and certain functions within the nitrogen cycle
 [@rocca2015]. In general, including aspects of microbial biodiversity
-(e.g., functional gene abundance or diversity) improve models of ecosystem
-function less than one third of the time, and increase variance explained by an
+(e.g., functional gene abundance or diversity) improves models of ecosystem
+function less than one third of the time and increases variance explained by an
 average of only 8 percentage points over environmental variables [@graham2016].
@@
 diseases such as Parkinson’s) are controlled by a single genetic locus
-[@kerem1989, @macdonald1992], most traits depend on a large number of genes
-that control variation in that phenotype [@hindorff2009, reich2001]. In
+[@kerem1989; @macdonald1992], most traits depend on a large number of genes
+that control variation in phenotype [@hindorff2009; @reich2001]. In
 addition, there is often no a priori expectation about which regions of the
@@
 causal relationships that are potentially confounded by complex patterns of
-covariation. There is often no strong expectations about which entities (i.e.,
+covariation. There is often no strong expectation about which entities (i.e.,
 which genomic regions or which microbial genes or lineages) are most likely to
-be causally-related to phenotype or function, and thus “agnostic” approaches
+be causally related to phenotype or function, and thus “agnostic” approaches
 are needed. For some ecosystem functions, it is possible that a single
@@
 uncultured and we do not know the function of most microbial taxa detected in
-environmental samples [@hug2016, @martiny2019]. In addition, there have
+environmental samples [@hug2016; @martiny2019]. In addition, there have
 been recent discoveries of functions in unexpected taxonomic groups, for
 example methanogenesis by fungi and cyanobacteria, a function previously
-considered restricted to the Archaea [@bizic-ionescu2019, @lenhart2012].
+considered restricted to the archaea [@bizic-ionescu2019; @lenhart2012].
 
@@
 function. Because of this, it might be prudent to look more agnostically at
-microbial communities to identify taxa, groups of taxa or genes that are
+microbial communities to identify taxa, groups of taxa, or genes that are
 important for predicting the rates of ecosystem functions rather than assuming
@@
 
-Beyond finding new physiologies in unexpected places, there are other reasons
+Beyond finding new physiologies in unexpected lineages, there are other reasons
 for looking agnostically. In the case of microbial functions, it may be that
@@
 could also be limited by the bacteria that produce the fermentative byproducts
-that methanogens depend on. Or there could be indirect limitation by organisms
+that methanogens use as substrates. Or there could be indirect limitation by organisms
 that liberate nitrogen or phosphorus into mineral forms. In other words, the
-influence of microbial communities on the rate ecosystem function could
+influence of microbial communities on the rate of ecosystem function could
 represent a complex metabolic network much like the regulation of gene
-expression in organisms that partially determine their phenotype. These broader
+expression in organisms that partially determines their phenotype. These broader
 patterns of biodiversity-ecosystem function relationships would be invisible to
-any study that focuses on putative functional groups without considering the
-reads that underlie those functional groups.
+any study that solely focuses on the most relevant functional group without
+considering the possible influence of other taxa.
 
@@
 population stratification due to geographic and environmental separation
-[@wright1943, @martiny2006]. This can lead to spurious associations between
+[@wright1943; @martiny2006]. This can lead to spurious associations between
 phenotypes and genetic markers that are at high frequency in isolated
@@
 population stratification (i.e., community similarity among ecosystems),
-although there are some exceptions [@meyer2019, @qin2012, @lloyd-price2019].
+although there are some exceptions [@meyer2019; @qin2012; @lloyd-price2019].
 Community similarity (the community analogue of shared ancestry among
 organisms) is not as tightly linked to geography or environment as is shared
-ancestry, and it could be very useful to account for these separately in
-microbial studies, especially if one is particularly interested in how
-composition alters function independent of the underlying environmental
-variation.
+ancestry. Therefore, it could be useful to account for these separately in
+microbial studies, particularly if one is interested in quantifying the effect
+of microbial communities on ecosystem function independent of these other factors.
 
@@
 of phenotype and shared ancestry (computed as the first one or more principal
-components of the genotype matrix) and then use the residuals from this model
+components of a genotype matrix) and then use the residuals from this model
 as the values for phenotype in a subsequent regression using the genotypes
@@
 flux. However, it might also be that that organism lives only in ecosystems
-that happen to have a high rate of ecosystem function. We would not be able to
-distinguish these outcomes with this simple correlation. If we then include
-environmental variables or environmental similarity as a covariate, we could
-start to ask, “Is the taxon uniquely associated with function independent of
-the environment?” By “uniquely associated”, we mean those taxa associated with
-function irrespective of environmental conditions, local community structure,
-or spatial proximity. This is not only a more specific question than is usually
-asked in microbial biodiversity-ecosystem function studies, but it is also a
-more appropriate one, especially if one is interested in how to incorporate
-microbial community data into ecosystem models. Finally, it is interesting on
-its own to understand whether microbial communities are selected by the
-underlying environmental conditions to produce a particular rate of ecosystem
-function (the conduit) or whether there is variation in community composition
-independent of the environment that explains some variation in ecosystem
-function.
+that happen to have a high rate of methane flux. In this scenario, we are
+unable to distinguish these possibilities. However, if we add environmental
+variables or environmental similarity as a covariate in our model, we can ask,
+“Is this taxon uniquely associated with function in a way that it is
+independent of the environment?” By “uniquely associated,” we mean those taxa
+associated with function irrespective of environmental conditions, local
+community structure, or spatial proximity. This slight reframing of the
+question could be especially rewarding for microbial biodiversity-ecosystem
+function research, particularly as it relates to incorporating microbial
+community data into ecosystem models.  Finally, it is interesting in its own right to
+understand whether microbial communities are selected by the underlying
+environmental conditions to produce a particular rate of ecosystem function or
+whether community variation has functional consequences independent of the
+environment.
 
@@
 for applying genotype-phenotype mapping to microbial communities [@meyer2019]. In our
-reanalysis, we do not intend to challenge the conclusions from that paper, but
+reanalysis, we do not intend to challenge the conclusions of that paper, but
 instead we want to demonstrate how to perform this type of study for microbial
@@
 under different concentrations of methane to identify the rates of specific
-methane cycling pathways. For this example, we will analyze data from just one
+methane cycling pathways. In this example, we analyze data from just one
 of these pathways, high-affinity methane oxidation (the oxidation of
 atmospheric concentrations of methane), which we will refer to simply as
-“methane oxidation”. In addition, we only include amplicon sequences from the
+“methane oxidation.” In addition, we only include amplicon sequences from the
 DNA-inferred community and not the RNA-inferred community, both of which are
-presented in the original paper (@meyer2019). The data we analyzed include methane
+presented in the original paper [@meyer2019]. The data we analyzed include methane
 oxidation rates, amplicon sequence variants (ASVs; @callahan2017) generated using DADA2
 and inferred from 16S rRNA gene sequences, _pmoA_ abundance estimates (via qPCR),
-latitude and longitude, and four environmental covariates (soil moisture, bulk
+and four environmental covariates (soil moisture, bulk
 density, carbon, and nitrogen).
@@
 Analyses were conducted in the R statistical environment using the phyloseq
-package [@mcmurdie2013, @rcoreteam2019]. The relative abundances of ASVs were corrected using the variance
-stabilizing transformation from DESeq2 [@love2014,
-@mcmurdie2014]. We first test typical measures of
-microbial structure including functional gene abundance and community richness,
-which was estimated using the breakaway package [@willis2015]. We then demonstrate
-significant covariation between community structure (estimated as Bray-Curtis
-distance using vegan), environmental variation (euclidean distance), and
-geographic distance (euclidean distance) using Mantel tests [@bray1957, @oksanen2019]. Finally, we
+package [@mcmurdie2013; @rcoreteam2019]. The relative abundances of ASVs were
+corrected using the variance stabilizing transformation from DESeq2 [@love2014;
+@mcmurdie2014]. We first test typical measures of microbial community structure including
+functional gene abundance and community richness, which was estimated using the
+breakaway package [@willis2015]. We then demonstrate significant covariation
+between community structure (estimated as Bray-Curtis distance using vegan),
+environmental variation (euclidean distance), and geographic distance
+(euclidean distance) using Mantel tests [@bray1957; @oksanen2019]. Finally, we
 present one approach to identifying taxa which are significantly associated
@@
 models using varComp to test the relationship between relative abundance of
-each ASV and methane oxidation rate [@kang2010, @qu2013]. To illustrate how
-including different
-covariates (environmental, geographic, and community) can result in different
-conclusions about which taxa are associated with function, we fit this model
-with and without random effects variance components for environmental
-similarity, geographic site ID, and Bray-Curtis similarity. Significant taxa
-were determined by controlling the false discovery rate at q-value < 0.05 [@storey2002].
-Figures were created using ggplot2 [@wickham2019].
+each ASV and methane oxidation rate [@kang2010; @qu2013]. To illustrate how
+including different covariates (environmental, geographic, and community) can
+result in different conclusions about which taxa are associated with function,
+we fit this model with and without random effects variance components for
+environmental similarity, geographic site ID, and Bray-Curtis similarity.
+Significant taxa were determined by controlling the false discovery rate at
+q-value < 0.05 [@storey2002].  Figures were created using ggplot2
+[@wickham2019].
 
-### Results & Discussion
+### Results & discussion
 
@@
 group abundance or community richness as it relates to ecosystem function. In
-our case, methane oxidation rate was not significantly correlated with the pmoA
-gene abundance or 16S read richness (Table 1, Figure 1). To demonstrate that
+our case, methane oxidation rate was not significantly correlated with _pmoA_
+gene abundance or 16S rRNA gene-based taxonomic richness (Table 1, Figure 1).
+To demonstrate that
 the covariance structure of the data might alter our conclusions about which
@@
 indicates substantial spatial and environmental structuring of microbial
-populations. This means that the presence or abundance of certain genetic
-markers (taxa) will be elevated in specific ecosystems. In this case,
+populations. This suggests that the presence or abundance of certain taxa will
+be elevated in specific ecosystems. In this case,
 high-affinity methane oxidation is typically greater in upland ecosystems than
 in wetland ecosystems and so any taxa differentially abundant in uplands will
-be correlated with methane oxidation regardless of their involvement in that
-process. It is necessary to control for this structuring to rigorously identify
+tend to be correlated with methane oxidation regardless of their involvement in that
+process. It is necessary to control for this stratification to rigorously identify
 associations between taxa and function.
@@
 covariates including environment, geography, and community. After controlling
-the false discovery rate, 460 unique 16S reads were identified as significantly
+the false discovery rate, 460 unique ASVs were identified as significantly
 correlated with function when no covariates were included in the model. We
@@
 these sets of taxa represent different versions of the biodiversity-ecosystem
-function mapping question. For example, controlling for environmental variation
-tests for taxa that are related to function independent of environmental
-conditions. Similarly, controlling for geographic distance tests for taxa that
-are related to function independent of geographic proximity; geographic
-proximity could influence function via unmeasured environmental variation that
-is spatially structured or differences in dispersal history among sites due to
-physical distance. In the model that controlled for all three covariates
-(community, environment, and geography), only six ASVs were significantly
-correlated with methane oxidation rate (Figure 3). These taxa could be useful
-indicators for methane oxidation rate across space and different ecosystems. In
-addition, researchers could manipulate their abundance in order to better
-understand the biological drivers of variation in methane oxidation.
+function mapping question. For example, by attempting to control for
+environmental variation statistically, we can identify taxa whose traits may
+contribute to variation in function that is independent of environmental
+conditions. Similarly, by controlling for geographic distance among samples we
+can reduce the likelihood that the taxa we identify are only related to
+function because of an association with unmeasured environmental variation that
+is spatially structured or because of differences in dispersal history among sites. In the
+model that controlled for all three covariates (community, environment, and
+geography), only six ASVs were significantly correlated with methane oxidation
+rate (Figure 3). These taxa could be useful indicators of methane oxidation
+rate across space and different ecosystems. Researchers could elaborate on
+these findings using targeted cultivation and manipulative experiments to
+further understand their contribution to methane oxidation.
 
 Notably, these six taxa fall into three genera and one class with cultured
-representatives that are not known to consume methane [@belova2009, @chang2011,
-@domeignoz-horta2019, @fritz2004]. These taxa could be
+representatives that are not known to consume methane [@belova2009; @chang2011;
+@domeignoz-horta2019; @fritz2004]. These taxa could be
 related to ecosystem function in multiple ways. The most interesting
@@
 
-This approach is powerful because it accounts for the covariances that arise
-from the fundamental ecological processes that drive community assembly and
-underlie biogeographical patterns [@martiny2006, @vellend2010]. It is also
-powerful because it is
-focused on a much more specific version of the “biodiversity-ecosystem
-function” question, i.e., it asks “which taxa are uniquely associated with
-function?” By “uniquely associated,” we mean those taxa associated with
-function irrespective of environmental conditions, local community structure,
-or spatial proximity. This is not only a more specific question than is usually
-asked in microbial biodiversity-ecosystem function studies, but it is also a
-more appropriate one, especially if one is interested in how to incorporate
-microbial community data into ecosystem models. These models depend on
-identifying taxa that add explanatory power beyond that provided by other
-factors, such as environmental conditions. Earth system models would benefit
-most from microbial bioindicators that capture variation unexplained by other,
-more easily measured variables (such as certain environmental conditions). The
-approach we describe is an important step toward identifying such biomarkers.
-
-## Caveats and future directions
+## Caveats & future directions
 
@@
 similar aggregate bioindicator could be generated for ecosystems that would
-summarize the probability of having a particular ecosystem function or rate.
-This would be accomplished by measuring the abundance of the taxa identified in
-an association study and determining their association with the rate of an
-ecosystem function. Alternatively, the identified taxa could be incorporated
-into a structural equation model in an attempt to better understand the
-individual effects and interactions among taxa as they contribute to the rate
-of ecosystem function. This might give an indication of the relative importance
-of different taxa as compared to other factors, such as environmental
-variables.
+summarize the probability of the rate or occurrence of a particular ecosystem
+function.  This would be accomplished by measuring the abundance of the taxa
+identified in an association study and determining their association with the
+rate of an ecosystem function. Alternatively, the identified taxa could be
+incorporated into a structural equation model in an attempt to better
+understand the individual effects and interactions among taxa as they
+contribute to the rate of ecosystem function [@grace2010]. This might give an
+indication of the relative importance of different taxa as compared to other
+factors, such as environmental variables, and also identify underlying latent
+variables that explain variation in ecosystem function.
 
-Ultimately, the relationships identified in any comparative mapping study
-must be verified. For genotype-phenotype studies in organisms there are
-multiple ways that this verification is
-accomplished. In some cases, organisms can be artificially selected for a
-particular phenotype (e.g., through experimental evolution in an environment
-that favors the phenotype of interest) and the genetic changes that occur in
-response to selection can be compared to those identified via mapping studies
-(such as GWAS). An analogous approach for microbial biodiversity-ecosystem
-function studies would be to apply artificial ecosystem selection (sensu @swenson2000)
-on a given function and compare the taxa that change in response to selection
-with those identified via a comparative approach (such as the one illustrated
-in our example).
+Ultimately, the relationships identified in any comparative mapping study must
+be verified. For genotype-phenotype studies in organisms there are multiple
+ways that this verification is accomplished. In some cases, organisms can be
+artificially selected for a particular phenotype (e.g., through experimental
+evolution in an environment that favors the phenotype of interest) and the
+genetic changes that occur in response to selection can be compared to those
+identified via mapping studies. An analogous approach for
+microbial biodiversity-ecosystem function studies would be to apply artificial
+ecosystem selection (sensu @swenson2000) on a given function and compare the
+taxa (or genes) that change in response to selection with those identified via
+a comparative approach (such as the one illustrated in our example).
 
-The most common way that loci identified in a genotype-phenotype mapping study are verified is
-through manipulative genetics. The identified loci can be knocked out or
-over-expressed and the effect on phenotype compared to that predicted from
-mapping studies. In the case of microbial biodiversity studies, it may be
-possible to inhibit a particular functional group through the use of a specific
-antimicrobial or a chemical inhibitor or by using phages with high
-host-specificity [@maxson2016], but this is not generally possible. In some cases, we
-may be able to isolate a microorganism of interest in pure culture and add it
-back to an ecosystem, transiently increasing its abundance (roughly analogous
-to “overexpressing” a gene). A greater focus on culture-based approaches could
-increase the success of these kinds of microbial enrichments. Finally,
-synthetic communities (contrived assemblages of microorganisms) may be the most
-powerful way to test hypotheses about microbial biodiversity-ecosystem function
-relationships, but currently these approaches are limited by the small number
-of taxa that can be routinely cultured from most environments.
+The most common way that loci identified in a genotype-phenotype mapping study
+are verified is through manipulative genetics. The identified loci can be
+knocked out or over-expressed and the effect on phenotype compared to that
+predicted from mapping studies. In the case of microbial biodiversity studies,
+it may be possible to inhibit a particular functional group through the use of
+specific antimicrobial or chemical inhibitors or by using phages that exhibit high
+host-specificity [@maxson2016; @koskella2013], but this is not generally
+possible. In some cases, we may be able to isolate a microorganism of interest
+in pure culture and add it back to an ecosystem, transiently increasing its
+abundance (roughly analogous to “overexpressing” a gene). A greater focus on
+culture-based approaches could increase the success of these kinds of microbial
+enrichments. Finally, synthetic communities (contrived assemblages of
+microorganisms) may be the most powerful way to test hypotheses about microbial
+biodiversity-ecosystem function relationships, but currently these approaches
+are limited by the small number of taxa that can be routinely cultured from
+most environments (but see @schnyder2018 and @berg2018).
 
@@
 mapping approach we describe, some in common with organismal mapping studies
-and others unique. For example,
-simple linear models such as the variance component model used in this study
-are typical for genetics studies, but may not be the best way to identify
-correlations for microbiomes due to the unique challenges of microbial data.
-These are zero-inflated compositional data with differences in sequencing
-depth across samples. We have addressed these challenges by using a variance
-stabilizing transformation, but other models that test differential
-abundance and differential variance that can control for differences in
-sequencing depth and are robust to zero-inflation might be more appropriate
-(e.g., CORNCOB @martin2019).  Alternatively, clustering reads at higher
-taxonomic levels could provide a more continuous variation in taxon abundances
-across ecosystems by reducing the number of “bins.” The tradeoff for this is
-that it introduces bias based on the completeness of our taxonomic databases
-and the accuracy of 16S-based taxonomic assignment and it removes reads that
-lack a taxonomic assignment. Here, we chose to test ASVs at the level of the
-individual read so as not to bias our results by losing reads that lack a
-taxonomic assignment at a particular level.
+and others unique. For example, simple linear models such as the variance
+component model used in this study are typical for genetics studies, but may
+not be the best way to identify correlations for microbiomes due to the unique
+challenges of microbial data.  Marker gene and metagenome sequences are
+inherently compositional, reads are often absent from most samples (i.e., they
+are zero-inflated), and differences in sequencing depth make it difficult to
+compare relative abundances across samples, challenges that are not faced by
+population geneticists.  We have addressed these challenges by using a variance
+stabilizing transformation, but other models that test differential abundance
+and differential variance that can control for differences in sequencing depth
+and are robust to zero-inflation might be more appropriate (e.g., @martin2019).
+Clustering reads at higher taxonomic levels could circumvent zero-inflation by
+providing more continuous variation in taxon abundances across ecosystems.
+However, this approach introduces biases based on the completeness of
+taxonomic databases, the accuracy of 16S-based taxonomic assignment, and
+the removal of reads that lack a taxonomic assignment (although, newer
+approaches to taxonomic classification might help [@shah2019]). Alternatively,
+decreasing the threshold of sequence similarity to cluster reads without
+taxonomy could be analogous to aggregating at higher taxonomic levels, but
+it is uncertain whether these larger aggregates of taxa have any
+trait conservatism related to function. Here, we chose to test ASVs at
+the level of the individual read so as not to bias our results in these ways.
+Finally, we have applied this approach to ASVs inferred from 16S rRNA
+gene sequences, but any unit of microbiome data such as metagenomic reads or
+metatranscriptomic mRNA reads could be tested in an association study.
 
 Experimentally, future studies could improve on our example by sampling a more
-homogeneous set of ecosystems. Our survey includes grasslands, pastures,
-forests, peatlands, and mineral soil wetlands among others. These ecosystems
-represent a range of moisture conditions that could regulate the abundance and
-activity of methane oxidizers and access to methane and oxygen, which methane
-oxidizers rely on. While this captured substantial variation in methane
-oxidation rates, sampling from such diverse ecosystems could result in spurious
-associations between taxa and function. For example, taxa differentially
-abundant in upland ecosystems that are unrelated to methane oxidation might
-appear correlated simply due to their presence in those ecosystems with high
-oxidation rates. Future studies could try restricting their search to a more
-homogeneous population of ecosystems specific to the question at hand.
+homogeneous set of ecosystems. Our survey includes an especially broad
+assortment of ecosystems, including grasslands, plantations, forests,
+peatlands, and mineral soil wetlands among others. These ecosystems represent a
+range of moisture conditions that could regulate the abundance and activity of
+methane oxidizers and access to methane and oxygen, which methane oxidizers
+rely on. While this captured substantial variation in methane oxidation rates,
+sampling from such diverse ecosystems could result in spurious associations
+between taxa and function. For example, taxa differentially abundant in upland
+ecosystems that are unrelated to methane oxidation might appear correlated
+simply due to their presence in those ecosystems with high oxidation rates.
+Future studies could try restricting their search to a more homogeneous
+population of ecosystems specific to the question at hand.
 
@@
 correlations between diversity and ecosystem function. However, the abundances
-of microbial functional groups (as currently defined) are often not good
+of microbial functional groups (as currently defined) are often poor
 predictors of ecosystem function and commonly do not add substantial
@@
     value = as_paragraph(as_i("pmoA"), " copy number")) %>% 
+  compose(i = 2, j = 1, 
+    value = as_paragraph("Richness")) %>% 
   compose(part = "header", j = 4, 
@@
 
-Table 1. Functional gene abundance and ASV richness are not good predictors of
-methane oxidation rate. Linear models predicting methane oxidation rate from
-measures of microbial community structure.
+Table 1. Estimates for the linear relationship between methane oxidation and
+two measures of microbial community structure: _pmoA_ functional gene abundance and
+ASV richness.
 
@@
 Table 3. Number of significant taxa after including each set of covariates in a
-variance component model. Removed and Added columns are relative to the
-no-covariate model. Significance is determined by controlling the false
+variance component model. "Removed" and "Added" columns are relative to the
+no-covariate model. Significance was determined by controlling the false
 discovery rate at q-values < 0.05.
 
-```{r methane, out.width="768px", out.height="364px"}
+```{r methane, out.width="768px", out.height="364px", include=FALSE}
 knitr::include_graphics('../figures/trad_cors.pdf')
@@
 abundance or ASV richness. Correlations between community attributes and
-ecosystem function. A) Abundance of the functional gene pmoA (n = 42) and B)
+ecosystem function. A) Abundance of the functional gene _pmoA_ (n = 42) and B)
 ASV richness (n = 44). Lines represent the ordinary least squares regression
@@
 
-```{r structure, out.width="960px", out.height="1152px"}
+```{r structure, out.width="960px", out.height="1152px", include=FALSE}
 knitr::include_graphics('../figures/pcoa_multi.pdf')
@@
 
-```{r taxa, out.width="768px", out.height="384px"}
+```{r taxa, out.width="768px", out.height="384px", include=FALSE}
 knitr::include_graphics('../figures/effect_sizes.pdf')
